### 2011-03-01 inst/doc/Foreword.tex
added demo and devel vignettes, version bumped to 0.1

\section*{Foreword}

\Rpackage{MSnbase} is in an early development %%(package version \Sexpr{packageVersion("MSnbase")}, 
(see section \ref{sec:sessionInfo} for details about packages and version used in this vignette). 
Although main data structures have been thought out and are meant to be compatible with other existing 
mature infrastructure in the Bioconductor project, changes may occur in the future. 
Current functionality will evolve and new ones will be added. 
Although at an early stage, this package is released in the hope that it may foster  
new developments in proteomics data analysis within \R by providing a common infrastructure. 
Several package developers related to mass spectrometry and proteomics data met during the 
Bioconductor Developer Meeting Europe\footnote{\url{http://bioconductor.org/help/course-materials/2010/HeidelbergNovember2010/}} 
held in Heidelberg in November 2010, and agreed to combine efforts. 
This library is one attempt to do so.

### 2011-03-01 inst/doc/MSnbase-demo.Rnw
added demo and devel vignettes, version bumped to 0.1

%\VignetteIndexEntry{Base Functions and Classes for MS-based Proteomics }
%\VignetteKeywords{Mass Spectrometry, MS, MSMS, Proteomics, Infrastructure, Bioinformatics, quantitative }
%\VignettePackage{MSnbase-demo}
\documentclass{article}
\usepackage{tikz}
\usepackage[nogin]{Sweave}
\usepackage{hyperref}
\usepackage[authoryear,round]{natbib}

\SweaveOpts{echo=T,eval=T,fig=F,results=verbatim,cache=F,tikz=T,external=F,prefix.string=sweave-cache/figs/fig}

\newcommand{\R}{\texttt{R} }
\newcommand{\Rfunction}[1]{{\texttt{#1}}}
\newcommand{\Robject}[1]{{\texttt{#1}}}
\newcommand{\Rpackage}[1]{{\textit{#1}}}
\newcommand{\email}[1]{\href{mailto:#1}{\normalfont\texttt{#1}}}
%% colors
\definecolor{Red}{rgb}{0.7,0,0}
\definecolor{Blue}{rgb}{0,0,0.8}

\hypersetup{%
  hyperindex = {true},
  colorlinks = {true},
  linktocpage = {true},
  plainpages = {false},
  linkcolor = {Blue},
  citecolor = {Blue},
  urlcolor = {Red},
  pdfstartview = {Fit},
  pdfpagemode = {UseOutlines},
  pdfview = {XYZ null null null}
}

\input{Author.tex}

\begin{document}

\title{\Rpackage{MSnbase}: labelled MS2 data pre-processing, visualisation and quantification.}

\maketitle

%% Abstract and keywords %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vskip 0.3in minus 0.1in
\hrule
\begin{abstract}
  This vignette describes the functionality implemented in the \Rpackage{MSnbase} package. 
  \Rpackage{MSnbase} aims at (1) facilitating the upload, processing, visualisation and 
  quantification of mass spectrometry data into the \R environment \citep{Rstat} by providing 
  specific data classes and methods and (2) enabling the utilisation of throughput-high data 
  analysis pipelines provided by the Bioconductor \citep{Gentleman2004} project. 
\end{abstract}
\textit{Keywords}: Mass Spectrometry (MS), proteomics, infrastructure, bioinformatics, quantitative.
\vskip 0.1in minus 0.05in
\hrule
\vskip 0.2in minus 0.1in
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\tableofcontents

%% <<setup,cache=FALSE>>=
%%  setCacheDir("sweave-cache/values")
%% @

<<environment,cache=FALSE,echo=FALSE>>=
library(MSnbase)
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{Foreword.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}\label{sec:intro}

\Rpackage{MSnbase} aims are providing a reproducible research framework to the proteomics data analysis. 
It should allow researcher to easily mine mass spectrometry data, explore the data and its statistical properties 
and visually display these.

\Rpackage{MSnbase} also aims at being compatible with the infrastructure implemented in Bioconductor, 
in particular \Rpackage{Biobase}. As such, classes 
developed specifically for proteomics mass spectrometry data are based on the \Robject{eSet} 
and \Robject{Expression} classes. The main goal is to assure seamless compatibility with 
existing meta data structure, accessor methods and normalisation techniques.

This vignette illustrates \Rpackage{MSnbase} utility using a dummy data sets provided with the package 
without describing the underlying data structures. More details can be found in the 
package, classes, method and function documentations. A description of the classes is provided 
in the \texttt{MSnbase-development} vignette.

\input{NoteAboutSpeedAndMemory.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Importing raw data}\label{sec:msnexp}

Raw data is stored in an \Robject{MSnExp} object, that stores all the spectra in an experiment, 
as defined by one or multiple raw data files.

\Rpackage{MSnbase} imports raw MS data stored in \texttt{mzXML} format \citep{Pedrioli2004}. 
More formats, notably \texttt{mzML} \citep{Martens2010}, will be added at a later stage. Either one of 
MS1 or MS2 spectra can be loaded at a time by setting the \texttt{msLevel} parameter accordingly.

<<MSnExp,cache=FALSE,echo=TRUE>>=
library(MSnbase)
filename <- dir(system.file(package="MSnbase",dir="extdata"),full.name=TRUE,pattern="mzXML$")
experiment <- readMzXMLData(filename, msLevel = 2, verbose=FALSE)
experiment
@

<<experiment-size,echo=FALSE,cache=FALSE>>=
sz <- sum(sapply(assayData(experiment),object.size)) + object.size(experiment)
sz <- round(sz/(1024^2),2)
@

As illustrated above, showing the experiment textually displays it's content: 
\begin{itemize}
  \item Information about the raw data, i.e. the spectra.
  \item Specific information about the experiment processing\footnote{%%
    this part will be automatically updated when the object is 
    modified with it's methods, as illustrated later} and package version.
    This slot can be accessed with the \Rfunction{processingData} method.
  \item Other meta data, including experimental phenotype, 
    file name(s) used to import the data, protocol data, 
    information about features (individual spectra here), 
    experiment data and proteomics data. Most of these are implemented as 
    in the \Robject{eSet} class and are described in more details in their 
    respective manual pages. See also \texttt{?pSet} and \texttt{MSnExp}
    and references therein for additional background information.

    The proteomics-specific meta data in slot \Robject{proteomicsData} is a 
    new addition from the \Rpackage{MSnbase} package.
    It contains MIAPE data (Minimum Information About a Proteomics Experiment, \citep{Taylor2007}) 
    and can be accessed with the \Rfunction{proteomicsData} method. 
    See \texttt{?MIAPE} for more details.
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Spectra objects}\label{sec:spectra}

The raw data is composed of the \Sexpr{length(experiment)} MS spectra. 
The spectra are named individually 
(\Sexpr{paste(paste(head(featureNames(experiment)),collapse=", "),", ...",sep="")})
and stored in a \Robject{environment}. They can be accessed individually with 
\texttt{experiment[["X10"]]} or as a list with \texttt{spectra(experiment)}.
As we have loaded our experiment specifying \texttt{msLevel=2}, 
the spectra will all be of level 2 (or higher, if available).

<<Spectrum,cache=FALSE,echo=TRUE>>=
sp <- experiment[["X66"]]
sp
@

Attributes of individual spectra or of all spectra of an experiment can be 
accessed with their respective methods: \Rfunction{precursorCharge} for the 
precursor charge, \Rfunction{rtime} for the retention time, \Rfunction{mz} 
for the MZ values, \Rfunction{intensity} for the inensities, \ldots see the 
respecitve class manuals for more details.

<<accessors,cache=FALSE,echo=TRUE>>=
peaksCount(sp)
head(peaksCount(experiment))
rtime(sp)
head(rtime(experiment))
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Reporter ions}\label{sec:reporterions}

Reporter ions are defined with the \Robject{ReporterIons} class. 
Specific peaks of interested are defined by a MZ value, a with 
around the expected MZ and a name (and optionally a colour for plotting, 
see section \ref{sec:plotting}). \Robject{ReporterIons} instances 
are required to quantitate reporter peaks in \Robject{MSnExp} 
experiments. Instances for the most commonly used isobaric tags like 
iTRAQ 4-plex and 8-plex and TMT tags are already defined in 
\Rpackage{MSnbase}. See \texttt{?ReporterIons} for details about how 
to generate new \Robject{ReporterIons} objects.

<<ReporterIons>>=
data(iTRAQ4)
iTRAQ4
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Plotting raw data}\label{sec:plotting}

Spectra can be plotted individually or as part of (subsetted) experiments 
with the \Robject{plot} method. Full spectra can be plotted 
(using \texttt{full=TRUE}), specific reporter ions of interest 
(by specifing with reporters with \texttt{reporters=iTRAQ4} for instance) 
or both (figure \ref{fig:spectrum-plot}).

\begin{figure}[!ht]
<<spectrum-plot,fig=T,echo=T,tikz=T,external=T,width=5,height=5>>=
plot(sp,reporters=iTRAQ4,full=TRUE)
@ 
\caption{Raw MS2 spectrum with detail about reporter ions.}
\label{fig:spectrum-plot}
\end{figure}

It is also possible to plot all spectra of an experiment (figure \ref{fig:msnexp-plot}). 
Lets start by extracting all spectra that have the sample precursor MZ value than \Robject{sp} 
into a separate experiment, using the \Rfunction{extractPrecSpectra} method. 
It takes an \Robject{MSnExp} experiment and a precursor MZ value as parameters.

<<extractPrecSpec,echo=T,cache=F>>=
exp2 <- extractPrecSpectra(experiment,precursorMz(experiment)["X66"])
exp2
@

We see that \Sexpr{length(exp2)} spectra have the same MZ value. There two spectra 
can be represented together by plotting the \Robject{MSnExp} object 
(figure \ref{fig:msnexp-plot}).

\begin{figure}[!ht]
<<msnexp-plot,fig=T,echo=T,tikz=T,external=T,width=5,height=6>>=
plot(exp2,reporters=iTRAQ4,full=FALSE)
@ 
\caption{Experiment-wide raw MS2 spectra.}
\label{fig:msnexp-plot}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Data processing}\label{sec:processing}

There are several methods implemented to perform basic data manipulation. 
Low intensity peaks can be set to 0 with the \Rfunction{removePeaks} method 
from spectra or whole experiments. The intensity below which peaks are removed 
is defined by the \Robject{t} parameter. \texttt{t} can be specified directly 
as a numeric. The default value is the character \texttt{"min"}, that will remove 
all peaks equal to the lowest intensity peaks in the spectrum.

<<removePeaks,echo=TRUE,cache=FALSE>>=
experiment <- removePeaks(experiment,t=3,verbose=FALSE)
## total ion current
tic(sp)
tic(experiment[["X66"]])
@

\begin{figure}[!ht]
<<spectrum-plot2,fig=T,echo=T,tikz=T,external=T,width=5,height=5>>=
plot(experiment[["X66"]],reporters=iTRAQ4,full=TRUE)
@ 
\caption{Same spectrum as on figure \ref{fig:spectrum-plot}, after setting peaks <= 5 to 0.}
\label{fig:spectrum-plot2}
\end{figure}

Unlike the name might suggest, the \Rfunction{removePeaks} method does not remove 
peaks from the spectrum. To effectively remove 0 intensity peaks from spectra, 
and reduce the size of the data set, one can use the \Rfunction{clean} method.

<<clean,echo=TRUE,cache=FALSE>>=
## number of peaks
peaksCount(sp)
peaksCount(experiment[["X66"]])
experiment <- clean(experiment,verbose=FALSE)
peaksCount(experiment[["X66"]])
@

Another useful manipulation method is \Rfunction{trimMz}, that taken as 
parameters and \Robject{MSnExp} (or a \Robject{Spectrum}) and a numeric 
\Robject{mzlim}. MZ values smaller to \texttt{min(mzlim)} 
or greater then \texttt{max(mzmax)} are discarded. This method is 
particularely useful when one wants to concentrate on a specific MZ range, 
as for reporter ions quantification, as is generally results in substantial 
reduction of data size. Compare the size of the full trimmed experiment to 
the original \Sexpr{sz} Mb.

<<trimMz,echo=TRUE,cache=FALSE>>=
range(mz(sp))
experiment <- trimMz(experiment,mzlim=c(112,120))
range(mz(experiment[["X66"]]))
experiment
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Quantitation}

Quantitation is performed on specific peaks in the spectra, that are specified 
with an \Robject{ReporterIon} object. A specific peak is defined by it's expected 
\Robject{mz} value and is expected to be found within 
\Robject{mz} $\pm$ \Robject{width}. If the peak reaches outside, a warning will 
be issued. If no data is found, \Robject{NA} is returned.

<<reporters,echo=TRUE,cache=FALSE>>=
mz(iTRAQ4)
width(iTRAQ4)
@

The \Rfunction{quantify} method takes an experiment, 
a character describing the quantification 
\Robject{method} and the \Robject{reporters} to be quantified as parameters. 
Additionally, a progress bar can be dispaying when setting 
the \Robject{verbose} paramter to \Robject{TRUE}.
Three quantification methods are implemented: 
\Robject{trapezoidation} returns the area under the peak of interest, 
\Robject{max} returns the apex of the peak and 
\Robject{sum} returns the sum of all intensities of the peak.

The \Rfunction{quantify} method returns \Robject{MSnSet} objects, that 
extend the well-known \Robject{ExpressionSet} objects defined in the 
\Rpackage{Biobase} package. The assay data is a matrix of dimensions 
$n \times m$ matrix, where $m$ is the  number of features/spectra 
originally in the \Robject{MSnExp} used as parameter in 
\Robject{quantify} and $m$ is the number of reporter ions. 
The meta data is directly inherited from the \Robject{MSnExp} instance.

<<quantify,echo=TRUE,cache=FALSE,keep.source=T>>=
qnt <- quantify(experiment,
                method="trap",
                reporters=iTRAQ4,
                verbose=FALSE)
qnt
head(exprs(qnt))
@

A \Robject{MSnSet} object is mean to be compatible with further downstream 
packages for data normalisation and statistical analysis. 
Below is an example of applying 
variance stabilization normalisation \citep{Huber2002} 
to the iTRAQ reporter intensities \citep{Karp2010}.

<<vsn,echo=TRUE,cache=FALSE>>=
library(vsn)
qnt.vsn <- vsn2(qnt)
@

The results of figure 3 of \cite{Karp2010} are reproduced with 
the dummy experiment and show on figure \ref{fig:cv-plot}.

<<cvdata,echo=FALSE,cache=FALSE>>=
sd1 <- apply(log2(exprs(qnt))+10,1,sd)
mn1 <- apply(log2(exprs(qnt))+10,1,mean)
cv1 <- sd1/mn1
sd2 <- apply(exprs(qnt.vsn)+10,1,sd)
mn2 <- apply(exprs(qnt.vsn)+10,1,mean)
cv2 <- sd2/mn2
library(zoo)
rmed1 <- rollapply(cv1,7,function(x) median(x,na.rm=TRUE))
rmed2 <- rollapply(cv2,7,function(x) median(x,na.rm=TRUE))
@

\begin{figure}[!ht]
<<cv-plot,fig=T,echo=F,tikz=T,external=F,width=5,height=5>>=
plot(order(mn1),cv1,col="steelblue",pch=19,
     ylab="CV",xlab="Rank mean")
points(order(mn2),cv2,col="red",pch=19)
grid()
lines(seq(1,70,by=(70/length(rmed1))),rmed1,col="steelblue",lwd=2)
lines(seq(1,70,by=(70/length(rmed2))),rmed2,col="red"      ,lwd=2)
## legend("topright",
##        legend=c("log2","vsn","running\nmedian"),
##        pch=c(19,19,NA),
##        col=c("steelblue","red","black"),
##        lwd=c(0,0,2),
##        bty="n",
##        cex=.7)
@ 
\caption{CV versus signal intensity comparison for log2 (red) and vsn (red) transformed data. Lines indicate running CV medians.}
\label{fig:cv-plot}
\end{figure}

\clearpage

\section{Session information}\label{sec:sessionInfo} 
<<label=sessioninfo,results=tex,echo=FALSE,cache=FALSE>>=
toLatex(sessionInfo())
@

\bibliographystyle{plainnat}
\bibliography{MSnbase}

\end{document}

### 2011-03-01 inst/doc/MSnbase-development.Rnw
added demo and devel vignettes, version bumped to 0.1

%\VignetteIndexEntry{MSnbase development}
%\VignetteKeywords{Mass Spectrometry, Proteomics, Infrastructure }
%\VignettePackage{MSnbase-development}
\documentclass{article}
\usepackage{tikz}
\usepackage[nogin]{Sweave}
\usepackage{hyperref}
\usepackage[authoryear,round]{natbib}

\SweaveOpts{echo=T,eval=T,fig=F,results=verbatim,cache=F,tikz=T,external=T,prefix.string=sweave-cache/figs/fig}

\newcommand{\R}{\texttt{R} }
\newcommand{\Rfunction}[1]{{\texttt{#1}}}
\newcommand{\Robject}[1]{{\texttt{#1}}}
\newcommand{\Rpackage}[1]{{\textit{#1}}}
\newcommand{\email}[1]{\href{mailto:#1}{\normalfont\texttt{#1}}}
%% colors
\definecolor{Red}{rgb}{0.7,0,0}
\definecolor{Blue}{rgb}{0,0,0.8}

\hypersetup{%
  hyperindex = {true},
  colorlinks = {true},
  linktocpage = {true},
  plainpages = {false},
  linkcolor = {Blue},
  citecolor = {Blue},
  urlcolor = {Red},
  pdfstartview = {Fit},
  pdfpagemode = {UseOutlines},
  pdfview = {XYZ null null null}
}

\input{Author.tex}

\begin{document}

\title{\Rpackage{MSnbase} development}

\maketitle

%% Abstract and keywords %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vskip 0.3in minus 0.1in
\hrule
\begin{abstract}
  This vignette describes the classes implemented in \Rpackage{MSnbase} package. 
  It is intended as a starting point for developers or users who would like to 
  learn more or further develop/extend \Robject{pSet}.

\end{abstract}
\textit{Keywords}: Mass Spectrometry (MS), proteomics, infrastructure.
\vskip 0.1in minus 0.05in
\hrule
\vskip 0.2in minus 0.1in
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\tableofcontents

%% <<setup,cache=FALSE>>=
%% setCacheDir("sweave-cache/values")
%% @

<<environment,cache=FALSE,echo=FALSE>>=
library(MSnbase)
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{Foreword.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

This document is not replacement for the individual manual pages, that  
slots of the \Rpackage{MSnbase} classes. It is a centralised high-level 
description of the package design.

\Rpackage{MSnbase} aims at being compatible with the \Rpackage{Biobase} 
infrastructure \cite{Gentleman2004}. 
Many meta data structures that are used in \Robject{eSet} and associated classes are 
also used here. As such, knowledge of the \textit{Biobase development and the new eSet} 
vignette\footnote{%%
  The vignette can directly be accessed with 
  \texttt{vignette("BiobaseDevelopment",package="Biobase")} 
  once \Rpackage{Biobase} is loaded.} would be beneficial.

The initial goal is to use the \Rpackage{MSnbase} infrastructure for labelled quantitation 
using reporter ions (iTRAQ \citep{Ross2004} and TMT \citep{Thompson2003}). Spectral counting 
should be trivial to apply with current features, as long as identification data is at hand. 
Currently, no effort is invested to streamline label-free quantitative proteomics, although 
some effort has been done to keep the infrastructure flexible enough to accommodate more 
designs.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{\Rpackage{MSnbase} classes}

All classes have a \Robject{.\_\_classVersion\_\_} slot, of class \Robject{Versioned} 
from the \Robject{Biobase} package. This slot documents the class version for any instance 
to be used for debugging and object update purposes. Any change in a class implementation 
should trigger a version change.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{\Robject{pSet}: a virtual class for raw mass spectrometry data and meta data}
\label{sec:pset}

This virtual class aims coordinate contain mass spectrometry 
data, i.e spectra, and meta data. It is based on the \Robject{eSet} for genomic 
data. The main difference with \Robject{eSet} that the \Robject{assayData} slot is an 
environment containing any number of \Robject{Spectrum} instances (see section \ref{sec:spectrum}).

Two new slots are introduced, namely \Robject{proteomicsData}, that contains 
\Robject{MIAPE} instances (see section \ref{sec:miape}) and 
\Robject{processingData}, that contains \Robject{MSnProcess} instances 
(see section \ref{sec:msnprocess}). The \Robject{annotation} slot has not 
been implemented, as no prior feature annotation is known in shotgun proteomics.

<<pSet>>=
getClass("pSet")
@

\paragraph{Future work} Currently, few setters have been implemented.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{\Robject{MSnExp}: a class MS experiments}
\label{sec:msnexp}

\Robject{MSnExp} extends \Robject{pSet} to 
an is used to store MS experiments. It does not add any new slots 
to \Robject{pSet}. Accessors and setters are all inherited from 
\Robject{pSet} and new ones should be implemented for \Robject{pSet}. 
Methods that manipulate actual data in experiments are implemented 
for \Robject{MSnExp}.

<<MSnExp>>=
getClass("MSnExp")
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{\Robject{MSnSet}: a class quantitative proteomics data}

This class stores quantitation data and meta data 
after running \Robject{quantify} on an \Robject{MSnExp} object. The 
quantitative data is in form of a $n \times m$ matrix, where $m$ is the 
number of features/spectra originally in the \Robject{MSnExp} used as 
parameter in \Robject{quantify} and $m$ is the number of reporter ions 
(see section \ref{sec:reporterions}). This thus prompted to directly extend 
the \Robject{ExpressionSet} and add two proteomics-specific annotation slots
introduced in the \Robject{pSet} class, namely \Robject{proteomicsData} for 
\Robject{MIAPE} objects (see section \ref{sec:miape}) and 
\Robject{processingData} for objects of class \Robject{MSnProcess} 
(see section \ref{sec:msnprocess}). The \Robject{annotation} slot, 
inherited from \Robject{eSet} is not used.

<<MSnSet>>=
getClass("MSnSet")
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{\Robject{MSnProcess}: a class for logging processing meta data}
\label{sec:msnprocess}

This class aims at recording specific manipulations 
applied to \Robject{MSnExp} or \Robject{MSnSet} instances. The \Robject{processing} 
slot is a \Robject{character} vector that describes major processing. Most 
other slots are of class \Robject{logical} that indicate whether the data has been 
centroided, smoothed, \ldots although many of the functionality is not implemented yet.
Any new processing that is implemented should be documented logged here.

It also documents the raw data file from which the data originates (\Robject{files} 
slot) and the \Rpackage{MSnbase} version that was installed when the \Robject{MSnProcess} 
instance was originally created.

<<MSnProcess>>=
getClass("MSnProcess")
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{\Robject{MIAPE}: Minimum Information About a Proteomics Experiment}
\label{sec:miape}

The Minimum Information About a Proteomics 
Experiment \citep{Taylor2007} \Robject{MIAPE} class is currently
only composed by a free text \Robject{description} attribute. This 
class will be expanded in future versions. 
Note that the MIAPE standard information that overlaps with MIAME 
requirements should directly be documented in the \Robject{experimentData} 
slots (see also note on page \pageref{misc:whymiame}).

Raw data is currently imported from \texttt{mzXML} files \citep{Pedrioli2004} 
using the \Rfunction{xcms:::rampRawData} 
and \Rfunction{xcms:::rampRawDataMSn} functions from the 
\Rpackage{xcms} package \citep{Smith2006}.
These functions do not give access to the meta data. New importer functions are 
under development (see for instance \Rpackage{mzR}\footnote{%%
  \url{https://github.com/sneumann/mzR/blob/master/DESCRIPTION}}) that will 
hopefully give programmatic access to meta data stored in the data file to 
populate the \Robject{MIAPE} object.

<<MIAPE>>=
getClass("MIAPE")
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{\Robject{Spectrum} \it{et al.}: classes for MS spectra}
\label{sec:spectrum}

\Robject{Spectrum} is a virtual class that defines 
common attributes to all types of spectra. MS1 and MS2 specific attributes 
are defined in the \Robject{Spectrum1} and \Robject{Spectrum2} classes, 
that directly extend \Robject{Spectrum}.

The choices of attributes has been dictated by the \Rfunction{xcms:::rampRawData} 
and \Rfunction{xcms:::rampRawDataMSn} functions and what data from the 
\texttt{mzXML} file they gave access to. 
It is expected that some hopefully minor changes might 
come up here when migrating to other data import packages, that allow random access 
to \texttt{mzXML} data and support \texttt{mzML} \citep{Martens2010}.

<<Spectrum>>=
getClass("Spectrum")
@

<<Spectrum>>=
getClass("Spectrum1")
@

<<Spectrum>>=
getClass("Spectrum2")
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{\Robject{ReporterIons}: a class for isobaric tags}
\label{sec:reporterions}

The iTRAQ and TMT (or any other peak of interest) 
are implemented with \Robject{ReporterIons}, 
that essentially defines an expected MZ position for the peak and a width around 
this value as well a names for the reporters.

<<ReporterIons>>=
getClass("ReporterIons")
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{\Robject{NAnnotatedDataFrame}: multiplexed \Robject{AnnotatedDataFrame}s}
\label{sec:nannotateddataframe}

The simple expansion of the \Robject{AnnotatedDataFrame} 
classes adds the \Robject{multiplex} and \Robject{multiLabel} slots to 
document the number and names of multiplexed samples.

<<ReporterIons>>=
getClass("NAnnotatedDataFrame")
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Miscellaneous}

\paragraph{Unit tests} \Rpackage{MSnbase} implements unit tests with 
the \Rpackage{testthat} package.

\paragraph{Processing methods} Methods that process raw data, i.e. spectra 
should be implemented for \Robject{Spectrum} objects first and then 
\texttt{eapply}'ed (or similar) to the \Robject{assayData} slot of an 
\Robject{MSnExp} instance in the specific method.

\paragraph{Why this MIAME slot?} \label{misc:whymiame} While it would have been 
possible to transfer all data stored in \Rpackage{Biobase}'s \Robject{MIAME} 
to a new \Robject{MIAPE} class and use the latter for \Robject{experimentData} slots in 
the \Robject{pSet} class, it would not have been possible to directly transfer this 
to \Robject{MSnSet} instances, as \Robject{MSnSet} classes directly inherit 
from the \Robject{ExpressionSet}, whose \Robject{experimentData} slot must be of class 
\Robject{MIAME}.

\input{NoteAboutSpeedAndMemory.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Session information}
\label{sec:sessionInfo}

<<label=sessioninfo,results=tex,echo=FALSE,cache=FALSE>>=
toLatex(sessionInfo())
@

\bibliographystyle{plainnat}
\bibliography{MSnbase}

\end{document}

### 2011-03-01 inst/doc/NoteAboutSpeedAndMemory.tex
added demo and devel vignettes, version bumped to 0.1

\paragraph{Speed and memory requirements} Raw mass spectrometry file are generally several 
hundreds of MB large and most of this is used for binary raw spectrum data. As such, 
\Robject{MSnExp} can easily grow very large and thus require large amounts of RAM. 
This requirement may be tackled in the future by avoiding to load the raw data into memory
and using random access to the content of \texttt{mzXML} data files on demand. 
When focusing on reporter ion quantitation, a direct solution for this is to trim the 
spectra using the \Robject{trimMz} method to select the area of interest and thus 
substantially reduce the size of the \Robject{Spectrum} objects.

The independent handling of spectra is ideally suited for parallel processing. 
This will be added soon.

### 2011-03-09 inst/doc/MSnbase-demo.Rnw
LazyData and NEWS
MSnProcess has not initialize(), changed MSnbase-demo and man to reflect truncated dummy data (size requirements)

@@
 
-\Rpackage{MSnbase} imports raw MS data stored in \texttt{mzXML} format \citep{Pedrioli2004}. 
-More formats, notably \texttt{mzML} \citep{Martens2010}, will be added at a later stage. Either one of 
-MS1 or MS2 spectra can be loaded at a time by setting the \texttt{msLevel} parameter accordingly.
+\Rpackage{MSnbase} imports raw MS data stored in \texttt{mzXML} 
+format \citep{Pedrioli2004}. More formats, notably \texttt{mzML} \citep{Martens2010}, 
+will be added at a later stage. 
+Either one of MS1 or MS2 spectra can be loaded at a time by setting the \texttt{msLevel} parameter accordingly.
 
@@
 library(MSnbase)
-filename <- dir(system.file(package="MSnbase",dir="extdata"),full.name=TRUE,pattern="mzXML$")
+filename <- dir(system.file(package="MSnbase",dir="extdata"),
+                full.name=TRUE,pattern="mzXML$")
+print(filename)
 experiment <- readMzXMLData(filename, msLevel = 2, verbose=FALSE)
@@
 <<Spectrum,cache=FALSE,echo=TRUE>>=
-sp <- experiment[["X66"]]
+sp <- experiment[["X43"]]
 sp
@@
 <<ReporterIons>>=
-data(iTRAQ4)
 iTRAQ4
@@
 <<extractPrecSpec,echo=T,cache=F>>=
-exp2 <- extractPrecSpectra(experiment,precursorMz(experiment)["X66"])
+exp2 <- extractPrecSpectra(experiment,precursorMz(experiment)["X43"])
 exp2
@@
 
-We see that \Sexpr{length(exp2)} spectra have the same MZ value. There two spectra 
-can be represented together by plotting the \Robject{MSnExp} object 
-(figure \ref{fig:msnexp-plot}).
+We see that \Sexpr{length(exp2)} spectra have the same MZ value. 
+There two spectra can be represented together by plotting the 
+\Robject{MSnExp} object (figure \ref{fig:msnexp-plot}).
 
@@
 tic(sp)
-tic(experiment[["X66"]])
+tic(experiment[["X43"]])
 @
@@
 <<spectrum-plot2,fig=T,echo=T,tikz=T,external=T,width=5,height=5>>=
-plot(experiment[["X66"]],reporters=iTRAQ4,full=TRUE)
+plot(experiment[["X43"]],reporters=iTRAQ4,full=TRUE)
 @ 
-\caption{Same spectrum as on figure \ref{fig:spectrum-plot}, after setting peaks <= 5 to 0.}
+\caption{Same spectrum as on figure \ref{fig:spectrum-plot}, 
+  after setting peaks <= 3 to 0.}
 \label{fig:spectrum-plot2}
@@
 
-Unlike the name might suggest, the \Rfunction{removePeaks} method does not remove 
-peaks from the spectrum. To effectively remove 0 intensity peaks from spectra, 
-and reduce the size of the data set, one can use the \Rfunction{clean} method.
+Unlike the name might suggest, the \Rfunction{removePeaks} method does 
+not remove peaks from the spectrum. To effectively remove 0 intensity 
+peaks from spectra, and reduce the size of the data set, one can use 
+the \Rfunction{clean} method.
 
@@
 peaksCount(sp)
-peaksCount(experiment[["X66"]])
+peaksCount(experiment[["X43"]])
 experiment <- clean(experiment,verbose=FALSE)
-peaksCount(experiment[["X66"]])
+peaksCount(experiment[["X43"]])
 @
@@
 experiment <- trimMz(experiment,mzlim=c(112,120))
-range(mz(experiment[["X66"]]))
+range(mz(experiment[["X43"]]))
 experiment

### 2011-03-09 R/DataClasses.R

@@ -15,17 +15,12 @@ setClass("MSnProcess",
            MSnbaseVersion="character"),
          contains=c("Versioned"),
          prototype = prototype(
-           new("Versioned", versions=c(MSnProcess="0.1.1")),
+           new("Versioned", versions=c(MSnProcess="0.1.2")),
            processing=character(),
            files=character(),
            trimmed=numeric(),
            removedPeaks=character(),
-           ## Put this in an initialise() method!
-           MSnbaseVersion=ifelse(
-             is.na(packageDescription("MSnbase",fields="Version")),
-             "0.0.0",
-             as.character(packageDescription("MSnbase",fields="Version")))
-           )
+           MSnbaseVersion=character()) ## set in initialize()
          )
 
 #################################################################

### 2011-03-09 R/methods-MSnProcess.R

@@ -1,5 +1,13 @@
 ##################################################################
 ## Methods for MSnProcess class
+setMethod("initialize","MSnProcess",          
+          function(.Object,...) {
+            .Object <- callNextMethod(.Object,...)
+            .Object@MSnbaseVersion <- as.character(packageDescription("MSnbase",fields="Version"))
+            if (validObject(.Object))
+              return(.Object)
+          })
+
 setMethod("show","MSnProcess",function(object) show.MSnProcess(object))
 
 setMethod("fileNames",

### 2011-03-09 inst/tests/test_MSnExp.R

@@ -3,49 +3,46 @@ context("MSnExp class")
 test_that("readMzXMLData and dummy MSnExp msLevel 2 instance", {
   file <- dir(system.file(package="MSnbase",dir="extdata"),full.name=TRUE,pattern="mzXML$")
   aa <- readMzXMLData(file,verbose=FALSE)
-  expect_that(class(aa)=="MSnExp",is_true())
+  expect_true(class(aa)=="MSnExp")
   ## checking slots and methods
-  expect_that(length(aa),equals(70))
+  expect_equal(length(aa),54)
   expect_that(nrow(header(aa)),equals(length(aa)))
   expect_that(names(header(aa)),
               equals(c("index","file","retention.time",
                        "precursor.mz","peaks.count","tic",
                        "ms.level","charge","collision.energy")))
   ## MS levels
-  expect_that(length(msLevel(aa)),equals(70))
-  expect_that(unique(msLevel(aa)),equals(2))
-  expect_that(length(MSnbase:::ms1scan(aa)),equals(70))
-  expect_that(length(unique(MSnbase:::ms1scan(aa))),equals(5))
+  expect_equal(length(msLevel(aa)),54)
+  expect_equal(unique(msLevel(aa)),2)
+  expect_equal(length(MSnbase:::ms1scan(aa)),54)
+  expect_equal(length(unique(MSnbase:::ms1scan(aa))),3)
   ## Precursor MZ
-  expect_that(length(precursorMz(aa)),equals(70))
+  expect_equal(length(precursorMz(aa)),54)
   expect_that(precursorMz(aa)[1],is_a("numeric"))
-  expect_that(length(unique(precursorMz(aa))),equals(11))
-  expect_that(range(precursorMz(aa)),
-              equals(c(424.76651001,630.33178711)))
-  expect_that(as.numeric(sort(precursorMz(aa))[1]),equals(424.76651001)) ## [*]
+  expect_equal(length(unique(precursorMz(aa))),7)
+  expect_equal(range(precursorMz(aa)),
+               c(425.77908325,630.33178711))
+  expect_equal(as.numeric(sort(precursorMz(aa))[1]),425.77908325) ## [*]
   ## Retention time
-  expect_that(length(rtime(aa)),equals(70))
+  expect_equal(length(rtime(aa)),54)
[…]

### 2011-03-15 inst/doc/MSnbase-demo.Rnw
sorted out zoo requirement in vignette

@@
 library(zoo)
-rmed1 <- rollapply(cv1,7,function(x) median(x,na.rm=TRUE))
-rmed2 <- rollapply(cv2,7,function(x) median(x,na.rm=TRUE))
+## rmed1 <- rollapply(cv1,7,function(x) median(x,na.rm=TRUE))
+## rmed2 <- rollapply(cv2,7,function(x) median(x,na.rm=TRUE))
+##
+## Calling directly rollapply.zoo to make it zoo_1.6-4 compatible.
+## The above requires zoo >= 1.7-0, which is as of 15 March 2011
+## not yet available on CRAN (only on r-forge).
+rmed1 <- zoo:::rollapply.zoo(cv1,7,function(x) median(x,na.rm=TRUE))
+rmed2 <- zoo:::rollapply.zoo(cv2,7,function(x) median(x,na.rm=TRUE))
 @

### 2011-03-21 inst/doc/MSnbase-demo.Rnw
MSnSet now extends eSet and has MIAPE expressionData slot; MIAPE extended with MIAPE-MS slots

@@
 <<environment,cache=FALSE,echo=FALSE>>=
-library(MSnbase)
+library("MSnbase")
 @
@@
 
-\Rpackage{MSnbase} aims are providing a reproducible research framework to the proteomics data analysis. 
+\Rpackage{MSnbase} aims are providing a reproducible research framework to proteomics data analysis. 
 It should allow researcher to easily mine mass spectrometry data, explore the data and its statistical properties 
@@
 
-Raw data is stored in an \Robject{MSnExp} object, that stores all the spectra in an experiment, 
+Raw data is contained in \Robject{MSnExp} objects, that stores all the spectra of an experiment, 
 as defined by one or multiple raw data files.
@@
 \Rpackage{MSnbase} imports raw MS data stored in \texttt{mzXML} 
-format \citep{Pedrioli2004}. More formats, notably \texttt{mzML} \citep{Martens2010}, 
-will be added at a later stage. 
-Either one of MS1 or MS2 spectra can be loaded at a time by setting the \texttt{msLevel} parameter accordingly.
+format \citep{Pedrioli2004} with the function \Rfunction{readMzXMLData} function. 
+More formats, notably \texttt{mzML} \citep{Martens2010}, will be added at a later stage. 
+Either MS1 or MS2 spectra can be loaded at a time by setting the \texttt{msLevel} parameter accordingly.
 
 <<MSnExp,cache=FALSE,echo=TRUE>>=
-library(MSnbase)
+library("MSnbase")
 filename <- dir(system.file(package="MSnbase",dir="extdata"),
@@
     this part will be automatically updated when the object is 
-    modified with it's methods, as illustrated later} and package version.
+    modified with it's \textit{ad hoc} methods, as illustrated later} and package version.
     This slot can be accessed with the \Rfunction{processingData} method.
@@
     file name(s) used to import the data, protocol data, 
-    information about features (individual spectra here), 
-    experiment data and proteomics data. Most of these are implemented as 
+    information about features (individual spectra here) and
+    experiment data. Most of these are implemented as 
     in the \Robject{eSet} class and are described in more details in their 
-    respective manual pages. See also \texttt{?pSet} and \texttt{MSnExp}
+    respective manual pages. See \texttt{?MSnExp}
     and references therein for additional background information.
 
-    The proteomics-specific meta data in slot \Robject{proteomicsData} is a 
-    new addition from the \Rpackage{MSnbase} package.
-    It contains MIAPE data (Minimum Information About a Proteomics Experiment, \citep{Taylor2007}) 
-    and can be accessed with the \Rfunction{proteomicsData} method. 
+    The experiment meta data associated with an \Robject{MSnExp} experiment
+    is of class \Robject{MIAPE}. It stores general information about the 
+    experiment as well as MIAPE (Minimum Information About a Proteomics Experiment, 
+    \citep{Taylor2007, Taylor2008}) information.
+    This meta-data can be accessed with the \Rfunction{experimentData} method. 
+    When available, a summary of MIAPE-MS data can be printed with the \Rfunction{msInfo} method.
     See \texttt{?MIAPE} for more details.
@@
 for the MZ values, \Rfunction{intensity} for the inensities, \ldots see the 
-respecitve class manuals for more details.
+\Robject{Spectrum}, \Robject{Spectrum1} and \Robject{Spectrum2} manuals for more details.
 
@@
 Reporter ions are defined with the \Robject{ReporterIons} class. 
-Specific peaks of interested are defined by a MZ value, a with 
+Specific peaks of interest are defined by a MZ value, a with 
 around the expected MZ and a name (and optionally a colour for plotting, 
 see section \ref{sec:plotting}). \Robject{ReporterIons} instances 
-are required to quantitate reporter peaks in \Robject{MSnExp} 
+are required to quantify reporter peaks in \Robject{MSnExp} 
 experiments. Instances for the most commonly used isobaric tags like 
@@
 (by specifing with reporters with \texttt{reporters=iTRAQ4} for instance) 
-or both (figure \ref{fig:spectrum-plot}).
+or both (see figure \ref{fig:spectrum-plot}).
 
@@
 @ 
-\caption{Raw MS2 spectrum with detail about reporter ions.}
+\caption{Raw MS2 spectrum with details about reporter ions.}
 \label{fig:spectrum-plot}
@@
 It is also possible to plot all spectra of an experiment (figure \ref{fig:msnexp-plot}). 
-Lets start by extracting all spectra that have the sample precursor MZ value than \Robject{sp} 
+Lets start by extracting all spectra that have the same precursor MZ value than \Robject{sp} 
 into a separate experiment, using the \Rfunction{extractPrecSpectra} method. 
-It takes an \Robject{MSnExp} experiment and a precursor MZ value as parameters.
+This method takes an \Robject{MSnExp} experiment and a precursor MZ value as parameters.
 
 <<extractPrecSpec,echo=T,cache=F>>=
-exp2 <- extractPrecSpectra(experiment,precursorMz(experiment)["X43"])
+exp2 <- extractPrecSpectra(experiment,precursorMz(sp))
 exp2
@@
 We see that \Sexpr{length(exp2)} spectra have the same MZ value. 
-There two spectra can be represented together by plotting the 
-\Robject{MSnExp} object (figure \ref{fig:msnexp-plot}).
+These can be visulaised together by plotting the 
+\Robject{MSnExp} object, as illustrated on figure \ref{fig:msnexp-plot}.
 
@@
 Low intensity peaks can be set to 0 with the \Rfunction{removePeaks} method 
-from spectra or whole experiments. The intensity below which peaks are removed 
-is defined by the \Robject{t} parameter. \texttt{t} can be specified directly 
+from spectra or whole experiments. The intensity threshold below which peaks are removed 
+is defined by the \Robject{t} parameter. \Robject{t} can be specified directly 
 as a numeric. The default value is the character \texttt{"min"}, that will remove 
-all peaks equal to the lowest intensity peaks in the spectrum.
+all peaks equal to the lowest non null intensity in any spectrum. 
+We observe the effect of the \Rfunction{removePeaks} method by comparing 
+total ion count (i.e. the total intensity in a spectrum) with the \Rfunction{tic} method
+before (object \Robject{sp}) and after (object \Robject{experiment}). 
+Figure \ref{fig:spectrum-plot2} illustrated spectrum \texttt{X43} after low intensity 
+peaks removal.
 
@@
 Unlike the name might suggest, the \Rfunction{removePeaks} method does 
-not remove peaks from the spectrum. To effectively remove 0 intensity 
+not actually remove peaks from the spectrum; they are set to 0. This 
+can be checked using the \Rfunction{peaksCount} method, that returns 
+the number of peaks (including 0 intensity peaks) in a spectrum.
+To effectively remove 0 intensity 
 peaks from spectra, and reduce the size of the data set, one can use 
@@
 
-Another useful manipulation method is \Rfunction{trimMz}, that taken as 
+Another useful manipulation method is \Rfunction{trimMz}, that takes as 
 parameters and \Robject{MSnExp} (or a \Robject{Spectrum}) and a numeric 
-\Robject{mzlim}. MZ values smaller to \texttt{min(mzlim)} 
+\Robject{mzlim}. MZ values smaller then \texttt{min(mzlim)} 
 or greater then \texttt{max(mzmax)} are discarded. This method is 
 particularely useful when one wants to concentrate on a specific MZ range, 
-as for reporter ions quantification, as is generally results in substantial 
+as for reporter ions quantification, and generally results in substantial 
 reduction of data size. Compare the size of the full trimmed experiment to 
 the original \Sexpr{sz} Mb.
+\label{timMz-example}
 
@@
 @
+
+As can be seen above, all processing performed on the experiment is recorded 
+and displayed as integral part of the experiment object.
 
@@
 The \Rfunction{quantify} method returns \Robject{MSnSet} objects, that 
-extend the well-known \Robject{ExpressionSet} objects defined in the 
-\Rpackage{Biobase} package. The assay data is a matrix of dimensions 
+extend the well-known \Robject{eSet} class defined in the 
+\Rpackage{Biobase} package. \Robject{MSnSet} instances are very similar 
+to \Robject{ExpressionSet} objects, except for the experiment meta-data 
+that captures MIAPE specific information.
+The assay data is a matrix of dimensions 
 $n \times m$ matrix, where $m$ is the  number of features/spectra 
 originally in the \Robject{MSnExp} used as parameter in 
-\Robject{quantify} and $m$ is the number of reporter ions. 
+\Robject{quantify} and $m$ is the number of reporter ions, that 
+can be accessed with the \Rfunction{exprs} method.
 The meta data is directly inherited from the \Robject{MSnExp} instance.
@@
 
-A \Robject{MSnSet} object is mean to be compatible with further downstream 
+A \Robject{MSnSet} object is meant to be compatible with further downstream 
 packages for data normalisation and statistical analysis. 
@@
 <<vsn,echo=TRUE,cache=FALSE>>=
-library(vsn)
-qnt.vsn <- vsn2(qnt)
+library("vsn")
+qnt.vsn <- vsn2(exprs(qnt))
 @
@@
 cv2 <- sd2/mn2
-library(zoo)
+library("zoo")
 ## rmed1 <- rollapply(cv1,7,function(x) median(x,na.rm=TRUE))
@@
 @ 
-\caption{CV versus signal intensity comparison for log2 (red) and vsn (red) transformed data. Lines indicate running CV medians.}
+\caption{CV versus signal intensity comparison for log2 (blue) and vsn (red) transformed data. Lines indicate running CV medians.}
 \label{fig:cv-plot}

### 2011-03-21 inst/doc/MSnbase-development.Rnw
MSnSet now extends eSet and has MIAPE expressionData slot; MIAPE extended with MIAPE-MS slots

@@
 
-This document is not replacement for the individual manual pages, that  
-slots of the \Rpackage{MSnbase} classes. It is a centralised high-level 
+This document is not a replacement for the individual manual pages, that  
+document the slots of the \Rpackage{MSnbase} classes. It is a centralised high-level 
 description of the package design.
@@
 
-This virtual class aims coordinate contain mass spectrometry 
-data, i.e spectra, and meta data. It is based on the \Robject{eSet} for genomic 
-data. The main difference with \Robject{eSet} that the \Robject{assayData} slot is an 
+This virtual class is the main container for mass spectrometry 
+data, i.e spectra, and meta data. It is based on the \Robject{eSet} implementation for genomic 
+data. The main difference with \Robject{eSet} is that the \Robject{assayData} slot is an 
 environment containing any number of \Robject{Spectrum} instances (see section \ref{sec:spectrum}).
 
-Two new slots are introduced, namely \Robject{proteomicsData}, that contains 
-\Robject{MIAPE} instances (see section \ref{sec:miape}) and 
-\Robject{processingData}, that contains \Robject{MSnProcess} instances 
-(see section \ref{sec:msnprocess}). The \Robject{annotation} slot has not 
+One new slot is introduced, namely \Robject{processingData}, 
+that contains one \Robject{MSnProcess} instance  (see section \ref{sec:msnprocess}). 
+ and the \Robject{experimentData} slot is now expected to contain
+\Robject{MIAPE} data (see section \ref{sec:miape}).
+The \Robject{annotation} slot has not 
 been implemented, as no prior feature annotation is known in shotgun proteomics.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{\Robject{MSnExp}: a class MS experiments}
+\subsection{\Robject{MSnExp}: a class for MS experiments}
 \label{sec:msnexp}
 
-\Robject{MSnExp} extends \Robject{pSet} to 
-an is used to store MS experiments. It does not add any new slots 
+\Robject{MSnExp} extends \Robject{pSet} to store MS experiments. 
+It does not add any new slots 
 to \Robject{pSet}. Accessors and setters are all inherited from 
@@
 Methods that manipulate actual data in experiments are implemented 
-for \Robject{MSnExp}.
+for \Robject{MSnExp} objects.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{\Robject{MSnSet}: a class quantitative proteomics data}
+\subsection{\Robject{MSnSet}: a class for quantitative proteomics data}
 
@@
 parameter in \Robject{quantify} and $m$ is the number of reporter ions 
-(see section \ref{sec:reporterions}). This thus prompted to directly extend 
-the \Robject{ExpressionSet} and add two proteomics-specific annotation slots
-introduced in the \Robject{pSet} class, namely \Robject{proteomicsData} for 
-\Robject{MIAPE} objects (see section \ref{sec:miape}) and 
+(see section \ref{sec:reporterions}).
+
+This prompted to keep a similar implementation as 
+the \Robject{ExpressionSet} class, while adding the proteomics-specific annotation slot
+introduced in the \Robject{pSet} class, namely 
 \Robject{processingData} for objects of class \Robject{MSnProcess} 
-(see section \ref{sec:msnprocess}). The \Robject{annotation} slot, 
-inherited from \Robject{eSet} is not used.
+(see section \ref{sec:msnprocess}).
+
+The \Robject{MSnSet} class extends the virtual \Robject{eSet} class 
+to provide compatibility for \Robject{ExpressionSet}-like behaviour. 
+The experiment meta-data in  \Robject{experimentData} is 
+also of class \Robject{MIAPE} (see section \ref{sec:miape}).
+The \Robject{annotation} slot, inherited from \Robject{eSet} is not used.
 
@@
 centroided, smoothed, \ldots although many of the functionality is not implemented yet.
-Any new processing that is implemented should be documented logged here.
+Any new processing that is implemented should be documented and logged here.
 
 It also documents the raw data file from which the data originates (\Robject{files} 
-slot) and the \Rpackage{MSnbase} version that was installed when the \Robject{MSnProcess} 
-instance was originally created.
+slot) and the \Rpackage{MSnbase} version that was in use when the \Robject{MSnProcess} 
+instance, and hence the \Robject{MSnExp}/\Robject{MSnSet} objects, were originally created.
 
@@
 The Minimum Information About a Proteomics 
-Experiment \citep{Taylor2007} \Robject{MIAPE} class is currently
-only composed by a free text \Robject{description} attribute. This 
-class will be expanded in future versions. 
-Note that the MIAPE standard information that overlaps with MIAME 
-requirements should directly be documented in the \Robject{experimentData} 
-slots (see also note on page \pageref{misc:whymiame}).
+Experiment \citep{Taylor2007, Taylor2008} \Robject{MIAPE} class 
+describes the experiment, including contact details, 
+information about the mass spectrometer and control 
+and analysis software.
 
-Raw data is currently imported from \texttt{mzXML} files \citep{Pedrioli2004} 
+Raw data is currently imported from \texttt{mzXML} files \citep{Pedrioli2004} <
 using the \Rfunction{xcms:::rampRawData} 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{\Robject{Spectrum} \it{et al.}: classes for MS spectra}
+\subsection{\Robject{Spectrum} \textit{et al.}: classes for MS spectra}
 \label{sec:spectrum}
@@
 The iTRAQ and TMT (or any other peak of interest) 
-are implemented with \Robject{ReporterIons}, 
+are implemented \Robject{ReporterIons} instances, 
 that essentially defines an expected MZ position for the peak and a width around 
@@
 
-\paragraph{Why this MIAME slot?} \label{misc:whymiame} While it would have been 
-possible to transfer all data stored in \Rpackage{Biobase}'s \Robject{MIAME} 
-to a new \Robject{MIAPE} class and use the latter for \Robject{experimentData} slots in 
-the \Robject{pSet} class, it would not have been possible to directly transfer this 
-to \Robject{MSnSet} instances, as \Robject{MSnSet} classes directly inherit 
-from the \Robject{ExpressionSet}, whose \Robject{experimentData} slot must be of class 
-\Robject{MIAME}.
+%% \paragraph{Why this MIAME slot?} \label{misc:whymiame} While it would have been 
+%% possible to transfer all data stored in \Rpackage{Biobase}'s \Robject{MIAME} 
+%% to a new \Robject{MIAPE} class and use the latter for \Robject{experimentData} slots in 
+%% the \Robject{pSet} class, it would not have been possible to directly transfer this 
+%% to \Robject{MSnSet} instances, as \Robject{MSnSet} classes directly inherit 
+%% from the \Robject{ExpressionSet}, whose \Robject{experimentData} slot must be of class 
+%% \Robject{MIAME}.
 

### 2011-03-21 R/DataClasses.R

@@ -28,12 +28,67 @@ setClass("MSnProcess",
 ## See online documentation for more information.
 setClass("MIAPE",
          representation = representation(
-           description="character"),
+           title="character",
+           url="character",
+           ##  Publication details
+           abstract="character",
+           pubMedIds="character",
+           ## Other useful slots, from MIAME
+           samples="list",
+           preprocessing="list",
+           other="list",
+           ##########################
+           ## Based on MIAPE-MS 2.24
+           ## 1. General features - (a) Global descriptors
+           dateStamp="character",
+           ##    Responsible person
+           name="character",
+           lab="character",
+           contact="character",
+           ##    Instrument details
+           instrumentModel="character",
+           instrumentManufacturer="character",
+           instrumentCustomisations="character",
+           ## 1. General features - (b) Control and analysis software
+           softwareName="character",
+           softwareVersion="character",
+           switchingCriteria="character",
+           isolationWidth="numeric",
+           parameterFile="character",
+           ## 2. Ion sources -- will be updated to
+           ##                   provided details specific to
+           ##                   different sources
+           ionSource="character", ## ESI, MALDI, ...
+           ionSourceDetails="character", 
+           ## 3. Post-source componentry
+           analyser="character", ## Quad, TOF, Trap, ...
+           analyserDetails="character",
+           ## 3. Post-source componentry - (d) Collision cell
+           collisionGas="character",
+           collisionPressure="numeric",
+           collisionEnergy="character",
+           ## 3. Post-source component — (f) Detectors
+           detectorType="character", 
+           detectorSensitivity="character"
+           ## 4. Spectrum and peak list generation and annotation 
+           ##    (a) Spectrum description 
+           ##    (b) Peak list generation
+           ##    (c) Quantitation for selected ions
+           ## -- see Spectrum and MSnProcess objects           
+           ),
          contains = c("MIAxE"),
          prototype = prototype(
-           new("Versioned", versions=c(classVersion("MIAxE"), MIAME="0.0.2")),
-           description="Will contain more MIAPE data."
-           )
+           new("Versioned", versions=c(classVersion("MIAxE"), MIAPE="0.2.0")),
+           name="",
+           lab="",
+           contact="",
+           title="",
+           abstract="",
+           url="",
+           pubMedIds="",
+           samples=list(),
[…]

### 2011-03-21 R/functions-MIAPE.R

@@ -1,3 +1,72 @@
 show.MIAPE <- function(object) {
-  cat("proteomicsData: ",object@description,"\n")
+  if (any(c(length(object@instrumentModel)>0,
+            length(object@instrumentManufacturer)>0,
+            length(object@instrumentCustomisations)>0))) {
+    cat("Instrument : \n")
+    cat("  Model:",object@object@instrumentModel,"\n")
+    cat("  Manufacturer:",object@object@instrumentManufacturer,"\n")
+    cat("  Customisations:",object@object@instrumentCustomisations,"\n")
+    cat("  Use 'msInfo(object)' for more MIAPE-MS information.\n")
+  }
+  ## code from original MIAPE show method
+  tmp <- c("samples","preprocessing")
+  Index <-c(length(object@samples) > 0,
+            length(object@preprocessing) > 0)
+  cat("Experiment data\n")
+  cat("  Experimenter name:",object@name,"\n")
+  cat("  Laboratory:",object@lab,"\n")
+  cat("  Contact information:",object@contact,"\n")
+  cat("  Title:",object@title,"\n")
+  cat("  URL:",object@url,"\n")
+  pmids <- pubMedIds(object)
+  cat("  PMIDs:",pmids,"\n")
+  if(length(object@abstract) > 0 && all(object@abstract!=""))
+    cat("\n  Abstract: A",length(strsplit(object@abstract," ")[[1]]),
+        "word abstract is available. Use 'abstract' method.\n")
+  else
+    cat("  No abstract available.\n")
+  if(any(Index))
+    cat("  Information is available on:", paste(tmp[Index],collapse=", "),"\n")
+  nO = notes(object)
+  if (length(nO) > 0) {
+    cat("  notes:\n" )
+    if( is.list(nO) ) {
+      nms = names(nO)
+      pw = options("width")[[1]] - 6
+      for(i in 1:length(nO) ) {
+        cat("   ", nms[i], ":", sep="")
+        cat("     ", strbreak(nO[[i]], width=pw, exdent=0), sep="\n      ")
+      }
+    }
+  }
+}
+
+msinfo.MIAPE <- function(object) {
+  cat("MIAPE-MS information:\n")
+  cat(" 1. General features:\n")
+  cat("  Date stamp: ",object@dateStamp,"\n")
+  cat("  Contact: ",object@contact,"\n")
+  cat("  Name: ",object@name,"\n")
+  cat("  Laboratory: ",object@lab,"\n")
[…]

### 2011-03-21 R/methods-MIAPE.R

@@ -1,3 +1,44 @@
+## TODO getters and setters for MIAPE-MS data
+
 ##################################################################
 ## Methods for MIAPE class
 setMethod("show","MIAPE",function(object) show.MIAPE(object))
+
+setMethod("msInfo","MIAPE",
+          function(object) msInfo.MIAPE(object))
+
+
+setMethod("abstract","MIAPE",function(object) object@abstract)
+setMethod("samples","MIAPE",function(object) object@samples)
+setMethod("pubMedIds","MIAPE",function(object) object@pubMedIds)
+
+setReplaceMethod("pubMedIds","MIAPE",function(object,value){
+   object@pubMedIds = value
+   object
+})
+
+setMethod("otherInfo","MIAPE",function(object) object@other)
+
+setMethod("expinfo","MIAPE",
+   function(object) {
+      tmp <- c(object@name, object@lab, object@contact, object@title, object@url)
+    names(tmp) <- c("name","lab","contact","title","url")
+    return(tmp)
+   }
+)
+
+setMethod("notes", signature(object="MIAPE"),
+          function(object) object@other)
+
+setReplaceMethod("notes", signature(object="MIAPE", value="list"),
+                 function(object, value) {
+                     object@other <- value
+                     object
+                 })
+
+setReplaceMethod("notes", signature(object="MIAPE", value="character"),
+                 function(object, value) {
+                     object@other <- append(object@other, value)
+                     object
+                 })
+

### 2011-03-21 R/methods-pSet.R

@@ -169,17 +169,18 @@ setMethod("spectra","MSnExp",function(object) {
 ##                    return(object)
 ##                  })
 
-setMethod("proteomicsData","MSnExp",function(object) object@proteomicsData)
-setMethod("proteomicsData<-","MSnExp",
-          function(object,value="MIAPE") object@proteomicsData <- value)
-setReplaceMethod("proteomicsData",
-                 signature(object="MSnExp",
-                           value="MIAPE"),
-                 function(object, value) {
-                   object@proteomicsData = value
-                   if (validObject(object))
-                     return(object)
-                 })
+## No proteomicsData slot anymore since MSnbase 0.2.0 - experimentData is not MIAPE
+## setMethod("proteomicsData","MSnExp",function(object) object@proteomicsData)
+## setMethod("proteomicsData<-","MSnExp",
+##           function(object,value="MIAPE") object@proteomicsData <- value)
+## setReplaceMethod("proteomicsData",
+##                  signature(object="MSnExp",
+##                            value="MIAPE"),
+##                  function(object, value) {
+##                    object@proteomicsData = value
+##                    if (validObject(object))
+##                      return(object)
+##                  })
 
 setMethod("sampleNames",
           signature(object="pSet"),
@@ -231,6 +232,10 @@ setMethod("fvarLabels",
           signature=signature(object="pSet"),
           function(object) varLabels(featureData(object)))
 setMethod("experimentData", signature(object="pSet"), function(object) object@experimentData)
+
+setMethod("msInfo","pSet",
+          function(object) msInfo(experimentData(object)))
+
 setMethod("description", signature(object="pSet"),
           function(object, ...) {
             experimentData(object)

### 2011-03-23 R/methods-ReporterIons.R
no artefacts in package anymore + 'fake' vignettes (see NEWS for more minor changes)

@@ -6,8 +6,7 @@ setMethod("[","ReporterIons",
 setMethod("length","ReporterIons",function(x) length(x@mz))
 setMethod("mz","ReporterIons", function(object) object@mz)
 
-## Using x as defined in the IRanges::width generic
-setMethod("width","ReporterIons", function(x) x@width) 
+setMethod("width","ReporterIons", function(object) object@width) 
 
 setMethod("reporterColours","ReporterIons", function(object) object@col)
 setMethod("reporterColors" ,"ReporterIons", function(object) object@col)

### 2011-03-24 inst/doc/NoteAboutSpeedAndMemory.tex
Added the MSnbase package to the repository.
git-svn-id: file:///home/git/hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/MSnbase@54004 bc3139a8-67e5-0310-9ffc-ced21a209358

@@
 hundreds of MB large and most of this is used for binary raw spectrum data. As such, 
-\Robject{MSnExp} can easily grow very large and thus require large amounts of RAM. 
+data containers can easily grow very large and thus require large amounts of RAM. 
 This requirement may be tackled in the future by avoiding to load the raw data into memory
-and using random access to the content of \texttt{mzXML} data files on demand. 
+and using random access to the content of \texttt{mzXML}/\texttt{mzML} data files on demand. 
 When focusing on reporter ion quantitation, a direct solution for this is to trim the 
 spectra using the \Robject{trimMz} method to select the area of interest and thus 
-substantially reduce the size of the \Robject{Spectrum} objects.
+substantially reduce the size of the \Robject{Spectrum} objects. This is illustrated on page 
+\pageref{trimMz-example} of the \texttt{MSnbase-demo} vignette.
 

### 2011-03-24 R/functions-MSnExp.R

@@ -1,0 +1,206 @@
+## mergeSpectra <- function(object, ## MSnExp object
+##                          fun=sum,
+##                          verbose=TRUE) {
+##   spectra <- spectra(object)
+##   prec <- sapply(spectra,function(x) x@precursorMz)
+##   uprec <- unique(prec)
+##   luprec <- length(uprec)
+##   if (verbose) 
+##     cat(luprec,"unique precursors for",length(spectra),"spectra\n")
+##   newSpectra <- vector("list",length=luprec)
+##   if (verbose)
+##     pb <- txtProgressBar(min = 0, max = luprec, style = 3)
+##   for (i in 1:luprec) {
+##     if (verbose)
+##       setTxtProgressBar(pb, i)   
+##     pi <- uprec[i]
+##     sel <- prec %in% pi
+##     l <- spectra[sel]
+##     ## unique M/Z values
+##     mzs <- unique(unlist(lapply(l,function(x) x@mz)))
+##     ## vectors of intensitues (ints) and number of times
+##     ## a given itensity is found (icounts)
+##     ints <- icounts <- numeric(length(mzs))
+##     names(icounts) <- mzs
+##     allints <- unlist(lapply(l,function(x) x@intensity))
+##     names(allints) <- unlist(lapply(l,function(x) x@mz))
+##     for (j in 1:length(mzs)) {
+##       k <- as.character(mzs[j])
+##       ints[j] <- fun(allints[names(allints) %in% k])
+##       icounts[j] <- sum(names(allints) %in% k)
+##     }
+##     o <- order(mzs)
+##     newSpectra[[i]] <- new("Spectrum",
+##                            merged=which(sel),
+##                            rt=unique(sapply(l,function(x) x@rt)),
+##                            msLevel=unique(sapply(l,function(x) x@msLevel)),
+##                            precursorMz=pi,
+##                            peaksCount=length(ints),
+##                            intensity=ints[o],
+##                            mz=mzs[o])
+##     validObject(newSpectra[[i]])
+##   }
+##   if (verbose)
+##     close(pb)
+##   object@process@processing <- c(object@process@processing,
+##                                  paste("Precursor ions with identical M/Z merged:",
+##                                        date()))
+##   object@process@merged <- TRUE
+##   ## TODO update and set featureData
+##   return(new("MSnExp",
+##              assayData=list2env(newSpectra),
+##              metadata=object@metadata,
+##              process=object@process))
[…]

### 2011-03-24 R/methods-MSnExp.R

@@ -1,0 +1,144 @@
+##################################################################
+## Methods for MSnExp class
+
+setMethod("show",
+          signature=signature(object="MSnExp"),
+          function(object) {
+            cat("Object of class \"",class(object),"\"\n",sep="")
+            cat(" Object size in memory: ")
+            if (length(assayData(object))==0) {
+              sz <- 0
+            } else {
+              sz <- sum(sapply(assayData(object),object.size)) + object.size(object)
+            }
+            cat(round(sz/(1024^2),2),"Mb\n")
+            cat("- - - Spectra data - - -\n")
+            if (sz==0) {
+              cat(" none\n")
+            } else {
+              msnLevels <- unique(msLevel(object))
+              cat(" MSn level(s):",msnLevels,"\n")
+              if (all(msLevel(object)>1)) {
+                cat(" Number of MS1 acquisitions:",length(unique(ms1scan(object))),"\n")
+                cat(" Number of MS2 scans:",length(ls(assayData(object))),"\n")
+                msnPrecMz <- precursorMz(object)
+                nbPrecIons <- length(msnPrecMz)
+                cat(" Number of precursor ions:",nbPrecIons,"\n")
+                if (nbPrecIons>0) {
+                  cat("",length(unique(msnPrecMz)),"unique MZs\n")
+                  cat(" Precursor MZ's:",paste(signif(range(msnPrecMz),5),collapse=" - "),"\n")
+                }
+                msnMz <- unname(mz(object)) ## unnaming list to avoid
+                                            ## huge time oberhead in range(msnMz)
+                msnMzRange <- round(range(msnMz),2) 
+                cat(" MSn M/Z range:",msnMzRange,"\n")
+              } else {
+                cat(" Number of MS1 scans:",length(spectra(object)),"\n")
+              }
+              msnRt <- rtime(object)
+              if (length(msnRt)>0) {
+                rtr <- range(msnRt)
+                cat(" MSn retention times:",formatRt(rtr[1]),"-",formatRt(rtr[2]),"minutes\n")
+              }
+            }
+            show(processingData(object))
+            cat("- - - Meta data  - - -\n")
+            Biobase:::.showAnnotatedDataFrame(phenoData(object),
+                                              labels=list(object="phenoData"))
+            cat("Loaded from:\n")
+            files <- processingData(object)@files
+            if (length(files)>0) {
+              for (i in 1:length(files)) {
+                f <- basename(files[i])
+                cat(" ",f,"\n")
+              }
+            } else {
[…]

### 2011-03-24 R/utils.R

@@ -1,0 +1,220 @@
+##' Converts seconds to min:sec format
+##'
+##' This function is used to convert retention times,
+##' expressed in seconds, in the more human friendly
+##' format mm:sec.
+##'
+##' @title Format Retention Time
+##' @usage formatRt(rt)
+##' @param rt retention in in seconds
+##' @return a character string in mm:ss format 
+##' @examples
+##' formatRt(1524)
+##' @author Laurent Gatto
+##' @export 
+##' @keywords chron utilities
+formatRt <- function(rt) {
+  min <- floor(rt/60)
+  sec <- round(rt-(min*60))
+  return(paste(min,":",sec,sep=""))
+}
+
+##' Convert spectrum data to an Spectrum object
+##'
+##' This function converts the i'th spectrum of a list produced
+##' by \code{xcms:::rampRawData} (for \code{rawToSpectrum1}) or
+##' \code{xcms:::rampRawDataMSn} (for \code{rawToSpectrum2}) to
+##' object of class \code{"\linkS4class{Spectrum1}"} or
+##' \code{"\linkS4class{Spectrum1}"}, respectively.
+##' Rudimentary data processing like low intensity peaks and
+##' zero intensity data points removal can optionally already be
+##' applied at this stage.
+##' 
+##' @title Raw spectrum data to Spectrum object 
+##' @aliases rawToSpectrum2
+##' @usage
+##' rawToSpectrum1(x,i,removePeaks,clean,updatePeaksCount,verbose)
+##' rawToSpectrum2(x,i,removePeaks,clean,updatePeaksCount,verbose)
+##' @param x a list as returned by \code{xcms:::rampRawData} (for \code{rawToSpectrum1})
+##' or \code{xcms:::rampRawDataMSn} (for \code{rawToSpectrum2}).
+##' @param i a numeric value (integer) indicating which spectrum to extract.
+##' @param removePeaks a numeric value to define low-intensity peaks that should
+##' be removed; indicate 0 (default) to keep data as is.
+##' @param clean a logical value specifying whether MZ data points with 0 intensity
+##' (if for instance removePeaks > 1) should be completely removed.
+##' @param updatePeaksCount a logical value indicating if peak counts should
[…]

### 2011-03-24 inst/doc/Foreword.tex
Added the MSnbase package to the repository.
git-svn-id: https://hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/MSnbase@54004 bc3139a8-67e5-0310-9ffc-ced21a209358

@@
 mature infrastructure in the Bioconductor project, changes may occur in the future. 
-Current functionality will evolve and new ones will be added. 
+Current functionality will evolve and new one will be added. 
 Although at an early stage, this package is released in the hope that it may foster  
 new developments in proteomics data analysis within \R by providing a common infrastructure. 
-Several package developers related to mass spectrometry and proteomics data met during the 
+Several package developers working with mass spectrometry and proteomics data met at the 
 Bioconductor Developer Meeting Europe\footnote{\url{http://bioconductor.org/help/course-materials/2010/HeidelbergNovember2010/}} 

### 2011-03-24 inst/doc/MSnbase-demo/MSnbase-demo.Rnw
Added the MSnbase package to the repository.
git-svn-id: https://hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/MSnbase@54004 bc3139a8-67e5-0310-9ffc-ced21a209358

--- 2011-03-21 inst/doc/MSnbase-demo.Rnw
@@
-%\VignetteIndexEntry{Base Functions and Classes for MS-based Proteomics }
+%\VignetteIndexEntry{Base Functions and Classes for MS-based Proteomics}
 %\VignetteKeywords{Mass Spectrometry, MS, MSMS, Proteomics, Infrastructure, Bioinformatics, quantitative }
@@
 \begin{figure}[!ht]
-<<spectrum-plot,fig=T,echo=T,tikz=T,external=T,width=5,height=5>>=
+<<spectrum-plot,fig=T,echo=T,tikz=T,external=F,width=5,height=5>>=
 plot(sp,reporters=iTRAQ4,full=TRUE)
@@
 \begin{figure}[!ht]
-<<msnexp-plot,fig=T,echo=T,tikz=T,external=T,width=5,height=6>>=
+<<msnexp-plot,fig=T,echo=T,tikz=T,external=F,width=5,height=6>>=
 plot(exp2,reporters=iTRAQ4,full=FALSE)
@@
 \begin{figure}[!ht]
-<<spectrum-plot2,fig=T,echo=T,tikz=T,external=T,width=5,height=5>>=
+<<spectrum-plot2,fig=T,echo=T,tikz=T,external=F,width=5,height=5>>=
 plot(experiment[["X43"]],reporters=iTRAQ4,full=TRUE)

### 2011-03-24 R/functions-MSnProcess.R

@@ -1,0 +1,6 @@
+show.MSnProcess <- function(object) {
+  cat("- - - Processing information - - -\n")
+  for (proc in object@processing)
+    cat(proc,"\n")
+  cat(" MSnbase version:",object@MSnbaseVersion,"\n")
+}

### 2011-03-24 R/methods-NAnnotatedDataFrame.R

@@ -1,0 +1,26 @@
+##################################################################
+## Methods for NAnnotatedDataFrame class
+setMethod("dim", "NAnnotatedDataFrame", function(x) {
+  d <- c(dim(pData(x)),
+         length(x@multiplex))
+  names(d) <- c(dimLabels(x),"multipleNames")
+  return(d)
+})
+
+setMethod("show",
+          signature=signature(object="NAnnotatedDataFrame"),
+          function(object) {
+            callNextMethod(object)
+            cat("  Multiplexing: ")
+            if (length(object@multiplex)>0)
+                cat(object@multiplex," - ",object@multiLabels,"\n",sep="")
+            else
+              cat("none\n")
+          })
+
+
+setMethod("multiplex", "NAnnotatedDataFrame",
+          function(object) object@multiplex)
+
+setMethod("multiLabels", "NAnnotatedDataFrame",
+          function(object) object@multiLabels)

### 2011-04-26 inst/doc/MSnbase-demo/MSnbase-demo.Rnw
Quantification by trapezoidation now has strict argument. Removed many warnings in quantify. Updated vignette with illustrative plots for quantitation methods and spectrum pre-processing. See NEWS for more details.
git-svn-id: https://hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/MSnbase@55087 bc3139a8-67e5-0310-9ffc-ced21a209358

@@
 the \Rfunction{clean} method. 
+The effect of the \Rfunction{removePeaks} and \Rfunction{clean} methods 
+are illustrated on figure \ref{fig:preproc} on page \pageref{fig:preproc}.
 
@@
 
+<<preprosp,cache=FALSE,echo=FALSE>>=
+int <- c(0,1,1,3,1,1,0,0,0,1,3,7,3,1,0)
+mz <- c(113.9,114.0,114.05,114.1,114.15,114.2,114.25,
+        114.3,114.35,114.4,114.42,114.48,114.5,114.55,114.6)
+ppsp <- new("Spectrum2",intensity=int,mz=mz,centroided=FALSE)
+p1 <- plot(ppsp,full=TRUE) + theme_gray(5) + 
+  geom_point(size=3,alpha=I(1/3)) +
+  geom_hline(yintercept=3,linetype=2) + 
+  opts(title="Original spectrum") 
+p2 <- plot(removePeaks(ppsp,t=3),full=TRUE) + 
+  theme_gray(5) +
+  geom_point(size=3,alpha=I(1/3)) + 
+  geom_hline(yintercept=3,linetype=2) + 
+  opts(title="Peaks < 3 removed") 
+p3 <- plot(clean(removePeaks(ppsp,t=3)),full=TRUE) + 
+  theme_gray(5) +
+  geom_point(size=3,alpha=I(1/3)) + 
+  geom_hline(yintercept=3,linetype=2) + 
+  opts(title="Peaks < 3 removed and cleaned") 
+@
+
+\begin{figure}[p]
+  \begin{center}
+<<preproc-plot,fig=T,echo=F,tikz=T,external=F,width=3,height=6>>=
+grid.newpage()
+pushViewport(viewport(layout = grid.layout(3, 1)))
+vplayout <- function(x, y)
+  viewport(layout.pos.row = x, layout.pos.col = y)
+print(p1, vp=vplayout(1,1))
+print(p2, vp=vplayout(2,1))
+print(p3, vp=vplayout(3,1))
+@ 
+\caption{This figure illustrated the effect or the \Rfunction{removePeaks} 
+  and \Rfunction{clean} methods. The left-most spectrum displays two peaks, 
+  of max height 3 and 7 respectively. The middle spectrum shows the result 
+  of calling \Rfunction{removePeaks} with argument \Robject{t=3}, which sets 
+  all data points of the first peak, whose maximum height is smaller or equal 
+  to \Robject{t} to 0. The second peak is unaffected. Calling \Rfunction{clean} 
+  after \Rfunction{removePeaks} effectly deletes successive 0 intensities from 
+  the spectrum, as shown on the right plot. }
+\label{fig:preproc}
+  \end{center}
+\end{figure}
+
 Another useful manipulation method is \Rfunction{trimMz}, that takes as 
@@
 the original \Sexpr{sz} Mb.
-\label{timMz-example}
+\label{trimMz-example}
 
@@
 with an \Robject{ReporterIon} object. A specific peak is defined by it's expected 
-\Robject{mz} value and is expected to be found within 
-\Robject{mz} $\pm$ \Robject{width}. If the peak reaches outside, a warning will 
-be issued. If no data is found, \Robject{NA} is returned.
+\Robject{mz} value and is expected to be found within \Robject{mz} $\pm$ \Robject{width}.  
+If no data is found, \Robject{NA} is returned.
 
@@
 
-The \Rfunction{quantify} method takes an experiment, 
-a character describing the quantification 
-\Robject{method} and the \Robject{reporters} to be quantified as parameters. 
+The \Rfunction{quantify} method takes the following paramters: 
+an \Robject{MSnExp} experiment, 
+a character describing the quantification \Robject{method}, 
+the \Robject{reporters} to be quantified 
+and a \Robject{strict} logical defining whether data points 
+ranging outside of \Robject{mz} $\pm$ \Robject{width} should be considered 
+for quantitation. 
 Additionally, a progress bar can be dispaying when setting 
-the \Robject{verbose} paramter to \Robject{TRUE}.
-Three quantification methods are implemented: 
+the \Robject{verbose} parameter to \Robject{TRUE}.
+Three quantification methods are implemented, as illustrated 
+on figure \ref{fig:quant-methods}: 
 \Robject{trapezoidation} returns the area under the peak of interest, 
@@
 \Robject{sum} returns the sum of all intensities of the peak. 
+See \texttt{?quantify} for more details.
+
+<<simplesp,cache=FALSE,echo=FALSE>>=
+int <- c(0,1,1,3,1,1,0)
+mz <- c(113.9,114.0,114.05,114.1,114.15,114.2,114.25)
+ssp <- new("Spectrum2",intensity=int,mz=mz,centroided=FALSE)
+p <- plot(ssp,full=TRUE)
+p <- p + theme_gray(5)
+@
+
+\begin{figure}[p]
+<<quantitation-plot,fig=T,echo=F,tikz=T,external=F,width=5,height=5>>=
+grid.newpage()
+pushViewport(viewport(layout = grid.layout(2, 2)))
+vplayout <- function(x, y)
+  viewport(layout.pos.row = x, layout.pos.col = y)
+print(p + opts(title="Quantitation using 'sum'") + geom_point(size=3,alpha=I(1/3),colour="red"), vp=vplayout(1,1))
+print(p + opts(title="Quantitation using 'max'") + geom_point(aes(x=114.1,y=3),alpha=I(1/3),colour="red",size=3), vp=vplayout(1,2))
+print(p + opts(title="Trapezoidation and strict=FALSE") + geom_polygon(alpha=I(1/5),fill="red"),vp=vplayout(2,1))
+print(p + opts(title="Trapezoidation and strict=TRUE") + 
+      geom_polygon(aes(x=c(114.05,114.05,114.1,114.15,114.15),y=c(0,1,3,1,0)),fill="red",alpha=I(1/5)),vp=vplayout(2,2))
+@ 
+\caption{The different quantitation methods are illustrated above. 
+  Quantitation using \Robject{sum} sums all the data points in the peaks to 
+  produce, for ths example, \Sexpr{quantify(sp,iTRAQ4[1],method="sum")[[1]]}, 
+  whereas method \Robject{max} only uses the peak's maximum intensity, 
+  \Sexpr{quantify(sp,iTRAQ4[1],method="max")[[1]]}. 
+  \Robject{Trapezoidation} calculates the area under the peak taking 
+  the full with into account (using \Robject{strict=FALSE} gives
+  \Sexpr{round(quantify(ssp,iTRAQ4[1],method="trap",strict=FALSE)[[1]],3)}) 
+  or only the width as defined by the reporter 
+  (using \Robject{strict=TRUE} gives
+  \Sexpr{round(quantify(ssp,iTRAQ4[1],method="trap",strict=TRUE)[[1]],3)}). }
+\label{fig:quant-methods}
+\end{figure}
+
+\bigskip
 
@@
 The assay data is a matrix of dimensions 
-$n \times m$ matrix, where $m$ is the  number of features/spectra 
+$n \times m$, where $m$ is the  number of features/spectra 
 originally in the \Robject{MSnExp} used as parameter in 
@@
                 reporters=iTRAQ4,
+                strict=FALSE,
                 verbose=FALSE)

### 2011-04-26 R/AllGenerics.R

@@ -33,7 +33,7 @@ setGeneric("tic",function(object) standardGeneric("tic"))
 
 setGeneric("fromFile",function(object) standardGeneric("fromFile"))
 
-setGeneric("quantify",function(object,method,...) standardGeneric("quantify"))
+setGeneric("quantify",function(object,...) standardGeneric("quantify"))
 setGeneric("curveStats",function(object,reporters,...) standardGeneric("curveStats"))
 
 setGeneric("qual",function(object) standardGeneric("qual"))

### 2011-04-26 R/functions-Spectrum.R

@@ -17,29 +17,43 @@ clean.Spectrum <- function(spectrum,updatePeaksCount=TRUE) {
   return(spectrum)
 }
 
-quantify.Spectrum <- function(spectrum,method,reporters) {
+quantify.Spectrum <- function(spectrum,method,
+                              reporters,strict) {
   ## Parameters:
   ##  spectrum: object of class Spectrum
+  ##  method: methods for how to quantify the peaks
   ##  reporters: object of class ReporterIons
+  ##  strict: use full width of peaks or a limited range around apex
   ## Return value:
   ##  a names list of length 2 with
   ##   peakQuant: named numeric of length length(reporters)
-  ##   curveStats: a length(reporters)x7 data frame 
+  ##   curveStats: a length(reporters) x 7 data frame 
   peakQuant <- vector("numeric",length(reporters))
   names(peakQuant) <- reporterNames(reporters)
   curveStats <- c()
   for (i in 1:length(reporters)) {
     ## Curve statistics
-    dfr <- curveData(spectrum,reporters[i])
+    dfr <- MSnbase:::curveData(spectrum,reporters[i])    
     ##  dfr:     mz int
     ##  1  114.1023   0
     ##  2  114.1063   2
     ##  3  114.1102   3
     ##  4  114.1142   4
     ##  ...
+    if (strict) {
+      lowerMz <- round(mz(reporters[i]) - width(reporters[i]),3)
+      upperMz <- round(mz(reporters[i]) + width(reporters[i]),3)
+      selMz <- (dfr$mz >= lowerMz) & (dfr$mz <= upperMz)
+      dfr <- dfr[selMz,]
+      dfr <- rbind(c(min(dfr$mz),0),
+                   dfr,
+                   c(max(dfr$mz),0))
+    }
     maxInt <- max(dfr$int)
     nMaxInt <- sum(dfr$int==maxInt)
     baseLength <- nrow(dfr)
+    if (strict)
+      baseLength <- baseLength-2
     mzRange <- range(dfr$mz)
     precMz <- precursorMz(spectrum)
     if (length(precMz)!= 1)
@@ -50,27 +64,27 @@ quantify.Spectrum <- function(spectrum,method,reporters) {
                           reporters[i]@mz,precMz))
     ## Quantification
     if (method=="trapezoidation") {
+      if (nrow(dfr)==1) {
+        if (!is.na(dfr$int)) 
[…]

### 2011-04-26 R/methods-Spectrum.R

@@ -83,13 +83,14 @@ setMethod("trimMz",
           function(object,mzlim,...) trimMz.Spectrum(object,mzlim))
 
 setMethod("quantify",
-          signature=signature("Spectrum","character"),
+          signature=signature("Spectrum"),
           function(object,
                    method=c("trapezoidation","max","sum"),
-                   reporters) {
+                   reporters,
+                   strict=FALSE) {
             if (!inherits(reporters,"ReporterIons"))
               stop("Argument 'reporters' must inherit from 'ReporterIons' class.")
-            quantify.Spectrum(object,match.arg(method),reporters)
+            quantify.Spectrum(object,match.arg(method),reporters,strict)
           })
 
 setMethod("curveStats","Spectrum",

### 2011-04-26 R/methods-pSet.R
Quantification by trapezoidation now has strict argument. Removed many warnings in quantify. Updated vignette with illustrative plots for quantitation methods and spectrum pre-processing. See NEWS for more details.
git-svn-id: file:///home/git/hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/MSnbase@55087 bc3139a8-67e5-0310-9ffc-ced21a209358

@@ -6,7 +6,7 @@ setValidity("pSet", function(object) {
     msg <- validMsg(msg,
                     "assayData must contain Spectrum objects.")
   msl <- msLevel(object)
-  if (length(unique(msl))!=1) 
+  if (length(unique(msl))>1) 
     warning(paste("Different MS levels in ",class(object),
                   " object:",unique(msl)))
   ## checking number of spectra in assayData and

### 2011-05-05 inst/doc/MSnbase-demo/MSnbase-demo.Rnw
added combineFeatures example in demo vignette
git-svn-id: file:///home/git/hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/MSnbase@55274 bc3139a8-67e5-0310-9ffc-ced21a209358

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Quantitation}
+\section{From spectra to quantitative expression data}
+
+\subsection{Reporter ions quantitation}\label{sec:quant}
 
@@
 
+\subsection{Normalisation}\label{sec:norm}
+
 A \Robject{MSnSet} object is meant to be compatible with further downstream 
@@
 
-The results of figure 3 of \cite{Karp2010} are reproduced with 
+Figure 3 of \cite{Karp2010} is reproduced with 
 the dummy experiment and show on figure \ref{fig:cv-plot}.
@@
 
-\clearpage
+\subsection{Feature aggregation}
+
+The above quantitation and normalisation has been performed on quantitative data 
+obtained from individual spectra. However, the biological unit of interest is not 
+the spectrum but the peptide or the protein. As such, it is important to be able to 
+summarise features that belong to a same group, i.e. spectra from one peptide, 
+peptides that originate from one protein, or directly combine all spectra that have 
+been uniquely associated to one protein.
+
+\Rpackage{MSnbase} provides one function, \Rfunction{combineFeatures}, that allows 
+to aggregate features stored in an \Robject{MSnSet} using build-in or user 
+defined summary function and return a new \Robject{MSnSet} instance. 
+The function has three main argument here described below. 
+Additional details can be found in the function manual.
+
+<<makeGroups1,echo=FALSE,cache=FALSE>>=
+gb <- as.factor(fData(qnt)$precursor.mz)
+@
+
+\Rfunction{combineFeatures}'s first argument, \Robject{object}, is an instance of class \Robject{MSnSet}, 
+as has been created in the section \ref{sec:quant} for instance. 
+The second argument, \Robject{groupBy}, is a \Robject{factor} than has as many elements as 
+there are features in the \Robject{MSnSet} \Robject{object} argument. The features corresponding 
+to the \Robject{groupBy} levels will be aggregated so that the resulting \Robject{MSnSet} output 
+will have \Rfunction{length(levels(groupBy))} features. Here, we will combine individual 
+MSMS spectra based on their precursor mass-to-charge ratio. This information has been saved 
+in the \Robject{MSnSet}'s \Robject{featureData} slot during the quantitation (see section \ref{sec:quant}). 
+As shown below, this will result in \Sexpr{length(table(gb))} new aggregated features.
+
+<<makeGroups2,echo=TRUE,cache=FALSE>>=
+names(fData(qnt))
+head(fData(qnt)$precursor.mz)
+gb <- as.factor(fData(qnt)$precursor.mz)
+table(gb)
+@
+
+The third argument, \Robject{fun}, 
+defined how to combine the features. Predefined functions are readily available and can be 
+specified as strings 
+(\Robject{fun="mean"}, \Robject{fun="median"}, \Robject{fun="sum"}, 
+\Robject{fun="weigthed.mean"} or \Robject{fun="medianpolish"} to compute respectively the mean, 
+media, sum, weighted mean or median polish of the features to be aggregated). 
+Alternatively, is is possible to supply userdefined functions with 
+\Robject{fun=function(x) \{ ... \}}. We will use the \Robject{median} here.
+
+<<combineFeatures,echo=TRUE,cache=FALSE>>=
+qnt2 <- combineFeatures(qnt,groupBy=gb,fun="median")
+qnt2
+@
+
+Other intended ways to define aggregation groups is to import 
+identification data into R and collate it to the \Robject{MSnSet} 
+feature data, and subsequently use peptide sequences or protein assignments 
+to combine features.
 

### 2011-05-06 inst/doc/MSnbase-demo/MSnbase-demo.Rnw
added incomplete dissociation and spectral counting sections to demo vignette and mention bio-sig-proteomics list
git-svn-id: https://hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/MSnbase@55300 bc3139a8-67e5-0310-9ffc-ced21a209358

@@
 print(filename)
-experiment <- readMzXMLData(filename, msLevel = 2, verbose=FALSE)
-experiment
+raw.experiment <- readMzXMLData(filename, msLevel = 2, verbose=FALSE)
+raw.experiment
 @
@@
 <<experiment-size,echo=FALSE,cache=FALSE>>=
-sz <- sum(sapply(assayData(experiment),object.size)) + object.size(experiment)
+sz <- sum(sapply(assayData(raw.experiment),object.size)) + 
+  object.size(raw.experiment)
 sz <- round(sz/(1024^2),2)
@@
 
-The raw data is composed of the \Sexpr{length(experiment)} MS spectra. 
+The raw data is composed of the \Sexpr{length(raw.experiment)} MS spectra. 
 The spectra are named individually 
-(\Sexpr{paste(paste(head(featureNames(experiment)),collapse=", "),", ...",sep="")})
+(\Sexpr{paste(paste(head(featureNames(raw.experiment)),collapse=", "),", ...",sep="")})
 and stored in a \Robject{environment}. They can be accessed individually with 
-\texttt{experiment[["X10"]]} or as a list with \texttt{spectra(experiment)}.
+\texttt{raw.experiment[["X10"]]} or as a list with \texttt{spectra(raw.experiment)}.
 As we have loaded our experiment specifying \texttt{msLevel=2}, 
@@
 <<Spectrum,cache=FALSE,echo=TRUE>>=
-sp <- experiment[["X43"]]
+sp <- raw.experiment[["X43"]]
 sp
@@
 peaksCount(sp)
-head(peaksCount(experiment))
+head(peaksCount(raw.experiment))
 rtime(sp)
-head(rtime(experiment))
+head(rtime(raw.experiment))
 @
@@
 <<extractPrecSpec,echo=T,cache=F>>=
-exp2 <- extractPrecSpectra(experiment,precursorMz(sp))
+exp2 <- extractPrecSpectra(raw.experiment,precursorMz(sp))
 exp2
@@
 <<removePeaks,echo=TRUE,cache=FALSE>>=
-experiment <- removePeaks(experiment,t=3,verbose=FALSE)
+experiment <- removePeaks(raw.experiment,t=3,verbose=FALSE)
 ## total ion current
@@
 
-Quantitation is performed on specific peaks in the spectra, that are specified 
+Quantitation is performed on fixed peaks in the spectra, that are specified 
 with an \Robject{ReporterIon} object. A specific peak is defined by it's expected 
-\Robject{mz} value and is expected to be found within \Robject{mz} $\pm$ \Robject{width}.  
+\Robject{mz} value and is searched for within \Robject{mz} $\pm$ \Robject{width}.  
 If no data is found, \Robject{NA} is returned.
@@
 plot(order(mn1),cv1,col="steelblue",pch=19,
-     ylab="CV",xlab="Rank mean")
+     ylab="CV",xlab="Rank mean",
+     cex.axis=.7,cex.lab=.8)
 points(order(mn2),cv2,col="red",pch=19)
 grid()
-lines(seq(1,70,by=(70/length(rmed1))),rmed1,col="steelblue",lwd=2)
-lines(seq(1,70,by=(70/length(rmed2))),rmed2,col="red"      ,lwd=2)
+lines(seq(1,70,by=(70/length(rmed1))),rmed1,col="steelblue",lwd=1)
+lines(seq(1,70,by=(70/length(rmed2))),rmed2,col="red"      ,lwd=1)
 ## legend("topright",
@@
 \subsection{Feature aggregation}
+\label{sec:feataggregation}
 
@@
 
+\subsection{Label-free MS2 quantitation}
+\label{sec:spectralcounting}
+
+Note that if samples are not multiplexed, label-free MS2 quantitation is possible 
+using \Rpackage{MSnbase}.
+Once individual spectra have been assigned to peptides and proteins, it becomes 
+straightforward to estimate protein quantities using the spectral counting method, 
+as illustrated in section \ref{sec:feataggregation}, when the \Robject{groupBy} 
+argument is defined.
+
+\section{Quantitative assessment of incomplete dissociation}
+\label{sec:incompdissoc}
+
+Quantitation using isobaric reporter tags assumes complete dissociation between
+the reporter group (red on figure \ref{fig:itraqchem}), 
+balance group (blue) 
+and peptide (the peptide reactive group is drawn in green). 
+However, incomplete dissociation does occur and 
+results in an isobaric tag (i.e reporter and balance groups) specific peaks.
+
+\begin{figure}[h]
+  \begin{center}
+    \includegraphics[width=4cm]{./itraqchem.pdf}
+\caption{ iTRAQ 4-plex isobaric tags reagent consist of three parts: 
+  (1) a charged reporter group (MZ of 114, 115, 116 and 117) that is unique to each of the four reagents (red),
+  (2) an uncharged mass balance group (28-31 Da) (blue)and 
+  (3) a peptide reactive group (NHS ester) that binds to the peptide. 
+  In case of incomplete dissociation, the reporter and balance groups produce a specific peaks 
+  at MZ 145.
+}
+\label{fig:itraqchem}
+  \end{center}
+\end{figure}
+
+\Rpackage{MSnbase} provides, among others, a \Robject{ReporterIons} object for iTRAQ 4-plex that includes the 
+145 peaks, called \Robject{iTRAQ5}. This can then be used to quantify the experiment as show in section \ref{sec:quant} 
+to estimate incomplete dissociation for each spectrum.
+
+<<incompdiss,echo=TRUE,cache=FALSE,keep.source=T>>=
+iTRAQ5
+incompdiss <- quantify(raw.experiment,
+                       method="trap",
+                       reporters=iTRAQ5,
+                       strict=FALSE,
+                       verbose=FALSE)
+head(exprs(incompdiss))
+@
+
+Figure \ref{fig:incompdiss} compares these intensities for the whole experiment.
+
+\begin{figure}[!hbt]
+  \begin{center}
+<<incompdiss-plot,fig=T,echo=F,tikz=T,external=F,width=4,height=3.5>>=
+dfr <- melt(exprs(incompdiss))
+colnames(dfr) <- c("feature","reporters","intensity")
+qplot(x=reporters,y=intensity,data=dfr,geom="boxplot") + theme_gray(8)
+@ 
+\caption{ Boxplot comparing intensities of the 4 reporter ions an the 
+  incomplete dissociation specific peak. }
+\label{fig:incompdiss}
+  \end{center}
+\end{figure}
+
+\clearpage
+
 \section{Session information}\label{sec:sessionInfo} 

### 2011-05-06 inst/doc/Foreword.tex
added incomplete dissociation and spectral counting sections to demo vignette and mention bio-sig-proteomics list
git-svn-id: file:///home/git/hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/MSnbase@55300 bc3139a8-67e5-0310-9ffc-ced21a209358

@@
 This library is one attempt to do so.
+
+\bigskip
+
+You are wellcome to contact me for questions, bugs, typos or suggestions about \Rpackage{MSnbase}. 
+If you wish to reach a broader audience for general questions about proteomics analysis using \texttt{R}, 
+you may want to use the Bioconductor proteomics special interest 
+group\footnote{\url{https://stat.ethz.ch/mailman/listinfo/bioc-sig-proteomics}}.

### 2011-05-09 inst/doc/MSnbase-demo/MSnbase-demo.Rnw
added/documented MSnSet purity correction method
git-svn-id: https://hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/MSnbase@55356 bc3139a8-67e5-0310-9ffc-ced21a209358

@@
 
+\subsection{Purity correction}
+\label{sec:purcor}
+
+Impurities in the reporter reagents can bias the results and can be corrected when 
+manufacturers provide correction coefficients. 
+These generally come as percentages of each reporter ion that have masses differing 
+by -2, -1, +1 and +2 Da from the nominal reporter ion mass due to isotopic variants.  
+The \Rfunction{purityCorrect} method applies such correction to \Robject{MSnSet} 
+instances. It also requires a square matric as second argument, \Robject{impurities}, 
+that defines the repaltive percentage of reporter in the quantified each peak. 
+See \texttt{?purityCorrect} for more details.
+
+<<purityCorrect,echo=TRUE,cache=FALSE>>=
+impurities <- matrix(c(0.929,0.059,0.002,0.000,
+                       0.020,0.923,0.056,0.001,
+                       0.000,0.030,0.924,0.045,
+                       0.000,0.001,0.040,0.923),
+                     nrow=4)
+qnt2.crct <- purityCorrect(qnt2,impurities)
+head(exprs(qnt2))
+head(exprs(qnt2.crct))
+@
+
 \subsection{Label-free MS2 quantitation}

### 2011-05-09 R/utils.R

@@ -203,3 +203,12 @@ utils.plot2d <- function(object,z=c("tic","file","peaks.count","charge")) {
   print(p)
   invisible(p)
 }
+
+
+makeImpuritiesMatrix <- function(x) {
+  M <- diag(x)
+  colnames(M) <- paste("reporter",1:x,sep=".")
+  rownames(M) <- paste("% reporter",1:x)
+  corrfactors <- edit(M)
+  invisible(corrfactors)
+}

### 2011-05-09 inst/tests/test_MSnSet.R

@@ -33,3 +33,12 @@ test_that("Combine MSnSet features", {
   expect_true(all(fData(bb)[,1]==c("A","B")))
   expect_true(all(fData(bb)[,2]==c("A.1","B.6")))
 })
+
+test_that("Purity correction", {
+  file <- dir(system.file(package="MSnbase",dir="extdata"),full.name=TRUE,pattern="mzXML$")
+  aa <- readMzXMLData(file,verbose=FALSE)
+  msnset <- quantify(aa,method="trap",reporters=iTRAQ4)
+  impurity0 <- matrix(c(1,0,0,0,0,1,0,0,0,0,1,0,0,0,0,1),ncol=4)
+  pc <- purityCorrect(msnset,impurity0)
+  expect_true(all(exprs(pc)==exprs(msnset)))
+})

### 2011-05-11 inst/doc/MSnbase-demo/MSnbase-demo.Rnw
new makeImpuritiesMatrix, itraqdata and writeMgfData
git-svn-id: https://hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/MSnbase@55442 bc3139a8-67e5-0310-9ffc-ced21a209358

@@
 format \citep{Pedrioli2004} with the function \Rfunction{readMzXMLData} function. 
+See \Rfunction{?readMzXMLData} for more details.
 More formats, notably \texttt{mzML} \citep{Martens2010}, will be added at a later stage. 
 Either MS1 or MS2 spectra can be loaded at a time by setting the \texttt{msLevel} parameter accordingly.
+In this document, we will use a slightly bigger dataset, \Robject{itraqdata}, provided with 
+\Rpackage{MSnbase}. It includes feature metadata, accessible with the \Rfunction{fData} method. 
+The metadata includes identification data for the \Sexpr{length(itraqdata)} MS2 spectra.
 
@@
 library("MSnbase")
-filename <- dir(system.file(package="MSnbase",dir="extdata"),
-                full.name=TRUE,pattern="mzXML$")
-print(filename)
-raw.experiment <- readMzXMLData(filename, msLevel = 2, verbose=FALSE)
-raw.experiment
+itraqdata
+head(fData(itraqdata))
 @
@@
 <<experiment-size,echo=FALSE,cache=FALSE>>=
-sz <- sum(sapply(assayData(raw.experiment),object.size)) + 
-  object.size(raw.experiment)
+sz <- sum(sapply(assayData(itraqdata),object.size)) + 
+  object.size(itraqdata)
 sz <- round(sz/(1024^2),2)
@@
 
-The raw data is composed of the \Sexpr{length(raw.experiment)} MS spectra. 
+The raw data is composed of the \Sexpr{length(itraqdata)} MS spectra. 
 The spectra are named individually 
-(\Sexpr{paste(paste(head(featureNames(raw.experiment)),collapse=", "),", ...",sep="")})
+(\Sexpr{paste(paste(head(featureNames(itraqdata)),collapse=", "),", ...",sep="")})
 and stored in a \Robject{environment}. They can be accessed individually with 
-\texttt{raw.experiment[["X10"]]} or as a list with \texttt{spectra(raw.experiment)}.
+\texttt{itraqdata[["X10"]]} or as a list with \texttt{spectra(itraqdata)}.
 As we have loaded our experiment specifying \texttt{msLevel=2}, 
@@
 <<Spectrum,cache=FALSE,echo=TRUE>>=
-sp <- raw.experiment[["X43"]]
+sp <- itraqdata[["X1"]]
 sp
@@
 peaksCount(sp)
-head(peaksCount(raw.experiment))
+head(peaksCount(itraqdata))
 rtime(sp)
-head(rtime(raw.experiment))
+head(rtime(itraqdata))
 @
@@
 
-It is also possible to plot all spectra of an experiment (figure \ref{fig:msnexp-plot}). 
-Lets start by extracting all spectra that have the same precursor MZ value than \Robject{sp} 
-into a separate experiment, using the \Rfunction{extractPrecSpectra} method. 
-This method takes an \Robject{MSnExp} experiment and a precursor MZ value as parameters.
-
-<<extractPrecSpec,echo=T,cache=F>>=
-exp2 <- extractPrecSpectra(raw.experiment,precursorMz(sp))
-exp2
+<<bsasel,echo=FALSE>>=
+bsasel <- fData(itraqdata)$ProteinAccession=="BSA"
 @
 
-We see that \Sexpr{length(exp2)} spectra have the same MZ value. 
-These can be visulaised together by plotting the 
+It is also possible to plot all spectra of an experiment (figure \ref{fig:msnexp-plot}). 
+Lets start by subsetting the \Robject{itraqdata} experiment using the protein accession 
+numbers included in the feature metadata, and keep the \Sexpr{sum(bsasel)} from 
+the \textit{BSA} protein.
+
+<<subset,echo=TRUE>>=
+sel <- fData(itraqdata)$ProteinAccession=="BSA"
+bsa <- itraqdata[sel]
+bsa
+as.character(fData(bsa)$ProteinAccession)
+@
+
+%% Lets start by extracting all spectra that have the same precursor MZ value than \Robject{sp} 
+%% into a separate experiment, using the \Rfunction{extractPrecSpectra} method. 
+%% This method takes an \Robject{MSnExp} experiment and a precursor MZ value as parameters.
+%% <<extractPrecSpec,echo=T,cache=F>>=
+%% exp2 <- extractPrecSpectra(raw.experiment,precursorMz(sp))
+%% exp2
+%% @
+
+These can then be visualised together by plotting the 
 \Robject{MSnExp} object, as illustrated on figure \ref{fig:msnexp-plot}.
@@
 <<msnexp-plot,fig=T,echo=T,tikz=T,external=F,width=5,height=6>>=
-plot(exp2,reporters=iTRAQ4,full=FALSE)
+plot(bsa,reporters=iTRAQ4,full=FALSE)
 @ 
-\caption{Experiment-wide raw MS2 spectra.}
+\caption{Experiment-wide raw MS2 spectra. The y-axes of the individual 
+  spectra are automatically rescaled to the same range. }
 \label{fig:msnexp-plot}
@@
 total ion count (i.e. the total intensity in a spectrum) with the \Rfunction{tic} method
-before (object \Robject{sp}) and after (object \Robject{experiment}). 
-Figure \ref{fig:spectrum-plot2} illustrated spectrum \texttt{X43} after low intensity 
-peaks removal.
+before (object \Robject{itraqdata}) and after (object \Robject{experiment}) for 
+spectrum \texttt{X55}.
 
 <<removePeaks,echo=TRUE,cache=FALSE>>=
-experiment <- removePeaks(raw.experiment,t=3,verbose=FALSE)
+experiment <- removePeaks(itraqdata,t=400,verbose=FALSE)
 ## total ion current
-tic(sp)
-tic(experiment[["X43"]])
+tic(itraqdata[["X55"]])
+tic(experiment[["X55"]])
 @
@@
 \begin{figure}[!ht]
-<<spectrum-plot2,fig=T,echo=T,tikz=T,external=F,width=5,height=5>>=
-plot(experiment[["X43"]],reporters=iTRAQ4,full=TRUE) 
+<<spectrum-plot2,fig=T,echo=F,tikz=T,external=F,width=5,height=2.5>>=
+grid.newpage()
+pushViewport(viewport(layout = grid.layout(1, 2)))
+vplayout <- function(x, y)
+  viewport(layout.pos.row = x, layout.pos.col = y)
+p1 <- plot(itraqdata[["X55"]],full=TRUE) + theme_gray(5) 
+p2 <- plot(experiment[["X55"]],full=TRUE) + theme_gray(5) 
+print(p1,vp=vplayout(1,1))
+print(p2,vp=vplayout(1,2))
 @ 
-\caption{Same spectrum as on figure \ref{fig:spectrum-plot}, 
-  after setting peaks <= 3 to 0.}
+\caption{Same spectrum before (left) and after setting peaks <= 400 to 0.}
 \label{fig:spectrum-plot2}
@@
 ## number of peaks
-peaksCount(sp)
-peaksCount(experiment[["X43"]])
+peaksCount(itraqdata[["X55"]])
+peaksCount(experiment[["X55"]])
 experiment <- clean(experiment,verbose=FALSE)
-peaksCount(experiment[["X43"]])
+peaksCount(experiment[["X55"]])
 @
@@
 pushViewport(viewport(layout = grid.layout(3, 1)))
-vplayout <- function(x, y)
-  viewport(layout.pos.row = x, layout.pos.col = y)
 print(p1, vp=vplayout(1,1))
@@
 <<trimMz,echo=TRUE,cache=FALSE>>=
-range(mz(sp))
+range(mz(itraqdata[["X55"]]))
 experiment <- trimMz(experiment,mzlim=c(112,120))
-range(mz(experiment[["X43"]]))
+range(mz(experiment[["X55"]]))
 experiment
 @ 
-
 As can be seen above, all processing performed on the experiment is recorded 
@@
 
+If no peak is detected for a reporter ion peak, the respective 
+quantitation value is set to \texttt{NA}. In our case, there 
+is \Sexpr{sum(is.na(exprs(qnt)))} such case in row 
+\Sexpr{which(is.na(exprs(qnt))) %% nrow(qnt)}.
+We will remove the offending line, as we do not expect any missing  peaks.
+
+<<removeNa,echo=TRUE>>=
+sum(is.na(exprs(qnt)))
+whichRow <- which(is.na(exprs(qnt))) %% nrow(qnt)
+qnt <- qnt[-whichRow,]
+sum(is.na(exprs(qnt)))
+@
+
 \subsection{Normalisation}\label{sec:norm}
@@
 <<makeGroups1,echo=FALSE,cache=FALSE>>=
-gb <- as.factor(fData(qnt)$precursor.mz)
+gb <- fData(qnt)$ProteinAccession
 @
@@
 will have \Rfunction{length(levels(groupBy))} features. Here, we will combine individual 
-MSMS spectra based on their precursor mass-to-charge ratio. This information has been saved 
-in the \Robject{MSnSet}'s \Robject{featureData} slot during the quantitation (see section \ref{sec:quant}). 
+MS2 spectra based on the protein they originate from. 
 As shown below, this will result in \Sexpr{length(table(gb))} new aggregated features.
@@
 <<makeGroups2,echo=TRUE,cache=FALSE>>=
-names(fData(qnt))
-head(fData(qnt)$precursor.mz)
-gb <- as.factor(fData(qnt)$precursor.mz)
+gb <- fData(qnt)$ProteinAccession
 table(gb)
@@
 
-Other intended ways to define aggregation groups is to import 
-identification data into R and collate it to the \Robject{MSnSet} 
-feature data, and subsequently use peptide sequences or protein assignments 
-to combine features.
+%% Other intended ways to define aggregation groups is to import 
+%% identification data into R and collate it to the \Robject{MSnSet} 
+%% feature data, and subsequently use peptide sequences or protein assignments 
+%% to combine features.
 
@@
 iTRAQ5
-incompdiss <- quantify(raw.experiment,
+incompdiss <- quantify(itraqdata,
                        method="trap",

### 2011-05-11 R/AllGenerics.R
new makeImpuritiesMatrix, itraqdata and writeMgfData
git-svn-id: file:///home/git/hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/MSnbase@55442 bc3139a8-67e5-0310-9ffc-ced21a209358

@@ -57,3 +57,5 @@ setGeneric("multiLabels",function(object) standardGeneric("multiLabels"))
 
 setGeneric("plot2d",function(object,z,...) standardGeneric("plot2d"))
 ##setGeneric("plot3d",function(object,z,...) standardGeneric("plot3d"))
+
+setGeneric("writeMgfData",function(object,...) standardGeneric("writeMgfData"))

### 2011-05-13 R/functions-Spectrum.R
updated and exported normalise method and vignette updates
git-svn-id: https://hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/MSnbase@55561 bc3139a8-67e5-0310-9ffc-ced21a209358

@@ -229,4 +229,10 @@ trimMz.Spectrum <- function(x,mzlim,updatePeaksCount=TRUE) {
   return(x)
 }
 
-
+normalise.Spectrum <- function(object,method) {
+  ints <- intensity(object)
+  maxInt <- max(ints)
+  normInts <- ints/maxInt
+  object@intensity <- normInts
+  return(object)
+}

### 2011-05-13 R/methods-MSnExp.R

@@ -83,7 +83,7 @@ setMethod("clean",
           signature=signature("MSnExp"),
           function(object,verbose=TRUE) clean.MSnExp(object,verbose))
 
-setMethod("removePeaks","MSnExp",
+setMethod("removePeaks",signature("MSnExp"),
           function(object,t,verbose=TRUE) removePeaks.MSnExp(object,t,verbose))
 
 
@@ -143,3 +143,14 @@ setMethod("extractPrecSpectra",
 setMethod("extractSpectra",
           signature=signature(object="MSnExp",selected="logical"),
           function(object,selected) extractSpectra.MSnExp(object,selected))
+
+setMethod("normalise","MSnExp",
+          function(object,method="max",...) {
+            if (method!="max")
+              warning("Normalising experiment with 'max'.")
+            normalise.MSnExp(object,method="max")
+        })
+
+setMethod("normalize","MSnExp",
+          function(object,method="max",...)
+          normalise(object,method))

### 2011-05-13 inst/doc/MSnbase-demo/MSnbase-demo.Rnw
updated and exported normalise method and vignette updates
git-svn-id: file:///home/git/hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/MSnbase@55561 bc3139a8-67e5-0310-9ffc-ced21a209358

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Importing raw data}\label{sec:msnexp}
+
+\section{Data structure and content}\label{sec:data}
+
+\subsection{Importing experiments}\label{sec:msnexp}
 
@@
 Either MS1 or MS2 spectra can be loaded at a time by setting the \texttt{msLevel} parameter accordingly.
-In this document, we will use a slightly bigger dataset, \Robject{itraqdata}, provided with 
+In this document, we will use a slightly bigger data set, \Robject{itraqdata}, provided with 
 \Rpackage{MSnbase}. It includes feature metadata, accessible with the \Rfunction{fData} method. 
@@
     is of class \Robject{MIAPE}. It stores general information about the 
-    experiment as well as MIAPE (Minimum Information About a Proteomics Experiment, 
-    \citep{Taylor2007, Taylor2008}) information.
+    experiment as well as MIAPE (Minimum Information About a Proteomics Experiment) 
+    information \citep{Taylor2007, Taylor2008}.
     This meta-data can be accessed with the \Rfunction{experimentData} method. 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Spectra objects}\label{sec:spectra}
+\subsection{Spectra objects}\label{sec:spectra}
 
@@
 and stored in a \Robject{environment}. They can be accessed individually with 
-\texttt{itraqdata[["X10"]]} or as a list with \texttt{spectra(itraqdata)}.
+\texttt{itraqdata[["X1"]]} or \texttt{itraqdata[[1]]}, 
+or as a list with \texttt{spectra(itraqdata)}.
 As we have loaded our experiment specifying \texttt{msLevel=2}, 
@@
 precursor charge, \Rfunction{rtime} for the retention time, \Rfunction{mz} 
-for the MZ values, \Rfunction{intensity} for the inensities, \ldots see the 
+for the MZ values, \Rfunction{intensity} for the intensities, \ldots see the 
 \Robject{Spectrum}, \Robject{Spectrum1} and \Robject{Spectrum2} manuals for more details.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Reporter ions}\label{sec:reporterions}
+\subsection{Reporter ions}\label{sec:reporterions}
 
@@
 
-Spectra can be plotted individually or as part of (subsetted) experiments 
+Spectra can be plotted individually or as part of (subset) experiments 
 with the \Robject{plot} method. Full spectra can be plotted 
 (using \texttt{full=TRUE}), specific reporter ions of interest 
-(by specifing with reporters with \texttt{reporters=iTRAQ4} for instance) 
+(by specifying with reporters with \texttt{reporters=iTRAQ4} for instance) 
 or both (see figure \ref{fig:spectrum-plot}).
@@
 \caption{Experiment-wide raw MS2 spectra. The y-axes of the individual 
-  spectra are automatically rescaled to the same range. }
+  spectra are automatically rescaled to the same range. 
+  See section \ref{sec:norm} to rescale peaks identically. }
 \label{fig:msnexp-plot}
@@
 
+<<prepare-spectrum-plot2,echo=FALSE>>=
+p1 <- plot(itraqdata[["X55"]],full=TRUE) + theme_gray(5) 
+p2 <- plot(experiment[["X55"]],full=TRUE) + theme_gray(5) 
+@
+
 \begin{figure}[!ht]
@@
   viewport(layout.pos.row = x, layout.pos.col = y)
-p1 <- plot(itraqdata[["X55"]],full=TRUE) + theme_gray(5) 
-p2 <- plot(experiment[["X55"]],full=TRUE) + theme_gray(5) 
 print(p1,vp=vplayout(1,1))
@@
   to \Robject{t} to 0. The second peak is unaffected. Calling \Rfunction{clean} 
-  after \Rfunction{removePeaks} effectly deletes successive 0 intensities from 
+  after \Rfunction{removePeaks} effectively deletes successive 0 intensities from 
   the spectrum, as shown on the right plot. }
@@
 or greater then \texttt{max(mzmax)} are discarded. This method is 
-particularely useful when one wants to concentrate on a specific MZ range, 
+particularly useful when one wants to concentrate on a specific MZ range, 
 as for reporter ions quantification, and generally results in substantial 
@@
 Quantitation is performed on fixed peaks in the spectra, that are specified 
-with an \Robject{ReporterIon} object. A specific peak is defined by it's expected 
+with an \Robject{ReporterIons} object. A specific peak is defined by it's expected 
 \Robject{mz} value and is searched for within \Robject{mz} $\pm$ \Robject{width}.  
@@
 
-The \Rfunction{quantify} method takes the following paramters: 
+The \Rfunction{quantify} method takes the following parameters: 
 an \Robject{MSnExp} experiment, 
@@
 for quantitation. 
-Additionally, a progress bar can be dispaying when setting 
+Additionally, a progress bar can be displaying when setting 
 the \Robject{verbose} parameter to \Robject{TRUE}.
@@
   Quantitation using \Robject{sum} sums all the data points in the peaks to 
-  produce, for ths example, \Sexpr{quantify(sp,iTRAQ4[1],method="sum")[[1]]}, 
+  produce, for this example, \Sexpr{quantify(ssp,iTRAQ4[1],method="sum")[[1]]}, 
   whereas method \Robject{max} only uses the peak's maximum intensity, 
-  \Sexpr{quantify(sp,iTRAQ4[1],method="max")[[1]]}. 
+  \Sexpr{quantify(ssp,iTRAQ4[1],method="max")[[1]]}. 
   \Robject{Trapezoidation} calculates the area under the peak taking 
@@
 A \Robject{MSnSet} object is meant to be compatible with further downstream 
-packages for data normalisation and statistical analysis. 
-Below is an example of applying 
-variance stabilization normalisation \citep{Huber2002} 
-to the iTRAQ reporter intensities \citep{Karp2010}.
+packages for data normalisation and statistical analysis. There is also a 
+\Rfunction{normalise} method for expression sets. The method takes and instance 
+of class \Robject{MSnSet} as first argument, and a character to describe the 
+\Robject{method} to be used:
 
-<<vsn,echo=TRUE,cache=FALSE>>=
-library("vsn")
-qnt.vsn <- vsn2(exprs(qnt))
+\begin{description}
+\item{\Robject{quantiles}} Applies quantile normalisation \citep{Bolstad03} as 
+  implemented in the \Rfunction{normalize.quantiles} function of the 
+  \Rpackage{preprocessCore} package. 
+\item{\Robject{quantiles.robust}} Applies robust quantile normalisation 
+  \citep{Bolstad03} as implemented in the 
+  \Rfunction{normalize.quantiles.robust} function of the 
+  \Rpackage{preprocessCore} package. 
+\item{\Robject{vsn}} Applies variance stabilisation normalization \citep{Huber2002} 
+  as implemented in the \Rfunction{vsn2} function of the 
+  \Rpackage{vsn} package. 
+\item{\Robject{max}} Each feature's reporter intensity is divided by the 
+  maximum of the reporter ions intensities.
+\item{\Robject{sum}} Each feature's reporter intensity is divided by the 
+  sum of the reporter ions intensities.
+\end{description}
+
+<<normalise,echo=TRUE,cache=FALSE>>=
+qnt.max <- normalise(qnt,"max")
+qnt.sum <- normalise(qnt,"sum")
+qnt.quant <- normalise(qnt,"quantiles")
+qnt.qrob <- normalise(qnt,"quantiles.robust")
+qnt.vsn <- normalise(qnt,"vsn")
 @
 
-Figure 3 of \cite{Karp2010} is reproduced with 
-the dummy experiment and show on figure \ref{fig:cv-plot}.
+The effect of these are illustrated on figure \ref{fig:norm-plot} and
+figure \ref{fig:cv-plot} reproduces figure 3 of \cite{Karp2010} that described 
+the application of vsn on iTRAQ reporter data.
+
+\begin{figure}[p]
+<<norm-plot,fig=T,echo=F,tikz=T,external=F,width=5,height=7>>=
+.plot <- function(x,ttl=NULL) {
+  boxplot(exprs(x),
+          main=ifelse(is.null(ttl),processingData(x)@processing[6],ttl),
+          cex.main=.8,
+          cex.lab=.5,
+          cex.axis=.5,
+          cex=.8)
+  grid()
+}
+oldmar <- par()$mar
+par(mfrow=c(3,2),mar=c(2.9,2.9,2.9,1))
+.plot(qnt,ttl="Non-normalised data")
+.plot(qnt.max)
+.plot(qnt.sum)
+.plot(qnt.quant)
+.plot(qnt.qrob)
+.plot(qnt.vsn)
+@ 
+\caption{ Comparison of the normalisation \Robject{MSnSet} methods. 
+  Note that vsn also glog-transforms the intensities. }
+\label{fig:norm-plot}
+\end{figure}
 
@@
 cv2 <- sd2/mn2
+dfr <- rbind(data.frame(rank=order(mn1),cv=cv1,norm="raw"),
+             data.frame(rank=order(mn2),cv=cv2,norm="vsn"))
 library("zoo")
@@
 rmed2 <- zoo:::rollapply.zoo(cv2,7,function(x) median(x,na.rm=TRUE))
+dfr2 <- 
+  rbind(data.frame(x=seq(1,70,by=(70/length(rmed1))),
+                   y=rmed1,norm="raw"),
+        data.frame(x=seq(1,70,by=(70/length(rmed1))),
+                   y=rmed2,norm="vsn"))
+
+p <- qplot(rank,cv,data=dfr,col=norm) + 
+  geom_line(data=dfr2,aes(x=x,y=y,colour=norm)) +
+  theme_gray(7)
 @
 
+\begin{center}
 \begin{figure}[!ht]
-<<cv-plot,fig=T,echo=F,tikz=T,external=F,width=5,height=5>>=
-plot(order(mn1),cv1,col="steelblue",pch=19,
-     ylab="CV",xlab="Rank mean",
-     cex.axis=.7,cex.lab=.8)
-points(order(mn2),cv2,col="red",pch=19)
-grid()
-lines(seq(1,70,by=(70/length(rmed1))),rmed1,col="steelblue",lwd=1)
-lines(seq(1,70,by=(70/length(rmed2))),rmed2,col="red"      ,lwd=1)
-## legend("topright",
-##        legend=c("log2","vsn","running\nmedian"),
-##        pch=c(19,19,NA),
-##        col=c("steelblue","red","black"),
-##        lwd=c(0,0,2),
-##        bty="n",
-##        cex=.7)
+<<cv-plot,fig=T,echo=F,tikz=T,external=F,width=5,height=4>>=
+print(p)
 @ 
-\caption{CV versus signal intensity comparison for log2 (blue) and vsn (red) transformed data. Lines indicate running CV medians.}
+\caption{CV versus signal intensity comparison for log2 and vsn
+  transformed data. Lines indicate running CV medians.}
 \label{fig:cv-plot}
+\end{figure}
+\end{center}
+
+Note that it is also possible to normalise individual spectra or whole experiments with 
+the \Rfunction{normalise} method using the \Robject{max} method. 
+This will rescale all peaks between 0 and 1. 
+To visualise the relative reporter peaks, one should this first trim the spectra 
+using method \Rfunction{trimMz} as illustrated in section \ref{sec:processing}, 
+then normalise the \Robject{MSnExp} with \Rfunction{normalise} using 
+\Robject{method="max"} as illustrated above and plot the data using \Rfunction{plot} 
+(figure \ref{fig:msnexp-norm-plot}).
+
+<<prepare-msnset-nrom-plot,cache=FALSE,echo=FALSE>>=
+p <- plot(normalise(experiment[bsasel],"max"),reporters=iTRAQ4,full=FALSE)
+p <- p + theme_gray(7)
+@
+
+\begin{figure}[p]
+<<msnexp-norm-plot,fig=T,echo=F,tikz=T,external=F,width=5,height=6>>=
+print(p)
+@ 
+\caption{Experiment-wide normalised MS2 spectra. The y-axes of the individual 
+  spectra is now rescaled between 0 and 1 (highest peak), as opposed to 
+  figure \ref{fig:msnexp-plot}. }
+\label{fig:msnexp-norm-plot}
 \end{figure}
@@
 defined summary function and return a new \Robject{MSnSet} instance. 
-The function has three main argument here described below. 
-Additional details can be found in the function manual.
+The three main arguments are described below. 
+Additional details can be found in the method documentation.
 
@@
 table(gb)
+length(unique(gb))
 @
@@
 (\Robject{fun="mean"}, \Robject{fun="median"}, \Robject{fun="sum"}, 
-\Robject{fun="weigthed.mean"} or \Robject{fun="medianpolish"} to compute respectively the mean, 
+\Robject{fun="weighted.mean"} or \Robject{fun="medianpolish"} to compute respectively the mean, 
 media, sum, weighted mean or median polish of the features to be aggregated). 
-Alternatively, is is possible to supply userdefined functions with 
+Alternatively, is is possible to supply user defined functions with 
 \Robject{fun=function(x) \{ ... \}}. We will use the \Robject{median} here.
@@
 The \Rfunction{purityCorrect} method applies such correction to \Robject{MSnSet} 
-instances. It also requires a square matric as second argument, \Robject{impurities}, 
-that defines the repaltive percentage of reporter in the quantified each peak. 
+instances. It also requires a square matrix as second argument, \Robject{impurities}, 
+that defines the relative percentage of reporter in the quantified each peak. 
 See \texttt{?purityCorrect} for more details.
@@
 
+%% \section{Conclusion}\ref{sec:ccl}
+
 \clearpage

### 2011-05-17 R/plotting-MSnExp.R
misc minor fixes/typos, added plot parameter to plot methods, new meanSdPlot method
git-svn-id: https://hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/MSnbase@55646 bc3139a8-67e5-0310-9ffc-ced21a209358

@@ -1,4 +1,4 @@
-plot.MSnExp <- function(object,reporters,full=FALSE) {
+plot.MSnExp <- function(object,reporters,full=FALSE,plot=TRUE) {
   i <- NULL # to satisfy codetools
   ## plot.MSnExp: no visible binding for global variable ‘i’
   spectraList <- spectra(object)
@@ -33,7 +33,8 @@ plot.MSnExp <- function(object,reporters,full=FALSE) {
 ##    + geom_vline(xintercept=c(reporters@mz+reporters@width,
 ##                   reporters@mz-reporters@width),col="grey")
   }
-  print(p)
+  if (plot)
+    print(p)
   invisible(p)
 }
 

### 2011-05-17 inst/doc/MSnbase-demo/MSnbase-demo.Rnw
misc minor fixes/typos, added plot parameter to plot methods, new meanSdPlot method
git-svn-id: file:///home/git/hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/MSnbase@55646 bc3139a8-67e5-0310-9ffc-ced21a209358

@@
 precursor charge, \Rfunction{rtime} for the retention time, \Rfunction{mz} 
-for the MZ values, \Rfunction{intensity} for the intensities, \ldots see the 
+for the MZ values, \Rfunction{intensity} for the intensities, ... see the 
 \Robject{Spectrum}, \Robject{Spectrum1} and \Robject{Spectrum2} manuals for more details.
@@
 @
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Section
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%\section{Quality control}\label{sec:qc}
 
@@
 \section{Data processing}\label{sec:processing}
+
+\subsection{Cleaning spectra}\label{sec:clean}
 
@@
 \end{figure}
+
+\subsection{Focusing on specific MZ values}\label{sec:trim}
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Section
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %% \section{Conclusion}\ref{sec:ccl}

### 2011-05-17 R/methods-MSnSet.R

@@ -85,7 +85,7 @@ setMethod("purityCorrect",
             if (ncol(impurities)!=nrow(impurities))
               stop("Impurity matrix must be a square matrix")
             if (ncol(object)!=ncol(impurities))
-              stop("Impurity matrix show be",ncol(object),"by",ncol(object))
+              stop("Impurity matrix show be ",ncol(object)," by ",ncol(object))
             .purcor <- function(x,.impurities=impurities) {
               keep <- !is.na(x)
               if (sum(keep)>1) 
@@ -138,3 +138,10 @@ setMethod("processingData",
 
 setMethod("msInfo","MSnSet",
           function(object) msInfo(experimentData(object)))
+
+setMethod("meanSdPlot",
+          signature="MSnSet",
+          definition =
+          function(x, ranks=TRUE, xlab = ifelse(ranks, "rank(mean)", "mean"),
+                   ylab = "sd", pch  = ".", plot = TRUE, ...)
+          vsn::meanSdPlot(exprs(x), ranks=ranks, xlab=xlab, ylab=ylab, pch=pch, plot=plot, ...))

### 2011-05-17 R/methods-Spectrum.R

@@ -135,12 +135,10 @@ setReplaceMethod("centroided",
                  })
 
 setMethod("normalise","Spectrum",
-          function(object,method="max",...) {
-            if (method!="max")
-              warning("Normalising spectrum with 'max'.")
-            normalise.Spectrum(object,method="max")
+          function(object,method=c("max","sum"),...) {
+            normalise.Spectrum(object,method=match.arg(method))
         })
 
 setMethod("normalize","Spectrum",
-          function(object,method="max",...)
+          function(object,method=c("max","sum"),...)
           normalise(object,method=method))

### 2011-05-17 R/plotting-Spectrum2.R

@@ -1,6 +1,7 @@
 plot.Spectrum2 <- function(spectrum,
                            reporters=NULL,full=FALSE,
                            centroided,
+                           plot=TRUE,                          
                            w1,w2) {
   if (!full & is.null(reporters))
     stop("Please provide repotrer ions if you do not want a full spectrum.")
@@ -71,12 +72,15 @@ plot.Spectrum2 <- function(spectrum,
   }
   ## plotting
   if (full) {
-    print(p+title,vp=mainvp)
-    if (!is.null(reporters)) 
-      print(reps,vp=subvp)
+    if (plot) {
+      print(p+title,vp=mainvp)
+      if (!is.null(reporters))
+        print(reps,vp=subvp)
+    }
     invisible(p+title)
   } else {
-    print(reps+title,vp=mainvp)
+    if (plot)
+      print(reps+title,vp=mainvp)
     invisible(reps+title)
   }
 }

### 2011-05-27 inst/doc/MSnbase-demo/MSnbase-demo.Rnw
new vignette sections (customising plots and QC), plot2D and plotDensity, preprocSelection[Table]
git-svn-id: file:///home/git/hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/MSnbase@55883 bc3139a8-67e5-0310-9ffc-ced21a209358

@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Section
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Spectra objects}\label{sec:spectra}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Section
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Reporter ions}\label{sec:reporterions}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%\section{Quality control}\label{sec:qc}
+\section{Plotting raw data}\label{sec:plotting}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Section
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Plotting raw data}\label{sec:plotting}
+\subsection{Default plots}\label{sec:defaultplots}
 
@@
 
-\begin{figure}[!ht]
+\begin{figure}[p]
 <<msnexp-plot,fig=T,echo=T,tikz=T,external=F,width=5,height=6>>=
@@
 \end{figure}
+
+\subsection{Customising your plots}\label{sec:customplots}
+
+The \Rpackage{MSnbase} \Rfunction{plot} methods have a logical \Robject{plot} parameter 
+(default is \Robject{TRUE}), that specifies if the plot should be printed to the current 
+device. A plot object is also (invisibly) returned, so that it can be saved as a variable 
+for later use or for customisation.
+
+\Rpackage{MSnbase} uses the \Rpackage{ggplot2} package to generate figures, 
+which can subsequently easily be customised. 
+More details about \Rpackage{ggplot2} can be found in \cite{ggplot2} (especially chapter 8) 
+and on \href{http://had.co.nz/ggplot2/}{http://had.co.nz/ggplot2/}. 
+Finally, if a plot object has been saved in a variable \Robject{p}, 
+it is possible to obtain a summary about the object with \Rfunction{summary(p)}. 
+To view the data frame used to generate the plot, use \Rfunction{p@data}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Section
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Quality control}\label{sec:qc}
+
+The current section is not executed dynamically for package size and processing time constrains. 
+The figures and tables have been generated with the respective methods and included statically 
+in the vignette for illustration purposes.
+
+\bigskip
+
+\Rpackage{MSnbase} allows easy and flexible access to the data, which allows to 
+visualise data features to assess it's quality. Some methods are readily available, 
+although many QC approaches will be experiment specific and users are encourage to 
+explore their data.
+
+The \Rfunction{plot2d} method takes one \Robject{MSnExp} instance as firs argument 
+to produce retention time \textit{vs.} precursor MZ scatter plots. Points represent 
+individual MS2 spectra and can be coloured based on 
+precursor charge (with second argument \Robject{z="charge"}), 
+total ion current (\Robject{z="tic"}),
+number of peaks in the MS2 spectra \Robject{z="peaks.count"}) or, 
+when multiple data files were loaded, file \Robject{z="file"}), 
+as illustrated on figure \ref{fig:plot2d}. 
+The lower right panel is produced for only a subset of proteins. 
+See the method documentation for more details.
+
+\begin{figure}[htb]
+  \begin{center}
+    \includegraphics[width=\linewidth]{./plot2d-figure.png}
+\caption{ Illustration of the \Rfunction{plot2d} output. } 
+\label{fig:plot2d}
+  \end{center}
+\end{figure}
+
+The \Rfunction{plotDensity} method illustrates the distribution 
+of several parameters of interest (see figure \ref{fig:plotDensity}).
+Similarly to \Rfunction{plot2d}, the first argument is an \Robject{MSnExp} instance. 
+The second is one of \Robject{precursor.mz}, \Robject{peaks.count} or \Robject{tic}, 
+whose density will be plotted. 
+An optional third argument specifies whether the x axes should be logged.
+
+\begin{figure}[htb]
+  \begin{center}
+    \includegraphics[width=\linewidth]{./plotDensity-figure.png}
+\caption{ Illustration of the \Rfunction{plotDensity} output. } 
+\label{fig:plotDensity}
+  \end{center}
+\end{figure}
+
+%% xtable(preprocSelectionTable(velos))
+%% % latex table generated in R 2.14.0 by xtable 1.5-6 package
+%% % Thu May 19 10:22:29 2011
+%% \begin{table}[ht]
+%% \begin{center}
+%% \begin{tabular}{rr}
+%%   \hline
+%%  & x \\ 
+%%   \hline
+%% 1 & 6836 \\ 
+%%   2 &  59 \\ 
+%%   3 &   5 \\ 
+%%    \hline
+%% \end{tabular}
+%% \end{center}
+%% \end{table}
+
+%% matplot on iTRAQ5 quantitation (see allquant code chunk)
+
+\bigskip
+
+In section \ref{sec:incompdissoc} on page \pageref{sec:incompdissoc}, 
+we illustrate how to assess incomplete reporter ion dissociation.
 

### 2011-06-13 inst/doc/MSnbase-demo/MSnbase-demo.Rnw
new plotting-dataframe.R file
git-svn-id: https://hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/MSnbase@56143 bc3139a8-67e5-0310-9ffc-ced21a209358

@@
 
+\subsection{Purity correction}
+\label{sec:purcor}
+
+Impurities in the reporter reagents can bias the results and can be corrected when 
+manufacturers provide correction coefficients. 
+These generally come as percentages of each reporter ion that have masses differing 
+by -2, -1, +1 and +2 Da from the nominal reporter ion mass due to isotopic variants.  
+The \Rfunction{purityCorrect} method applies such correction to \Robject{MSnSet} 
+instances. It also requires a square matrix as second argument, \Robject{impurities}, 
+that defines the relative percentage of reporter in the quantified each peak. 
+See \texttt{?purityCorrect} for more details.
+
+<<purityCorrect,echo=TRUE,cache=FALSE>>=
+impurities <- matrix(c(0.929,0.059,0.002,0.000,
+                       0.020,0.923,0.056,0.001,
+                       0.000,0.030,0.924,0.045,
+                       0.000,0.001,0.040,0.923),
+                     nrow=4)
+qnt.crct <- purityCorrect(qnt,impurities)
+head(exprs(qnt))
+head(exprs(qnt.crct))
+@
+
+The \Rfunction{makeImpuritiesMatrix} can be used to create impurity matrices. 
+It opens a rudimentary spreadsheet that can be directly edited.
+
 \subsection{Normalisation}\label{sec:norm}
@@
 
-Note that it is also possible to normalise individual spectra or whole experiments with 
+Note that it is also possible to normalise individual spectra or whole \Robject{MSnExp} experiments with 
 the \Rfunction{normalise} method using the \Robject{max} method. 
@@
 %% to combine features.
-
-\subsection{Purity correction}
-\label{sec:purcor}
-
-Impurities in the reporter reagents can bias the results and can be corrected when 
-manufacturers provide correction coefficients. 
-These generally come as percentages of each reporter ion that have masses differing 
-by -2, -1, +1 and +2 Da from the nominal reporter ion mass due to isotopic variants.  
-The \Rfunction{purityCorrect} method applies such correction to \Robject{MSnSet} 
-instances. It also requires a square matrix as second argument, \Robject{impurities}, 
-that defines the relative percentage of reporter in the quantified each peak. 
-See \texttt{?purityCorrect} for more details.
-
-<<purityCorrect,echo=TRUE,cache=FALSE>>=
-impurities <- matrix(c(0.929,0.059,0.002,0.000,
-                       0.020,0.923,0.056,0.001,
-                       0.000,0.030,0.924,0.045,
-                       0.000,0.001,0.040,0.923),
-                     nrow=4)
-qnt2.crct <- purityCorrect(qnt2,impurities)
-head(exprs(qnt2))
-head(exprs(qnt2.crct))
-@
 

### 2011-06-13 R/plotting-dataframe.R

@@ -1,0 +1,35 @@
+plot2d.header <- function(object, ## MSnExp header
+                          z=c("tic","file","peaks.count","charge"),
+                          alpha, plot) {
+  z <- match.arg(z)
+  stopifnot(c("retention.time","precursor.mz",z) %in% names(object))
+  peaks.count <- charge <- retention.time <- precursor.mz <- NULL # to satisfy codetools
+  p <- ggplot(object,aes(retention.time,precursor.mz)) + labs(colour=z)
+  switch(z,
+         tic = p <- p + geom_point(aes(colour=tic), alpha=alpha),
+         peaks.count = p <- p + geom_point(aes(colour=peaks.count), alpha=alpha),
+         file = p <- p + geom_point(aes(colour=as.factor(file)), alpha=alpha),
+         charge = p <- p + geom_point(aes(colour=as.factor(charge)), alpha=alpha))
+  if (plot)
+    print(p)
+  invisible(p)
+}
+
+plotDensity.header <- function(object, ## MSnExp header
+                               z=c("precursor.mz","peaks.count","tic"),
+                               log, plot) {
+  z <- match.arg(z)
+  stopifnot(c("charge",z) %in% names(object))
+  peaks.count <- charge <- tic <- precursor.mz <- NULL # to satisfy codetools
+  object$charge <- as.factor(object$charge)
+  switch(z,
+         precursor.mz = p <- ggplot(object,aes(precursor.mz)) + xlab("Precursor M/Z"),
+         peaks.count = p <- ggplot(object,aes(peaks.count)) + xlab("Peaks count"),
+         tic = p <- ggplot(object,aes(tic)) + xlab("Total ion current"))
+  p <- p + geom_freqpoly(aes(group=charge,colour=charge))
+  if (log)
+    p <- p + scale_x_log10()
+  if (plot)
+    print(p)
+  invisible(p)
+}

### 2011-06-26 inst/doc/MSnbase-demo/MSnbase-demo.Rnw
added plotMzDelta paragraph in demo vignette
git-svn-id: file:///home/git/hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/MSnbase@56414 bc3139a8-67e5-0310-9ffc-ced21a209358

@@
 %% matplot on iTRAQ5 quantitation (see allquant code chunk)
+
+\bigskip
+
+The \Rfunction{plotMzDelta} method\footnote{The code to generate the 
+  histograms has been contributed by Guangchuang Yu from Jinan University, 
+  China.} implements the M/Z delta plot from \cite{Foster11} 
+The M/Z delta plot illustrates the suitability of MS2 spectra for
+identification by plotting the M/Z differences of the most intense
+peaks. The resulting histogram should optimally shown outstanding
+bars at amino acid residu masses. More details and parameters 
+are described in the method documentation (\Rfunction{?plotMzDelta}).
+Figure \ref{fig:plotMzDelta-pride12011} has been generated using 
+the PRIDE experiment 12011, as in \cite{Foster11}.
+
+\begin{figure}[htb]
+  \begin{center}
+    \includegraphics[width=.7\linewidth]{./plotMzDelta-pride12011.pdf}
+\caption{ Illustration of the \Rfunction{plotMzDelta} output for the PRIDE experiment 12011, 
+  as in figure 4A from \cite{Foster11}. } 
+\label{fig:plotMzDelta-pride12011}
+  \end{center}
+\end{figure}
 

### 2011-09-08 inst/doc/MSnbase-demo/MSnbase-demo.Rnw
new io functions and doc
 o added a readMSnSet function <2011-09-08 Thu>
 o added a write.MSnSet("MSnSet") method  <2011-09-08 Thu>
 o vignette/man updates - mainly data import section documenting 
   readMSData and readMgfData <2011-09-08 Thu>



git-svn-id: file:///home/git/hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/MSnbase@57904 bc3139a8-67e5-0310-9ffc-ced21a209358

@@
 \newcommand{\Robject}[1]{{\texttt{#1}}}
-\newcommand{\Rpackage}[1]{{\textit{#1}}}
+%%\newcommand{\Rpackage}[1]{{\textit{#1}}}
+\newcommand{\Rpackage}[1]{{\mbox{\normalfont\textsf{#1}}}}
 \newcommand{\email}[1]{\href{mailto:#1}{\normalfont\texttt{#1}}}
@@
 
-\subsection{Importing experiments}\label{sec:msnexp}
+\subsection{Importing experiments}\label{sec:io}
+
+\Rpackage{MSnbase} is able to import raw MS data stored in one of the \texttt{XML}-based formats 
+as well as peak lists in the \texttt{mfg} format\footnote{%%
+Mascot Generic Format -- \url{http://www.matrixscience.com/help/data\_file\_help.html\#GEN}}
+
+\paragraph{Raw data} The \texttt{XML}-based formats, \texttt{mzXML} \citep{Pedrioli2004}, 
+\texttt{mzData} \citep{Orchard2007} and \texttt{mzML} \citep{Martens2010} can be imported 
+with the \Rfunction{readMSData} function, as illstrated below 
+(see \Rfunction{?readMSData} for more details).
+
+<<readdata, echo=TRUE, cache=FALSE>>=
+file <- dir(system.file(package="MSnbase",dir="extdata"),
+            full.name=TRUE,
+            pattern="mzXML$")
+rawdata <- readMSData(file, msLevel=2, verbose=FALSE)
+@
+
+Either MS1 or MS2 spectra can be loaded at a time by setting the \texttt{msLevel} parameter accordingly.
+In this document, we will use the \Robject{itraqdata} data set, provided with 
+\Rpackage{MSnbase}. It includes feature metadata, accessible with the \Rfunction{fData} accessor. 
+The metadata includes identification data for the \Sexpr{length(itraqdata)} MS2 spectra.
+
+\paragraph{Peak lists} Peak lists can often be exported after spectrum processing 
+from vendor-specific software and are also used as input to search engines. 
+Peak lists in \texttt{mgf} format can be imported with the function 
+\Rfunction{readMgfData} (see \Rfunction{?readMgfData} for details)
+to create experiment objects. 
+Experiments or individual spectra can be exported to an \texttt{mgf} file with 
+the \Rfunction{writeMgfData} methods (see \Rfunction{?writeMgfData} for details and examples).
+
+\bigskip
+
+See also section \ref{sec:io2} to import quantitative data stored in spreadsheets into \R 
+for further processing using \Rpackage{MSnbase}.
+
+\subsection{MS experiments}\label{sec:msnexp}
 
@@
 as defined by one or multiple raw data files.
-
-\Rpackage{MSnbase} imports raw MS data stored in \texttt{mzXML} 
-format \citep{Pedrioli2004} with the function \Rfunction{readMzXMLData} function. 
-See \Rfunction{?readMzXMLData} for more details.
-More formats, notably \texttt{mzML} \citep{Martens2010}, will be added at a later stage. 
-Either MS1 or MS2 spectra can be loaded at a time by setting the \texttt{msLevel} parameter accordingly.
-In this document, we will use a slightly bigger data set, \Robject{itraqdata}, provided with 
-\Rpackage{MSnbase}. It includes feature metadata, accessible with the \Rfunction{fData} method. 
-The metadata includes identification data for the \Sexpr{length(itraqdata)} MS2 spectra.
 
@@
 
-\subsection{Purity correction}
-\label{sec:purcor}
+\subsection{Importing quantitation data}\label{sec:io2}
 
-Impurities in the reporter reagents can bias the results and can be corrected when 
-manufacturers provide correction coefficients. 
+If quantitation data is already available as a spreadsheet, it can be 
+imported, along with additional optional feature and sample (pheno) meta data, 
+with the \Rfunction{readMSnSet} function. This function takes the 
+respective text-based spreadsheet (comma- or tab-separated) file names 
+as argument to create a valid \Robject{MSnSet} instance.
+
+Note that the quantitation data of \Robject{MSnSet} objects can also be 
+exported to a text-based spreadsheet file using the \Rfunction{write.exps} 
+method.
+
+\subsection{Peak adjustments}\label{sec:purcor}
+
+\paragraph{Single peak adjustment} In certain cases, peak intensities need to be adjusted 
+as a result of peak interferance. For example, the $+1$ peak of the phenylalanine (F, Phe) 
+immonium ion (with m/z 120.03) inteferes with the 121.1 TMT reporter ion. 
+Below, we calculate the relative intensity of the +1 peaks compared to the main peak using the 
+\Rpackage{Rdispo} package.
+
+<<pheplus1, echo=TRUE, cache=FALSE>>=
+library(Rdisop)
+## Phenylalanine immonium ion
+Fim <- getMolecule("C8H10N")
+getMass(Fim)
+isotopes <- getIsotope(Fim)
+F1 <- isotopes[2,2]
+F1
+@
+
+If desired, one can thus specifically quantify the F immonium ion in the MS2 spectrum, 
+estimate the intensity of the +1 ion (\Sexpr{round(F1,4)}\% of the F peak) and substract 
+this calculated value from the 121.1 TMT reporter intensity.
+
+The above principle can also be generalised for a set of overlapping peaks, 
+as described below.
+
+\paragraph{Reporter ions purity correction} Impurities in the reporter reagents can also 
+bias the results and can be corrected when manufacturers provide correction coefficients. 
 These generally come as percentages of each reporter ion that have masses differing 

### 2011-09-20 R/DataClasses.R
Spectrum2 ms1scan slot updated to precScanNum, using mzR::header()$precursorScanNum to populate. Updated and exported readIspyData.
 o removed (internal) Mascot query link column in readIspyData 
   to work with latest ouput version <2011-09-12 Mon>
 o removed the fillUp part in readIspyData <2011-09-14 Wed>
 o exported readIspyData <2011-09-20 Tue>
 o removed link to proteomics sig list <2011-09-20 Tue>
 o removed url in DESCRIPTION <2011-09-20 Tue>
 o Spectrum2 slot ms1scan renamed to precScanNum and populated using 
   mzR::header()$precursorScanNum. Updated affected methods/functions 
   a

@@ -214,7 +214,7 @@ setClass("Spectrum",
 setClass("Spectrum2",
          representation = representation(
            merged="numeric",
-           ms1scan="integer",
+           precScanNum="integer",
            precursorMz="numeric",
            precursorIntensity = "numeric",
            precursorCharge = "integer",
@@ -225,7 +225,7 @@ setClass("Spectrum2",
                versions=c(classVersion("Spectrum"), Spectrum2="0.1.0")),
            merged = 1,
            acquisitionNum = integer(),
-           ms1scan = integer(),
+           precScanNum = integer(),
            precursorMz = numeric(),
            precursorIntensity = numeric(),
            msLevel = as.integer(2),

### 2011-09-20 R/methods-MSnExp.R

@@ -19,7 +19,7 @@ setMethod("show",
               msnLevels <- unique(msLevel(object))
               cat(" MSn level(s):",msnLevels,"\n")
               if (all(msLevel(object)>1)) {
-                cat(" Number of MS1 acquisitions:",length(unique(ms1scan(object))),"\n")
+                cat(" Number of MS1 acquisitions:",length(unique(precScanNum(object))),"\n")
                 cat(" Number of MS2 scans:",length(ls(assayData(object))),"\n")
                 msnPrecMz <- precursorMz(object)
                 nbPrecIons <- length(msnPrecMz)
@@ -113,13 +113,6 @@ setMethod("clean",
 setMethod("removePeaks",signature("MSnExp"),
           function(object,t,verbose=TRUE) removePeaks.MSnExp(object,t,verbose))
 
-
-setMethod("ms1scan","MSnExp",
-          function(object) {
-            if (msLevel(object)[1]>1) 
-              return(unlist(eapply(assayData(object), ms1scan)))
-            stop("This experiment contains MS1 spectra.")
-          })
 
 setMethod("trimMz",
           signature=signature("MSnExp","numeric"),

### 2011-09-20 inst/doc/Foreword.tex
fixed typo in vignette foreword
git-svn-id: https://hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/MSnbase@55303 bc3139a8-67e5-0310-9ffc-ced21a209358
Spectrum2 ms1scan slot updated to precScanNum, using mzR::header()$precursorScanNum to populate. Updated and exported readIspyData.
 o removed (internal) Mascot query link column in readIspyData 
   to work with latest ouput version <2011-09-12 Mon>
 o removed the fillUp part in readIspyData <2011-09-14 Wed>
 o exported readIspyData <2011-09-20 Tue>
 o removed link to proteomics sig list <2011-09-20 Tue>
 o removed url in DESCRIPTION <2011-0

--- 2011-03-24 inst/doc/Foreword.tex
@@
 This library is one attempt to do so.
+
+\bigskip
+
+You are welcome to contact me for questions, bugs, typos or suggestions about \Rpackage{MSnbase}. 
+If you wish to reach a broader audience for general questions about proteomics analysis using \texttt{R}, 
+you may want to use the Bioconductor mailing 
+list\footnote{\url{https://stat.ethz.ch/mailman/listinfo/bioconductor}}.
+%% proteomics special interest 
+%% group\footnote{\url{https://stat.ethz.ch/mailman/listinfo/bioc-sig-proteomics}}.

### 2011-09-28 inst/doc/MSnbase-demo/MSnbase-demo.Rnw
readMgfData update and new MSnbase-io vignette
 o in readMgfData, SCANS is not used to populate 
   acquisitionNum anymore, as several scans might 
   have been combined upstreams <2011-09-26 Mon>
 o added fillUp function in utils.R and exported <2011-09-27 Tue>
 o new MSnbase-io vignette <2011-09-28 Wed>




git-svn-id: file:///home/git/hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/MSnbase@58519 bc3139a8-67e5-0310-9ffc-ced21a209358

@@
 \newcommand{\Robject}[1]{{\texttt{#1}}}
-%%\newcommand{\Rpackage}[1]{{\textit{#1}}}
 \newcommand{\Rpackage}[1]{{\mbox{\normalfont\textsf{#1}}}}
@@
   This vignette describes the functionality implemented in the \Rpackage{MSnbase} package. 
-  \Rpackage{MSnbase} aims at (1) facilitating the upload, processing, visualisation and 
+  \Rpackage{MSnbase} aims at (1) facilitating the import, processing, visualisation and 
   quantification of mass spectrometry data into the \R environment \citep{Rstat} by providing 

### 2011-09-28 inst/doc/MSnbase-io/MSnbase-io.Rnw
new vignette files
git-svn-id: https://hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/MSnbase@58520 bc3139a8-67e5-0310-9ffc-ced21a209358

%\VignetteIndexEntry{MSnbase IO capabilities}
%\VignetteKeywords{Mass Spectrometry, Proteomics, Infrastructure }
%\VignettePackage{MSnbase-io}
\documentclass{article}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows,shadows,fit}
\usepackage[nogin]{Sweave}
\usepackage{hyperref}
\usepackage[authoryear,round]{natbib}

\SweaveOpts{echo=T,eval=T,fig=F,results=verbatim,cache=F,tikz=T,external=T,prefix.string=sweave-cache/figs/fig}

\newcommand{\R}{\texttt{R} }
\newcommand{\Rfunction}[1]{{\texttt{#1}}}
\newcommand{\Robject}[1]{{\texttt{#1}}}
\newcommand{\Rpackage}[1]{{\mbox{\normalfont\textsf{#1}}}}
\newcommand{\email}[1]{\href{mailto:#1}{\normalfont\texttt{#1}}}
%% colors
\definecolor{Red}{rgb}{0.7,0,0}
\definecolor{Blue}{rgb}{0,0,0.8}

\hypersetup{%
  hyperindex = {true},
  colorlinks = {true},
  linktocpage = {true},
  plainpages = {false},
  linkcolor = {Blue},
  citecolor = {Blue},
  urlcolor = {Red},
  pdfstartview = {Fit},
  pdfpagemode = {UseOutlines},
  pdfview = {XYZ null null null}
}

\input{Author.tex}

\begin{document}

%% pgf settup
\pgfdeclarelayer{background}
\pgfdeclarelayer{foreground}
\pgfsetlayers{background,main,foreground}
% Define block styles
\tikzstyle{input} = [rectangle, draw, fill=blue!20,
  text width=6em, text centered, rounded corners, minimum height=4em]
\tikzstyle{fun} = [rectangle, draw, fill=white, drop shadow, 
  text width=7em, text centered, rounded corners, minimum height=2em]
\tikzstyle{obj} = [rectangle, draw, fill=red!20, 
  text width=5em, text centered, rounded corners, minimum height=5em]

\title{\Rpackage{MSnbase} input/output capabilities}

\maketitle

%% Abstract and keywords %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vskip 0.3in minus 0.1in
\hrule
\begin{abstract}
  This vignette describes \Rpackage{MSnbase}'s input and output 
  capabilites.
\end{abstract}
\textit{Keywords}: Mass Spectrometry (MS), proteomics, infrastructure, IO.
\vskip 0.1in minus 0.05in
\hrule
\vskip 0.2in minus 0.1in
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\tableofcontents

%% <<setup,cache=FALSE>>=
%% setCacheDir("sweave-cache/values")
%% @

%% <<environment,cache=FALSE,echo=FALSE>>=
%% library(MSnbase)
%% @

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%\input{Foreword.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Overview}

\Rpackage{MSnbase}'s aims are to facilitate the reproducible analysis of 
mass spectrometry data within the \R environment, from raw data import and processing, 
feature quantification, quantification and statistical analysis of the results.
Several input functions are provided to allow execution of all of several 
of these steps. Intermediate or final results can also be saved or exported. 
These capabilites are presented below.

\section{Data input}

\paragraph{Raw data} Data stored in one of the published \texttt{XML}-based 
formats. i.e. \texttt{mzXML} \citep{Pedrioli2004}, \texttt{mzData} \citep{Orchard2007} 
or \texttt{mzML} \citep{Martens2010}, 
can be imported with the \Rfunction{readMSData} method, which makes use of 
the \Rpackage{mzR} package to create \Robject{MSnExp} objects. 
The files can be in profile or centroided mode. 
See \Rfunction{?readMSData} for details.

\paragraph{Peak lists} Peak lists in the \texttt{mgf} 
format\footnote{\url{http://www.matrixscience.com/help/data\_file\_help.html\#GEN}} 
can be imported using the \Rfunction{readMgfData}. 
In this case, the peak data has generally been pre-processed by other software.
See \Rfunction{?readMgfData} for details.

\paragraph{Quantitation data} Third party software can be used to generate 
quantitative data and exported as a spreadsheet (generally comma or tab separated format). 
This data as well as any additional metadata can be imported with the 
\Rfunction{readMSnSet} function. See \Rfunction{?readMSnSet} for details.

\begin{figure}[!htb] %% input
  \begin{center}
    \begin{tikzpicture}[node distance = 2cm, auto]
      % Place nodes
      \node [input] (raw) {Raw data in and open \texttt{XML} format};
      \node [fun, left of=raw, node distance=4cm] (readMSData) {\Rfunction{readMSData}};
      \node [input, below of=raw] (mgf) {Peak list in \texttt{mgf} format};
      \node [fun, left of=mgf, node distance=4cm] (readMgfData) {\Rfunction{readMgfData}};
      \node [input, below of=mgf] (spreadsheet) {Quantitation data as a spreadsheet};
      \node [fun, left of=spreadsheet, node distance=4cm] (readMSnSet) {\Rfunction{readMSnSet}};
      \node [obj, left of=readMSData, yshift=-2.5mm, node distance=4cm] (MSnExp) {\Robject{MSnExp}};
      \node [obj, left of=readMSnSet, node distance=4cm] (MSnSet) {\Robject{MSnSet}};
      \node [fun, below of=MSnExp] (quantify) {\Rfunction{quantify}};
      % Background
      \begin{pgfonlayer}{background}
        \node [fill=yellow!20,rounded corners, draw=black!50, dashed, fit=(MSnExp) (quantify) (MSnSet)] {};
      \end{pgfonlayer}
      % Draw edges
      \draw  (raw) -- (readMSData);
      \draw  (mgf) -- (readMgfData);
      \draw  (spreadsheet) -- (readMSnSet);
      \draw [->] (readMSData) -- (MSnExp);
      \draw [->] (readMgfData) -- (MSnExp);
      \draw [->] (readMSnSet) -- (MSnSet);
      \draw  (MSnExp) -- (quantify);
      \draw [->] (quantify) -- (MSnSet);
    \end{tikzpicture}
    \caption{Illustration of \texttt{MSnbase} input capabilities. 
      The white and red boxes represent \R functions/methods and objects respectively. 
      The blue boxes represent different disk storage formats.}
    \label{fig:input}
  \end{center}
\end{figure}

\section{Data output}

\paragraph{RData files} \R objects can most easily be stored on disk with the \Rfunction{save} function.
It creates compressed binary images of the data representation that can later be read back from the file 
with the \Rfunction{load} function.

\paragraph{Peak lists} \Robject{MSnExp} instances as well as individual spectra can be written 
as \texttt{mgf} files with the \Rfunction{writeMgfData} method. Note that the metadata in the original 
\R object can not be included in the file. See \Rfunction{?writeMgfData} for details.

\paragraph{Quantitation data} Quantitation data can be exported to spreadsheet files with 
the \Rfunction{write.exprs} method. Feature metadata can be appended to the feature 
intenstiy values. See \Rfunction{?writeMgfData} for details.

\begin{figure}[!htb] %% out
  \begin{center}
    \begin{tikzpicture}[node distance = 2cm, auto]
      % Place nodes
      \node [input] (raw) {Raw data in and open \texttt{XML} format};
      \node [input, below of=raw] (mgf) {Peak list in \texttt{mgf} format};
      \node [fun, right of=mgf, node distance=4cm, yshift=1cm] (writeMgfData) {\Rfunction{writeMgfData}};
      \node [input, below of=mgf] (spreadsheet) {Quantitation data as a spreadsheet};
      \node [fun, right of=spreadsheet, node distance=4cm] (writeexprs) {\Rfunction{write.exprs}};

      \node [obj, right of=writeMgfData, yshift=1.25cm, node distance=4cm] (MSnExp) {\Robject{MSnExp}};
      \node [obj, right of=writeexprs, node distance=4cm] (MSnSet) {\Robject{MSnSet}};
      \node [fun, below of=MSnExp] (quantify) {\Rfunction{quantify}};
      % Background
      \begin{pgfonlayer}{background}
        \node [fill=yellow!20,rounded corners, draw=black!50, dashed, fit=(MSnExp) (quantify) (MSnSet)] {};
      \end{pgfonlayer}
      % Draw edges
      \draw  (MSnExp) -- (writeMgfData);
      \draw [->] (writeMgfData) -- (mgf);
      \draw (MSnSet) -- (writeexprs);
      \draw [->] (writeexprs) -- (spreadsheet);
      \draw  (MSnExp) -- (quantify);
      \draw [->] (quantify) -- (MSnSet);
    \end{tikzpicture}
    \caption{Illustration of \texttt{MSnbase} output capabilities.
      The white and red boxes represent \R functions/methods and objects respectively. 
      The blue boxes represent different disk storage formats.}
    \label{fig:output}
  \end{center}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% \section{Session information}
%% \label{sec:sessionInfo}

%% <<label=sessioninfo,results=tex,echo=FALSE,cache=FALSE>>=
%% toLatex(sessionInfo())
%% @

\bibliographystyle{plainnat}
\bibliography{MSnbase}

\end{document}

### 2011-09-29 inst/doc/MSnbase-io/MSnbase-io.Rnw
typo in io vignette
git-svn-id: https://hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/MSnbase@58522 bc3139a8-67e5-0310-9ffc-ced21a209358
rephrasing in io vignette
git-svn-id: file:///home/git/hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/MSnbase@58554 bc3139a8-67e5-0310-9ffc-ced21a209358

@@
 feature quantification, quantification and statistical analysis of the results.
-Several input functions are provided to allow execution of all of several 
-of these steps. Intermediate or final results can also be saved or exported. 
+Data import functions for several formats are provided and intermediate or final 
+results can also be saved or exported. 
 These capabilites are presented below.
@@
       % Place nodes
-      \node [input] (raw) {Raw data in and open \texttt{XML} format};
+      \node [input] (raw) {Raw data in an open \texttt{XML} format};
       \node [fun, left of=raw, node distance=4cm] (readMSData) {\Rfunction{readMSData}};
@@
       % Place nodes
-      \node [input] (raw) {Raw data in and open \texttt{XML} format};
+      \node [input] (raw) {Raw data in an open \texttt{XML} format};
       \node [input, below of=raw] (mgf) {Peak list in \texttt{mgf} format};

### 2011-10-03 inst/doc/MSnbase-demo/MSnbase-demo.Rnw
Added cache capabilities to pSet, readMzXMLData deprecated
 o added .cache slot to pSet class <2011-10-03 Mon>
 o added pSet initialize method to set .cache and 
   .cache$level as '0'. <2011-10-03 Mon>
 o Check that level is defined in .cache and env is locked 
   in pSet validity method. <2011-10-03 Mon>
 o updated itraqdata.RData  <2011-10-03 Mon>
 o Deprecated readMzXMLData, added defunct.Rd <2011-10-03 Mon>
 o changed readMzXMLData to readMSData in man <2011-10-03 Mon>
 o updated show MSnExp for speed <2011-10-03 Mon>
 o updated read*Data is support cache = [0|1]. <2011-10-03 Mon>
 o upda

@@
 <<readdata, echo=TRUE, cache=FALSE>>=
-file <- dir(system.file(package="MSnbase",dir="extdata"),
-            full.name=TRUE,
-            pattern="mzXML$")
+file <- dir(system.file(package="MSnbase", dir="extdata"),
+            full.names=TRUE, pattern="mzXML$")
 rawdata <- readMSData(file, msLevel=2, verbose=FALSE)

### 2011-10-03 R/cache.R

@@ -1,0 +1,86 @@
+testCacheArg <- function(cache, maxCache=3) {
+  ## Function used to test the value of a 'cache'
+  ## parameter in a read*Data function
+  ## Parameters:
+  ##  cache: value of the cache argument to test
+  ##  maxCache: max value allowed. Generally
+  ##            3, but could be less, depending
+  ##            on the input data. maxCache is 1
+  ##            for readMgfData for instance.
+  ## Value: valid (possibly updated) cache value
+  if (!is.numeric(cache))
+    stop("'cache' must be numeric.")
+  if (cache < 0 | cache > maxCache) {
+    warning("cache must be [0:",maxCache,"]!")
+    if (cache < 0) {
+      warning("Setting cache to 0.")
+      cache <- 0
+    } else {
+      warning("Setting cache to ",maxCache,".")
+      cache <- maxCache
+    }
+  }
+  return(cache)
+}
+
+newCacheEnv <- function(assaydata, level=0, lock=TRUE) {
+  ## Set the .cache slot of a pSet object.
+  ## Parameters
+  ##  assaydata: environment - pSet assaydata slot
+  ##  level: numeric - cache level
+  ##  lock: logical - lock env and bindings (default is TRUE)
+  ## Return:
+  ##  A new cache environment
+  cacheEnv <- new.env()
+  assign("level", level, cacheEnv)
+  if (level >= 1) {
+    assign("rangePrecursorMz", range(unname(eapply(assaydata,precursorMz))), cacheEnv)
+    assign("rangeMz", range(unname(eapply(assaydata,mz))), cacheEnv)
+    assign("rangeRtime", range(unname(eapply(assaydata,rtime))), cacheEnv)
+  }
+  ## levels 2 and 3 not yet implemented
+  if (lock)
+    lockEnvironment(cacheEnv, bindings=TRUE)
+  return(cacheEnv)
+}
+
+setCacheEnv <- function(object, level, lock=TRUE) {
+  ## Set the .cache slot of a pSet object.
+  ## Parameters
+  ##  object: MSnExp instance
+  ##  level: numeric - cache level
+  ##  lock: logical - lock env and bindings (default is TRUE)
+  ## Return:
+  ##  Invisibly returns the updated env.
+  ## Side effect:
+  ##  The object@.cache is updated
+  object@.cache <- new.env()
+  assign("level", level, object@.cache) 
+  if (level >= 1) {
[…]

### 2011-10-03 R/readMzXMLData.R

@@ -1,11 +1,16 @@
 readMzXMLData <- function(files,
-                          pdata=NULL,
-                          msLevel=2,
-                          verbose=TRUE,
-                          centroided=FALSE,
-                          smoothed=FALSE,
-                          removePeaks=0,
-                          clean=FALSE) {
+                          pdata = NULL,
+                          msLevel = 2,
+                          verbose = TRUE,
+                          centroided = FALSE,
+                          smoothed = FALSE,
+                          removePeaks = 0,
+                          clean = FALSE,
+                          cache = 1) {
+  msg <- c("The 'readMzXMLData' function is deprecated\n",
+          "Please use 'readMSData' instead.")
+  .Deprecated(msg=msg)
+  msLevel <- as.integer(msLevel)
   ## Opening file handles to mzXML files
   n <- length(files)
   if (n<1)
@@ -87,7 +92,7 @@ readMzXMLData <- function(files,
   nms <- paste("X",seq_len(length(spectra)),sep="")
   names(spectra) <- nms
 
-  spectra.env <- list2env(spectra)
+  assaydata <- list2env(spectra)
 
   ## here, a multiplex number (possibly 1) should be used
   ## to generate th NAnnotatedDataFrame object to afterwards,
@@ -100,18 +105,19 @@ readMzXMLData <- function(files,
   fdata <- new("AnnotatedDataFrame",
                data=data.frame(spectrum=1:length(spectra),
                  row.names=nms))
-  fdata <- fdata[ls(spectra.env)] ## reorder features
-  
+  fdata <- fdata[ls(assaydata)] ## reorder features
+  ## cache levels 2 and 3 not yet implemented
+  cache <- testCacheArg(cache,maxCache=1)
+  .cacheEnv <- newCacheEnv(assaydata, cache, lock=TRUE)
   ## Create and return 'MSnPeaks' object
   if (verbose)
     cat("Creating 'MSnExp' object\n")
-  
   toReturn <- new("MSnExp",
-                  assayData = spectra.env,
+                  assayData = assaydata,
                   phenoData = pdata,
                   featureData = fdata,
-                  processingData = process)
-  
+                  processingData = process,
+                  .cache = .cacheEnv)
   if (validObject(toReturn))
     return(toReturn)
 }

### 2011-10-03 inst/tests/test_MSnExp.R
Added cache capabilities to pSet, readMzXMLData deprecated
 o added .cache slot to pSet class <2011-10-03 Mon>
 o added pSet initialize method to set .cache and 
   .cache$level as '0'. <2011-10-03 Mon>
 o Check that level is defined in .cache and env is locked 
   in pSet validity method. <2011-10-03 Mon>
 o updated itraqdata.RData  <2011-10-03 Mon>
 o Deprecated readMzXMLData, added defunct.Rd <2011-10-03 Mon>
 o changed readMzXMLData to readMSData in man <2011-10-03 Mon>
 o updated show MSnExp for speed <2011-10-03 Mon>
 o updated read*Data is support cache = [0|1]. <2011-10-03 Mon>
 o upda

@@ -2,7 +2,8 @@ context("MSnExp class")
 
 test_that("MSnExp validity", {
   expect_true(validObject(new("MSnExp")))
-  })
+  expect_true(validObject(itraqdata))
+})
 
 test_that("readMzXMLData and dummy MSnExp msLevel 2 instance", {
   file <- dir(system.file(package="MSnbase",dir="extdata"),full.name=TRUE,pattern="mzXML$")

### 2011-12-06 R/cache.R
bugs fixed in readMSData and disabling cache
git-svn-id: file:///home/git/hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/MSnbase@61147 bc3139a8-67e5-0310-9ffc-ced21a209358

@@ -11,12 +11,12 @@ testCacheArg <- function(cache, maxCache=3) {
   if (!is.numeric(cache))
     stop("'cache' must be numeric.")
   if (cache < 0 | cache > maxCache) {
-    warning("cache must be [0:",maxCache,"]!")
+    warning("cache must be [0:", maxCache,"]!")
     if (cache < 0) {
       warning("Setting cache to 0.")
       cache <- 0
     } else {
-      warning("Setting cache to ",maxCache,".")
+      warning("Setting cache to ", maxCache, ".")
       cache <- maxCache
     }
   }
@@ -35,7 +35,7 @@ setCacheEnv <- function(assaydata, level=0, lock=TRUE) {
   assign("level", level, cacheEnv)
   if (level >= 1) { ## levels 2 and 3 not yet implemented
     ## precursor MZ
-    precMz <- unname(eapply(assaydata,precursorMz))
+    precMz <- unname(eapply(assaydata, precursorMz))
     assign("rangePrecursorMz", range(precMz), cacheEnv)
     assign("nPrecursorMz", length(precMz), cacheEnv)
     assign("uPrecursorMz", length(unique(precMz)), cacheEnv)

### 2011-12-06 inst/tests/test_io.R

@@ -6,38 +6,38 @@ test_that("Compatibility between writeMgfData and readMgfData", {
   ## no spectra order issue here, as < 10 spectra
   ## see [read|write]MgfData for details
   d1 <- itraqdata[1:3]
-  writeMgfData(d1,file=tf)
+  writeMgfData(d1, file=tf)
   d2 <- readMgfData(tf, verbose=FALSE)
-  expect_true(all(sapply(1:3, function(i) all.equal(as.data.frame(d1[[i]]),as.data.frame(d1[[i]])))))
-  expect_true(all.equal(precursorMz(d1),precursorMz(d1)))
+  expect_true(all(sapply(1:3, function(i) all.equal(as.data.frame(d1[[i]]), as.data.frame(d1[[i]])))))
+  expect_true(all.equal(precursorMz(d1), precursorMz(d1)))
   unlink(tf)
 })
 
 test_that("Compare readMzXMLData and readMSData output", {
   file <- dir(system.file(package="MSnbase", dir="extdata"),
-              full.name=TRUE,
+              full.names=TRUE,
               pattern="mzXML$")
   expect_warning(aa <- readMzXMLData(file, verbose=FALSE))
   bb <- readMSData(file, verbose=FALSE)  
   ## comparing data
-  expect_true(all.equal(header(aa),header(bb)))
-  expect_true(all.equal(intensity(aa),intensity(bb)))
-  expect_true(all.equal(mz(aa),mz(bb)))  
+  expect_true(all.equal(header(aa), header(bb)))
+  expect_true(all.equal(intensity(aa), intensity(bb)))
+  expect_true(all.equal(mz(aa), mz(bb)))  
 })
 
 test_that("Testing write.exprs and readMSnSet", {
   data(itraqdata)
   tf <- tempfile()
   x <- quantify(itraqdata, reporters=iTRAQ4, method="max", verbose=FALSE)
-  colchars <- c("ProteinAccession","PeptideSequence","retention.time","precursor.mz")
-  write.exprs(x,file=tf)  
+  colchars <- c("ProteinAccession", "PeptideSequence", "retention.time", "precursor.mz")
+  write.exprs(x, file=tf)  
   y <- readMSnSet(tf)
-  expect_true(all.equal(exprs(x),exprs(y)))
+  expect_true(all.equal(exprs(x), exprs(y)))
   unlink(tf)
[…]

### 2012-01-27 R/methods-MSnExp.R
CITATION, NoteAboutSpeedAndMemory, expander header output, check dfr nrow whem plotting
CHANGES IN VERSION 1.3.6
-------------------------
 o Updated NoteAboutSpeedAndMemory since parallel processing 
   has been added. <2011-12-18 Sun>
 o Added CITATION <2012-01-27 Fri>
 o Added information to header output: acquisition number 
   and precursor intensity <2012-01-27 Fri>
 o Added a test in plot.Spectrum2 for empty dataframe <2012-01-27 Fri>



git-svn-id: https://hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/MSnbase@62340 bc3139a8-67e5-0310-9ffc-ced21a209358

@@ -163,7 +163,7 @@ setMethod("quantify",
             ## has msLevel>1, all have
             if (msLevel(object)[1]<2) 
               stop("No quantification for MS1 data implemented.")
-            quantify.MSnExp(object,match.arg(method),reporters,strict,verbose)
+            quantify.MSnExp(object, match.arg(method), reporters, strict, verbose)
           })
 
 setMethod("curveStats","MSnExp",

### 2012-01-27 R/plotting-Spectrum2.R
CITATION, NoteAboutSpeedAndMemory, expander header output, check dfr nrow whem plotting
CHANGES IN VERSION 1.3.6
-------------------------
 o Updated NoteAboutSpeedAndMemory since parallel processing 
   has been added. <2011-12-18 Sun>
 o Added CITATION <2012-01-27 Fri>
 o Added information to header output: acquisition number 
   and precursor intensity <2012-01-27 Fri>
 o Added a test in plot.Spectrum2 for empty dataframe <2012-01-27 Fri>



git-svn-id: file:///home/git/hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/MSnbase@62340 bc3139a8-67e5-0310-9ffc-ced21a209358

@@ -1,12 +1,13 @@
 plot.Spectrum2 <- function(spectrum,
-                           reporters=NULL,full=FALSE,
+                           reporters = NULL,
+                           full = FALSE,
                            centroided,
-                           plot=TRUE,                          
-                           w1,w2) {
+                           plot = TRUE,                          
+                           w1, w2) {
   if (!full & is.null(reporters))
-    stop("Please provide repotrer ions if you do not want a full spectrum.")
+    stop("Please provide reporter ions if you do not want a full spectrum.")
   if (!is.null(reporters)) {
-    if (class(reporters)!="ReporterIons")
+    if (class(reporters) != "ReporterIons")
       stop("Reporters must be of class 'ReporterIons'.")
   }
   if (missing(centroided))
@@ -21,66 +22,81 @@ plot.Spectrum2 <- function(spectrum,
   if (missing(w2))
     w2 <- 0.01
   ## preparing full spectrum plot p
-  dfr <- data.frame(i=i,mtc=mtc,width=w1)
-  mainvp <- viewport(width=1,height=1,x=0.5,y=0.5)
-  title <- opts(title=paste("Precursor M/Z",round(precursorMz(spectrum),2)))
+  dfr <- data.frame(i = i, mtc = mtc, width = w1)
+  if (nrow(dfr) == 0)
+    stop("No data to be plotted in full scan")
+  mainvp <- viewport(width = 1, height = 1, x = 0.5, y = 0.5)
+  title <- opts(title = paste("Precursor M/Z",
+                  round(precursorMz(spectrum), 2)))
   if (centroided) {
-    p <- ggplot(dfr,aes(x=mtc,y=i,width=width)) + opts(legend.position = "none") +
-      labs(x="M/Z",y="Intensity") +
-        geom_bar(stat="identity",position="identity")
+    p <- ggplot(dfr, aes(x = mtc, y = i, width = width)) +
+      opts(legend.position = "none") +
+        labs(x = "M/Z", y = "Intensity") +
+          geom_bar(stat = "identity", position = "identity")
   } else {
-    p <- ggplot(dfr,aes(x=mtc,y=i)) + opts(legend.position = "none") +
-      labs(x="M/Z",y="Intensity") +
+    p <- ggplot(dfr, aes(x = mtc,y = i)) + opts(legend.position = "none") +
+      labs(x = "M/Z", y = "Intensity") +
         geom_line() ## + geom_point(alpha=I(1/10))
   }
   ## data reporters plot reps
[…]

### 2012-01-31 inst/doc/MSnbase-demo/MSnbase-demo.Rnw
bugs fixed in readMSData and disabling cache
git-svn-id: file:///home/git/hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/MSnbase@61147 bc3139a8-67e5-0310-9ffc-ced21a209358
ggplot2 v0.9.0 related fixes
git-svn-id: https://hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/MSnbase@62406 bc3139a8-67e5-0310-9ffc-ced21a209358

@@
 library("MSnbase")
+library("grid")
+library("reshape")
 @
@@
   viewport(layout.pos.row = x, layout.pos.col = y)
-print(p + opts(title="Quantitation using 'sum'") + geom_point(size=3,alpha=I(1/3),colour="red"), vp=vplayout(1,1))
-print(p + opts(title="Quantitation using 'max'") + geom_point(aes(x=114.1,y=3),alpha=I(1/3),colour="red",size=3), vp=vplayout(1,2))
-print(p + opts(title="Trapezoidation and strict=FALSE") + geom_polygon(alpha=I(1/5),fill="red"),vp=vplayout(2,1))
-print(p + opts(title="Trapezoidation and strict=TRUE") + 
-      geom_polygon(aes(x=c(114.05,114.05,114.1,114.15,114.15),y=c(0,1,3,1,0)),fill="red",alpha=I(1/5)),vp=vplayout(2,2))
+print(p + opts(title = "Quantitation using 'sum'") +
+      geom_point(size = 3, alpha = I(1/3), colour = "red"),
+      vp = vplayout(1, 1))
+print(p + opts(title = "Quantitation using 'max'") +
+      geom_point(aes(x = 114.1, y = 3), alpha = I(1/3), colour = "red", size = 3),
+      vp = vplayout(1, 2))
+print(p +
+      opts(title = "Trapezoidation and strict=FALSE") +
+      geom_polygon(alpha = I(1/5), fill = "red"),
+      vp = vplayout(2, 1))
+print(p +
+      opts(title = "Trapezoidation and strict=TRUE") + 
+      geom_polygon(aes(x = c(NA, 114.05, 114.05, 114.1, 114.15, 114.15, NA),
+                       y=c(NA, 0, 1, 3, 1, 0, NA)), fill = "red", alpha = I(1/5)),
+      vp = vplayout(2,2))
 @ 
@@
 the reporter group (red on figure \ref{fig:itraqchem}), 
-balance group (blue) 
-and peptide (the peptide reactive group is drawn in green). 
+balance group (blue) and peptide (the peptide reactive group is drawn in green). 
 However, incomplete dissociation does occur and 

### 2012-01-31 R/functions-MSnSet.R

@@ -26,12 +26,12 @@ normalise.MSnSet <- function(object,method,...) {
 
 combineMatrixFeatures <- function(matr,    ## matrix
                                    groupBy, ## factor
-                                   fun=c("mean",
+                                   fun = c("mean",
                                      "median",
                                      "weighted.mean",
                                      "sum",
                                      "medpolish"),
-                                   ...,    ## additional arguments to fun
+                                  ...,    ## additional arguments to fun
                                   verbose=TRUE) {
   if (is.character(fun)) {
     ## Using a predefined function
@@ -40,7 +40,7 @@ combineMatrixFeatures <- function(matr,    ## matrix
       summarisedFeatures <- by(matr,
                                groupBy,
                                function(x) {
-                                 medpol <- medpolish(x,trace.iter=verbose,...)
+                                 medpol <- medpolish(x, trace.iter = verbose, ...)
                                  return(medpol$overall+medpol$col)
                                })
     } else if (fun == "weighted.mean") {
@@ -54,12 +54,14 @@ combineMatrixFeatures <- function(matr,    ## matrix
       summarisedFeatures <- apply(matr,2,
                                   function(x) {
                                     .data <- data.frame(x=x,groupBy,w=w)
-                                    ddply(.data,"groupBy",summarise,
-                                          wmn=weighted.mean(x,w))
+                                    ddply(.data,
+                                          "groupBy",
+                                          summarise,
+                                          wmn = weighted.mean(x,w))
                                   })
-      summarisedFeatures <- do.call(cbind,summarisedFeatures)
+      summarisedFeatures <- do.call(cbind, summarisedFeatures)
       rn <- summarisedFeatures[,1]
-      summarisedFeatures <- summarisedFeatures[,grep("wmn",colnames(summarisedFeatures))]
+      summarisedFeatures <- summarisedFeatures[, grep("wmn", colnames(summarisedFeatures))]
       colnames(summarisedFeatures) <- colnames(matr)
       rownames(summarisedFeatures) <- rn
       return(summarisedFeatures)
@@ -67,21 +69,21 @@ combineMatrixFeatures <- function(matr,    ## matrix
       ## using either 'sum', 'mean', 'median'
       summarisedFeatures <- by(matr,
                                groupBy,
-                               function(x) apply(x,2,eval(parse(text=fun)),...))
[…]

### 2012-01-31 R/plotting-MSnExp.R

@@ -93,8 +93,8 @@ plotMzDelta.MSnExp <- function(object,            ## MSnExp object
     ## TODO - better than setting precMzWidth statically
     ## would be to get the peaks based on it m/z value
     ## and then find it's upper/lower m/z limits to set to 0
-    sp <- utils.removePrecMz(sp, precMz, precMzWidth)
-    delta <- c(delta, utils.getMzDelta(sp, percentage))
+    sp <- MSnbase:::utils.removePrecMz(sp, precMz, precMzWidth)
+    delta <- c(delta, MSnbase:::utils.getMzDelta(sp, percentage))
   }
   if (verbose) {
     close(pb)

### 2012-02-03 inst/doc/MSnbase-demo/MSnbase-demo.Rnw
MSnSet combine method and related updates/changes
CHANGES IN VERSION 1.3.8
-------------------------
 o Modified trimMz warning to report acquisition number <2012-02-01 Wed>
 o add 'experimentData(object, value) <- ' method for signature eSet
   and MIAPE <2012-02-02 Thu>
 o combine methods for MIAPE instances <2012-02-02 Thu>
 o combine methods for MSnProcess instances <2012-02-02 Thu>
 o changed qual drop warning into message in combineFeatures,
   updated test_MSnSet accordingly <2012-02-02 Thu>
 o new updateFvarLabels and updateSampleNames function <2012-02-03 Fri>
 o combine method for MS

@@
 print(p + opts(title = "Quantitation using 'max'") +
-      geom_point(aes(x = 114.1, y = 3), alpha = I(1/3), colour = "red", size = 3),
+      geom_point(aes(x = 114.1, y = 3),
+                 alpha = I(1/18),
+                 colour = "red", size = 3),
       vp = vplayout(1, 2))
@@
       vp = vplayout(2,2))
+
 @ 

### 2012-02-03 R/methods-pSet.R

@@ -183,10 +183,10 @@ setMethod("fromFile","pSet",
 setMethod("header",
           signature("pSet","missing"),
           function(object) {
-            if (any(msLevel(object)<2))
+            if (any(msLevel(object) < 2))
               stop("header() only works for MS levels > 1.")
             tbl <- table(fromFile(object))
-            idx <- as.numeric(unlist(apply(tbl,1,function(x) 1:x)))
+            idx <- as.numeric(unlist(apply(tbl, 1, function(x) 1:x)))
             return(data.frame(cbind(index = idx,
                                     file = fromFile(object),
                                     retention.time = rtime(object),
@@ -203,21 +203,21 @@ setMethod("header",
 setMethod("header",
           signature=c("pSet","numeric"),
           function(object, scans) {
-            if (any(msLevel(object)<2))
+            if (any(msLevel(object) < 2))
               stop("header() only works for MS levels > 1.")
             tbl <- table(fromFile(object))
-            idx <- as.numeric(unlist(apply(tbl,1,function(x) 1:x)))
+            idx <- as.numeric(unlist(apply(tbl, 1, function(x) 1:x)))
             ## OK for length(scans) > 1 -- slow for 1 scan
-            hdfr <- data.frame(cbind(index=idx,
-                                     file=fromFile(object),
-                                     retention.time=rtime(object),
-                                     precursor.mz=precursorMz(object),
-                                     peaks.count=peaksCount(object),
-                                     tic=tic(object),
-                                     ms.level=msLevel(object),
-                                     charge=precursorCharge(object),
-                                     collision.energy=collisionEnergy(object)))          
-            return(hdfr[scans,])
+            hdfr <- data.frame(cbind(index = idx,
+                                     file = fromFile(object),
+                                     retention.time = rtime(object),
+                                     precursor.mz = precursorMz(object),
+                                     peaks.count = peaksCount(object),
+                                     tic = tic(object),
+                                     ms.level = msLevel(object),
+                                     charge = precursorCharge(object),
+                                     collision.energy = collisionEnergy(object)))          
+            return(hdfr[scans, ])
           })
 
 ##################################################################
@@ -399,7 +399,7 @@ setMethod("processingData",
 ## - - - - - - - - - - - - - - - - - - - - - -
 ## Other TODO (based on eSet):
 
-## setMethod("combine","pSet",
+## setMethod("combine","pSet", ## rather a combine(MSnSet, )
 
[…]

### 2012-02-03 inst/tests/test_MSnSet.R

@@ -13,12 +13,11 @@ test_that("Combine MSnSet features", {
               data=data.frame(
                 A=rep(c("A","B"),each=5),
                 B=paste(rep(c("A","B"),each=5),1:10,sep="."))))
-  expect_warning(bb <- combineFeatures(aa,factor(rep(letters[1:2],each=5)),"mean"))
-  expect_warning(cc <- combineFeatures(aa,factor(rep(letters[1:2],each=5)),"sum"))
-  expect_warning(dd <- combineFeatures(aa,factor(rep(letters[1:2],each=5)),"median"))
-  expect_warning(ee <- combineFeatures(aa,factor(rep(letters[1:2],each=5)), "weighted.mean", w=rep(1,10)))
-  ff <- combineFeatures(aa,factor(rep(letters[1:2], each = 5)),
-                        "medpolish", verbose=FALSE)
+  bb <- combineFeatures(aa,factor(rep(letters[1:2],each=5)),"mean")
+  cc <- combineFeatures(aa,factor(rep(letters[1:2],each=5)),"sum")
+  dd <- combineFeatures(aa,factor(rep(letters[1:2],each=5)),"median")
+  ee <- combineFeatures(aa,factor(rep(letters[1:2],each=5)), "weighted.mean", w=rep(1,10))
+  ff <- combineFeatures(aa,factor(rep(letters[1:2], each = 5)), "medpolish", verbose=FALSE)
   expect_equal(exprs(bb),
                matrix(c(4,10,4,10), ncol = 2),
                tolerance = .001,

### 2012-02-03 R/functions-MSnExp.R
MSnSet combine method and related updates/changes
CHANGES IN VERSION 1.3.8
-------------------------
 o Modified trimMz warning to report acquisition number <2012-02-01 Wed>
 o add 'experimentData(object, value) <- ' method for signature eSet
   and MIAPE <2012-02-02 Thu>
 o combine methods for MIAPE instances <2012-02-02 Thu>
 o combine methods for MSnProcess instances <2012-02-02 Thu>
 o changed qual drop warning into message in combineFeatures,
   updated test_MSnSet accordingly <2012-02-02 Thu>
 o new updateFvarLabels and updateSampleNames function <2012-02-03 Fri>
 o combine method for MS

@@ -140,7 +140,7 @@ clean.MSnExp <- function(object,verbose=TRUE) {
     return(object)
 }
 
-quantify.MSnExp <- function(object,method,reporters,strict,verbose) {
+quantify.MSnExp <- function(object, method, reporters, strict, verbose) {
   ## Display progress bar with eapply
   ## TODO - test if using eapply is more efficient in terms of mem/cpu usage
   ## if (verbose) {
@@ -158,11 +158,11 @@ quantify.MSnExp <- function(object,method,reporters,strict,verbose) {
   ## } else {
   ##   peakData <- eapply(assayData(object),quantify,method,reporters)
   ## }
-  ifelse(verbose,progress <- "text",progress <- "none")
+  ifelse(verbose, progress <- "text", progress <- "none")
   spectraList <- spectra(object)
   ## Quantification -- creating exprs for assayData slot
-  if (length(spectraList)==1) {
-    peakData <- quantify(spectraList[[1]],method,reporters,strict)
+  if (length(spectraList) == 1) {
+    peakData <- quantify(spectraList[[1]], method, reporters, strict)
     .exprs <- t(peakData$peakQuant)
     .qual <- t(peakData$curveData)
   } else {
@@ -172,37 +172,37 @@ quantify.MSnExp <- function(object,method,reporters,strict,verbose) {
       parallel <- TRUE
     }
     peakData <- llply(spectraList, quantify, method, reporters, strict,
-                      .progress=progress, .parallel=parallel)
-    .exprs <- do.call(rbind,sapply(peakData,"[","peakQuant"))
-    .qual <- do.call(rbind,sapply(peakData,"[","curveStats"))
+                      .progress = progress, .parallel = parallel)
+    .exprs <- do.call(rbind, sapply(peakData, "[", "peakQuant"))
+    .qual <- do.call(rbind, sapply(peakData, "[", "curveStats"))
   }
-  rownames(.exprs) <- sub(".peakQuant","",rownames(.exprs))
-  rownames(.qual) <- sub(".curveStats","",rownames(.qual))  
+  rownames(.exprs) <- sub(".peakQuant", "", rownames(.exprs))
+  rownames(.qual) <- sub(".curveStats", "", rownames(.qual))  
   ## Updating MSnprocess slot
   object@processingData@processing <- c(object@processingData@processing,
                                         paste(reporters@name,
-                                              ifelse(strict," (strict) "," "),
-                                              "quantification by ",method,
[…]

### 2012-02-03 R/functions-Spectrum.R

@@ -229,9 +229,9 @@ trimMz.Spectrum <- function(x,mzlim,updatePeaksCount=TRUE) {
   mzmax <- max(mzlim)
   sel <- (x@mz >= mzmin) & (x@mz <= mzmax)
   if (sum(sel)==0) {
-    warning(paste("No data points between ",mzmin," and ",mzmax,
-                  " for precursor ",precursorMz(x),
-                  ".\nLeaving data as is.",sep=""))
+    warning(paste("No data points between ", mzmin, " and ", mzmax,
+                  " for spectrum with acquisition number ", acquisitionNum(x),
+                  ".\n Leaving data as is.", sep=""))
     return(x)
   }
   x@mz <- x@mz[sel]

### 2012-02-03 inst/doc/MSnbase-demo/MSnbase-demo.Rnw
Update vignette with 'combine' section + minor fixes.
CHANGES IN VERSION 1.3.9
-------------------------
 o updated demo vignette <2012-02-03 Fri>
 o sorting numeric subsets in "[" pSet, as unsorted 
   numerical indexes fails  <2012-02-03 Fri>
 o Added match.arg in combineFeatures so that 
   a unique default value (the first) is used when no 
   fun is specified <2012-02-03 Fri>




git-svn-id: https://hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/MSnbase@62533 bc3139a8-67e5-0310-9ffc-ced21a209358

@@
 
-\Rpackage{MSnbase} aims are providing a reproducible research framework to proteomics data analysis. 
+\Rpackage{MSnbase} \citep{Gatto2012} aims are providing a reproducible research framework to proteomics data analysis. 
 It should allow researcher to easily mine mass spectrometry data, explore the data and its statistical properties 
@@
 <<readdata, echo=TRUE, cache=FALSE>>=
-file <- dir(system.file(package="MSnbase", dir="extdata"),
-            full.names=TRUE, pattern="mzXML$")
-rawdata <- readMSData(file, msLevel=2, verbose=FALSE)
+file <- dir(system.file(package = "MSnbase", dir = "extdata"),
+            full.names = TRUE, pattern = "mzXML$")
+rawdata <- readMSData(file, msLevel = 2, verbose = FALSE)
 @
@@
 <<spectrum-plot,fig=T,echo=T,tikz=T,external=F,width=5,height=5>>=
-plot(sp,reporters=iTRAQ4,full=TRUE) 
+plot(sp, reporters = iTRAQ4, full = TRUE) 
 @ 
@@
 <<bsasel,echo=FALSE>>=
-bsasel <- fData(itraqdata)$ProteinAccession=="BSA"
+bsasel <- fData(itraqdata)$ProteinAccession == "BSA"
 @
@@
 <<subset,echo=TRUE>>=
-sel <- fData(itraqdata)$ProteinAccession=="BSA"
+sel <- fData(itraqdata)$ProteinAccession == "BSA"
 bsa <- itraqdata[sel]
@@
 <<msnexp-plot,fig=T,echo=T,tikz=T,external=F,width=5,height=6>>=
-plot(bsa,reporters=iTRAQ4,full=FALSE)
+plot(bsa, reporters = iTRAQ4, full = FALSE)
 @ 
@@
 <<removePeaks,echo=TRUE,cache=FALSE>>=
-experiment <- removePeaks(itraqdata,t=400,verbose=FALSE)
+experiment <- removePeaks(itraqdata, t = 400, verbose = FALSE)
 ## total ion current
@@
 <<prepare-spectrum-plot2,echo=FALSE>>=
-p1 <- plot(itraqdata[["X55"]],full=TRUE) + theme_gray(5) 
-p2 <- plot(experiment[["X55"]],full=TRUE) + theme_gray(5) 
+p1 <- plot(itraqdata[["X55"]], full = TRUE) + theme_gray(5) 
+p2 <- plot(experiment[["X55"]], full = TRUE) + theme_gray(5) 
 @
@@
 peaksCount(experiment[["X55"]])
-experiment <- clean(experiment,verbose=FALSE)
+experiment <- clean(experiment, verbose = FALSE)
 peaksCount(experiment[["X55"]])
@@
 range(mz(itraqdata[["X55"]]))
-experiment <- trimMz(experiment,mzlim=c(112,120))
+experiment <- trimMz(experiment, mzlim = c(112,120))
 range(mz(experiment[["X55"]]))
@@
 qnt <- quantify(experiment,
-                method="trap",
-                reporters=iTRAQ4,
-                strict=FALSE,
-                verbose=FALSE)
+                method = "trap",
+                reporters = iTRAQ4,
+                strict = FALSE,
+                verbose = FALSE)
 qnt
@@
 whichRow <- which(is.na(exprs(qnt))) %% nrow(qnt)
-qnt <- qnt[-whichRow,]
+qnt <- qnt[-whichRow, ]
 sum(is.na(exprs(qnt)))
@@
 isotopes <- getIsotope(Fim)
-F1 <- isotopes[2,2]
+F1 <- isotopes[2, 2]
 F1
@@
 <<purityCorrect,echo=TRUE,cache=FALSE>>=
-impurities <- matrix(c(0.929,0.059,0.002,0.000,
-                       0.020,0.923,0.056,0.001,
-                       0.000,0.030,0.924,0.045,
-                       0.000,0.001,0.040,0.923),
-                     nrow=4)
-qnt.crct <- purityCorrect(qnt,impurities)
+impurities <- matrix(c(0.929, 0.059, 0.002, 0.000,
+                       0.020, 0.923, 0.056, 0.001,
+                       0.000, 0.030, 0.924, 0.045,
+                       0.000, 0.001, 0.040, 0.923),
+                     nrow = 4)
+qnt.crct <- purityCorrect(qnt, impurities)
 head(exprs(qnt))
@@
 <<normalise,echo=TRUE,cache=FALSE>>=
-qnt.max <- normalise(qnt,"max")
-qnt.sum <- normalise(qnt,"sum")
-qnt.quant <- normalise(qnt,"quantiles")
-qnt.qrob <- normalise(qnt,"quantiles.robust")
-qnt.vsn <- normalise(qnt,"vsn")
+qnt.max <- normalise(qnt, "max")
+qnt.sum <- normalise(qnt, "sum")
+qnt.quant <- normalise(qnt, "quantiles")
+qnt.qrob <- normalise(qnt, "quantiles.robust")
+qnt.vsn <- normalise(qnt, "vsn")
 @
@@
 <<prepare-msnset-nrom-plot,cache=FALSE,echo=FALSE>>=
-p <- plot(normalise(experiment[bsasel],"max"),reporters=iTRAQ4,full=FALSE)
+p <- plot(normalise(experiment[bsasel], "max"),
+          reporters = iTRAQ4, full = FALSE)
 p <- p + theme_gray(7)
@@
 <<combineFeatures,echo=TRUE,cache=FALSE>>=
-qnt2 <- combineFeatures(qnt,groupBy=gb,fun="median")
+qnt2 <- combineFeatures(qnt, groupBy = gb, fun = "median")
 qnt2
@@
 incompdiss <- quantify(itraqdata,
-                       method="trap",
-                       reporters=iTRAQ5,
-                       strict=FALSE,
-                       verbose=FALSE)
+                       method = "trap",
+                       reporters = iTRAQ5,
+                       strict = FALSE,
+                       verbose = FALSE)
 head(exprs(incompdiss))
@@
   \begin{center}
-<<incompdiss-plot,fig=T,echo=F,tikz=T,external=F,width=4,height=3.5>>=
+<<incompdiss-plot,fig=T,echo=F,tikz=T,external=F,width=5,height=2.5>>=
 dfr <- melt(exprs(incompdiss))
-colnames(dfr) <- c("feature","reporters","intensity")
-qplot(x=reporters,y=intensity,data=dfr,geom="boxplot") + theme_gray(8)
-@ 
-\caption{ Boxplot comparing intensities of the 4 reporter ions an the 
-  incomplete dissociation specific peak. }
+colnames(dfr) <- c("feature", "reporters", "intensity")
+dfr$reporters <- sub("iTRAQ5.", "", dfr$reporters)
+repsum <- apply(exprs(incompdiss), 1, function(x) sum(x[1:4]))
+dfr2 <- data.frame(iTRAQ1to4 = repsum, iTRAQ5 = exprs(incompdiss)[,5])
+p1 <- ggplot(data = dfr, aes(x = reporters,y = log10(intensity))) +
+  geom_boxplot() +
+  theme_gray(6)
+p2 <- ggplot(data = dfr2, aes(x = log10(iTRAQ1to4), y = log10(iTRAQ5))) +
+  geom_point(alpha = I(1/2)) +
+  geom_abline(intercept = 0, slope = 1, linetype = "dotted") +
+  stat_smooth(method = lm, se = FALSE) +
+  xlab(label = expression(log[10]~sum~114~to~117)) +
+  ylab(label = expression(log[10]~145)) +
+  theme_gray(6)
+grid.newpage()
+pushViewport(viewport(layout = grid.layout(1, 2)))
+print(p1, vp = vplayout(1, 1))
+print(p2, vp = vplayout(1, 2))
+@ 
+\caption{ Boxplot and scatterplot comparing intensities of the 4 reporter ions 
+  (or their sum, on the right) and the incomplete dissociation specific peak. }
 \label{fig:incompdiss}
@@
 \end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Section
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Combining \Robject{MSnSet} instances}\label{sec:combine}
+
+Default MSnSet feature names (\Robject{X1}, \Robject{X2}, \ldots) and sample names 
+(\Robject{iTRAQ4.114}, \Robject{iTRAQ4.115}, \Robject{iTRAQ4.116}, \ldots) are not very informative.
+The features and samples of these anonymous quantitative data-sets should be updated prior 
+to additional downstream analysis.
+
+Lets create two \Robject{MSnSet}s from the same raw data to simulate two different 
+independent experiments. As done previously (see section \ref{sec:feataggregation}), 
+we combine the specta based on the proteins they have been identified to belong to.
+Features can thus naturally be named using protein accession numbers.
+Alternatively, if peptide sequences would have been used as grouping factor in 
+\Rfunction{combineFeatures}, then these would be good feature name candidates.
+
+<<make2exps, echo=TRUE, cache=FALSE>>=
+set.seed(1)
+i <- sample(length(itraqdata), 35)
+j <- sample(length(itraqdata), 35)
+exp1 <- quantify(itraqdata[i], reporters = iTRAQ4, verbose = FALSE)
+exp2 <- quantify(itraqdata[j], reporters = iTRAQ4, verbose = FALSE)
+table(featureNames(exp1) %in% featureNames(exp2))
+exp1 <- combineFeatures(exp1, groupBy = fData(exp1)$ProteinAccession)
+exp2 <- combineFeatures(exp2, groupBy = fData(exp2)$ProteinAccession)
+head(featureNames(exp1))
+featureNames(exp1) <- fData(exp1)$ProteinAccession
+featureNames(exp2) <- fData(exp2)$ProteinAccession
+head(featureNames(exp1))
+@
+
+Setting the feature names is critical when multiple experiments are to be combined, as 
+this is done using common features as defined by their names (see below).
+
+\bigskip
+
+Sample names should also be updated to replace reporter names with relevant identifiers; 
+the individual reporter data is stored in the \Robject{phenoData} and is not lost.
+A convenience function \Rfunction{updateSampleNames} is provided to append the \Robject{MSnSet}'s 
+variable name to the already defined names, although in general, biologically relevant 
+identifiers are preferred.
+
+<<renameSamples, echo=TRUE, cache=FALSE>>=
+sampleNames(exp1)
+exp1 <- updateSampleNames(exp1)
+sampleNames(exp1)
+sampleNames(exp1) <- c("Ctrl1", "Cond1", "Ctrl2", "Cond2")
+sampleNames(exp2) <- c("Ctrl3", "Cond3", "Ctrl4", "Cond4")
+@
+
+At this stage, it is not yet possible to combine the two experiments, 
+because their feature data is not compatible; they share the same feature 
+variable labels, i.e. the feature data column names 
+(\Sexpr{paste(head(fvarLabels(exp1), n = 3), collapse=", ")}, \ldots), 
+but the content is different because the original data was, in particular 
+all the spectrum data. Feature data with identical labels (columns in the data frame) 
+and names (row in the data frame) are expected to have the same data and produce an
+error if not conform.
+
+<<fdatanames, echo=TRUE, cache=FALSE>>=
+all(fvarLabels(exp1) == fvarLabels(exp1))
+fData(exp1)["BSA", 1:4]
+fData(exp2)["BSA", 1:4]
+@
+
+Instead of removing these overlapping and different feature data columns, 
+one can use a second convenience function, \Rfunction{updateFvarLabels}, 
+to update feature labels based on the experiements variable name and maintain all the metadata.
+
+<<renameFvars, echo=TRUE, cache=FALSE>>=
+exp1 <- updateFvarLabels(exp1)
+exp2 <- updateFvarLabels(exp2)
+head(fvarLabels(exp1))
+head(fvarLabels(exp2))
+@
+
+It is now possible to combine \Robject{exp1} and \Robject{exp2} with the 
+\Rfunction{combine} method. The new experiment will contain the union of 
+the feature names of the individual experiments with missing values 
+inserted appropriately.
+
+<<combine, echo=TRUE, cache=FALSE>>=
+exp12 <- combine(exp1, exp2)
+dim(exp12)
+pData(exp12)
+exprs(exp12)[25:28, ]
+exp12
+@
+
+In summary, when experiments need to be combined, one needs to 
+(1) clarify the sample names using \Rfunction{updateSampleNames} or,
+better manually, for biological relevance and (2) update the feature 
+data variable labels with \Rfunction{updateFvarLabels}. 
+The individual experiments (there can be more than 2) can then easily be combined 
+with the \Rfunction{combine} method while retaining the meta-data.
 
@@
 \end{document}
+
+% LocalWords:  renameSamples

### 2012-02-03 inst/doc/Foreword.tex
Update vignette with 'combine' section + minor fixes.
CHANGES IN VERSION 1.3.9
-------------------------
 o updated demo vignette <2012-02-03 Fri>
 o sorting numeric subsets in "[" pSet, as unsorted 
   numerical indexes fails  <2012-02-03 Fri>
 o Added match.arg in combineFeatures so that 
   a unique default value (the first) is used when no 
   fun is specified <2012-02-03 Fri>




git-svn-id: file:///home/git/hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/MSnbase@62533 bc3139a8-67e5-0310-9ffc-ced21a209358

@@
 
-\Rpackage{MSnbase} is in an early development %%(package version \Sexpr{packageVersion("MSnbase")}, 
-(see section \ref{sec:sessionInfo} for details about packages and version used in this vignette). 
-Although main data structures have been thought out and are meant to be compatible with other existing 
-mature infrastructure in the Bioconductor project, changes may occur in the future. 
-Current functionality will evolve and new one will be added. 
-Although at an early stage, this package is released in the hope that it may foster  
-new developments in proteomics data analysis within \R by providing a common infrastructure. 
-Several package developers working with mass spectrometry and proteomics data met at the 
-Bioconductor Developer Meeting Europe\footnote{\url{http://bioconductor.org/help/course-materials/2010/HeidelbergNovember2010/}} 
-held in Heidelberg in November 2010, and agreed to combine efforts. 
-This library is one attempt to do so.
+%% \Rpackage{MSnbase} is in an early development %%(package version \Sexpr{packageVersion("MSnbase")}, 
+%% (see section \ref{sec:sessionInfo} for details about packages and version used in this vignette). 
+%% Although main data structures have been thought out and are meant to be compatible with other existing 
+%% mature infrastructure in the Bioconductor project, changes may occur in the future. 
+%% Current functionality will evolve and new one will be added. 
+%% Although at an early stage, this package is released in the hope that it may foster  
+%% new developments in proteomics data analysis within \R by providing a common infrastructure. 
+%% Several package developers working with mass spectrometry and proteomics data met at the 
+%% Bioconductor Developer Meeting Europe\footnote{\url{http://bioconductor.org/help/course-materials/2010/HeidelbergNovember2010/}} 
+%% held in Heidelberg in November 2010, and agreed to combine efforts. 
+%% This library is one attempt to do so.
 
-\bigskip
+\Rpackage{MSnbase} is under active developed; current functionality is evolving and 
+new features will be added. This software is free and open-source software. 
+If you use it, please support the project by citing it in publications:
+
+\begin{quote}
+  Laurent Gatto and Kathryn S. Lilley. \textit{MSnbase - an R/Bioconductor 
+    package for isobaric tagged mass spectrometry data visualization,
+    processing and quantitation.} Bioinformatics 28, 288-289 (2011).
+\end{quote}
 

### 2012-02-13 R/readWriteMgfData.R
typo in vignette
git-svn-id: https://hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/MSnbase@62736 bc3139a8-67e5-0310-9ffc-ced21a209358

@@ -14,6 +14,7 @@
                con <- file(description = con,
                            open = "at",
                            blocking = TRUE)
+               on.exit(close(con))
              }
              if (!inherits(con, "connection"))
                stop("'con' is not a proper connection!")
@@ -23,7 +24,7 @@
              ## write spectrum
              writeLines(COM, con = con)
              writeMgfContent(object, TITLE = TITLE, con = con)
-             close(con)
+             ## close(con)
            })
 
 setMethod("writeMgfData",
@@ -41,6 +42,7 @@ setMethod("writeMgfData",
               con <- file(description = con,
                           open = "at",
                           blocking = TRUE)
+              on.exit(close(con))
             }
             if (!inherits(con, "connection"))
               stop("'con' is not a proper connection!")
@@ -53,7 +55,7 @@ setMethod("writeMgfData",
                    writeMgfContent,
                    TITLE = NULL,
                    con = con)            
-            close(con)            
+            ## close(con)            
           })
 
 writeMgfContent <- function(sp, TITLE = NULL, con) {

### 2012-02-16 R/methods-MSnSet.R
is.na.MSnSet and vignette/doc updates
git-svn-id: https://hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/MSnbase@62825 bc3139a8-67e5-0310-9ffc-ced21a209358

@@ -302,7 +302,7 @@ setMethod("topN", signature(object = "MSnSet"),
 
 
 setMethod("plotNA", signature(object = "MSnSet"), 
-          function(object, pNA = .5, ...) {
+          function(object, pNA = .5) {
             if (pNA > 1)
               pNA <- 1
             if (pNA < 0)
@@ -314,7 +314,7 @@ setMethod("plotNA", signature(object = "MSnSet"),
 
 
 setMethod("plotNA", signature(object = "matrix"), 
-          function(object, pNA = .5, ...) {
+          function(object, pNA = .5) {
             if (pNA > 1)
               pNA <- 1
             if (pNA < 0)
@@ -324,7 +324,7 @@ setMethod("plotNA", signature(object = "matrix"),
           })
 
 setMethod("filterNA", signature(object = "matrix"), 
-          function(object, pNA = .5, ...) {
+          function(object, pNA = .5) {
             if (pNA > 1)
               pNA <- 1
             if (pNA < 0)
@@ -343,7 +343,7 @@ setMethod("filterNA", signature(object = "matrix"),
 
 
 setMethod("filterNA", signature(object = "MSnSet"), 
-          function(object, pNA = .5, ...) {
+          function(object, pNA = .5) {
             if (pNA > 1)
               pNA <- 1
             if (pNA < 0)
@@ -360,3 +360,4 @@ setMethod("filterNA", signature(object = "MSnSet"),
               return(ans)
           })
 
+is.na.MSnSet <- function(x) is.na(exprs(x))

### 2012-02-16 inst/doc/MSnbase-demo/MSnbase-demo.Rnw
typo in vignette
git-svn-id: file:///home/git/hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/MSnbase@62736 bc3139a8-67e5-0310-9ffc-ced21a209358
is.na.MSnSet and vignette/doc updates
git-svn-id: file:///home/git/hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/MSnbase@62825 bc3139a8-67e5-0310-9ffc-ced21a209358

@@
 <<removeNa,echo=TRUE>>=
-sum(is.na(exprs(qnt)))
-whichRow <- which(is.na(exprs(qnt))) %% nrow(qnt)
+table(is.na(qnt))
+whichRow <- which(is.na((qnt))) %% nrow(qnt)
 qnt <- qnt[-whichRow, ]
@@
 @
+
+The \Rfunction{filterNA} method allows to filter features based on the proportion 
+of missing data their carry, specified using the \Robject{pNA} argument. 
+The same result as above could have been achieved with 
+\Rfunction{qnt <- filterNA(qnt, pNA = 0)}, 
+specifying that we allow 0 percent of missing data. 
+See also the \Rfunction{plotNA} method to get an overview of the completeness 
+of a data set.
 
@@
 <<fdatanames, echo=TRUE, cache=FALSE>>=
-all(fvarLabels(exp1) == fvarLabels(exp1))
+stopifnot(all(fvarLabels(exp1) == fvarLabels(exp2)))
 fData(exp1)["BSA", 1:4]

### 2012-02-25 inst/doc/MSnbase-demo/MSnbase-demo.Rnw
more MSnSet combine, new updateFeatureNames
 o new updateFeatureNames function <2012-02-17 Fri>
 o updated vignettes to illustrate vertical/horizontal combine <2012-02-17 Fri>
 o typo in normalised.Rd <2012-02-19 Sun>
 o TODO combine unit tests




git-svn-id: https://hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/MSnbase@63146 bc3139a8-67e5-0310-9ffc-ced21a209358

@@
 
-Default MSnSet feature names (\Robject{X1}, \Robject{X2}, \ldots) and sample names 
-(\Robject{iTRAQ4.114}, \Robject{iTRAQ4.115}, \Robject{iTRAQ4.116}, \ldots) are not very informative.
-The features and samples of these anonymous quantitative data-sets should be updated prior 
-to additional downstream analysis.
-
-Lets create two \Robject{MSnSet}s from the same raw data to simulate two different 
-independent experiments. As done previously (see section \ref{sec:feataggregation}), 
+Combining mass spectrometry runs can be done in two different ways depending on the 
+nature of these runs. If the runs represent repeated measures of identical samples, 
+for instance multiple fractions, the data has to be combine along the row of the 
+quantitation matrix: all the features (along the rows) represent measurements of the 
+same set of samples (along the columns). In this situation, described in section 
+\ref{sec:comb1}, two experiments of dimensions 
+$n_1$ (rows) by $m$ (columns and $n_2$ by $m$ will produce 
+a new experiment of dimensions $n_1 + n_2$ by $m$.
+
+When however, different sets of samples have been analysed in different mass 
+spectrometry runs, the data has 
+to be combined along the columns of the quantitation matrix: some features will be 
+shared across experiments and should thus be aligned on a same row in the new
+data set, whereas unique 
+features to one experiment should be set as missing in the other one. 
+In this situation, described in section \ref{sec:comb2}, 
+two experiments of dimensions $n_1$ by $m_1$ and $n_2$ by $m_2$ will 
+produce a new experiment of dimensions 
+$unique_{n_1} + unique_{n_2} + shared_{n_1, n_2}$ by $m_1 + m_2$. 
+The two first terms of the first dimension will be complemented by \Robject{NA} values.
+
+Default \Robject{MSnSet} feature names (\Robject{X1}, \Robject{X2}, \ldots) and sample names 
+(\Robject{iTRAQ4.114}, \Robject{iTRAQ4.115}, \Robject{iTRAQ4.116}, \ldots) are not informative.
+The features and samples of these anonymous quantitative data-sets should be updated 
+before being combined, to guide how to meaningfully merge them.
+
+\subsection{Combining identical samples}\label{sec:comb1}
+
+To simulate this situation, let us use quantiation data from the \Robject{itraqdata} 
+object that is provided with the package as experiment 1 and the data from the 
+\Robject{rawdata} \Robject{MSnExp} instance created at the very beginning of this 
+document. Both experiments share the \emph{same} default iTRAQ 4-plex reporter names 
+as default sample names, and will thus automatically be combined along rows.
+
+<<make2exps1, echo=TRUE, cache=FALSE>>=
+exp1 <- quantify(itraqdata, reporters = iTRAQ4, verbose = FALSE)
+sampleNames(exp1)
+exp2 <- quantify(rawdata, reporters = iTRAQ4, verbose = FALSE)
+sampleNames(exp2)
+@
+
+It important to note that the features of these independent experiments share 
+the same default anonymous names: X1, X2, X3, \ldots, 
+that however represent quantitation of distinct physical analytes. 
+If the experiments were to be combined as is, it would result in an error because
+data points for the same \emph{feature} name (say \Robject{X1}) and the 
+same \emph{sample name} (say \Robject{iTRAQ4.114}) have different values. 
+We thus first update the feature names to explicitate that they originate from 
+different experiment and represent quantitation from different spectra 
+using the convenience function \Rfunction{updateFeatureNames}.
+Note that updating the names of one experiment would suffice here.
+
+<<updateFnames, echo=TRUE, cache=FALSE>>=
+head(featureNames(exp1))
+exp1 <- updateFeatureNames(exp1)
+head(featureNames(exp1))
+head(featureNames(exp2))
+exp2 <- updateFeatureNames(exp2)
+head(featureNames(exp2))
+@
+
+The two experiments now share the same sample names and have different feature names 
+and will be combined along the row. Note that all meta-data is correctly 
+combined along the quantitation values.
+
+<<comb1, echo=TRUE, cache=FALSE>>=
+exp12 <- combine(exp1, exp2)
+dim(exp1)
+dim(exp2)
+dim(exp12)
+@
+
+\subsection{Combine different samples}\label{sec:comb2}
+
+Lets now create two \Robject{MSnSet}s from the same raw data to simulate two different 
+independent experiments that share some features. 
+As done previously (see section \ref{sec:feataggregation}), 
 we combine the specta based on the proteins they have been identified to belong to.
@@
 
-<<make2exps, echo=TRUE, cache=FALSE>>=
+<<make2exps2, echo=TRUE, cache=FALSE>>=
 set.seed(1)
@@
 
-Sample names should also be updated to replace reporter names with relevant identifiers; 
+Sample names should also be updated to replace anonymous reporter names with relevant identifiers; 
 the individual reporter data is stored in the \Robject{phenoData} and is not lost.
@@
 At this stage, it is not yet possible to combine the two experiments, 
-because their feature data is not compatible; they share the same feature 
+because their feature data is not compatible yet; they share the same feature 
 variable labels, i.e. the feature data column names 
 (\Sexpr{paste(head(fvarLabels(exp1), n = 3), collapse=", ")}, \ldots), 
-but the content is different because the original data was, in particular 
-all the spectrum data. Feature data with identical labels (columns in the data frame) 
+but the part of the content is different because the original data was (in particular 
+all the spectrum centric data: identical peptides in different runs will have different 
+retention times, precursor intensities, \ldots). 
+Feature data with identical labels (columns in the data frame) 
 and names (row in the data frame) are expected to have the same data and produce an
@@
 
-Instead of removing these overlapping and different feature data columns, 
+Instead of removing these identical feature data columns, 
 one can use a second convenience function, \Rfunction{updateFvarLabels}, 
-to update feature labels based on the experiements variable name and maintain all the metadata.
+to update feature labels based on the experiements variable name 
+and maintain all the metadata.
 
@@
 
-It is now possible to combine \Robject{exp1} and \Robject{exp2} with the 
-\Rfunction{combine} method. The new experiment will contain the union of 
+It is now possible to combine \Robject{exp1} and \Robject{exp2}, 
+including all the meta-data, with the \Rfunction{combine} method. 
+The new experiment will contain the union of 
 the feature names of the individual experiments with missing values 
@@
 
-In summary, when experiments need to be combined, one needs to 
-(1) clarify the sample names using \Rfunction{updateSampleNames} or,
+\bigskip
+
+In summary, when experiments with different samples need to be combined 
+(along the columns), one needs to 
+(1) clarify the sample names using \Rfunction{updateSampleNames} or
 better manually, for biological relevance and (2) update the feature 
@@
 
+If runs for the same sample (different fractions for example) need to be combines, 
+one needs to (1) differentiate the feature provenance with 
+\Rfunction{updateFeatureNames} prior to use \Rfunction{combine}.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2012-04-02 R/plotting-MSnSet.R
Misc changes
CHANGES IN VERSION 1.5.1
-------------------------
 o updated NA warning message in readIspyData <2012-03-05 Mon>
 o fixed combineFeatures/combineMatrixFeatures for 1 sample <2012-03-20 Tue>
 o image method for MSnSet instances <2012-03-20 Tue>
 o fixed bug in plotNA (first t was wring) <2012-03-21 Wed>
 o import plot from stats4 <2012-03-21 Wed>
 o using reshape2 <2012-03-30 Fri>




git-svn-id: file:///home/git/hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/MSnbase@64784 bc3139a8-67e5-0310-9ffc-ced21a209358

@@ -21,7 +21,10 @@ plotNA.matrix <- function(X, pNA) {
   pNA <- pNA[1]
   calcNApp <- function(x) {
     nbna <- sum(is.na(x))
-    nbcells <- prod(dim(x))
+    if (is.null(dim(x)))
+      nbcells <- length(x)
+    else 
+      nbcells <- prod(dim(x))
     return(nbna/nbcells)
   }
   ocol <- apply(X,2, function(x) sum(is.na(x)))
@@ -69,3 +72,24 @@ plotNA.matrix <- function(X, pNA) {
   print(p)
   invisible(p)
 }
+
+
+setMethod("image", "MSnSet",
+          function(x, 
+                   yticks = 10,
+                   x.cex.axis = .75,
+                   y.cex.axis = .75,
+                   ...) {
+            lab <- sampleNames(x)
+            x <- exprs(x)
+            nc <- ncol(x)
+            nr <- nrow(x)
+            image(t(x), xaxt = "n", yaxt = "n", ...)
+            axis(1, seq(0,1, 1/(nc - 1)),
+                 labels = lab,
+                 cex.axis = x.cex.axis)
+            yticks <- seq(0, 1, 1/(yticks-1)) * nr
+            axis(2, seq(0,1, 1/(length(yticks) - 1)),
+                 labels = round(yticks, 0),
+                 cex.axis = y.cex.axis)
+          })

### 2012-04-02 R/readIspyData.R

@@ -1,17 +1,17 @@
-readIspyData <- function(file="ispy_results.tsv",
-                         uniquePeps=TRUE,
-                         pep=0.05,
-                         na.rm=TRUE,
-                         min.int=0,
-                         reporters=19:23,
-                         ## skipFillUp=24,
-                         ## fillUp=FALSE,
-                         keepAll=FALSE,
-                         verbose=TRUE) {
+readIspyData <- function(file = "ispy_results.tsv",
+                         uniquePeps = TRUE,
+                         pep = 0.05,
+                         na.rm = TRUE,
+                         min.int = 0,
+                         reporters = 19:23,
+                         ## skipFillUp = 24,
+                         ## fillUp = FALSE,
+                         keepAll = FALSE,
+                         verbose = TRUE) {
   if (verbose)
     cat("Reading table\n")
-  tab <- read.csv(file,header=TRUE,sep="\t",fill=TRUE,
-                  colClasses=c(
+  tab <- read.csv(file, header = TRUE, sep = "\t", fill = TRUE,
+                  colClasses = c(
                     "numeric",       # Index
                     "factor",        # ProteinAccession 
                     "character",     # ProteinDescription
@@ -32,7 +32,7 @@ readIspyData <- function(file="ispy_results.tsv",
                     "numeric",       # PosteriorErrorProbability
                     rep("numeric",length(reporters)),
                     "numeric"))      # precursorrelativesignal
-  names(tab) <- gsub("\\.\\.","pc",gsub("_","",names(tab)))
+  names(tab) <- gsub("\\.\\.","pc", gsub("_","",names(tab)))
   names(tab)[ncol(tab)] <- "PrecursorRelativeSignal"
   ## This column was called MascotQuery, then changed to BrowserData
   ## - just ignore it
@@ -93,11 +93,11 @@ readIspyData <- function(file="ispy_results.tsv",
   .featureData$VariableModifications <- .featureData$VariableModifications[,drop=TRUE]
   if (any(is.na(.featureData))) {
     whichCols <- apply(.featureData,2,function(x) any(is.na(x)))
-    warning(paste("NA values in featureData column(s)",
-                  names(.featureData)[whichCols]))
+    nacolnames <- paste(names(.featureData)[whichCols], collapse = ", ")
+    warning(paste0("NA values in featureData column(s) ", nacolnames))
   }
   if (any(is.na(.exprs)))
-    warning("NA values in .exprs.")
[…]

### 2012-04-19 inst/doc/MSnbase-demo/MSnbase-demo.Rnw
Misc changes
CHANGES IN VERSION 1.5.1
-------------------------
 o updated NA warning message in readIspyData <2012-03-05 Mon>
 o fixed combineFeatures/combineMatrixFeatures for 1 sample <2012-03-20 Tue>
 o image method for MSnSet instances <2012-03-20 Tue>
 o fixed bug in plotNA (first t was wring) <2012-03-21 Wed>
 o import plot from stats4 <2012-03-21 Wed>
 o using reshape2 <2012-03-30 Fri>




git-svn-id: https://hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/MSnbase@64784 bc3139a8-67e5-0310-9ffc-ced21a209358
new version 1.5.2
CHANGES IN VERSION 1.5.2
-------------------------
 o tran

@@
 library("grid")
-library("reshape")
+library("reshape2")
 @
@@
                 strict = FALSE,
+                parallel = FALSE,
                 verbose = FALSE)
@@
                        strict = FALSE,
+                       parallel = FALSE,
                        verbose = FALSE)

### 2012-04-19 R/methods-pSet.R

@@ -31,8 +31,11 @@ setValidity("pSet", function(object) {
     msg <- validMsg(msg,"featureNames differ between assayData and featureData.")
   ## checking number of files in phenoData and
   ##          number of files in assayData
-  nfilespData   <- length(processingData(object)@files)
-  nfilesSpectra <- length(unique(eapply(assayData(object),fromFile)))
+  nfilesprocData   <- length(processingData(object)@files)
+  nfilesSpectra <- length(unique(unlist(eapply(assayData(object),fromFile))))
+  if (nfilesprocData != nfilesSpectra)
+    msg <- validMsg(msg, "unequal number of files in assayData and processingData.")
+  nfilespData <- nrow(pData(object))
   if (nfilespData != nfilesSpectra)
     msg <- validMsg(msg, "unequal number of files in assayData and phenoData.")  
   ## protocolData not checked yet - depends very much

### 2012-04-19 R/functions-MSnExp.R
new version 1.5.2
CHANGES IN VERSION 1.5.2
-------------------------
 o transposing and MSnSet does _silently_ drop the 
   protocolData now <2012-04-03 Tue>
 o fixed MSnExp pData creation for multiple files, feature names 
   have a .fileNumber extension now. <2012-04-19 Thu>
 o testing for uniqueness of files (filenames) in 
   readMSData <2012-04-19 Thu>
 o updated itraqdata.RData <2012-04-19 Thu>
 o added a paralle argument to quantify and using 
   paralle = FALSE in vignette, to avoid duplicated display
   of the command <2012-04-19 Thu>




git-svn-id: https://hedgehog.fhcrc.org/biocond

@@ -140,7 +140,7 @@ clean.MSnExp <- function(object,verbose=TRUE) {
     return(object)
 }
 
-quantify.MSnExp <- function(object, method, reporters, strict, verbose) {
+quantify.MSnExp <- function(object, method, reporters, strict, parallel, verbose) {
   ## Display progress bar with eapply
   ## TODO - test if using eapply is more efficient in terms of mem/cpu usage
   ## if (verbose) {
@@ -166,10 +166,8 @@ quantify.MSnExp <- function(object, method, reporters, strict, verbose) {
     .exprs <- t(peakData$peakQuant)
     .qual <- t(peakData$curveData)
   } else {
-    parallel <- FALSE
-    if (require(foreach) & require(doMC)) {
+    if (parallel & require(foreach) & require(doMC)) {
       registerDoMC()
-      parallel <- TRUE
     }
     peakData <- llply(spectraList, quantify, method, reporters, strict,
                       .progress = progress, .parallel = parallel)
@@ -206,7 +204,8 @@ quantify.MSnExp <- function(object, method, reporters, strict, verbose) {
     } else {
       ## Here, something more clever should be done, like replicating
       ## old phenoData variables length(reporters) times
-      warning("Unexpected number of samples in phenoData slot. Dropping it.")
+      if (verbose)
+        message("Original MSnExp and new MSnSet have different number of samples in phenoData. Dropping original.")
     }
   }
   msnset <- new("MSnSet",

### 2012-04-23 inst/doc/MSnbase-demo/MSnbase-demo.Rnw
CHANGES IN VERSION 1.5.3 -------------------------  o extractSpectra is now defunct <2012-04-20 Fri>  o caching full header in level 1; this is required when and MSnExp    instance with *many* spectra (created from many raw files) is    quantified - calling header(object) is a too big overhead compared    to actual reporter quantification. <2012-04-20 Fri>  o The header() method now uses the cached dataframe if level >= 1;    the (unexported) .header function can be used to generate the    dataframe using the assayData slot data. <2012-04-20 Fri>  o Setting processingData in MSnSet initialisat

@@
 precursor charge (with second argument \Robject{z="charge"}), 
-total ion current (\Robject{z="tic"}),
+total ion count (\Robject{z="ionCount"}),
 number of peaks in the MS2 spectra \Robject{z="peaks.count"}) or, 
@@
 Similarly to \Rfunction{plot2d}, the first argument is an \Robject{MSnExp} instance. 
-The second is one of \Robject{precursor.mz}, \Robject{peaks.count} or \Robject{tic}, 
+The second is one of \Robject{precursor.mz}, \Robject{peaks.count} or \Robject{ionCount}, 
 whose density will be plotted. 
@@
 We observe the effect of the \Rfunction{removePeaks} method by comparing 
-total ion count (i.e. the total intensity in a spectrum) with the \Rfunction{tic} method
-before (object \Robject{itraqdata}) and after (object \Robject{experiment}) for 
+total ion count (i.e. the total intensity in a spectrum) with the \Rfunction{ionCount} 
+method before (object \Robject{itraqdata}) and after (object \Robject{experiment}) for 
 spectrum \texttt{X55}.
@@
 ## total ion current
-tic(itraqdata[["X55"]])
-tic(experiment[["X55"]])
+ionCount(itraqdata[["X55"]])
+ionCount(experiment[["X55"]])
 @
@@
   boxplot(exprs(x),
-          main=ifelse(is.null(ttl),processingData(x)@processing[6],ttl),
+          main=ifelse(is.null(ttl),processingData(x)@processing[2],ttl),
           cex.main=.8,

### 2012-04-23 R/utils.R

@@ -315,3 +315,73 @@ subsetBy <- function(X, groups, byIdx) {
 }
 
 
+
+## Computes header from assay data by-passing cache
+.header <- function(object) {
+  if (length(object) == 0) {
+    hd <- data.frame()
+  } else {
+    ## tbl <- table(fromFile(object))
+    ## idx <- as.numeric(unlist(apply(tbl, 1, function(x) 1:x)))
+    hd <- list(file = fromFile(object),
+               retention.time = rtime(object),
+               precursor.mz = precursorMz(object),
+               precursor.intensity = precursorIntensity(object),
+               charge = precursorCharge(object),
+               peaks.count = peaksCount(object),
+               tic = tic(object),
+               ionCount = ionCount(object),
+               ms.level = msLevel(object),
+               acquisition.number = acquisitionNum(object),
+               collision.energy = collisionEnergy(object))
+    ## items are either a numeric or a list of integer() - keep former only
+    sel <- sapply(hd, function(i) !is.list(i))
+    hd <- as.data.frame(hd[sel])    
+  }
+  return(hd)
+}
+
+checkHeader <- function(object) {
+  if (object@.cache$level == 0) {
+    ans <- TRUE
+  } else {
+    cachedhd <- header(object)
+    fromdata <- .header(object)
+    ans <- identical(cachedhd, fromdata)
+  }
+  return(ans)
+}
+
+
+updateSpectrum2 <- function(x) {
+  ## Version 0.2.0 of Spectrum has now a tic slot (MSnbase 1.5.3)
+  newx <- new("Spectrum2")
+  newx@merged <- x@merged
+  newx@precScanNum <- x@precScanNum
+  newx@precursorMz <-   x@precursorMz
+  newx@precursorIntensity <- x@precursorIntensity
+  newx@precursorCharge <- x@precursorCharge
+  newx@collisionEnergy <- x@collisionEnergy   
+  newx@msLevel <- x@msLevel
+  newx@peaksCount <- x@peaksCount
+  newx@rt <- x@rt
+  newx@acquisitionNum <- x@acquisitionNum
+  newx@scanIndex <- x@scanIndex
+  newx@mz <- x@mz
+  newx@intensity <- x@intensity
+  newx@fromFile <- x@fromFile
+  newx@centroided <- x@centroided
+  newx@tic <- 0
+  if (validObject(newx))
+    return(newx)
+}
+
[…]

### 2012-04-23 inst/tests/test_topN.R

@@ -1,6 +1,5 @@
 context("topN method")
 
-
 test_that("topN for matrix", {
   M <- matrix(c(1.0, 2.1, 10.9, # group 1
                 2.1, 1.1, 2.3,

### 2012-04-23 R/methods-MSnSet.R
CHANGES IN VERSION 1.5.3 -------------------------  o extractSpectra is now defunct <2012-04-20 Fri>  o caching full header in level 1; this is required when and MSnExp    instance with *many* spectra (created from many raw files) is    quantified - calling header(object) is a too big overhead compared    to actual reporter quantification. <2012-04-20 Fri>  o The header() method now uses the cached dataframe if level >= 1;    the (unexported) .header function can be used to generate the    dataframe using the assayData slot data. <2012-04-20 Fri>  o Setting processingData in MSnSet initialisat

@@ -40,7 +40,9 @@ setMethod("initialize", "MSnSet",
                                         ...)
             } else stop("provide at most one of 'assayData' or 'exprs' to initialize MSnSet",
                         call.=FALSE)
-            Biobase:::.harmonizeDimnames(.Object)
+            .Object@processingData <- new("MSnProcess")            
+            if (validObject(.Object))
+              Biobase:::.harmonizeDimnames(.Object)
           })
 
 
@@ -170,14 +172,15 @@ t.MSnSet <- function(x) {
 setMethod("[", "MSnSet", function(x, i, j, ...) {
   .Object <- callNextMethod(...)
   ## subsetting qual - requires pData(x)$mz!
-  fn <- featureNames(.Object)
-  reps <- match(.Object$mz,x$mz)
-  qrows <- paste(rep(fn,each=length(reps)),reps,sep=".")
-  .Object@qual <- .Object@qual[qrows,]
+  ## fn <- featureNames(.Object)
+  ## reps <- match(.Object$mz,x$mz)
+  ## qrows <- paste(rep(fn,each=length(reps)),reps,sep=".")
+  ## .Object@qual <- .Object@qual[qrows,]
+  .Object@qual <- data.frame()
   .Object@processingData@processing <- c(.Object@processingData@processing,
                                          ifelse(missing(j),
-                                                paste("Features subsetted: ",date(),sep=""),
-                                                paste("Samples subsetted: ",date(),sep="")))
+                                                paste("Features subset: ",date(),sep=""),
+                                                paste("Samples subset: ",date(),sep="")))
   if (validObject(.Object))
     return(.Object)
 })

### 2012-05-01 inst/doc/MSnbase-demo/MSnbase-demo.Rnw
demo vignette update and MIAPE instrument info
CHANGES IN VERSION 1.5.4
-------------------------
 o automatically populating experiment data instrument info 
   while reading data <2012-04-30 Mon>
 o msInfo fixed and exported <2012-04-30 Mon>
 o update demo vignette with 14 fractions analysis paragraph <2012-05-01 Tue>




git-svn-id: file:///home/git/hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/MSnbase@65631 bc3139a8-67e5-0310-9ffc-ced21a209358

@@
 the \Rfunction{writeMgfData} methods (see \Rfunction{?writeMgfData} for details and examples).
+
+\bigskip
+
+\paragraph{Experiments with multiple runs} Although it is possible to load and process multiple 
+files serially and later merge the resulting quantitation data as show in section 
+\ref{sec:combine} (page \pageref{sec:combine}), it is also feasible to load several raw data files at once. 
+Here, we report the analysis of an LC-MSMS experiment were 
+14 liquid chromatography (LC) fractions were loaded using \Rfunction{readMSData} on a 
+32-cores servers with 128 Gb of RAM. 
+It took about 90 minutes to read the 14 uncentroided \texttt{mzXML} raw files (4.9 Gb on disk in total)
+and create a 3.3 Mb raw data object (an \Robject{MSnExp} instance, see next section). 
+Quantitation of 9 reporter ions (\Robject{iTRAQ9} object, see \ref{sec:reporterions}) 
+for 88690 features was performed in parallel on 16 processors and took 76 minutes. 
+The resulting quantitation data was only 22.1 Mb and could easily be further processed and 
+analysed on a standard laptop computer.
 

### 2012-05-01 R/readMSData.R

@@ -27,11 +27,13 @@ readMSData <- function(files,
   ioncount <- c()
   filenams <- filenums <- c()
   fullhd2 <- fullhdorder <- c()
+  .instrumentInfo <- list()
   for (f in files) {
     filen <- match(f, files)
     filenums <- c(filenums, filen)
     filenams <- c(filenams, f)
-    msdata <- mzR::openMSfile(f)
+    msdata <- mzR::openMSfile(f)    
+    .instrumentInfo <- c(.instrumentInfo, list(instrumentInfo(msdata)))
     fullhd <- mzR::header(msdata)    
     ifelse(msLevel == 1,
            spidx <- which(fullhd$msLevel == 1),
@@ -175,7 +177,28 @@ readMSData <- function(files,
                  spectrum=1:length(nms),
                  row.names=nms))
   fdata <- fdata[ls(assaydata)] ## reorder features
-  ## Create and return 'MSnPeaks' object
+  ## expriment data slot
+  if (length(.instrumentInfo) > 1) {
+    cmp <- sapply(.instrumentInfo[-1], function(x) identical(x, .instrumentInfo[[1]]))
+    if (!all(cmp)) {
+      wrn <- "According to the instrument information in your files, the data has been acquired on different instruments!"
+      warning(wrn)
+      .instrumentInfo <- c(manufacturer = wrn,
+                           model = wrn,
+                           ionisation = wrn,
+                           analyzer = wrn,
+                           detector = wrn)      
+    }
+  } else {
+    .instrumentInfo <- .instrumentInfo[[1]]
+  }
+  expdata <- new("MIAPE",
+                 instrumentManufacturer = .instrumentInfo$manufacturer,
+                 instrumentModel = .instrumentInfo$model,
+                 ionSource = .instrumentInfo$ionisation,
+                 analyser = .instrumentInfo$analyzer,
+                 detectorType = .instrumentInfo$detector)
+  ## Create and return 'MSnExp' object
   if (verbose)
     cat("Creating 'MSnExp' object\n")
   toReturn <- new("MSnExp",
@@ -183,6 +206,7 @@ readMSData <- function(files,
                   phenoData = pdata,
                   featureData = fdata,
                   processingData = process,
+                  experimentData = expdata,
                   .cache = .cacheEnv)
   if (validObject(toReturn))
     return(toReturn)

### 2012-05-01 R/functions-MIAPE.R
demo vignette update and MIAPE instrument info
CHANGES IN VERSION 1.5.4
-------------------------
 o automatically populating experiment data instrument info 
   while reading data <2012-04-30 Mon>
 o msInfo fixed and exported <2012-04-30 Mon>
 o update demo vignette with 14 fractions analysis paragraph <2012-05-01 Tue>




git-svn-id: https://hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/MSnbase@65631 bc3139a8-67e5-0310-9ffc-ced21a209358

@@ -1,11 +1,11 @@
 show.MIAPE <- function(object) {
-  if (any(c(length(object@instrumentModel)>0,
-            length(object@instrumentManufacturer)>0,
-            length(object@instrumentCustomisations)>0))) {
+  if (any(c(length(object@instrumentModel) > 0,
+            length(object@instrumentManufacturer) > 0,
+            length(object@instrumentCustomisations) > 0))) {
     cat("Instrument : \n")
-    cat("  Model:",object@object@instrumentModel,"\n")
-    cat("  Manufacturer:",object@object@instrumentManufacturer,"\n")
-    cat("  Customisations:",object@object@instrumentCustomisations,"\n")
+    cat("  Model:", object@instrumentModel,"\n")
+    cat("  Manufacturer:", object@instrumentManufacturer,"\n")
+    cat("  Customisations:", object@instrumentCustomisations,"\n")
     cat("  Use 'msInfo(object)' for more MIAPE-MS information.\n")
   }
   ## code from original MIAPE show method
@@ -13,15 +13,15 @@ show.MIAPE <- function(object) {
   Index <-c(length(object@samples) > 0,
             length(object@preprocessing) > 0)
   cat("Experiment data\n")
-  cat("  Experimenter name:",object@name,"\n")
-  cat("  Laboratory:",object@lab,"\n")
-  cat("  Contact information:",object@contact,"\n")
-  cat("  Title:",object@title,"\n")
-  cat("  URL:",object@url,"\n")
+  cat("  Experimenter name:", object@name,"\n")
+  cat("  Laboratory:", object@lab,"\n")
+  cat("  Contact information:", object@contact,"\n")
+  cat("  Title:", object@title,"\n")
+  cat("  URL:", object@url,"\n")
   pmids <- pubMedIds(object)
-  cat("  PMIDs:",pmids,"\n")
+  cat("  PMIDs:", pmids,"\n")
   if(length(object@abstract) > 0 && all(object@abstract!=""))
-    cat("\n  Abstract: A",length(strsplit(object@abstract," ")[[1]]),
+    cat("\n  Abstract: A", length(strsplit(object@abstract," ")[[1]]),
         "word abstract is available. Use 'abstract' method.\n")
   else
     cat("  No abstract available.\n")
@@ -41,32 +41,31 @@ show.MIAPE <- function(object) {
   }
 }
 
-msinfo.MIAPE <- function(object) {
[…]

### 2012-05-10 inst/doc/MSnbase-io/MSnbase-io.Rnw
CHANGES IN VERSION 1.5.6 -------------------------  o new ionSource, analyser, detectorType, title accessor methods    for MIAPE, pSet and MSnSet classes <2012-05-06 Sun>  o updated quantify example to use data(itraqdata) instead of    reading dummyiTRAQ.mzXML <2012-05-09 Wed>  o initial mzTab write support <2012-05-10 Thu>  o mzTab read support <2012-05-10 Thu>  o updated demo and io vignettes with mzTab info <2012-05-10 Thu>
git-svn-id: https://hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/MSnbase@65826 bc3139a8-67e5-0310-9ffc-ced21a209358

@@
 
-\tableofcontents
+%% \tableofcontents
 
@@
 
+\Rpackage{MSnbase} also supports the \texttt{mzTab} 
+format\footnote{\url{http://code.google.com/p/mztab/}}, a light-weight, tab-delimited 
+file format for proteomics data developed within the Proteomics Standards Initiative (PSI).
+\texttt{mzTab} files can be read into \R with \Rfunction{readMzTabData} to create and \Robject{MSnSet} instance.
+
 \begin{figure}[!htb] %% input
@@
       \node [input, below of=mgf] (spreadsheet) {Quantitation data as a spreadsheet};
+      \node [input, below of=spreadsheet] (mztab) {\texttt{mzTab} format};
       \node [fun, left of=spreadsheet, node distance=4cm] (readMSnSet) {\Rfunction{readMSnSet}};
+      \node [fun, left of=mztab, node distance=4cm] (readMzTabData) {\Rfunction{readMzTabData}};
       \node [obj, left of=readMSData, yshift=-2.5mm, node distance=4cm] (MSnExp) {\Robject{MSnExp}};
@@
       \draw [->] (quantify) -- (MSnSet);
+      \draw (mztab) -- (readMzTabData);
+      \draw [->] (readMzTabData) -- (MSnSet);
     \end{tikzpicture}
@@
 
+\Robject{MSnSet} instances can also be exported to \texttt{mzTab} files using the 
+\Rfunction{writeMzTabData} function.
+
 \begin{figure}[!htb] %% out
@@
       \node [input, below of=mgf] (spreadsheet) {Quantitation data as a spreadsheet};
+      \node [input, below of=spreadsheet] (mztab) {\texttt{mzTab} format};
       \node [fun, right of=spreadsheet, node distance=4cm] (writeexprs) {\Rfunction{write.exprs}};
-
+      \node [fun, right of=mztab, node distance=4cm] (writeMzTabData) {\Rfunction{writeMzTabData}};
       \node [obj, right of=writeMgfData, yshift=1.25cm, node distance=4cm] (MSnExp) {\Robject{MSnExp}};
@@
       \draw [->] (quantify) -- (MSnSet);
+      \draw (MSnSet) -- (writeMzTabData);
+      \draw [->] (writeMzTabData) -- (mztab);
     \end{tikzpicture}

### 2012-05-10 inst/doc/MSnbase-demo/MSnbase-demo.Rnw
CHANGES IN VERSION 1.5.6 -------------------------  o new ionSource, analyser, detectorType, title accessor methods    for MIAPE, pSet and MSnSet classes <2012-05-06 Sun>  o updated quantify example to use data(itraqdata) instead of    reading dummyiTRAQ.mzXML <2012-05-09 Wed>  o initial mzTab write support <2012-05-10 Thu>  o mzTab read support <2012-05-10 Thu>  o updated demo and io vignettes with mzTab info <2012-05-10 Thu>
git-svn-id: file:///home/git/hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/MSnbase@65826 bc3139a8-67e5-0310-9ffc-ced21a209358

@@
 See also section \ref{sec:io2} to import quantitative data stored in spreadsheets into \R 
-for further processing using \Rpackage{MSnbase}.
+for further processing using \Rpackage{MSnbase}. The \texttt{MSnbase-io} vignette 
+gives a general overview of \Rpackage{MSnbase}'s input/ouput capabilites.
 
@@
 
+\Rpackage{MSnbase} also supports the \texttt{mzTab} 
+format\footnote{\url{http://code.google.com/p/mztab/}}, a light-weight, tab-delimited 
+file format for proteomics data. \texttt{mzTab} files can be read into \R with 
+\Rfunction{readMzTabData} to create and \Robject{MSnSet} instance. 
+\Robject{MSnSet} objects can also be exported to \texttt{mzTab} with the 
+\Rfunction{writeMzTabData} function.
+
+See the \texttt{MSnbase-io} vignette for a general overview of 
+\Rpackage{MSnbase}'s input/ouput capabilites.
+
 \subsection{Peak adjustments}\label{sec:purcor}

### 2012-05-10 R/readWriteMzTab.R

@@ -1,0 +1,1129 @@
+## mzTab - Reporting Proteomics Results
+## ref: http://code.google.com/p/mztab/
+
+mzTabDate <- function() format(Sys.time(), "%Y-%m-%dT%H%M%Z")
+
+## Currently only supports export of unique units
+  
+writeCOM <- function(com, file = "")
+  quite <- sapply(paste("COM", com), function(x) cat(x, file, "\n"))
+
+##' \code{mzTab} is a light-weight, tab-delimited file format for proteomics data.
+##' It describes general metadata, protein, peptide and small molecule information
+##' (all of which are optinal), including quantitation and identification.
+##' The metadata section (MTD) can be generated from an \code{\linkS4class{MSnSet}}
+##' instance using \code{makeMTD}. The detailed description of all the parameters
+##' can be found in the \code{mzTab} specification document (see references).
+##' 
+##' @title Creates the mzTab metadata section
+##' @param x An instance of class \code{\linkS4class{MSnSet}}. 
+##' @param unitId A \code{character} of lenth 1 or NULL (default),  in
+##' which case \code{x}'s variable name will be used. This identifier
+##' references the item under study all sections.
+##' @param title A \code{character} of lenght 1 or \code{NULL}
+##' (default), in which case \code{title(x)} is used if available.
+##' @param mtdDescription A \code{character} of length 1 describing the unit
+##' or \code{NULL} (default) to ignore.
+##' @param sampleProcessing A \code{list} of (possibly multiple) valid
+##' \code{CVParam} objects or \code{NULL} (default) to ignore.
+##' @param instrumentSource A list of valid \code{CVParam} instances
+##' or \code{NULL} (default), in which case \code{ionSource(x)} is
+##' used to generate a \code{CVParam}.
+##' @param instrumentAnalyzer A list of valid \code{CVParam} instances
+##' or \code{NULL} (default), in which case \code{analyser(x)} is used
+##' to generate a \code{CVParam}.
+##' @param instrumentDetector A list of valid \code{CVParam} instances
+##' or \code{NULL} (default), in which case \code{detectorType(x)} is
+##' used to generate a \code{CVParam}.
+##' @param software A list of valid \code{CVParam} instances describing
+##' the ordered list of software used to process the data. 
+##' \code{NULL} (default) to ignore.
[…]

### 2012-05-31 R/methods-MSnSet.R
CHANGES IN VERSION 1.5.8 -------------------------  o added aa data in environment data.frame <2012-05-22 Tue>  o fixed MSnbase:::subsetBy (used by topN) when    ncol(object) == 1 <2012-05-25 Fri>  o new nQuants function <2012-05-31 Thu>  o minor vignettes updates <2012-05-31 Thu>
git-svn-id: file:///home/git/hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/MSnbase@66318 bc3139a8-67e5-0310-9ffc-ced21a209358

@@ -300,13 +300,15 @@ setMethod("topN", signature(object = "MSnSet"),
             idx <- by(exprs(object), groupBy, getTopIdx, n, fun, ...)
             fn <- subsetBy(featureNames(object), groupBy, idx)
             .eset <- subsetBy(exprs(object), groupBy, idx)
+            if (!is.matrix(.eset)) 
+              .eset <- matrix(.eset, ncol = 1)
             rownames(.eset) <- fn
             .proc <- processingData(object)
             .proc@processing <- c(.proc@processing,
                                   paste0("Selected top ", n,
                                          " features: ", date()))
             .fdata <- subsetBy(fData(object), groupBy, idx)
-            message("Dropping spectrum-level 'qual' slot.")
+            ## message("Dropping spectrum-level 'qual' slot.")
             ans <- new("MSnSet",
                        experimentData = experimentData(object),
                        processingData = .proc,
@@ -353,7 +355,7 @@ setMethod("filterNA", signature(object = "matrix"),
                        function(x) sum(is.na(x))/length(x))
             accept <- k <= pNA
             if (sum(accept) == 1) {              
-              ans <- matrix(object[accept, ], nrow=1)
+              ans <- matrix(object[accept, ], nrow = 1)
               rownames(ans) <- rownames(object)[accept]
             } else {
               ans <- object[accept, ]

### 2012-05-31 inst/doc/MSnbase-demo/MSnbase-demo.Rnw
CHANGES IN VERSION 1.5.8 -------------------------  o added aa data in environment data.frame <2012-05-22 Tue>  o fixed MSnbase:::subsetBy (used by topN) when    ncol(object) == 1 <2012-05-25 Fri>  o new nQuants function <2012-05-31 Thu>  o minor vignettes updates <2012-05-31 Thu>
git-svn-id: https://hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/MSnbase@66318 bc3139a8-67e5-0310-9ffc-ced21a209358

@@
 
-<<make2exps1, echo=TRUE, cache=FALSE>>=
-exp1 <- quantify(itraqdata, reporters = iTRAQ4, verbose = FALSE)
+<<makeexp12, echo=TRUE, cache=FALSE>>=
+exp1 <- quantify(itraqdata, reporters = iTRAQ4,
+                 parallel = FALSE, verbose = FALSE)
 sampleNames(exp1)
-exp2 <- quantify(rawdata, reporters = iTRAQ4, verbose = FALSE)
+exp2 <- quantify(rawdata, reporters = iTRAQ4,
+                 parallel = FALSE, verbose = FALSE)
 sampleNames(exp2)
@@
 j <- sample(length(itraqdata), 35)
-exp1 <- quantify(itraqdata[i], reporters = iTRAQ4, verbose = FALSE)
-exp2 <- quantify(itraqdata[j], reporters = iTRAQ4, verbose = FALSE)
+exp1 <- quantify(itraqdata[i], reporters = iTRAQ4,
+                 parallel = FALSE, verbose = FALSE)
+exp2 <- quantify(itraqdata[j], reporters = iTRAQ4,
+                 parallel = FALSE, verbose = FALSE)
 table(featureNames(exp1) %in% featureNames(exp2))
-exp1 <- combineFeatures(exp1, groupBy = fData(exp1)$ProteinAccession)
-exp2 <- combineFeatures(exp2, groupBy = fData(exp2)$ProteinAccession)
+exp1 <- combineFeatures(exp1,
+                        groupBy = fData(exp1)$ProteinAccession)
+exp2 <- combineFeatures(exp2,
+                        groupBy = fData(exp2)$ProteinAccession)
 head(featureNames(exp1))

### 2012-06-02 inst/doc/MSnbase-demo/MSnbase-demo.Rnw
CHANGES IN VERSION 1.5.9 -------------------------  o topN now properly updates processingData <2012-06-01 Fri>  o combineFeatures updates featureNames based on the    groupBy argument - updated demo vignette and man    accordinlgy <2012-06-01 Fri>  o additional parameters were not passed when normalise    using vsn <2012-06-01 Fri>
git-svn-id: https://hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/MSnbase@66357 bc3139a8-67e5-0310-9ffc-ced21a209358

@@
 head(featureNames(exp1))
-featureNames(exp1) <- fData(exp1)$ProteinAccession
-featureNames(exp2) <- fData(exp2)$ProteinAccession
 head(featureNames(exp1))
@@
 
-Setting the feature names is critical when multiple experiments are to be combined, as 
-this is done using common features as defined by their names (see below).
+The feature names are updated automatically by \Rfunction{combineFeatures}, using the \Robject{groupBy} argument. 
+Proper feature names, reflecting the nature of the features (spectra, peptides or proteins) is critical when multiple experiments are to be combined, as this is done using common features as defined by their names (see below).
 

### 2012-07-15 inst/doc/MSnbase-demo/MSnbase-demo.Rnw
typos in filterNA and demo vignette
git-svn-id: https://hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/MSnbase@67664 bc3139a8-67e5-0310-9ffc-ced21a209358

@@
 It took about 90 minutes to read the 14 uncentroided \texttt{mzXML} raw files (4.9 Gb on disk in total)
-and create a 3.3 Mb raw data object (an \Robject{MSnExp} instance, see next section). 
+and create a 3.3 Gb raw data object (an \Robject{MSnExp} instance, see next section). 
 Quantitation of 9 reporter ions (\Robject{iTRAQ9} object, see \ref{sec:reporterions}) 

### 2012-08-14 inst/doc/MSnbase-development/MSnbase-development.Rnw
readMgfData update and new MSnbase-io vignette
 o in readMgfData, SCANS is not used to populate 
   acquisitionNum anymore, as several scans might 
   have been combined upstreams <2011-09-26 Mon>
 o added fillUp function in utils.R and exported <2011-09-27 Tue>
 o new MSnbase-io vignette <2011-09-28 Wed>




git-svn-id: https://hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/MSnbase@58519 bc3139a8-67e5-0310-9ffc-ced21a209358
using knitr instead of pgfSweave and misc vignette updated
git-svn-id: https://hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/MSnbase@68442 bc3139a8-67e5-0310-9f

--- 2011-03-21 inst/doc/MSnbase-development.Rnw
@@
 
-\SweaveOpts{echo=T,eval=T,fig=F,results=verbatim,cache=F,tikz=T,external=T,prefix.string=sweave-cache/figs/fig}
-
 \newcommand{\R}{\texttt{R} }
@@
 \newcommand{\Robject}[1]{{\texttt{#1}}}
-\newcommand{\Rpackage}[1]{{\textit{#1}}}
+\newcommand{\Rpackage}[1]{{\mbox{\normalfont\textsf{#1}}}}
 \newcommand{\email}[1]{\href{mailto:#1}{\normalfont\texttt{#1}}}
@@
 
+\usepackage{geometry}
+\geometry{verbose,
+  tmargin = 2.5cm,
+  bmargin = 2.5cm,
+  lmargin = 3.5cm,
+  rmargin = 3.5cm}
+
+\usepackage{hyperref}
+\usepackage{breakurl}
 \hypersetup{%
+  pdfusetitle,
+  bookmarks = {true},
+  bookmarksnumbered = {true},
+  bookmarksopen = {true},
+  bookmarksopenlevel = 2,
+  unicode = {true},
+  breaklinks = {false},
   hyperindex = {true},
@@
 
+<<'setup', include = FALSE, cache = FALSE>>=
+library(knitr)
+opts_chunk$set(fig.align = 'center', 
+               fig.show = 'hold', 
+               par = TRUE,
+               prompt = FALSE,
+               comment = NA)
+options(replace.assign = TRUE, 
+        width = 65)
+knit_hooks$set(par = function(before, options, envir) {
+  if (before && options$fig.show != 'none') 
+    par(mar = c(4,4,.1,.1),
+        cex.lab = .95,
+        cex.axis = .9,
+        mgp = c(2,.7,0),
+        tcl = -.3)
+})
+@
+
 \tableofcontents
 
-%% <<setup,cache=FALSE>>=
-%% setCacheDir("sweave-cache/values")
-%% @
-
-<<environment,cache=FALSE,echo=FALSE>>=
-library(MSnbase)
+<<environment, cache=FALSE, echo=FALSE>>=
+suppressPackageStartupMessages(library(MSnbase))
 @
@@
 
-Raw data is currently imported from \texttt{mzXML} files \citep{Pedrioli2004} <
-using the \Rfunction{xcms:::rampRawData} 
-and \Rfunction{xcms:::rampRawDataMSn} functions from the 
-\Rpackage{xcms} package \citep{Smith2006}.
-These functions do not give access to the meta data. New importer functions are 
-under development (see for instance \Rpackage{mzR}\footnote{%%
-  \url{https://github.com/sneumann/mzR/blob/master/DESCRIPTION}}) that will 
-hopefully give programmatic access to meta data stored in the data file to 
-populate the \Robject{MIAPE} object.
+%% Raw data is currently imported from \texttt{mzXML} files \citep{Pedrioli2004} <
+%% using the \Rfunction{xcms:::rampRawData} 
+%% and \Rfunction{xcms:::rampRawDataMSn} functions from the 
+%% \Rpackage{xcms} package \citep{Smith2006}.
+%% These functions do not give access to the meta data. New importer functions are 
+%% under development (see for instance \Rpackage{mzR}\footnote{%%
+%%   \url{https://github.com/sneumann/mzR/blob/master/DESCRIPTION}}) that will 
+%% hopefully give programmatic access to meta data stored in the data file to 
+%% populate the \Robject{MIAPE} object.
 
@@
 
-The choices of attributes has been dictated by the \Rfunction{xcms:::rampRawData} 
-and \Rfunction{xcms:::rampRawDataMSn} functions and what data from the 
-\texttt{mzXML} file they gave access to. 
-It is expected that some hopefully minor changes might 
-come up here when migrating to other data import packages, that allow random access 
-to \texttt{mzXML} data and support \texttt{mzML} \citep{Martens2010}.
+%% The choices of attributes has been dictated by the \Rfunction{xcms:::rampRawData} 
+%% and \Rfunction{xcms:::rampRawDataMSn} functions and what data from the 
+%% \texttt{mzXML} file they gave access to. 
+%% It is expected that some hopefully minor changes might 
+%% come up here when migrating to other data import packages, that allow random access 
+%% to \texttt{mzXML} data and support \texttt{mzML} \citep{Martens2010}.
 
@@
 
-<<Spectrum>>=
+<<Spectrum1>>=
 getClass("Spectrum1")
@@
 
-<<Spectrum>>=
+<<Spectrum2>>=
 getClass("Spectrum2")
@@
 
-<<ReporterIons>>=
+<<NAnnotatedDF>>=
 getClass("NAnnotatedDataFrame")
@@
 
-\input{NoteAboutSpeedAndMemory.tex}
+%% \input{NoteAboutSpeedAndMemory.tex}
 
@@
 
-<<label=sessioninfo,results=tex,echo=FALSE,cache=FALSE>>=
+<<label=sessioninfo, results=tex, echo=FALSE, cache=FALSE>>=
 toLatex(sessionInfo())

### 2012-08-14 inst/doc/MSnbase-io/MSnbase-io.Rnw
using knitr instead of pgfSweave and misc vignette updated
git-svn-id: https://hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/MSnbase@68442 bc3139a8-67e5-0310-9ffc-ced21a209358

@@
 
+\usepackage{geometry}
+\geometry{verbose,
+  tmargin = 2.5cm,
+  bmargin = 2.5cm,
+  lmargin = 3.0cm,
+  rmargin = 3.0cm}
+
+\usepackage{hyperref}
+\usepackage{breakurl}
 \hypersetup{%
+  pdfusetitle,
+  bookmarks = {true},
+  bookmarksnumbered = {true},
+  bookmarksopen = {true},
+  bookmarksopenlevel = 2,
+  unicode = {true},
+  breaklinks = {false},
   hyperindex = {true},
@@
 
-\input{Author.tex}
-
-\begin{document}
-
-%% pgf settup
+%% pgf setup
 \pgfdeclarelayer{background}
@@
   text width=5em, text centered, rounded corners, minimum height=5em]
+
+\input{Author.tex}
+
+\begin{document}
 
@@
 
-%% <<setup,cache=FALSE>>=
-%% setCacheDir("sweave-cache/values")
-%% @
+\vspace{10mm}
 
-%% <<environment,cache=FALSE,echo=FALSE>>=
-%% library(MSnbase)
-%% @
+<<knitr, include=FALSE, cache=FALSE>>=
+library(knitr)
+opts_chunk$set(fig.align = 'center', 
+               fig.show = 'hold', 
+               par = TRUE,
+               prompt = TRUE,
+               comment = NA)
+options(replace.assign = TRUE, 
+        width = 70)
+knit_hooks$set(par = function(before, options, envir) {
+  if (before && options$fig.show != 'none') 
+    par(mar = c(4,4,.1,.1),
+        cex.lab = .95,
+        cex.axis = .9,
+        mgp = c(2,.7,0),
+        tcl = -.3)
+})
+@
 
@@
 
+\clearpage
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2012-08-14 inst/doc/MSnbase-demo/MSnbase-demo.Rnw
using knitr instead of pgfSweave and misc vignette updated
git-svn-id: file:///home/git/hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/MSnbase@68442 bc3139a8-67e5-0310-9ffc-ced21a209358

@@
 
-\SweaveOpts{echo=T,eval=T,fig=F,results=verbatim,cache=F,tikz=T,external=F,prefix.string=sweave-cache/figs/fig}
-
 \newcommand{\R}{\texttt{R} }
@@
 
-%% <<setup,cache=FALSE>>=
-%%  setCacheDir("sweave-cache/values")
-%% @
-
-<<environment,cache=FALSE,echo=FALSE>>=
-library("MSnbase")
+<<environment, cache=FALSE, echo=FALSE>>=
+suppressPackageStartupMessages(library("MSnbase"))
+suppressPackageStartupMessages(library("zoo"))
+suppressPackageStartupMessages(library("doMC"))
+suppressPackageStartupMessages(library("multicore"))
 library("grid")
 library("reshape2")
+@
+
+<<knitr, include=FALSE, cache=FALSE>>=
+library(knitr)
+opts_chunk$set(fig.align = 'center', 
+               fig.show = 'hold', 
+               par = TRUE,
+               prompt = TRUE,
+               comment = NA)
+options(replace.assign = TRUE, 
+        width = 68)
+knit_hooks$set(par = function(before, options, envir) {
+  if (before && options$fig.show != 'none') 
+    par(mar = c(4,4,.1,.1),
+        cex.lab = .95,
+        cex.axis = .9,
+        mgp = c(2,.7,0),
+        tcl = -.3)
+})
 @
@@
 \begin{figure}[!ht]
-<<spectrum-plot,fig=T,echo=T,tikz=T,external=F,width=5,height=5>>=
+<<spectrum-plot, dev='tikz', fig.width=4, fig.height=4, fig.keep='high'>>=
 plot(sp, reporters = iTRAQ4, full = TRUE) 
@@
 
-<<bsasel,echo=FALSE>>=
+<<bsasel, echo=FALSE>>=
 bsasel <- fData(itraqdata)$ProteinAccession == "BSA"
@@
 \begin{figure}[p]
-<<msnexp-plot,fig=T,echo=T,tikz=T,external=F,width=5,height=6>>=
-plot(bsa, reporters = iTRAQ4, full = FALSE)
+<<msnexp-plot, dev='tikz', fig.width=3.7, fig.height=4.5, fig.keep='last'>>=
+plot(bsa, reporters = iTRAQ4, full = FALSE) + theme_gray(8)
 @ 
@@
 \begin{figure}[htb]
-  \begin{center}
     \includegraphics[width=\linewidth]{./plot2d-figure.png}
@@
 \label{fig:plot2d}
-  \end{center}
 \end{figure}
@@
 \begin{figure}[htb]
-  \begin{center}
     \includegraphics[width=\linewidth]{./plotDensity-figure.png}
@@
 \label{fig:plotDensity}
-  \end{center}
 \end{figure}
@@
 method before (object \Robject{itraqdata}) and after (object \Robject{experiment}) for 
-spectrum \texttt{X55}.
+spectrum \texttt{X55}. The respective spectra are shown on 
+figure \ref{fig:spectrum-clean-plot} (page \pageref{fig:spectrum-clean-plot}).
 
@@
 
-<<prepare-spectrum-plot2,echo=FALSE>>=
-p1 <- plot(itraqdata[["X55"]], full = TRUE) + theme_gray(5) 
-p2 <- plot(experiment[["X55"]], full = TRUE) + theme_gray(5) 
-@
-
 \begin{figure}[!ht]
-<<spectrum-plot2,fig=T,echo=F,tikz=T,external=F,width=5,height=2.5>>=
+<<spectrum-clean-plot, dev='tikz', fig.width=6, fig.height=3, echo=FALSE, fig.keep='high'>>=
+p1 <- plot(itraqdata[["X55"]], full = TRUE, plot = FALSE) + theme_gray(5) 
+p2 <- plot(experiment[["X55"]], full = TRUE, plot = FALSE) + theme_gray(5) 
 grid.newpage()
@@
 \caption{Same spectrum before (left) and after setting peaks <= 400 to 0.}
-\label{fig:spectrum-plot2}
+\label{fig:spectrum-clean-plot}
 \end{figure}
@@
 
-<<clean,echo=TRUE,cache=FALSE>>=
+<<clean, echo=TRUE, cache=FALSE>>=
 ## number of peaks
@@
 
-<<preprosp,cache=FALSE,echo=FALSE>>=
+<<preprosp, cache=FALSE, echo=FALSE>>=
 int <- c(0,1,1,3,1,1,0,0,0,1,3,7,3,1,0)
@@
 ppsp <- new("Spectrum2",intensity=int,mz=mz,centroided=FALSE)
-p1 <- plot(ppsp,full=TRUE) + theme_gray(5) + 
+p1 <- plot(ppsp, full = TRUE, plot = FALSE) + theme_gray(5) + 
   geom_point(size=3,alpha=I(1/3)) +
@@
   opts(title="Original spectrum") 
-p2 <- plot(removePeaks(ppsp,t=3),full=TRUE) + 
+p2 <- plot(removePeaks(ppsp,t=3), full=TRUE, plot = FALSE) + 
   theme_gray(5) +
@@
   opts(title="Peaks < 3 removed") 
-p3 <- plot(clean(removePeaks(ppsp,t=3)),full=TRUE) + 
+p3 <- plot(clean(removePeaks(ppsp,t=3)), full = TRUE, plot = FALSE) + 
   theme_gray(5) +
@@
 \begin{figure}[p]
-  \begin{center}
-<<preproc-plot,fig=T,echo=F,tikz=T,external=F,width=3,height=6>>=
+<<preproc-plot, dev='tikz', fig.width=3, fig.height=6, echo=FALSE>>=
 grid.newpage()
@@
 \label{fig:preproc}
-  \end{center}
 \end{figure}
@@
 
-<<simplesp,cache=FALSE,echo=FALSE>>=
+<<simplesp, cache=FALSE, echo=FALSE, fig.keep='none'>>=
 int <- c(0,1,1,3,1,1,0)
@@
 ssp <- new("Spectrum2",intensity=int,mz=mz,centroided=FALSE)
-p <- plot(ssp,full=TRUE)
+p <- plot(ssp, full=TRUE, plot=FALSE)
 p <- p + theme_gray(5)
@@
 \begin{figure}[p]
-<<quantitation-plot,fig=T,echo=F,tikz=T,external=F,width=5,height=5>>=
+<<quantitation-plot, dev='tikz',fig.width=5, fig.height=5, echo=FALSE>>=
 grid.newpage()
@@
 \begin{figure}[p]
-<<norm-plot,fig=T,echo=F,tikz=T,external=F,width=5,height=7>>=
+<<norm-plot, dev='tikz', fig.width=5, fig.height=7, echo=FALSE>>=
 .plot <- function(x,ttl=NULL) {
@@
 par(mfrow=c(3,2),mar=c(2.9,2.9,2.9,1))
-.plot(qnt,ttl="Non-normalised data")
-.plot(qnt.max)
-.plot(qnt.sum)
-.plot(qnt.quant)
-.plot(qnt.qrob)
-.plot(qnt.vsn)
+.plot(qnt, ttl = "Non-normalised data")
+.plot(qnt.max, ttl = "Maximum")
+.plot(qnt.sum, ttl = "Sum")
+.plot(qnt.quant, ttl = "Quantile")
+.plot(qnt.qrob, ttl = "Robust quantile")
+.plot(qnt.vsn, ttl = "vsn")
 @ 
@@
 
-\begin{center}
 \begin{figure}[!ht]
-<<cv-plot,fig=T,echo=F,tikz=T,external=F,width=5,height=4>>=
+<<cv-plot, dev='tikz', fig.width=5, fig.height=4, echo=FALSE>>=
 print(p)
@@
 \end{figure}
-\end{center}
 
@@
 
-<<prepare-msnset-nrom-plot,cache=FALSE,echo=FALSE>>=
+<<prepare-msnset-nrom-plot, cache=FALSE, echo=FALSE, keep.fig='none'>>=
 p <- plot(normalise(experiment[bsasel], "max"),
-          reporters = iTRAQ4, full = FALSE)
+          reporters = iTRAQ4, full = FALSE, plot = FALSE)
 p <- p + theme_gray(7)
@@
 \begin{figure}[p]
-<<msnexp-norm-plot,fig=T,echo=F,tikz=T,external=F,width=5,height=6>>=
+<<msnexp-norm-plot, dev='tikz', fig.width=5, fig.height=6, echo=FALSE>>=
 print(p)
@@
 \begin{figure}[!hbt]
-  \begin{center}
-<<incompdiss-plot,fig=T,echo=F,tikz=T,external=F,width=5,height=2.5>>=
+<<incompdiss-plot, dev='tikz', fig.width=5, fig.height=2.5, echo=FALSE, warning=FALSE>>=
 dfr <- melt(exprs(incompdiss))
@@
 @ 
-\caption{ Boxplot and scatterplot comparing intensities of the 4 reporter ions 
+\caption{Boxplot and scatterplot comparing intensities of the 4 reporter ions 
   (or their sum, on the right) and the incomplete dissociation specific peak. }
 \label{fig:incompdiss}
-  \end{center}
 \end{figure}
@@
 
-\clearpage
-
 \section{Session information}\label{sec:sessionInfo} 
-<<label=sessioninfo,results=tex,echo=FALSE,cache=FALSE>>=
+<<label='sessioninfo', results=tex, echo=FALSE, cache=FALSE>>=
 toLatex(sessionInfo())
@@
 \end{document}
-
-% LocalWords:  renameSamples

### 2012-08-16 inst/doc/MSnbase-demo/MSnbase-demo.Rnw
fillUp and tikzDevice for vignette
git-svn-id: file:///home/git/hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/MSnbase@68499 bc3139a8-67e5-0310-9ffc-ced21a209358

@@
 library(knitr)
+library(tikzDevice)                             
 opts_chunk$set(fig.align = 'center', 

### 2012-08-16 inst/doc/MSnbase-demo/MSnbase-demo.Rnw
not tikz usage anymore
git-svn-id: https://hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/MSnbase@68501 bc3139a8-67e5-0310-9ffc-ced21a209358

--- 2012-08-14 inst/doc/MSnbase-demo/MSnbase-demo.Rnw
@@
 \documentclass{article}
-\usepackage{tikz}
 \usepackage[nogin]{Sweave}
@@
 \begin{figure}[!ht]
-<<spectrum-plot, dev='tikz', fig.width=4, fig.height=4, fig.keep='high'>>=
+<<spectrum-plot, dev='pdf', fig.width=4, fig.height=4, fig.keep='high'>>=
 plot(sp, reporters = iTRAQ4, full = TRUE) 
@@
 \begin{figure}[p]
-<<msnexp-plot, dev='tikz', fig.width=3.7, fig.height=4.5, fig.keep='last'>>=
+<<msnexp-plot, dev='pdf', fig.width=3.7, fig.height=4.5, fig.keep='last'>>=
 plot(bsa, reporters = iTRAQ4, full = FALSE) + theme_gray(8)
@@
 \begin{figure}[!ht]
-<<spectrum-clean-plot, dev='tikz', fig.width=6, fig.height=3, echo=FALSE, fig.keep='high'>>=
+<<spectrum-clean-plot, dev='pdf', fig.width=6, fig.height=3, echo=FALSE, fig.keep='high'>>=
 p1 <- plot(itraqdata[["X55"]], full = TRUE, plot = FALSE) + theme_gray(5) 
@@
 \begin{figure}[p]
-<<preproc-plot, dev='tikz', fig.width=3, fig.height=6, echo=FALSE>>=
+<<preproc-plot, dev='pdf', fig.width=3, fig.height=6, echo=FALSE>>=
 grid.newpage()
@@
 \begin{figure}[p]
-<<quantitation-plot, dev='tikz',fig.width=5, fig.height=5, echo=FALSE>>=
+<<quantitation-plot, dev='pdf',fig.width=5, fig.height=5, echo=FALSE>>=
 grid.newpage()
@@
 \begin{figure}[p]
-<<norm-plot, dev='tikz', fig.width=5, fig.height=7, echo=FALSE>>=
+<<norm-plot, dev='pdf', fig.width=5, fig.height=7, echo=FALSE>>=
 .plot <- function(x,ttl=NULL) {
@@
 \begin{figure}[!ht]
-<<cv-plot, dev='tikz', fig.width=5, fig.height=4, echo=FALSE>>=
+<<cv-plot, dev='pdf', fig.width=5, fig.height=4, echo=FALSE>>=
 print(p)
@@
 \begin{figure}[p]
-<<msnexp-norm-plot, dev='tikz', fig.width=5, fig.height=6, echo=FALSE>>=
+<<msnexp-norm-plot, dev='pdf', fig.width=5, fig.height=6, echo=FALSE>>=
 print(p)
@@
 \begin{figure}[!hbt]
-<<incompdiss-plot, dev='tikz', fig.width=5, fig.height=2.5, echo=FALSE, warning=FALSE>>=
+<<incompdiss-plot, dev='pdf', fig.width=5, fig.height=2.5, echo=FALSE, warning=FALSE>>=
 dfr <- melt(exprs(incompdiss))

### 2012-09-07 inst/doc/MSnbase-demo/MSnbase-demo.Rnw

@@
 suppressPackageStartupMessages(library("zoo"))
+if (.Platform$OS.type != "windows")
   suppressPackageStartupMessages(library("doMC"))
@@
 @ 
+%% $
 
@@
 
-<<MSnExp,cache=FALSE,echo=TRUE>>=
+<<MSnExp, cache=FALSE, echo=TRUE>>=
 library("MSnbase")
@@
 
-<<experiment-size,echo=FALSE,cache=FALSE>>=
-sz <- sum(sapply(assayData(itraqdata),object.size)) + 
+<<experimentsize, echo=FALSE, cache=FALSE>>=
+sz <- sum(sapply(assayData(itraqdata), object.size)) + 
   object.size(itraqdata)
-sz <- round(sz/(1024^2),2)
+sz <- as.numeric(sz)
+sz <- round(sz/(1024^2), 2)
 @
@@
 
-<<Spectrum,cache=FALSE,echo=TRUE>>=
+<<Spectrum, cache=FALSE, echo=TRUE>>=
 sp <- itraqdata[["X1"]]
@@
 
-<<accessors,cache=FALSE,echo=TRUE>>=
+<<accessors, cache=FALSE, echo=TRUE>>=
 peaksCount(sp)
@@
 \begin{figure}[!ht]
-<<spectrum-plot, dev='pdf', fig.width=4, fig.height=4, fig.keep='high'>>=
+<<spectrumPlot, dev='pdf', fig.width=4, fig.height=4, fig.keep='high'>>=
 plot(sp, reporters = iTRAQ4, full = TRUE) 
@@
 
-<<bsasel, echo=FALSE>>=
-bsasel <- fData(itraqdata)$ProteinAccession == "BSA"
-@
+<<bsaSelect, eval=TRUE, echo=FALSE>>=
+## bsasel <- fData(itraqdata)$ProteinAccession == "BSA"
+bsasel <- 1:3
+@ 
+%% $
 
@@
 
-<<subset,echo=TRUE>>=
+<<subset, echo=TRUE>>=
 sel <- fData(itraqdata)$ProteinAccession == "BSA"
@@
 \begin{figure}[p]
-<<msnexp-plot, dev='pdf', fig.width=3.7, fig.height=4.5, fig.keep='last'>>=
+<<msnexpPlot, dev='pdf', fig.width=3.7, fig.height=4.5, fig.keep='last'>>=
 plot(bsa, reporters = iTRAQ4, full = FALSE) + theme_gray(8)
@@
 
-<<removePeaks,echo=TRUE,cache=FALSE>>=
+<<removePeaks, echo=TRUE, cache=FALSE>>=
 experiment <- removePeaks(itraqdata, t = 400, verbose = FALSE)
@@
   geom_hline(yintercept=3,linetype=2) + 
-  opts(title="Original spectrum") 
+  ggtitle("Original spectrum") 
 p2 <- plot(removePeaks(ppsp,t=3), full=TRUE, plot = FALSE) + 
@@
   geom_hline(yintercept=3,linetype=2) + 
-  opts(title="Peaks < 3 removed") 
+  ggtitle("Peaks < 3 removed") 
 p3 <- plot(clean(removePeaks(ppsp,t=3)), full = TRUE, plot = FALSE) + 
@@
   geom_hline(yintercept=3,linetype=2) + 
-  opts(title="Peaks < 3 removed and cleaned") 
+  ggtitle("Peaks < 3 removed and cleaned") 
 @
@@
 \begin{figure}[p]
-<<preproc-plot, dev='pdf', fig.width=3, fig.height=6, echo=FALSE>>=
+<<preprocPlot, dev='pdf', fig.width=3, fig.height=6, echo=FALSE>>=
 grid.newpage()
@@
 
-<<trimMz,echo=TRUE,cache=FALSE>>=
+<<trimMz, echo=TRUE, cache=FALSE>>=
 range(mz(itraqdata[["X55"]]))
@@
 
-<<reporters,echo=TRUE,cache=FALSE>>=
+<<reporters, echo=TRUE, cache=FALSE>>=
 mz(iTRAQ4)
@@
 \begin{figure}[p]
-<<quantitation-plot, dev='pdf',fig.width=5, fig.height=5, echo=FALSE>>=
+<<quantitationPlot, dev='pdf',fig.width=5, fig.height=5, echo=FALSE>>=
 grid.newpage()
@@
   viewport(layout.pos.row = x, layout.pos.col = y)
-print(p + opts(title = "Quantitation using 'sum'") +
+print(p + ggtitle("Quantitation using 'sum'") +
       geom_point(size = 3, alpha = I(1/3), colour = "red"),
       vp = vplayout(1, 1))
-print(p + opts(title = "Quantitation using 'max'") +
+print(p + ggtitle("Quantitation using 'max'") +
       geom_point(aes(x = 114.1, y = 3),
@@
 print(p +
-      opts(title = "Trapezoidation and strict=FALSE") +
+      ggtitle("Trapezoidation and strict=FALSE") +
       geom_polygon(alpha = I(1/5), fill = "red"),
@@
 print(p +
-      opts(title = "Trapezoidation and strict=TRUE") + 
+      ggtitle("Trapezoidation and strict=TRUE") + 
       geom_polygon(aes(x = c(NA, 114.05, 114.05, 114.1, 114.15, 114.15, NA),
@@
       vp = vplayout(2,2))
-
 @ 
@@
 
-<<quantify,echo=TRUE,cache=FALSE,keep.source=T>>=
+<<quantify, echo=TRUE, cache=FALSE, keep.source=TRUE>>=
 qnt <- quantify(experiment,
@@
 
-<<removeNa,echo=TRUE>>=
+<<removeNa, echo=TRUE>>=
 table(is.na(qnt))
@@
 
-<<purityCorrect,echo=TRUE,cache=FALSE>>=
+<<purityCorrect, echo=TRUE, cache=FALSE>>=
 impurities <- matrix(c(0.929, 0.059, 0.002, 0.000,
@@
 
-<<normalise,echo=TRUE,cache=FALSE>>=
+<<normalise, echo=TRUE, cache=FALSE>>=
 qnt.max <- normalise(qnt, "max")
@@
 \begin{figure}[p]
-<<norm-plot, dev='pdf', fig.width=5, fig.height=7, echo=FALSE>>=
+<<normPlot, dev='pdf', fig.width=5, fig.height=7, echo=FALSE>>=
 .plot <- function(x,ttl=NULL) {
@@
 
-<<cvdata,echo=FALSE,cache=FALSE>>=
+<<cvdata, echo=FALSE, cache=FALSE>>=
 sd1 <- apply(log2(exprs(qnt))+10,1,sd)
@@
 \begin{figure}[!ht]
-<<cv-plot, dev='pdf', fig.width=5, fig.height=4, echo=FALSE>>=
+<<cvPlot, dev='pdf', fig.width=5, fig.height=4, echo=FALSE>>=
 print(p)
@@
 
-<<prepare-msnset-nrom-plot, cache=FALSE, echo=FALSE, keep.fig='none'>>=
+<<prepareMsnsetNormPlot, cache=FALSE, echo=FALSE, keep.fig='none'>>=
 p <- plot(normalise(experiment[bsasel], "max"),
@@
 \begin{figure}[p]
-<<msnexp-norm-plot, dev='pdf', fig.width=5, fig.height=6, echo=FALSE>>=
+<<msnexpNormPlot, dev='pdf', fig.width=5, fig.height=6, echo=FALSE>>=
 print(p)
@@
 
-<<combineFeatures,echo=TRUE,cache=FALSE>>=
+<<combineFeatures, echo=TRUE, cache=FALSE>>=
 qnt2 <- combineFeatures(qnt, groupBy = gb, fun = "median")
@@
 
-<<incompdiss,echo=TRUE,cache=FALSE,keep.source=T>>=
+<<incompdiss, echo=TRUE, cache=FALSE, keep.source=TRUE>>=
 iTRAQ5
@@
 \begin{figure}[!hbt]
-<<incompdiss-plot, dev='pdf', fig.width=5, fig.height=2.5, echo=FALSE, warning=FALSE>>=
+<<incompdissPlot, dev='pdf', fig.width=5, fig.height=2.5, echo=FALSE, warning=FALSE>>=
 dfr <- melt(exprs(incompdiss))
@@
 head(featureNames(exp1))
-head(featureNames(exp1))
+head(featureNames(exp2))
 @

### 2012-10-02 inst/doc/MSnbase-demo/MSnbase-demo.Rnw
fixing vignettes
git-svn-id: https://hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/MSnbase@70084 bc3139a8-67e5-0310-9ffc-ced21a209358

@@
         width = 68)
-knit_hooks$set(par = function(before, options, envir) {
-  if (before && options$fig.show != 'none') 
-    par(mar = c(4,4,.1,.1),
-        cex.lab = .95,
-        cex.axis = .9,
-        mgp = c(2,.7,0),
-        tcl = -.3)
-})
+## knit_hooks$set(par = function(before, options, envir) {
+##   if (before && options$fig.show != 'none') 
+##     par(mar = c(4,4,.1,.1),
+##         cex.lab = .95,
+##         cex.axis = .9,
+##         mgp = c(2,.7,0),
+##         tcl = -.3)
+## })
 @

### 2012-10-02 inst/doc/MSnbase-development/MSnbase-development.Rnw
fixing vignettes
git-svn-id: file:///home/git/hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/MSnbase@70084 bc3139a8-67e5-0310-9ffc-ced21a209358

@@
         width = 65)
-knit_hooks$set(par = function(before, options, envir) {
-  if (before && options$fig.show != 'none') 
-    par(mar = c(4,4,.1,.1),
-        cex.lab = .95,
-        cex.axis = .9,
-        mgp = c(2,.7,0),
-        tcl = -.3)
-})
+## knit_hooks$set(par = function(before, options, envir) {
+##   if (before && options$fig.show != 'none') 
+##     par(mar = c(4,4,.1,.1),
+##         cex.lab = .95,
+##         cex.axis = .9,
+##         mgp = c(2,.7,0),
+##         tcl = -.3)
+## })
 @

### 2012-10-02 inst/doc/MSnbase-io/MSnbase-io.Rnw
fixing vignettes
git-svn-id: file:///home/git/hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/MSnbase@70084 bc3139a8-67e5-0310-9ffc-ced21a209358

@@
         width = 70)
-knit_hooks$set(par = function(before, options, envir) {
-  if (before && options$fig.show != 'none') 
-    par(mar = c(4,4,.1,.1),
-        cex.lab = .95,
-        cex.axis = .9,
-        mgp = c(2,.7,0),
-        tcl = -.3)
-})
+## knit_hooks$set(par = function(before, options, envir) {
+##   if (before && options$fig.show != 'none') 
+##     par(mar = c(4,4,.1,.1),
+##         cex.lab = .95,
+##         cex.axis = .9,
+##         mgp = c(2,.7,0),
+##         tcl = -.3)
+## })
 @

### 2012-12-14 inst/doc/MSnbase-demo/MSnbase-demo.Rnw

--- 2012-09-07 inst/doc/MSnbase-demo/MSnbase-demo.Rnw
@@
 \newcommand{\R}{\texttt{R} }
+\newcommand{\code}[1]{{\texttt{#1}}}
 \newcommand{\Rfunction}[1]{{\texttt{#1}}}
@@
         width = 68)
-knit_hooks$set(par = function(before, options, envir) {
-  if (before && options$fig.show != 'none') 
-    par(mar = c(4,4,.1,.1),
-        cex.lab = .95,
-        cex.axis = .9,
-        mgp = c(2,.7,0),
-        tcl = -.3)
-})
+## see ?evaluate::evaluate for details
+opts_knit$set(stop_on_error = 1L)
+
+## knit_hooks$set(par = function(before, options, envir) {
+##   if (before && options$fig.show != 'none') 
+##     par(mar = c(4,4,.1,.1),
+##         cex.lab = .95,
+##         cex.axis = .9,
+##         mgp = c(2,.7,0),
+##         tcl = -.3)
+## })
 @
@@
 \Sexpr{which(is.na(exprs(qnt))) %% nrow(qnt)}.
-We will remove the offending line, as we do not expect any missing peaks.
-
-<<removeNa, echo=TRUE>>=
+We will remove the offending line using the \Rfunction{filterNA} method.
+The \Robject{pNA} argument defines the percentage of accepted missing 
+values per feature. As we do not expect any missing peaks, we set it 
+to be 0 (which is also the detault value).
+
+<<filterNA, echo=TRUE>>=
 table(is.na(qnt))
+qnt <- filterNA(qnt, pNA = 0)
+sum(is.na(qnt))
+@
+
+The infrastructure around the \Robject{MSnSet} class allows flexible filtering 
+using the \Rfunction{[} sub-setting operator. Below, we mimic the behaviour of
+\Rfunction{filterNA(, pNA = 0)} by calculating the row indices that should be removed, 
+i.e. those that have at least on \Robject{NA} value and explicitly remove these row. 
+This method allows one to devise and easily apply any filtering strategy.
+
+<<removeNa, echo=TRUE, eval=FALSE>>=
 whichRow <- which(is.na((qnt))) %% nrow(qnt)
 qnt <- qnt[-whichRow, ]
-sum(is.na(exprs(qnt)))
-@
-
-The \Rfunction{filterNA} method allows to filter features based on the proportion 
-of missing data their carry, specified using the \Robject{pNA} argument. 
-The same result as above could have been achieved with 
-\Rfunction{qnt <- filterNA(qnt, pNA = 0)}, 
-specifying that we allow 0 percent of missing data. 
-See also the \Rfunction{plotNA} method to get an overview of the completeness 
-of a data set.
+@
+
+See also the \Rfunction{plotNA} method to obtain a graphical overview of 
+the completeness of a data set.
 
@@
                  parallel = FALSE, verbose = FALSE)
+exp1 <- droplevels(exp1) 
+exp2 <- droplevels(exp2)
 table(featureNames(exp1) %in% featureNames(exp2))
@@
 
-The feature names are updated automatically by \Rfunction{combineFeatures}, using the \Robject{groupBy} argument. 
-Proper feature names, reflecting the nature of the features (spectra, peptides or proteins) is critical when multiple experiments are to be combined, as this is done using common features as defined by their names (see below).
+The \Rfunction{droplevels} drops the unused \Robject{featureData} levels. 
+This is required to avoid passing absent levels as \Robject{groupBy} 
+in \Rfunction{combineFeatures}. 
+Alternatively, one could also use \Robject{factor(fData(exp1)\$ProteinAccession)} 
+as \Robject{groupBy} argument.
+
+The feature names are updated automatically by \Rfunction{combineFeatures}, 
+using the \Robject{groupBy} argument. 
+Proper feature names, reflecting the nature of the features (spectra, peptides or proteins) 
+is critical when multiple experiments are to be combined, as this is done using common 
+features as defined by their names (see below).
 
@@
 
+\section{MS$^E$ data processing}\label{sec:mse}
+
+\Rpackage{MSnbase} can also be used for MS$^E$ data independent acquisition from Waters instrument. 
+The MS$^E$ pipeline depends on the Bioconductor \Rpackage{synapter} package \citep{synapter} that 
+produces \Robject{MSnSet} instances for indvidual acquisitions. The \Rpackage{MSnbase} infrastructure can 
+subsequently be used to further combine experiments, as shown in section \ref{sec:comb2} and 
+apply \textit{top3} quantitation using the \Rfunction{topN} method.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2013-01-15 inst/doc/MSnbase-demo/MSnbase-demo.Rnw
new 'pattern' argument to filterNA
git-svn-id: file:///home/git/hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/MSnbase@71957 bc3139a8-67e5-0310-9ffc-ced21a209358
knity 1.0 compatibility
git-svn-id: https://hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/MSnbase@72561 bc3139a8-67e5-0310-9ffc-ced21a209358

@@
 \documentclass{article}
-\usepackage[nogin]{Sweave}
 \usepackage{hyperref}
@@
 
-<<quantify, echo=TRUE, cache=FALSE, keep.source=TRUE>>=
+<<quantify, echo=TRUE, cache=FALSE, tidy=FALSE>>=
 qnt <- quantify(experiment,
@@
 @
+
+The filtering criteria for \Rfunction{filterNA} can also be defined as a pattern
+of columns that can have missing values and columns that must not exhibit any. 
+See \Rfunction{?filterNA} for details and examples.
 
@@
 
-<<incompdiss, echo=TRUE, cache=FALSE, keep.source=TRUE>>=
+<<incompdiss, echo=TRUE, cache=FALSE, tidy=FALSE>>=
 iTRAQ5
@@
 \section{Session information}\label{sec:sessionInfo} 
-<<label='sessioninfo', results=tex, echo=FALSE, cache=FALSE>>=
+<<sessioninfo, results='asis', echo=FALSE, cache=FALSE>>=
 toLatex(sessionInfo())

### 2013-01-15 inst/doc/MSnbase-io/MSnbase-io.Rnw
knity 1.0 compatibility
git-svn-id: file:///home/git/hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/MSnbase@72561 bc3139a8-67e5-0310-9ffc-ced21a209358

--- 2012-08-14 inst/doc/MSnbase-io/MSnbase-io.Rnw
@@
 \usetikzlibrary{shapes,arrows,shadows,fit}
-\usepackage[nogin]{Sweave}
 \usepackage{hyperref}
@@
         width = 70)
-knit_hooks$set(par = function(before, options, envir) {
-  if (before && options$fig.show != 'none') 
-    par(mar = c(4,4,.1,.1),
-        cex.lab = .95,
-        cex.axis = .9,
-        mgp = c(2,.7,0),
-        tcl = -.3)
-})
 @
@@
 
-%% <<label=sessioninfo,results=tex,echo=FALSE,cache=FALSE>>=
+%% <<label=sessioninfo,results='asis',echo=FALSE,cache=FALSE>>=
 %% toLatex(sessionInfo())

### 2013-03-22 inst/doc/MSnbase-demo/MSnbase-demo.Rnw
Merge branch 'master' into local-hedgehog
git-svn-id: https://hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/MSnbase@74588 bc3139a8-67e5-0310-9ffc-ced21a209358
Merge branch 'master' into local-hedgehog
git-svn-id: file:///home/git/hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/MSnbase@74673 bc3139a8-67e5-0310-9ffc-ced21a209358

@@
 if (.Platform$OS.type != "windows")
-  suppressPackageStartupMessages(library("doMC"))
-suppressPackageStartupMessages(library("multicore"))
+  suppressPackageStartupMessages(require("doMC"))
+suppressPackageStartupMessages(require("multicore"))
 library("grid")
@@
 packages for data normalisation and statistical analysis. There is also a 
-\Rfunction{normalise} method for expression sets. The method takes and instance 
+\Rfunction{normalise} (also available as \Rfunction{normalize}) 
+method for expression sets. The method takes and instance 
 of class \Robject{MSnSet} as first argument, and a character to describe the 
@@
 \end{description}
+
+See \Rfunction{?normalise} for more methods. A \Rfunction{scale} method for \Robject{MSnSet} 
+instances, that relies on the \Rfunction{base::scale} function.
 

### 2013-04-29 inst/doc/MSnbase-io.Rnw
using knitr vignette builder and updates to io

--- 2013-01-15 inst/doc/MSnbase-io/MSnbase-io.Rnw
@@
+%\VignetteEngine{knitr}
 %\VignetteIndexEntry{MSnbase IO capabilities}
@@
 %\VignettePackage{MSnbase-io}
-\documentclass{article}
+
+\documentclass[12pt,a4paper,english]{scrartcl}
+\usepackage{amsmath,amsfonts,amssymb}
 \usepackage{tikz}
-\usetikzlibrary{shapes,arrows,shadows,fit}
 \usepackage{hyperref}
 \usepackage[authoryear,round]{natbib}
+\usepackage[auth-sc]{authblk}
+\usepackage{setspace}
+\onehalfspacing
+
+\usetikzlibrary{shapes,arrows,shadows,fit}
+
+% caption formatting
+\setcapindent{0em}
+\setkomafont{captionlabel}{\sffamily\bfseries}
+\setkomafont{caption}{\sffamily}
 
@@
 
-\input{Author.tex}
+\author{
+  Laurent Gatto\thanks{\email{lg390@cam.ac.uk}}
+}
+
+\affil{
+  Cambridge Center for Proteomics\\
+  University of Cambridge
+}
 
@@
 <<knitr, include=FALSE, cache=FALSE>>=
-library(knitr)
+library("knitr")
+suppressPackageStartupMessages(library("MSnbase"))
+suppressPackageStartupMessages(library("pRolocdata"))
 opts_chunk$set(fig.align = 'center', 
@@
 mass spectrometry data within the \R environment, from raw data import and processing, 
-feature quantification, quantification and statistical analysis of the results.
+feature quantification, quantification and statistical analysis of the results \citep{Gatto2012}.
 Data import functions for several formats are provided and intermediate or final 
@@
 
+\section{Creating Robject{MSnSet} from text spread sheets}
+
+This section describes the generation of \Robject{MSnSet} objects using 
+data available in a text-based spreadsheet. This entry point into \R and 
+\Rpackage{MSnbase} allows to import data processed by any of the 
+third party mass-spectrometry processing software available and proceed 
+with data exploration, normalisation and statistical analysis using 
+functions available in \R and the numerous Bioconductor packages.
+
+\subsection{A complete workflow}
+
+The following section describes a workflow that uses three input files to 
+create the \Robject{MSnSet}. These files respectively describe the quantitative 
+expression data, the sample meta-data and the feature meta-data. 
+It is taken from the \Rpackage{pRolco} tutorial and uses example files from 
+the \Rpackage{pRolocdata} package.
+
+\subsubsection{The original data file}\label{sec:orgcsv}
+
+We start by describing the  \texttt{csv} to be used as input using the \Rfunction{read.csv} function.
+
+<<readCsvData0>>=
+## The original data for replicate 1, available
+## from the pRolocdata package
+f0 <- dir(system.file("extdata", package = "pRolocdata"), 
+          full.names = TRUE, pattern = "pr800866n_si_004-rep1.csv")
+csv <- read.csv(f0)
+@
+
+The three first lines of the original spreadsheet, containing the data for replicate one, are illustrated below (using the function \Rfunction{head}).
+It contains \Sexpr{nrow(csv)} rows (proteins) and \Sexpr{ncol(csv)} columns, including protein identifiers, database accession numbers, gene symbols, reporter ion quantitation values, information related to protein identification, \ldots
+
+<<showOrgCsv>>=
+head(csv, n=3)
+@
+
+Below read in turn the spread sheets that contain the quantitation data (\texttt{exprsFile.csv}),
+feature meta-data (\texttt{fdataFile.csv}) and sample meta-data (\texttt{pdataFile.csv}).
+
+<<readCsvData1, tidy = FALSE>>=
+## The quantitation data, from the original data
+f1 <- dir(system.file("extdata", package = "pRoloc"), 
+          full.names = TRUE, pattern = "exprsFile.csv")
+exprsCsv <- read.csv(f1)
+## Feature meta-data, from the original data
+f2 <- dir(system.file("extdata", package = "pRoloc"), 
+          full.names = TRUE, pattern = "fdataFile.csv")
+fdataCsv <- read.csv(f2)
+## Sample meta-data, a new file
+f3 <- dir(system.file("extdata", package = "pRoloc"), 
+          full.names = TRUE, pattern = "pdataFile.csv")
+pdataCsv <- read.csv(f3)
+@
+
+\begin{description}
+
+\item[\texttt{exprsFile.csv}] containing the quantitation (expression) data for the \Sexpr{nrow(exprsCsv)} proteins and 4 reporter tags.
+<<showExprsFile>>=
+head(exprsCsv, n=3)
+@
+\item[\texttt{fdataFile.csv}] containing meta-data for the \Sexpr{nrow(fdataCsv)} features (here proteins).
+<<showFdFile>>=
+head(fdataCsv, n=3)
+@
+\item[\texttt{pdataFile.csv}] containing samples (here fractions) meta-data. This simple file has been created manually.
+<<showPdFile>>=
+pdataCsv
+@
+\end{description}
+
+The self-contained \Robject{MSnSet} can now easily be generated using the \Rfunction{readMSnSet} constructor, 
+providing the respective \texttt{csv} file names shown above and specifying that the data is comma-separated 
+(with \Robject{sep = ","}). Below, we call that object \Robject{res} and display its content.
+
+<<makeMSnSet, tidy = FALSE>>=
+library("MSnbase")
+res <- readMSnSet(exprsFile = f1,
+                  featureDataFile = f2,
+                  phenoDataFile = f3,
+                  sep = ",")
+res
+@
+
+\subsubsection{The \Robject{MSnSet} class}
+
+Although there are additional specific sub-containers for additional meta-data 
+(for instance to make the object MIAPE compliant), the feature 
+(the sub-container, or slot \Robject{featureData}) and sample (the \Robject{phenoData} slot) 
+are the most important ones. They need to meet the following validity requirements 
+(see figure \ref{fig:msnset}):
+
+\begin{itemize}
+\item the number of row in the expression/quantitation data and feature data must 
+  be equal and the row names must match exactly, and  
+\item the number of columns in the expression/quantitation data and number of row 
+  in the sample meta-data must be equal and the column/row names must match exactly. 
+\end{itemize}
+
+A detailed description of the \Robject{MSnSet} class is available by typing \Rfunction{?MSnSet} in the \R console.
+
+\begin{figure}[!hbt]
+\centering
+    \includegraphics[width=0.5\textwidth]{msnset.png}
+\caption{Dimension requirements for the respective expression, feature and sample meta-data slots. }
+\label{fig:msnset}
+\end{figure}
+
+The individual parts of this data object can be accessed with their respective accessor methods:
+
+\begin{itemize}
+\item the quantitation data can be retrieved with \Rfunction{exprs(res)}, 
+\item the feature meta-data with \Rfunction{fData(res)} and 
+\item the sample meta-data with \Rfunction{pData(res)}. 
+\end{itemize}
+
 \clearpage

### 2013-04-29 inst/doc/MSnbase-io.Rnw
add forword to io
export new readMSnSet2 function and new io section

@@
   This vignette describes \Rpackage{MSnbase}'s input and output 
-  capabilites.
+  capabilities.
 \end{abstract}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%\input{Foreword.tex}
+\input{Foreword.tex}
 
@@
 results can also be saved or exported. 
-These capabilites are presented below.
+These capabilities are presented below.
 
@@
 quantitative data and exported as a spreadsheet (generally comma or tab separated format). 
-This data as well as any additional metadata can be imported with the 
+This data as well as any additional meta-data can be imported with the 
 \Rfunction{readMSnSet} function. See \Rfunction{?readMSnSet} for details.
@@
 \paragraph{Peak lists} \Robject{MSnExp} instances as well as individual spectra can be written 
-as \texttt{mgf} files with the \Rfunction{writeMgfData} method. Note that the metadata in the original 
+as \texttt{mgf} files with the \Rfunction{writeMgfData} method. Note that the meta-data in the original 
 \R object can not be included in the file. See \Rfunction{?writeMgfData} for details.
@@
 \paragraph{Quantitation data} Quantitation data can be exported to spreadsheet files with 
-the \Rfunction{write.exprs} method. Feature metadata can be appended to the feature 
-intenstiy values. See \Rfunction{?writeMgfData} for details.
+the \Rfunction{write.exprs} method. Feature meta-data can be appended to the feature 
+intensity values. See \Rfunction{?writeMgfData} for details.
 
@@
 
-\section{Creating Robject{MSnSet} from text spread sheets}
+\clearpage
+
+\section{Creating \Robject{MSnSet} from text spread sheets}
 
@@
 
-\subsection{A complete workflow}
+\subsection{A complete work flow}
 
-The following section describes a workflow that uses three input files to 
+The following section describes a work flow that uses three input files to 
 create the \Robject{MSnSet}. These files respectively describe the quantitative 
 expression data, the sample meta-data and the feature meta-data. 
-It is taken from the \Rpackage{pRolco} tutorial and uses example files from 
+It is taken from the \Rpackage{pRoloc} tutorial and uses example files from 
 the \Rpackage{pRolocdata} package.
@@
 
-<<readCsvData0>>=
+<<readCsvData0, tidy = FALSE>>=
 ## The original data for replicate 1, available
@@
 f0 <- dir(system.file("extdata", package = "pRolocdata"), 
-          full.names = TRUE, pattern = "pr800866n_si_004-rep1.csv")
+          full.names = TRUE, 
+          pattern = "pr800866n_si_004-rep1.csv")
 csv <- read.csv(f0)
@@
 
-\clearpage
+\subsection{A shorter work flow}
+
+The \Rfunction{readMSnSet2} function provides a simplified import workforce. 
+It takes a single spreadsheet as input (default is \texttt{csv}) and extract the 
+columns identified by \Robject{ecol} to create the expression data, while the 
+others are used as feature meta-data. \Robject{ecol} can be a \Robject{character} with 
+the respective column labels or a numeric with their indices. In the former case, 
+it is important to make sure that the names match exactly. Special characters like 
+\texttt{'-'} or \texttt{'('} will be transformed by \R into \texttt{'.'} when the 
+\texttt{csv} file is read in.
+Optionally, one can also specify a column to be used as feature names. 
+Note that these must be unique to guarantee the final object validity.
+
+<<readMSnSet2>>=
+ecol <- paste("area", 114:117, sep = ".")
+fname <- "Protein.ID"
+eset <- readMSnSet2(f0, ecol, fname)
+eset
+@
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \section{Session information}
-%% \label{sec:sessionInfo}
+\section{Session information}
+\label{sec:sessionInfo}
 
-%% <<label=sessioninfo,results='asis',echo=FALSE,cache=FALSE>>=
-%% toLatex(sessionInfo())
-%% @
+<<label=sessioninfo,results='asis',echo=FALSE,cache=FALSE>>=
+toLatex(sessionInfo())
+@
 

### 2014-02-27 R/imputation.R
new impute docs and namespace

@@ -1,12 +1,3 @@
-##' Imputation of missing values
-##'
-##' @title Missing values imputation
-##' @param object An instance of class \code{"\linkS4class{MSnSet}"}.
-##' @param method The method of imputation to use. Default value is bpca.
-##' @references Oba et al., A Bayesian missing value estimation method for 
-##' gene expression profile data, Bioinformatics (2003) 19 (16): 2088-2096.
-##' @return An instance of class \code{"\linkS4class{MSnSet}"}.
-##' @author Laurent Gatto, Samuel Wieczorek, Vasile-Cosmin Lazar
 setMethod("impute", "MSnSet",
           function(object, method = c("bpca","knn"), ...) {
             method <- match.arg(method)
@@ -17,8 +8,9 @@ setMethod("impute", "MSnSet",
             }
             else if (method=="bpca"){
              nSamples <- dim(exprs(object))[2]
-             .resultBPCA = pca(exprs(object), method="bpca", nPcs=(nSamples-1), verbose=F)
-             exprs(object) = completeObs(.resultBPCA)
+             .resultBPCA <- pca(exprs(object), method = "bpca",
+                 nPcs = (nSamples-1), verbose = FALSE, ...)
+             exprs(object) <- completeObs(.resultBPCA)
             }
             impargs <- pairlist(...)
             object@processingData@processing <-

### 2014-03-24 tests/testthat/test_MSnSet.R
add idSummary,MSnExp,MSnSet

@@ -25,16 +25,16 @@ test_that("Combine MSnSet features (V)", {
                       B = paste(
                           rep(c("A","B"), each = 5),
                           1:10,
-                          sep = "."))))    
+                          sep = "."))))
     expect_true(validObject(aa))
     ## gb <- factor(rep(letters[1:2], each = 5))
     gb <- factor(rep(1:2, each = 5))
     bb <- MSnbase:::nologging(combineFeatures(aa, gb, "mean"), 1)
     bb2 <- MSnbase:::nologging(combineFeatures(aa, as.character(gb), "mean"), 1)
-    gbn <- rep(1:2, each = 5)    
+    gbn <- rep(1:2, each = 5)
     bb3 <- MSnbase:::nologging(combineFeatures(aa, gbn, "mean"), 1)
     expect_equal(bb, bb2)
-    expect_equal(bb, bb3)  
+    expect_equal(bb, bb3)
     cc <- combineFeatures(aa, gb, "sum")
     dd <- combineFeatures(aa, gb, "median")
     ee <- combineFeatures(aa, gb, "weighted.mean", w=rep(1,10))
@@ -63,8 +63,8 @@ test_that("Combine MSnSet features (V)", {
     expect_true(all(fData(bb)[,2] == c("A.1", "B.6")))
     gb2 <- factor(c("a", "c", "z", "a", "z", "b", "b", "a", "c", "a"))
     gb3 <- factor(rev(c("a", "c", "z", "a", "z", "b", "b", "a", "c", "a")))
-    fData(aa)$Z <- gb2  
-    zz <- combineFeatures(aa, gb2, fun = "sum")  
+    fData(aa)$Z <- gb2
+    zz <- combineFeatures(aa, gb2, fun = "sum")
     zz3 <- combineFeatures(aa[10:1, ], gb3, fun = "sum")
     expect_true(all.equal(exprs(zz), exprs(zz3)))
     expect_true(all.equal(fData(zz)[, 3:5],
@@ -83,7 +83,7 @@ test_that("Combine MSnSet features (V)", {
                           check.attributes = FALSE))
     expect_true(all.equal(featureNames(zz),
                           as.character(fData(zz)$Z),
-                          check.attributes = FALSE))  
+                          check.attributes = FALSE))
 })
 
 
@@ -103,7 +103,7 @@ test_that("Combine MSnSet features (L)", {
     names(L) <- LETTERS[1:5]
     expect_error(combineFeatures(ee, L))
 
[…]

### 2014-03-25 R/AllGenerics.R
merged master
git-svn-id: file:///home/git/hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/MSnbase@87858 bc3139a8-67e5-0310-9ffc-ced21a209358

@@ -17,7 +17,7 @@ setGeneric("precAcquisitionNum",function(object) standardGeneric("precAcquisitio
 setGeneric("precScanNum",function(object) standardGeneric("precScanNum"))
 setGeneric("rtime",function(object) standardGeneric("rtime"))
 setGeneric("msLevel",function(object) standardGeneric("msLevel"))
-setGeneric("collisionEnergy",function(object) standardGeneric("collisionEnergy")) 
+setGeneric("collisionEnergy",function(object) standardGeneric("collisionEnergy"))
 ## setGeneric("peaksCount",function(object) standardGeneric("peaksCount")) ## use mzR generic
 ## setGeneric("header",function(object) standardGeneric("header")) ## use mzR generic
 setGeneric("polarity", function(object) standardGeneric("polarity"))
@@ -87,4 +87,10 @@ setGeneric("impute",function(object, ...) standardGeneric("impute"))
 ## mzR
 setGeneric("chromatogram",function(object, ...) standardGeneric("chromatogram"))
 setGeneric("xic",function(object, ...) standardGeneric("xic"))
+
+## identification
+setGeneric("addIdentificationData", function(object, ...) standardGeneric("addIdentificationData"))
+setGeneric("idSummary", function(object, ...) standardGeneric("idSummary"))
+setGeneric("removeNoId", function(object, ...) standardGeneric("removeNoId"))
+setGeneric("removeMultipleAssignment", function(object, ...) standardGeneric("removeMultipleAssignment"))
 

### 2014-03-25 R/methods-MSnSet.R

@@ -40,7 +40,7 @@ setMethod("initialize", "MSnSet",
                                         ...)
             } else stop("provide at most one of 'assayData' or 'exprs' to initialize MSnSet",
                         call.=FALSE)
-            .Object@processingData <- new("MSnProcess")            
+            .Object@processingData <- new("MSnProcess")
             if (validObject(.Object))
               Biobase:::.harmonizeDimnames(.Object)
           })
@@ -58,7 +58,7 @@ setValidity("MSnSet", function(object) {
                       "number of rows in assayData and qual slots do not match.")
   }
   if (!inherits(experimentData(object),"MIAPE"))
-    msg <- validMsg(msg, 
+    msg <- validMsg(msg,
                     "experimentData slot in MSnSet must be 'MIAPE' object")
   if (is.null(msg)) TRUE else msg
 })
@@ -96,7 +96,7 @@ setMethod("scale", "MSnSet",
             if (!is.null(attr(e, "scaled:scale"))) {
               escale <- attr(e, "scaled:scale")
               attr(e, "scaled:scale") <- NULL
-              escale <- paste(round(escale, ), 3)              
+              escale <- paste(round(escale, ), 3)
               proc <- c(proc,
                         paste("Scaled",
                               paste0("[", escale, collapse = ", "), "]:",
@@ -105,18 +105,18 @@ setMethod("scale", "MSnSet",
             if (!is.null(attr(e, "scaled:center"))) {
               ecenter <- attr(e, "scaled:center")
               attr(e, "scaled:center") <- NULL
-              ecenter <- paste(round(ecenter, ), 3)              
+              ecenter <- paste(round(ecenter, ), 3)
               proc <- c(proc,
                         paste("Centered",
                               paste0("[", ecenter, collapse = ", "), "]:",
                               date()))
-            }            
+            }
             exprs(x) <- e
             x@processingData@processing <-
               c(x@processingData@processing,
                 proc)
             if (validObject(x))
-              return(x) 
+              return(x)
           })
 
 setMethod("purityCorrect",
@@ -128,14 +128,14 @@ setMethod("purityCorrect",
               stop("Impurity matrix should be ",ncol(object)," by ",ncol(object))
             .purcor <- function(x, .impurities = impurities) {
               keep <- !is.na(x)
-              if (sum(keep)>1) 
[…]

### 2014-03-25 R/methods-NAnnotatedDataFrame.R

@@ -2,8 +2,8 @@
 ## Methods for NAnnotatedDataFrame class
 setMethod("dim", "NAnnotatedDataFrame", function(x) {
   d <- c(dim(pData(x)),
-         length(x@multiplex))
-  names(d) <- c(dimLabels(x),"multipleNames")
+         x@multiplex)
+  names(d) <- c(dimLabels(x), "multiNames")
   return(d)
 })
 
@@ -12,8 +12,8 @@ setMethod("show",
           function(object) {
             callNextMethod(object)
             cat("  Multiplexing: ")
-            if (length(object@multiplex)>0)
-                cat(object@multiplex," - ",object@multiLabels,"\n",sep="")
+            if (length(object@multiplex) > 0)
+                cat(object@multiplex," - ", object@multiLabels, "\n", collapse=" ")
             else
               cat("none\n")
             invisible(NULL)

### 2014-03-25 R/methods-pSet.R

@@ -1,53 +1,67 @@
 setMethod("initialize",
           signature(.Object="pSet"),
           function(.Object, ..., .cache) {
-            if (missing(.cache)) {
-              .cache <- new.env()
-              assign("level", 0 ,.cache)
-              lockEnvironment(.cache)
-            }
-            callNextMethod(.Object, ..., .cache = .cache)
+              if (missing(.cache)) {
+                  .cache <- new.env()
+                  assign("level", 0 ,.cache)
+                  lockEnvironment(.cache)
+              }
+              callNextMethod(.Object, ..., .cache = .cache)
           })
 
 setValidity("pSet", function(object) {
-  msg <- validMsg(NULL, NULL)
-  if (!all(sapply(assayData(object),function(x) inherits(x,"Spectrum"))))
-    msg <- validMsg(msg,
-                    "assayData must contain 'Spectrum' objects.")
-  msl <- msLevel(object)
-  if (length(unique(msl))>1) 
-    warning(paste("Different MS levels in ",class(object),
-                  " object:",unique(msl)))
-  ## checking number of spectra in assayData and
-  ##          number of features in featureData
-  nspectra  <- length(assayData(object)) 
-  nfeatures <- nrow(featureData(object)) 
-  if (nspectra != nfeatures)
-    msg <- validMsg(msg,"unequal number of spectra in assayData and features in featureData.")
-  if (length(spectra(object)) != length(ls(assayData(object))))
-    msg <- validMsg(msg,"object size inconsistence using assayData() and spectra() methods.")  
-  if (!identical(featureNames(object), ## obtained as ls(assayData(object))
-                 featureNames(featureData(object))))
-    msg <- validMsg(msg,"featureNames differ between assayData and featureData.")
-  ## checking number of files in phenoData and
-  ##          number of files in assayData
-  nfilesprocData   <- length(processingData(object)@files)
-  nfilesSpectra <- length(unique(unlist(eapply(assayData(object),fromFile))))
-  if (nfilesprocData != nfilesSpectra)
-    msg <- validMsg(msg, "unequal number of files in assayData and processingData.")
-  nfilespData <- nrow(pData(object))
-  if (nfilespData != nfilesSpectra)
-    msg <- validMsg(msg, "unequal number of files in assayData and phenoData.")  
-  ## protocolData not checked yet - depends very much
-  ## on type of assay (MS1, MS2 quant, reporter ions, ...)
-  if (!cacheEnvIsLocked(object))
[…]

### 2014-03-25 R/quantitation-MS2-isobaric.R

@@ -1,0 +1,114 @@
+quantify_MSnExp <- function(object, method,
+                            reporters, strict,
+                            parallel, verbose) {
+    ## Display progress bar with eapply
+    ## TODO - test if using eapply is more efficient in terms of mem/cpu usage
+    ## if (verbose) {
+    ##   ._cnt <- 1
+    ##   pb <- txtProgressBar(min = 0, max = length(object), style = 3)
+    ##   ## Quantification -- creating exprs for assayData slot
+    ##   peakData <- eapply(assayData(object),function(x) {
+    ##     setTxtProgressBar(pb, ._cnt)
+    ##     ._cnt <<- ._cnt+1
+    ##     quantify(x,method,reporters)
+    ##   })
+    ##   close(pb)
+    ##   rm(pb)
+    ##   rm(._cnt)
+    ## } else {
+    ##   peakData <- eapply(assayData(object),quantify,method,reporters)
+    ## }
+    detectCores <- registerDoMC <- NULL ## no visible global function definition
+    ifelse(verbose, progress <- "text", progress <- "none")
+    if (.Platform$OS.type == "windows") {
+        parallel <- FALSE
+        if (verbose)
+            message("Parallel processing not yet supported on Windows.")
+    }
+    if (any(centroided(object)) & method == "trapezoidation")
+        warning("You are quantifying using 'trapezoidation' on centroided data!",
+                immediate. = TRUE)
+    spectraList <- spectra(object)
+    ## Quantification -- creating exprs for assayData slot
+    if (length(spectraList) == 1) {
+        peakData <- quantify(spectraList[[1]], method, reporters, strict)
+        .exprs <- t(peakData$peakQuant)
+        rownames(.exprs) <- featureNames(object)
+        ## .qual <- as.data.frame(peakData$curveStats)
+        ## rownames(.qual) <-
+        ##   paste(reporterNames(reporters),
+        ##         rep(featureNames(object), each = length(reporters)),
+        ##         sep = ".")
+        .qual <- data.frame()
+    } else {
+        if (parallel && require("foreach") && require("doMC") && require("parallel")) {
+            registerDoMC(cores = detectCores())
+        }
+        peakData <- llply(spectraList, quantify, method, reporters, strict,
+                          .progress = progress, .parallel = parallel)
+        .exprs <- do.call(rbind, sapply(peakData, "[", "peakQuant"))
[…]

### 2014-03-25 tests/testthat/test_NAnnotatedDataFrame.R

@@ -1,0 +1,22 @@
+context("NAnnotatedDataFrame")
+
+test_that("NAnnotatedDataFrame validity", {
+    na <- new("NAnnotatedDataFrame")
+    expect_true(validObject(na))
+    dnms <- c(rowNames = 0, columnNames = 0, multiNames = 1)
+    expect_true(all.equal(dim(na), dnms))
+    pd <- data.frame(pvarA = "A", pvarB = "B")
+    na <- new("NAnnotatedDataFrame", data = pd) 
+    expect_true(validObject(na))
+    dnms <- c(rowNames = 1, columnNames = 2, multiNames = 1)
+    expect_true(all.equal(dim(na), dnms))
+    na <- new("NAnnotatedDataFrame",
+              multiplex = 4, multiLabels = reporterNames(iTRAQ4),
+              data = pd)
+    expect_true(validObject(na))
+    dnms <- c(rowNames = 1, columnNames = 2, multiNames = 4)
+    expect_true(all.equal(dim(na), dnms))
+
+    expect_true(multiplex(na) == 4)
+    expect_true(all.equal(multiLabels(na), reporterNames(iTRAQ4)))
+})

### 2014-05-14 R/methods-MSnExp.R
fixes and mention BiocParallel in vignette

@@ -180,7 +180,8 @@ setMethod("trimMz",
 setMethod("quantify",
           signature = signature("MSnExp"),
           function(object,
-                   method = c("trapezoidation", "max", "sum",
+                   method = c(
+                       "trapezoidation", "max", "sum",
                        "SI", "SIgi", "SIn",
                        "SAF", "NSAF",
                        "count"),
@@ -191,7 +192,7 @@ setMethod("quantify",
                    verbose = TRUE,
                    ...) {
               if (!missing(parallel))
-                  message("Please use the BPPARAM argument to set a parallel framework.")
+                  message("Please use BPPARAM to set a parallel framework.")
               method <- match.arg(method)
               ## this assumes that if first spectrum has msLevel > 1, all have
               if (msLevel(object)[1] < 2)
@@ -201,12 +202,12 @@ setMethod("quantify",
                   if (!inherits(reporters, "ReporterIons"))
                       stop("Argument 'reporters' must inherit from 'ReporterIons' class.")
                   if (missing(BPPARAM)) {
-                      parallel <- bpparam()
+                      BPPARAM <- bpparam()
                       if (verbose)
                           message("Using default parallel backend: \n",
-                                  PBPARAM)
+                                  BPPARAM)
                   }
-                  quantify_MSnExp(object, method, reporters, strict, parallel)
+                  quantify_MSnExp(object, method, reporters, strict, BPPARAM, verbose)
               } else if (method == "count") {
                   count_MSnSet(object)
               } else {
@@ -218,7 +219,7 @@ setMethod("quantify",
           })
 
 setMethod("curveStats","MSnExp",
-          function(object, reporters, verbose=TRUE) {
+          function(object, reporters, verbose = TRUE) {
             ifelse(verbose,progress <- "text",progress <- "none")
             l <- llply(object@spectra, curveStats, reporters, .progress=progress)
             qdfr <- l[[1]]

### 2014-10-31 R/readWriteMzTab.R
fix conflict
git-svn-id: https://hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/MSnbase@96275 bc3139a8-67e5-0310-9ffc-ced21a209358

@@ -177,14 +177,14 @@ makeMTD <- function(x,
   if (is.null(instrumentSource)) {
     if (length(ionSource(x)) > 0) {
       instrumentSource <- ionSource(x)
-      instrumentSource <- olsQuery(instrumentSource, "PSI", exact = TRUE)
+      instrumentSource <- rols::olsQuery(instrumentSource, "PSI", exact = TRUE)
       if (length(instrumentSource) != 1) {
         instrumentSource <- NULL
         warning("No unique instrumentSource ontology term found.")
       } else {
         .label <- strsplit(names(instrumentSource), ":")[[1]][1]
         .accession <- names(instrumentSource)
-        .name <- term(.accession, .label)
+        .name <- rols::term(.accession, .label)
         instrumentSource <- new("CVParam", label = .label, accession = .accession, name = .name)
       }
     }
@@ -194,14 +194,14 @@ makeMTD <- function(x,
   if (is.null(instrumentAnalyzer)) {
     if (length(analyzer(x)) > 0) {
       instrumentAnalyzer <- analyzer(x)
-      instrumentAnalyzer <- olsQuery(instrumentAnalyzer, "MS", exact = TRUE)
+      instrumentAnalyzer <- rols::olsQuery(instrumentAnalyzer, "MS", exact = TRUE)
       if (length(instrumentAnalyzer) != 1) {
         instrumentAnalyzer <- NULL
         warning("No unique instrumentAnalyzer ontology term found.")
       } else {
         .label <- strsplit(names(instrumentAnalyzer), ":")[[1]][1]
         .accession <- names(instrumentAnalyzer)
-        .name <- term(.accession, .label)
+        .name <- rols::term(.accession, .label)
         instrumentAnalyzer <- new("CVParam",
                                   label = .label,
                                   accession = .accession,
@@ -214,14 +214,14 @@ makeMTD <- function(x,
   if (is.null(instrumentDetector)) {
     if (length(detectorType(x)) > 0) {
       instrumentDetector <- detectorType(x)
-      instrumentDetector <- olsQuery(instrumentDetector, "MS", exact = TRUE)
+      instrumentDetector <- rols::olsQuery(instrumentDetector, "MS", exact = TRUE)
       if (length(instrumentDetector) != 1) {
         instrumentDetector <- NULL
         warning("No unique instrumentDetector ontology term found.")
       } else {
[…]

### 2015-01-09 R/methods-Spectrum.R
Merge branch 'master' into local-hedgehog
git-svn-id: https://hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/MSnbase@98188 bc3139a8-67e5-0310-9ffc-ced21a209358

@@ -36,6 +36,11 @@ setMethod("plot", c("Spectrum", "Spectrum"),
             plotSpectrumVsSpectrum(list(x, y), ...)
           })
 
+setMethod("plot", c("Spectrum2", "character"),
+          function(x, y, ...) {
+              .plotSingleSpectrum(x, y, ...)
+          })
+
 setMethod("clean",
           signature=signature("Spectrum"),
           function(object, all = FALSE) clean_Spectrum(object, all))

### 2016-05-26 R/readMSData.R
Merge branch 'master' into devel
* master:
  fix map unit test
  remove useless getBins function
  don't export as.data.frame.MSmap - use as
  fix MSmap to dataframe coerce
  finish MSmap unit test
  update MSmap unittest
  update news
  MSmap unit testing and fix bug
  MSmap unit testing
  rearrange mzTab unit test - now counted in coverage
  update old googlecode url
  new gh dev version

From: Laurent <lg390@cam.ac.uk>

git-svn-id: file:///home/git/hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/MSnbase@117880 bc3139a8-67e5-0310-9ffc-ced21a209358

@@ -189,8 +189,6 @@ readMSData <- function(files,
                     basename(f),"\n",sep="")
                 pb <- txtProgressBar(min=0,max=length(spidx),style=3)
             }
-            ## this was fullhd$acquisitionNum -- check/wrong
-            ## ms1scanNums <- MSnbase:::getBins(fullhd$msLevel[spidx])
             scanNums <- fullhd[fullhd$msLevel == 2, "precursorScanNum"]
             if (length(scanNums)!=length(spidx))
                 stop("Number of spectra and precursor scan number do not match!")

### 2016-08-30 R/methods-Spectrum.R
Merge branch 'master' into devel
* master: (312 commits)
  close #153: update low/highMZ checks in validateOnDiskMSnSet
  suggest microbenchmark and suppress message in benchark vig
  Add unit test for issue #118
  Fix failing tests and example due to r120413 in Biobase
  Fix for loop that iterates over spectra to work with empty objects
  Spectra import without need to read mzML header
  Intermediate fix for issue #151
  add unit test for extractMgfSpectrum2Info
  fix warning in readMgfData if TITLE contains multiple =
  split MSnExp tests and new [[ error test
  fix OnDiskMSnExp [[ subsettin

@@ -1,39 +1,40 @@
 ##################################################################
 ## Methods for Spectrum class and children
-setMethod("initialize",
-          "Spectrum",
-          function(.Object,...,mz,intensity,peaksCount) {
-            if (!missing(mz) & !missing(intensity)) {
-              .Object <- callNextMethod(.Object,
-                                        ...,
-                                        intensity=intensity,
-                                        mz=mz,
-                                        peaksCount=length(mz))
-            } else if (!missing(mz) | !missing(intensity)) {
-              stop("'mz' and 'intensity' or none required.")
-            } else {
-              .Object <- callNextMethod(.Object,...)
-            }
-            if (validObject(.Object))
-              .Object
-          })
+## setMethod("initialize",
+##           "Spectrum",
+##           function(.Object, ..., mz, intensity, peaksCount) {
+##               if (xor(!missing(mz), !missing(intensity)))
+##                   stop("'mz' and 'intensity' or none required.")
+##               if (!missing(mz) & !missing(intensity)) {
+##                   o <- base::order(mz, method = "radix")
+##                   .Object <- callNextMethod(.Object,
+##                                             ...,
+##                                             mz = mz[o],
+##                                             intensity = intensity[o],
+##                                             peaksCount = length(mz))
+##               } else .Object <- callNextMethod(.Object, ...)
+##               if (.Object@tic == 0)
+##                   .Object@tic <- sum(.Object@intensity)
+##               if (validObject(.Object))
+##                   .Object
+##           })
 
 setMethod("show", "Spectrum",
           function(object) {
-            if (msLevel(object)==1) show_Spectrum1(object)
-            else show_Spectrum2(object)
-            invisible(NULL)
+              if (msLevel(object) == 1) show_Spectrum1(object)
+              else show_Spectrum2(object)
+              invisible(NULL)
           })
 
-setMethod("plot",c("Spectrum","missing"),
-          function(x,y,...) {
-            if (msLevel(x)==1) plot_Spectrum1(x,...)
-            else plot_Spectrum2(x,...)
+setMethod("plot", c("Spectrum", "missing"),
+          function(x, y, ...) {
+              if (msLevel(x) == 1) plot_Spectrum1(x, ...)
+              else plot_Spectrum2(x, ...)
           })
 
 setMethod("plot", c("Spectrum", "Spectrum"),
           function(x, y, ...) {
-            plotSpectrumVsSpectrum(list(x, y), ...)
[…]

### 2016-08-30 R/methods-all.equal.R

@@ -1,0 +1,42 @@
+setMethod("all.equal", c("MSnExp", "MSnExp"),
+          function(target, current, ...) {
+              target@processingData@processing <-
+                  current@processingData@processing <- NA_character_
+              callNextMethod(target, current, ...)
+          })
+
+## The function below work for in memory and on disk MSnExp instances
+## and for spectra lists.
+equalMSnExps <- function(inmem, ondisk, ...) {
+    if ((l <- length(inmem)) != length(ondisk))
+        return("Objects have different lengths.")
+    if (inherits(ondisk, "OnDiskMSnExp"))
+        ondisk <- spectra(ondisk)
+    if (inherits(inmem, "OnDiskMSnExp"))
+        inmem <- spectra(inmem)
+    ## otherwise, expected to be a list of spectra
+    for (i in seq_len(l)) {
+        if (!isTRUE(speq <- all.equal(ondisk[[i]], inmem[[i]], ...)))
+            return(paste("Spectra at position", i, "are different."))
+    }
+    return(TRUE)
+}
+
+setMethod("all.equal", c("MSnExp", "OnDiskMSnExp"),
+          function(target, current, ...) equalMSnExps(target, current, ...))
+
+setMethod("all.equal", c("OnDiskMSnExp", "MSnExp"),
+          function(target, current, ...) equalMSnExps(target, current, ...))
+
+setMethod("all.equal", c("OnDiskMSnExp", "OnDiskMSnExp"),
+          function(target, current, ...) {
+              current@processingData@processing <-
+                  target@processingData@processing <- NA_character_
+              sp1 <- spectra(target)
+              sp2 <- spectra(current)
+              msg <- equalMSnExps(sp1, sp2, ...)
+              ## call super to test other slots
+              msg <- c(msg, callNextMethod(target, current))
+              if (is.logical(msg) && all(msg)) TRUE
+              else msg
+          })

### 2016-08-30 R/readMSData.R

@@ -1,137 +1,25 @@
-.testReadMSDataInput <- function(e) {
-    if (!all(e$msLevel > 0))
-        stop("msLevel must be an integer > 0.")
-    if (length(e$files) < 1)
-        stop("At least one MS file is required.")
-    if (all(unique(e$files) != e$files))
-        stop("Non unique files provided as input. ")
-    extensions <- unique(toupper(sub("^.+\\.", "", e$files)))
-    if (length(extensions) > 1)
-        warning(paste("Reading different file formats in.",
-                      "This is untested and you are welcome to try it out.",
-                      "Please report back!", sep="\n"))
-    invisible(TRUE)
-}
-
-## readMSData2 <- function(files,
-##                        pdata = NULL,
-##                        msLevel = 2,
-##                        verbose = TRUE,
-##                        centroided = FALSE,
-##                        smoothed = FALSE,
-##                        removePeaks = 0,
-##                        clean = FALSE,
-##                        cache = 2) {
-## ## - deal with multiple files
-##   if (removePeaks > 0)
-##     stop("Can't remove peaks when cache level is > 1")
-##   if (clean > 0)
-##     stop("Can't clean peaks when cache level is > 1")
-##   MSnbase:::.testReadMSDataInput(environment())
-##   if (msLevel == 1) ## cache currently only works for MS2 level data
-##     cache <- 0
-##   msLevel <- as.integer(msLevel)
-##   mzr <- lapply(files, mzR::openMSfile)
-##   .instrumentInfo <- lapply(mzr, instrumentInfo)
-##   names(.instrumentInfo) <- names(mzr) <- seq_along(mzr)
-##   fullhd <- do.call(rbind, lapply(mzr, header))
-##   ioncount <- unlist(lapply(mzr, function(xx)
-##                             sapply(peaks(xx), function(pk) sum(pk[, 2]))))
-##   .n <- sapply(mzr, length)
-##   featnms <- lapply(.n, function(n) paste0("X", 1:n))
-##   featnms <- unlist(lapply(seq_along(featnms),
-##                            function(x) {
-##                              paste(featnms[[x]], x, sep=".")
-##                            }))
-##   rownames(fullhd) <- featnms
-##   fromfile <- unlist(mapply(rep, 1:length(.n), .n, SIMPLIFY = FALSE)) ## to be saved to cache
-##   newhd <- data.frame(file = fromfile,
-##                       retention.time = fullhd$retentionTime,
[…]

### 2016-08-30 R/utils.R

@@ -529,7 +529,7 @@ nologging <- function(object, n = 1) {
 
 logging <- function(object, msg, date. = TRUE) {
   if (date.)
-    msg <- paste0(msg, ": ", date())
+    msg <- paste0(msg, " [", date(), "]")
   object@processingData@processing <-
     c(object@processingData@processing, msg)
   if (validObject(object))
@@ -672,57 +672,61 @@ utils.leftJoin <- function(x, y, by, by.x=by, by.y=by,
 utils.mergeSpectraAndIdentificationData <- function(featureData, id,
                                                     fcol = c("file", "acquisition.number"),
                                                     icol = c("file", "acquisitionnum")) {
-  # mzR::acquisitionNum (stored in fData()[, "acquisition.number"] and
-  # mzID::acquisitionnum should be identical
+    ## mzR::acquisitionNum (stored in fData()[, "acquisition.number"] and
+    ## mzID::acquisitionnum should be identical
+    if (!all(fcol %in% colnames(featureData))) {
+        stop("The column(s) ", sQuote(fcol),
+             " are not all in the feature data.frame!")
+    }
 
-  if (!all(fcol %in% colnames(featureData))) {
-      stop("The column(s) ", sQuote(fcol),
-           " are not all in the feature data.frame!")
-  }
+    if (!all(icol %in% colnames(id))) {
+        stop("The column(s) ", sQuote(icol),
+             " are not all in the identification data.frame!")
+    }
 
-  if (!all(icol %in% colnames(id))) {
-    stop("The column(s) ", sQuote(icol),
-         " are not all in the identification data.frame!")
-  }
+    if (sum(fcol %in% colnames(featureData)) != sum(icol %in% colnames(id))) {
+        stop("The number of selected column(s) in the feature and identification ",
+             "data don't match!")
+    }
 
-  if (sum(fcol %in% colnames(featureData)) != sum(icol %in% colnames(id))) {
-    stop("The number of selected column(s) in the feature and identification ",
-         "data don't match!")
-  }
+    ## sort id data to ensure the best matching peptide is on top in case of
+    ## multiple matching peptides
+    o <- do.call("order", lapply(c(icol, "rank"), function(j)id[, j]))
+    id <- id[o, ]
 
-  ## sort id data to ensure the best matching peptide is on top in case of
-  ## multiple matching peptides
[…]

### 2016-08-30 src/Spectrum2_class.c

@@ -1,0 +1,428 @@
+#include "MSnbase.h"
+
+/************************************************************
+ * The Spectrum2 constructor in C.
+ *
+ */
+
+/* This constructor uses functions defined in Rinternals.h */
+static SEXP _new_Spectrum2(SEXP msLevel, SEXP peaksCount, SEXP rt,
+			   SEXP acquisitionNum, SEXP scanIndex, SEXP tic,
+			   SEXP mz, SEXP intensity, SEXP fromFile,
+			   SEXP centroided, SEXP smoothed, SEXP polarity,
+			   SEXP merged, SEXP precScanNum, SEXP precursorMz,
+			   SEXP precursorIntensity, SEXP precursorCharge,
+			   SEXP collisionEnergy)
+{
+  SEXP classdef, ans;
+  if (asInteger(msLevel) < 2)
+    error("_new_Spectrum2: msLevel should be >= 2, but I got: %d!\n", asInteger(msLevel));
+
+  PROTECT(classdef = R_getClassDef("Spectrum2"));
+  PROTECT(ans = R_do_new_object(classdef));
+  /* Fill with data; should I do PROTECT here for all slots too???*/
+  R_do_slot_assign(ans, install("msLevel"), msLevel);
+  R_do_slot_assign(ans, install("msLevel"), msLevel);
+  R_do_slot_assign(ans, install("peaksCount"), peaksCount);
+  R_do_slot_assign(ans, install("rt"), rt);
+  R_do_slot_assign(ans, install("acquisitionNum"), acquisitionNum);
+  R_do_slot_assign(ans, install("scanIndex"), scanIndex);
+  R_do_slot_assign(ans, install("tic"), tic);
+  R_do_slot_assign(ans, install("mz"), mz);
+  R_do_slot_assign(ans, install("intensity"), intensity);
+  R_do_slot_assign(ans, install("fromFile"), fromFile);
+  R_do_slot_assign(ans, install("centroided"), centroided);
+  R_do_slot_assign(ans, install("smoothed"), smoothed);
+  R_do_slot_assign(ans, install("polarity"), polarity);
+  R_do_slot_assign(ans, install("merged"), merged);
+  R_do_slot_assign(ans, install("precScanNum"), precScanNum);
+  R_do_slot_assign(ans, install("precursorMz"), precursorMz);
[…]

### 2016-08-30 tests/testthat/test_Spectrum.R

@@ -7,166 +7,182 @@ test_that("Spectrum validity", {
 
 
 test_that("Spectrum processing", {
-  int <- c(0,1,2,3,1,0,0,0,0,1,3,10,6,2,1,0,1,2,0,0,1,5,10,5,1)
-  sp <- new("Spectrum2",
-            intensity=int,
-            mz=1:length(int))
-  ## removePeaks
-  defaultT <- min(intensity(sp)[intensity(sp)>0])
-  sp2a <- removePeaks(sp)
-  sp2b <- removePeaks(sp,defaultT)
-  sp2c <- removePeaks(sp,3)
-  expect_that(identical(sp2a,sp2b),is_true())
-  expect_that(identical(sp,sp2b),is_true())
-  expect_that(identical(sp,sp2c),is_false())
-  expect_that(ionCount(sp),equals(sum(int)))
-  expect_that(peaksCount(sp2c),equals(peaksCount(sp)))
-  expect_that(ionCount(sp),equals(55))
-  expect_that(ionCount(sp2c),equals(45))
-  ## clean
-  sp3 <- clean(sp)
-  expect_that(ionCount(sp),equals(ionCount(sp3)))
-  expect_that(peaksCount(sp),equals(length(int)))
-  expect_that(peaksCount(sp3),equals(23))
-  ##trimMz
-  sp4 <- trimMz(sp,c(10,20))
-  expect_that(intensity(sp4),equals(int[10:20]))
-  expect_that(mz(sp4),equals(10:20))
-  expect_that(peaksCount(sp4),equals(length(10:20)))
-  expect_that(ionCount(sp4),equals(sum(int[10:20])))
+    int <- c(0, 1, 2, 3, 1, 0, 0, 0, 0, 1, 3, 10, 6, 2, 1, 0, 1, 2, 0,
+             0, 1, 5, 10, 5, 1)
+    sp <- new("Spectrum2",
+              intensity = int,
+              mz = 1:length(int))
+    centroided(sp) <- FALSE
+    ## removePeaks
+    defaultT <- min(intensity(sp)[intensity(sp)>0])
+    sp2a <- removePeaks(sp)
+    sp2b <- removePeaks(sp, defaultT)
+    sp2c <- removePeaks(sp, 3)
+    expect_true(identical(sp2a, sp2b))
+    expect_true(identical(sp, sp2b))
+    expect_false(identical(sp, sp2c))
[…]

### 2016-08-30 tests/testthat/test_iPQF.R

@@ -15,8 +15,8 @@ test_that("data is valid", {
 test_that("iPQF unit test", {
               data("msnset2")
               fl <- list.files(system.file("extdata", package = "MSnbase"),
-                               full.names=TRUE, pattern = "ipqfres.rda")
-              load(fl)
+                               full.names = TRUE, pattern = "ipqfres.rda")
+              load(fl) ## ipqf1 to 5
               ##
               res1 <- combineFeatures(msnset2,
                                       groupBy = fData(msnset2)$accession,
@@ -26,6 +26,7 @@ test_that("iPQF unit test", {
                                       ratio.calc = "sum",
                                       method.combine = FALSE)
               res1@processingData <- processingData(ipqf1)
+              res1@.__classVersion__ <- ipqf1@.__classVersion__
               expect_true(all.equal(res1, ipqf1))
 
               dflt <- combineFeatures(msnset2,
@@ -33,6 +34,7 @@ test_that("iPQF unit test", {
                                       redundancy.handler = "unique",
                                       fun = "iPQF")
               dflt@processingData <- processingData(ipqf1)
+              res1@.__classVersion__ <- dflt@.__classVersion__
               expect_true(all.equal(res1, dflt))
               
 
@@ -44,6 +46,7 @@ test_that("iPQF unit test", {
                                       ratio.calc = "sum",
                                       method.combine = TRUE)
               res2@processingData <- processingData(ipqf2)
+              res2@.__classVersion__ <- ipqf2@.__classVersion__
               expect_true(all.equal(res2, ipqf2))
 
               res3 <- combineFeatures(msnset2,
@@ -54,6 +57,7 @@ test_that("iPQF unit test", {
                                       ratio.calc = "sum",
                                       method.combine = FALSE)
               res3@processingData <- processingData(ipqf3)
+              res3@.__classVersion__ <- ipqf3@.__classVersion__
               expect_true(all.equal(res3, ipqf3))
 
               res4 <- combineFeatures(msnset2,
@@ -64,6 +68,7 @@ test_that("iPQF unit test", {
                                       ratio.calc = "none",
                                       method.combine = FALSE)
               res4@processingData <- processingData(ipqf4)
+              res4@.__classVersion__ <- ipqf4@.__classVersion__
               expect_true(all.equal(res4, ipqf4))
 
               res5 <- combineFeatures(msnset2,
@@ -74,5 +79,6 @@ test_that("iPQF unit test", {
[…]

### 2016-08-30 R/functions-Spectrum1.R
Merge branch 'master' into devel
* master: (312 commits)
  close #153: update low/highMZ checks in validateOnDiskMSnSet
  suggest microbenchmark and suppress message in benchark vig
  Add unit test for issue #118
  Fix failing tests and example due to r120413 in Biobase
  Fix for loop that iterates over spectra to work with empty objects
  Spectra import without need to read mzML header
  Intermediate fix for issue #151
  add unit test for extractMgfSpectrum2Info
  fix warning in readMgfData if TITLE contains multiple =
  split MSnExp tests and new [[ error test
  fix OnDiskMSnExp [[ subsettin

@@ -5,3 +5,250 @@ show_Spectrum1 <- function(spectrum) {
   cat(" Total ion count:", spectrum@peaksCount, "\n")
   cat(" Polarity:", spectrum@polarity, "\n")
 }
+
+############################################################
+## Constructor function for Spectrum1 objects. This one uses C-code and
+## is faster than a call to "new"
+Spectrum1 <- function(peaksCount = length(mz), rt = numeric(),
+                      acquisitionNum = NA_integer_, scanIndex = integer(),
+                      tic = 0, mz = numeric(), intensity = numeric(),
+                      fromFile = integer(), centroided = NA,
+                      smoothed = NA,
+                      polarity = NA_integer_){
+    if (tic == 0)
+        tic <- sum(intensity)
+    if (!is.integer(peaksCount)) peaksCount <- as.integer(peaksCount)
+    if (!is.double(rt)) rt <- as.double(rt)
+    if (!is.integer(acquisitionNum)) acquisitionNum <- as.integer(acquisitionNum)
+    if (!is.integer(scanIndex)) scanIndex <- as.integer(scanIndex)
+    if (!is.double(tic)) tic <- as.double(tic)
+    if (!is.double(mz)) mz <- as.double(mz)
+    if (!is.double(intensity)) intensity <- as.double(intensity)
+    if (!is.integer(fromFile)) fromFile <- as.integer(fromFile)
+    if (!is.logical(centroided)) centroided <- as.logical(centroided)
+    if (!is.logical(smoothed)) smoothed <- as.logical(smoothed)
+    if (!is.integer(polarity)) polarity <- as.integer(polarity)
+    res <- .Call("Spectrum1_constructor",
+                 1L, peaksCount, rt, acquisitionNum, scanIndex, tic, mz,
+                 intensity, fromFile, centroided, smoothed, polarity, TRUE,
+                 PACKAGE = "MSnbase")
+    return(res)
+}
+
+############################################################
+## Constructor for Spectrum1 that ensures ordering of M/Z-intensity
+## pairs by M/Z value (increasing).
+Spectrum1_mz_sorted <- function(peaksCount = length(mz), rt = numeric(),
+                                acquisitionNum = NA_integer_,
+                                scanIndex = integer(), tic = 0,
+                                mz = numeric(), intensity = numeric(),
+                                fromFile = integer(),
[…]

### 2016-08-30 R/methods-filters.R

@@ -1,0 +1,156 @@
+## TODO: implement filter[Mz|Rt|MsLevel|File|filterAcquisitionNum] for
+## MSnExp and OnDiskMSnExp classes
+
+setMethod("filterMsLevel", "MSnExp",
+          function(object, msLevel.) {
+              if (missing(msLevel.)) return(object)
+              msLevel. <- as.numeric(msLevel.)
+              object <- object[msLevel(object) %in% msLevel.]
+              object <- logging(object,
+                                paste("Filter: select MS level(s)",
+                                      paste(unique(msLevel.),
+                                            collapse = " ")))
+              object
+          })
+
+
+setMethod("filterRt", "MSnExp",
+          function(object, rt, msLevel.) {
+              if (missing(rt)) return(object)
+              if (length(rt) != 2 | !is.numeric(rt))
+                  stop("'rt' must be a numeric of length 2")
+              if (missing(msLevel.))
+                  msLevel. <- base::sort(unique(msLevel(object)))
+              msLevel. <- as.numeric(msLevel.)
+              ## MS levels that need to be filtered out
+              selms <- msLevel(object) %in% msLevel.
+              ## keep spectra that match MS levels and that are within
+              ## the retention time limits
+              sel1 <- rtime(object) >= rt[1] & rtime(object) <= rt[2] & selms
+              ## keep all spectra that don't match MS levels
+              sel2 <- !selms
+              object <- object[sel1 | sel2]
+              fltmsg <- paste0("Filter: select retention time [",
+                               paste0(rt, collapse = "-"),
+                               "] and MS level(s), ",
+                               paste(unique(msLevel.),
+                                     collapse = " "))
+              object <- logging(object, fltmsg)
+              object
+          })
+
+############################################################
+## filterMz for MSnExp
+setMethod("filterMz", "MSnExp",
+          function(object, mz, msLevel., ...) {
+              if (missing(mz))
+                  return(object)
+              if (!is.numeric(mz) & length(mz) != 2)
+                  stop("'mz' must be a numeric of length 2!")
+              if (missing(msLevel.)) {
+                  msLevel. <- base::sort(unique(msLevel(object)))
+              } else {
+                  if (!is.numeric(msLevel.))
+                      stop("'msLevel' must be numeric!")
+              }
+              ## Note: the msLevel. argument is passed down to the
+              ## filterMz_Spectrum function.
+              filtered <- eapply(assayData(object), filterMz, mz = mz,
[…]

### 2016-08-30 tests/testthat/test_readMSData2.R

@@ -1,0 +1,68 @@
+context("readMSData2")
+
+test_that("msLevel set correctly", {
+    f <- msdata::proteomics(full.names = TRUE)
+    x <- readMSData2(f, verbose = FALSE)
+    expect_equivalent(centroided(x), rep(NA, length(x)))
+    x <- readMSData2(f, centroided = TRUE, verbose = FALSE)
+    expect_equivalent(centroided(x), rep(TRUE, length(x)))
+    x <- readMSData2(f, centroided = FALSE, verbose = FALSE)
+    expect_equivalent(centroided(x), rep(FALSE, length(x)))
+    x <- readMSData2(f, centroided = c(FALSE, TRUE), verbose = FALSE)
+    expect_true(all(centroided(filterMsLevel(x, msLevel = 2))))
+    expect_true(all(!centroided(filterMsLevel(x, msLevel = 1))))
+    x2 <- readMSData2(f, centroided = c(FALSE, TRUE, NA, NA), verbose = FALSE)
+    expect_identical(centroided(x), centroided(x2))
+})
+
+############################################################
+## Load the required data files.
+.getMzMLFiles <- function(force.msdata = FALSE) {
+    ## Return the mzML files, the ones from the XXX package, or if run
+    ## locally, some of my test files.
+    HOST <- unlist(strsplit(system("hostname", intern = TRUE), split = ".",
+                            perl = FALSE, fixed = TRUE))[1]
+    if (HOST == "macbookjo" & !force.msdata) {
+        mzfiles <- dir("/Users/jo/R-workspaces/EURAC/2016/2016-04-21-PolarMetabolom/data/mzML/",
+                       pattern = "POS_C_O", full.names = TRUE)
+    } else {
+        require(msdata)
+        mzfiles <- c(system.file("microtofq/MM14.mzML", package = "msdata"),
+                     system.file("microtofq/MM8.mzML", package = "msdata"))
+    }
+    return(mzfiles)
+}
+
+mzf <- .getMzMLFiles(TRUE)[1:2]
+
+## Load the data with readMSData2
+odmse <- readMSData2(files = mzf, centroided = TRUE)
+
+test_that("Constructor performance and test for MS1 only", {
+    featDat <- fData(odmse)
+    featDat <- featDat[featDat$fileIdx == 1, ]
+    ## system.time(
[…]

### 2016-08-30 README.Rmd
remove pRoloc and pRolocdata, as travis complains MSnbase is not available for them

```{r env, echo=FALSE}
opts_chunk$set(comment=NA, echo=FALSE)
```

`MSnbase` is an R/Bioconductor package that provides infrastructure
for plotting, data manipulation and processing of MS-based proteomics
data. The project was started by
[Laurent Gatto](http://lgatto.github.io/) in October 2010 (Mon Oct 4
23:35:23 2010, according to the git log) and has, since then,
benefited from various contributions.

```{r}
auth <- packageDescription("MSnbase")$Author
ctrb <- sub("<lg390@cam.ac.uk> ", "",
            gsub("\t", " ", gsub("\n", " ", auth)))
cat(gsub("  ", " ", ctrb), "\n")
```

The official page is the Bioconductor landing page
([release](http://www.bioconductor.org/packages/release/bioc/html/MSnbase.html)
or
[devel](http://www.bioconductor.org/packages/devel/bioc/html/MSnbase.html)
versions). The [github page](https://github.com/lgatto/MSnbase) page
is for sharing, testing, issue tracking and forking/pulling
purposes. The
[Bioconductor svn server](https://hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/),
is the official repository for the working source code.

To cite the MSnbase package in publications, please use:

> Gatto L, Lilley KS. *`MSnbase` - an R/Bioconductor package for
> isobaric tagged mass spectrometry data visualization, processing and
> quantitation*. Bioinformatics. 2012 Jan
> 15;28(2):288-9. doi:[10.1093/bioinformatics/btr645](http://bioinformatics.oxfordjournals.org/content/28/2/288.long). Epub
> 2011 Nov 22.  PubMed
> [PMID:22113085](http://www.ncbi.nlm.nih.gov/pubmed/22113085).

### 2016-08-30 README.Rmd
improve readme

@@
 
+## Installation
+
+To install `MSnbase`
+
+```{r install, eval = FALSE}
+library("BiocInstaller")
+biocLite("MSnbase")
+```
+
+If you need the github version
+
+```{r installgh, eval = FALSE}
+biocLite("lgato/MSnbase")
+```
+
+## Questions
+
+General questions should be asked on the
+[Bioconductor support forum](https://support.bioconductor.org/), using
+`MSnbase` to tag the question.
+
+## Citation
+
 To cite the MSnbase package in publications, please use:

### 2016-08-30 README.Rmd
fix readme

@@
 ```{r env, echo=FALSE}
-opts_chunk$set(comment=NA, echo=FALSE)
+opts_chunk$set(comment=NA)
 ```
@@
 
-```{r}
+```{r, echo=FALSE}
 auth <- packageDescription("MSnbase")$Author
-ctrb <- sub("<lg390@cam.ac.uk> ", "",
-            gsub("\t", " ", gsub("\n", " ", auth)))
-cat(gsub("  ", " ", ctrb), "\n")
+auth <- sub("<lg390@cam.ac.uk> ", "", auth)
+auth <- gsub("\n\t", "\n", auth)
+auth <- gsub("\t", " ", auth)
+cat(auth, "\n")
 ```

### 2016-08-31 README.Rmd
udpate readme with build status

@@
 ```{r env, echo=FALSE}
-opts_chunk$set(comment=NA)
+knitr::opts_chunk$set(comment=NA)
 ```
@@
 ```{r installgh, eval = FALSE}
-biocLite("lgato/MSnbase")
+biocLite("lgatto/MSnbase")
 ```
+
+## Build status
+
+This repository, `master` branch
+[![Build Status](https://travis-ci.org/lgatto/MSnbase.svg?branch=master)](https://travis-ci.org/lgatto/MSnbase)
 
@@
 > [PMID:22113085](http://www.ncbi.nlm.nih.gov/pubmed/22113085).
+
+## Contributing
+
+Contributions to the package are more than welcome. If you want to
+contribute to this package, you should follow the same conventions as
+the rest of the functions whenever it makes sense to do so. Feel free
+to get in touch (preferable opening a
+[github issue](https://github.com/lgatto/MSnbase/issues/)) to discuss
+any suggestions. The
+[`MSnbase` development vignette](http://bioconductor.org/packages/devel/bioc/vignettes/MSnbase/inst/doc/MSnbase-development.pdf)
+gives some background on the class infrastructure.
+
+Please note that this project is released with a
+[Contributor Code of Conduct](https://github.com/lgatto/MSnbase/blob/master/CONDUCT.md). By
+participating in this project you agree to abide by its terms.

### 2016-09-01 README.Rmd
add codecov shield

@@
 
-This repository, `master` branch
-[![Build Status](https://travis-ci.org/lgatto/MSnbase.svg?branch=master)](https://travis-ci.org/lgatto/MSnbase)
+[![Build Status](https://travis-ci.org/lgatto/MSnbase.svg?branch=master)](https://travis-ci.org/lgatto/MSnbase) [![codecov.io](https://codecov.io/github/lgatto/MSnbase/coverage.svg?branch=master)](https://codecov.io/github/lgatto/MSnbase?branch=master)
 

### 2016-09-30 R/DataClasses.R
Merge branch 'master' into devel
* master:
  Added isolationWindow,MSnExp method
  MS3 support
  simplify SSD vs HDD comparison
  Update the benchmarking vignette.
  Add comparison SSD vs HDD to benchmarking vignette
  ignore makefile and travis when building
  new gh devel version

From: Laurent <lg390@cam.ac.uk>

git-svn-id: file:///home/git/hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/MSnbase@121600 bc3139a8-67e5-0310-9ffc-ced21a209358

@@ -248,9 +248,12 @@ setClass("Spectrum2",
          validity = function(object) {
              msg <- validMsg(NULL, NULL)
              msl <- object@msLevel
-             if (msl!=as.integer(2))
-                 msg <- validMsg(msg,paste("Object of class",class(object),
-                                           "but msLevel is",msl,sep=" "))
+             if (msl < as.integer(2))
+                 msg <- validMsg(msg,
+                                 paste0("Object of class ",
+                                        class(object),
+                                        " but msLevel is ", msl,
+                                        " (should be > 1)"))
              if (is.null(msg)) TRUE
              else msg
          })

### 2016-12-18 docs/articles/Foreword.tex
add pkgdown docs and swap Seb's given/familiy name

--- 2012-02-03 inst/doc/Foreword.tex
@@
 \end{quote}
-
-You are welcome to contact me for questions, bugs, typos or suggestions about \Rpackage{MSnbase}. 
-If you wish to reach a broader audience for general questions about proteomics analysis using \texttt{R}, 
-you may want to use the Bioconductor mailing 
-list\footnote{\url{https://stat.ethz.ch/mailman/listinfo/bioconductor}}.
-%% proteomics special interest 
-%% group\footnote{\url{https://stat.ethz.ch/mailman/listinfo/bioc-sig-proteomics}}.

### 2016-12-18 docs/articles/MSnbase-demo.Rnw
add pkgdown docs and swap Seb's given/familiy name

%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{Base Functions and Classes for MS-based Proteomics}
%\VignetteKeywords{Mass Spectrometry, MS, MSMS, Proteomics, Infrastructure, Bioinformatics, quantitative }
%\VignettePackage{MSnbase-demo}

\documentclass[12pt, oneside]{article}

<<style, eval=TRUE, echo=FALSE, results="asis">>=
BiocStyle::latex()
@

\bioctitle[\Biocpkg{MSnbase} introduction]{
  \Biocpkg{MSnbase}: labelled and label-free MS2 data pre-processing,
  visualisation and quantification.
}

\author{
  Laurent Gatto\footnote{\email{lg390@cam.ac.uk}} and Sebastian Gibb
}

\begin{document}

\maketitle

%% Abstract and keywords %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vskip 0.3in minus 0.1in
\hrule
\begin{abstract}
  This vignette describes the functionality implemented in the
  \Biocpkg{MSnbase} package.  \Biocpkg{MSnbase} aims at (1)
  facilitating the import, processing, visualisation and
  quantification of mass spectrometry data into the \R environment
  \cite{Rstat} by providing specific data classes and methods and (2)
  enabling the utilisation of throughput-high data analysis pipelines
  provided by the Bioconductor \cite{Gentleman2004} project.
\end{abstract}
\textit{Keywords}: Mass Spectrometry (MS), proteomics, infrastructure, quantitative.
\vskip 0.1in minus 0.05in
\hrule
\vskip 0.2in minus 0.1in
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newpage

\tableofcontents

<<environment, cache=FALSE, echo=FALSE>>=
suppressPackageStartupMessages(library("ggplot2"))
suppressPackageStartupMessages(library("MSnbase"))
suppressPackageStartupMessages(library("zoo"))
suppressPackageStartupMessages(require("Rdisop"))
suppressPackageStartupMessages(require("pRolocdata"))
suppressPackageStartupMessages(require("pRoloc"))
library("grid")
library("reshape2")
suppressPackageStartupMessages(library("BiocParallel"))
@

\newpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{Foreword.tex}

\input{Bugs.tex}

\newpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}\label{sec:intro}

\Biocpkg{MSnbase} \cite{Gatto2012} aims are providing a reproducible
research framework to proteomics data analysis.  It should allow
researcher to easily mine mass spectrometry data, explore the data and
its statistical properties and visually display these.

\Biocpkg{MSnbase} also aims at being compatible with the
infrastructure implemented in Bioconductor, in particular
\Biocpkg{Biobase}. As such, classes developed specifically for
proteomics mass spectrometry data are based on the \Rclass{eSet} and
\Rclass{ExpressionSet} classes. The main goal is to assure seamless
compatibility with existing meta data structure, accessor methods and
normalisation techniques.

This vignette illustrates \Biocpkg{MSnbase} utility using a dummy
data sets provided with the package without describing the underlying
data structures. More details can be found in the package, classes,
method and function documentations. A description of the classes is
provided in the \texttt{MSnbase-development} vignette.

\input{NoteAboutSpeedAndMemory.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Data structure and content}\label{sec:data}

\subsection{Importing experiments}\label{sec:io}

\Biocpkg{MSnbase} is able to import raw MS data stored in one of the
\texttt{XML}-based formats as well as peak lists in the \texttt{mfg}
format\footnote{%%
  Mascot Generic Format --
  \url{http://www.matrixscience.com/help/data\_file\_help.html\#GEN}}

\paragraph{Raw data} The \texttt{XML}-based formats, \texttt{mzXML}
\cite{Pedrioli2004}, \texttt{mzData} \cite{Orchard2007} and
\texttt{mzML} \cite{Martens2010} can be imported with the
\Rfunction{readMSData} function, as illustrated below (see
\Rfunction{?readMSData} for more details).

<<readdata, echo=TRUE, cache=FALSE, tidy=FALSE>>=
file <- dir(system.file(package = "MSnbase", dir = "extdata"),
            full.names = TRUE, pattern = "mzXML$")
rawdata <- readMSData(file, msLevel = 2, verbose = FALSE)
@
%% $

Only spectra of a give MS level can be loaded at a time by setting the
\texttt{msLevel} parameter accordingly.  In this document, we will use
the \Robject{itraqdata} data set, provided with \Biocpkg{MSnbase}. It
includes feature metadata, accessible with the \Rfunction{fData}
accessor.  The metadata includes identification data for the
\Sexpr{length(itraqdata)} MS2 spectra.

\paragraph{MSnbase 2.0 } The new major version of \Biocpkg{MSnbase}
uses a new \textit{on-disk} data storage model (see the
\textit{benchmarking} vignette for more details). The new data backend
is compatible with the orignal \textit{in-memory} model. To make use
of the new infrastructure, read your raw data using
\Rfunction{readMSData2}, rather than \Rfunction{readMSData}. All
existing operations work irrespective of the backend.

\paragraph{Peak lists} Peak lists can often be exported after spectrum
processing from vendor-specific software and are also used as input to
search engines.  Peak lists in \texttt{mgf} format can be imported
with the function \Rfunction{readMgfData} (see
\Rfunction{?readMgfData} for details) to create experiment objects.
Experiments or individual spectra can be exported to an \texttt{mgf}
file with the \Rfunction{writeMgfData} methods (see
\Rfunction{?writeMgfData} for details and examples).

\bigskip

\paragraph{Experiments with multiple runs} Although it is possible to
load and process multiple files serially and later merge the resulting
quantitation data as show in section \ref{sec:combine} (page
\pageref{sec:combine}), it is also feasible to load several raw data
files at once.  Here, we report the analysis of an LC-MSMS experiment
were 14 liquid chromatography (LC) fractions were loaded using
\Rfunction{readMSData} on a 32-cores servers with 128 Gb of RAM.  It
took about 90 minutes to read the 14 uncentroided \texttt{mzXML} raw
files (4.9 Gb on disk in total) and create a 3.3 Gb raw data object
(an \Rclass{MSnExp} instance, see next section).  Quantitation of 9
reporter ions (\Rclass{iTRAQ9} object, see \ref{sec:reporterions})
for 88690 features was performed in parallel on 16 processors and took
76 minutes.  The resulting quantitation data was only 22.1 Mb and
could easily be further processed and analysed on a standard laptop
computer.

Since verions 1.13.5, parallel support is provided by the
\Biocpkg{BiocParallel} and various backends including multicore
(forking), simple networf network of workstations (SNOW) using
sockets, forking or MPI among others.

\bigskip

See also section \ref{sec:io2} to import quantitative data stored in
spreadsheets into \R for further processing using
\Biocpkg{MSnbase}. The \texttt{MSnbase-io} vignette gives a general
overview of \Biocpkg{MSnbase}'s input/ouput capabilites.

\subsection{MS experiments}\label{sec:msnexp}

Raw data is contained in \Rclass{MSnExp} objects, that stores all the
spectra of an experiment, as defined by one or multiple raw data
files.

<<MSnExp, cache=FALSE, echo=TRUE>>=
library("MSnbase")
itraqdata
head(fData(itraqdata))
@

<<experimentsize, echo=FALSE, cache=FALSE>>=
sz <- sum(sapply(assayData(itraqdata), object.size)) +
  object.size(itraqdata)
sz <- as.numeric(sz)
sz <- round(sz/(1024^2), 2)
@

As illustrated above, showing the experiment textually displays it's
content:
\begin{itemize}
\item Information about the raw data, i.e. the spectra.
\item Specific information about the experiment processing\footnote{%%
    this part will be automatically updated when the object is
    modified with it's \textit{ad hoc} methods, as illustrated later}
  and package version.  This slot can be accessed with the
  \Rfunction{processingData} method.
\item Other meta data, including experimental phenotype, file name(s)
  used to import the data, protocol data, information about features
  (individual spectra here) and experiment data. Most of these are
  implemented as in the \Rclass{eSet} class and are described in more
  details in their respective manual pages. See \Rfunction{?MSnExp}
  and references therein for additional background information.

  The experiment meta data associated with an \Rclass{MSnExp}
  experiment is of class \Rclass{MIAPE}. It stores general
  information about the experiment as well as MIAPE (Minimum
  Information About a Proteomics Experiment) information
  \cite{Taylor2007, Taylor2008}.  This meta-data can be accessed with
  the \Rfunction{experimentData} method.  When available, a summary of
  MIAPE-MS data can be printed with the \Rfunction{msInfo} method.
  See \Rfunction{?MIAPE} for more details.
\end{itemize}

\subsection{Spectra objects}\label{sec:spectra}

The raw data is composed of the \Sexpr{length(itraqdata)} MS spectra.
The spectra are named individually
(\Sexpr{paste(paste(head(featureNames(itraqdata)),collapse=", "),",
  ...",sep="")}) and stored in a \Robject{environment}. They can be
accessed individually with \texttt{itraqdata[["X1"]]} or
\texttt{itraqdata[[1]]}, or as a list with
\texttt{spectra(itraqdata)}.  As we have loaded our experiment
specifying \texttt{msLevel=2}, the spectra will all be of level 2 (or
higher, if available).

<<Spectrum, cache=FALSE, echo=TRUE>>=
sp <- itraqdata[["X1"]]
sp
@

Attributes of individual spectra or of all spectra of an experiment
can be accessed with their respective methods:
\Rfunction{precursorCharge} for the precursor charge,
\Rfunction{rtime} for the retention time, \Rfunction{mz} for the MZ
values, \Rfunction{intensity} for the intensities, ... see the
\Rclass{Spectrum}, \Rclass{Spectrum1} and \Rclass{Spectrum2}
manuals for more details.

<<accessors, cache=FALSE, echo=TRUE>>=
peaksCount(sp)
head(peaksCount(itraqdata))
rtime(sp)
head(rtime(itraqdata))
@

\subsection{Reporter ions}\label{sec:reporterions}

Reporter ions are defined with the \Rclass{ReporterIons} class.
Specific peaks of interest are defined by a MZ value, a with around
the expected MZ and a name (and optionally a colour for plotting, see
section \ref{sec:plotting}). \Rclass{ReporterIons} instances are
required to quantify reporter peaks in \Rclass{MSnExp}
experiments. Instances for the most commonly used isobaric tags like
iTRAQ 4-plex and 8-plex and TMT 6- and 10-plex tags are already
defined in \Biocpkg{MSnbase}. See \Rfunction{?ReporterIons} for
details about how to generate new \Rclass{ReporterIons} objects.

<<ReporterIons>>=
iTRAQ4
TMT10
@

\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Plotting raw data}\label{sec:plotting}

\subsection{MS data space}\label{sec:msmaps}

The \Rclass{MSmap} class can be used to isolate specific slices of
interest from a complete MS acquisition by specifying $m/z$ and
retention time ranges. One needs a raw data file in a format supported
by \Biocpkg{mzR}'s \Rfunction{openMSfile} (\texttt{mzML},
\texttt{mzXML}, ...). Below we first download a raw data file from the
PRIDE repository and create\footnote{This code chunk is not evaluated
  to avoid repeated downloaded of the raw data file. The \Robject{M}
  map is provided with the package and loaded to evaluate subsequent
  code chunks.} an \Rclass{MSmap} containing all the MS$^1$ spectra
between acquired between 30 and 35 minutes and peaks between 521 and
523 $m/z$. See \Rfunction{?MSmap} for details.

<<msmap, eval=FALSE>>=
## downloads the data
library("rpx")
px1 <- PXDataset("PXD000001")
mzf <- pxget(px1, 6)

## reads the data 
ms <- openMSfile(mzf)
hd <- header(ms)

## a set of spectra of interest: MS1 spectra eluted
## between 30 and 35 minutes retention time
ms1 <- which(hd$msLevel == 1)
rtsel <- hd$retentionTime[ms1] / 60 > 30 &
    hd$retentionTime[ms1] / 60 < 35

## the map
M <- MSmap(ms, ms1[rtsel], 521, 523, .005, hd, zeroIsNA = TRUE)
@

<<msmaplaod, echo=FALSE>>=
mrda <- dir(system.file(package = "MSnbase", dir = "extdata"),
            full.names = TRUE, pattern = "M.rda$")
mrda2 <- dir(system.file(package = "MSnbase", dir = "extdata"),
            full.names = TRUE, pattern = "M2.rda$")
load(mrda)
load(mrda2)
@

<<msmaphow>>=
M
@

The \Robject{M} map object can be rendered as a heatmap with
\Rfunction{plot}, as shown on figure \ref{fig:mapheat}.

\begin{figure}[!ht]
\centering
<<mapheat, dev='pdf', fig.width=5, fig.height=5>>=
plot(M, aspect = 1, allTicks = FALSE)
@
\caption{Heat map of a chunk of the MS data.}
\label{fig:mapheat}
\end{figure}

One can also render the data in 3 dimension with the
\Rfunction{plot3D} function, as show on figure \ref{fig:map3d}.

\begin{figure}[!ht]
\centering
<<map3d1, dev='pdf', fig.width=4, fig.height=4>>=
plot3D(M)
@
\caption{3 dimensional represention of MS map data.}
\label{fig:map3d}
\end{figure}

To produce figure \ref{fig:map3d2}, we create a second \Rclass{MSmap}
object containing the first two MS$^1$ spectra of the first map
(object \Robject{M} above) and all intermediate MS$^2$ spectra and
display $m/z$ values between 100 and 1000.

<<msmap2, eval=FALSE>>=
i <- ms1[which(rtsel)][1]
j <- ms1[which(rtsel)][2]
M2 <- MSmap(ms, i:j, 100, 1000, 1, hd)
@

<<m2>>=
M2
@

\begin{figure}[!ht]
\centering
<<map3d2, dev='pdf', fig.width=4, fig.height=4>>=
plot3D(M2)
@
\caption{3 dimensional represention of MS map data. MS$^1$ and MS$^2$
  spectra are coloured in blue and magenta respectively.}
\label{fig:map3d2}
\end{figure}

\subsection{MS Spectra}\label{sec:specplots}

Spectra can be plotted individually or as part of (subset) experiments
with the \Rfunction{plot} method. Full spectra can be plotted (using
\texttt{full=TRUE}), specific reporter ions of interest (by specifying
with reporters with \texttt{reporters=iTRAQ4} for instance) or both
(see figure \ref{fig:spectrum-plot}).

\begin{figure}[!ht]
\centering
<<spectrumPlot, dev='pdf', fig.width=6, fig.height=4, fig.keep='high'>>=
plot(sp, reporters = iTRAQ4, full = TRUE)
@
\caption{Raw MS2 spectrum with details about reporter ions.}
\label{fig:spectrum-plot}
\end{figure}

<<bsaSelect, eval=TRUE, echo=FALSE>>=
## bsasel <- fData(itraqdata)$ProteinAccession == "BSA"
bsasel <- 1:3
@
%% $

It is also possible to plot all spectra of an experiment (figure
\ref{fig:msnexp-plot}).  Lets start by subsetting the
\Robject{itraqdata} experiment using the protein accession numbers
included in the feature metadata, and keep the \Sexpr{sum(bsasel)}
from the \textit{BSA} protein.

<<subset, echo=TRUE>>=
sel <- fData(itraqdata)$ProteinAccession == "BSA"
bsa <- itraqdata[sel]
bsa
as.character(fData(bsa)$ProteinAccession)
@

%% Lets start by extracting all spectra that have the same precursor MZ value than \Robject{sp}
%% into a separate experiment, using the \Rfunction{extractPrecSpectra} method.
%% This method takes an \Robject{MSnExp} experiment and a precursor MZ value as parameters.
%% <<extractPrecSpec,echo=T,cache=F>>=
%% exp2 <- extractPrecSpectra(raw.experiment,precursorMz(sp))
%% exp2
%% @

These can then be visualised together by plotting the
\Rclass{MSnExp} object, as illustrated on figure \ref{fig:msnexp-plot}.

\begin{figure}[p]
\centering
<<msnexpPlot, dev='pdf', fig.width=3.7, fig.height=4.5, fig.keep='last'>>=
plot(bsa, reporters = iTRAQ4, full = FALSE) + theme_gray(8)
@
\caption{Experiment-wide raw MS2 spectra. The y-axes of the individual
  spectra are automatically rescaled to the same range.
  See section \ref{sec:norm} to rescale peaks identically. }
\label{fig:msnexp-plot}
\end{figure}

\paragraph{Customising your plots}\label{sec:customplots}

The \Biocpkg{MSnbase} \Rfunction{plot} methods have a logical
\Rfunction{plot} parameter (default is \Robject{TRUE}), that specifies
if the plot should be printed to the current device. A plot object is
also (invisibly) returned, so that it can be saved as a variable for
later use or for customisation.

\Biocpkg{MSnbase} uses the \CRANpkg{ggplot2} package to generate
plots, which can subsequently easily be customised.  More details
about \CRANpkg{ggplot2} can be found in \cite{ggplot2} (especially
chapter 8) and on
\href{http://had.co.nz/ggplot2/}{http://had.co.nz/ggplot2/}.  Finally,
if a plot object has been saved in a variable \Robject{p}, it is
possible to obtain a summary of the object with
\Rfunction{summary(p)}.  To view the data frame used to generate the
plot, use \Rfunction{p@data}.

\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Tandem MS identification data}\label{sec:id}

\subsection{Adding identification data}

\Biocpkg{MSnbase} is able to integrate identification data from
\texttt{mzIdentML} \cite{Jones2012} files.

We first load two example files shipped with the \Biocpkg{MSnbase}
containing raw data (as above) and the corresponding identification
results respectively. The raw data is read with the
\Rfunction{readMSData}, as demonstrated above. As can be seen, the
default feature data only contain spectra numbers\footnote{More data
  about the spectra is of course available in an \Rclass{MSnExp}
  object, as illustrated in the previous sections. See also
  \Rfunction{?pSet} and \Rfunction{?MSnExp} for more details.}.

<<msnexpIdentification, echo=TRUE, cache=FALSE, tidy=FALSE>>=
## find path to a mzXML file
quantFile <- dir(system.file(package = "MSnbase", dir = "extdata"),
                 full.name = TRUE, pattern = "mzXML$")
## find path to a mzIdentML file
identFile <- dir(system.file(package = "MSnbase", dir = "extdata"),
                 full.name = TRUE, pattern = "dummyiTRAQ.mzid")
## create basic MSnExp
msexp <- readMSData(quantFile, verbose = FALSE)
head(fData(msexp), n = 2)
@

The \Rfunction{addIdentificationData} method takes an \Rclass{MSnExp}
instance (or an \Rclass{MSnSet} instance storing quantitation data,
see section \ref{sec:quant}) as first argument and one or multiple
\texttt{mzIdentML} file names (as a character vector) as second one and
updates the \Rclass{MSnExp} feature data using the identification
data read from the \texttt{mzIdentML} file(s).

<<msnexpIdentification2, echo=TRUE, cache=FALSE, tidy=FALSE>>=
## add identification information
msexp <- addIdentificationData(msexp, id = identFile,
                               verbose = FALSE)
head(fData(msexp), n = 2)
@

Finally we can use \Rfunction{idSummary} to summarise the percentage
of identified features per quantitation/identification pairs.

<<msnexpIdentification3, echo=TRUE, cache=FALSE, tidy=FALSE>>=
idSummary(msexp)
@

When identification data is present, and hence peptide sequences, one
can annotation fragment peaks on the MS2 figure by passing the peptide
sequence to the \Rfunction{plot} method.

<<fragplot>>=
itraqdata2 <- pickPeaks(itraqdata, verbose=FALSE)
i <- 14
s <- as.character(fData(itraqdata2)[i, "PeptideSequence"])
@

\begin{figure}[!ht]
\centering
<<fragplotfig, dev='pdf', fig.width=5, fig.height=5>>=
plot(itraqdata2[[i]], s, main = s)
@
\caption{Annotated MS2 spectrum.}
\label{fig:fragplot}
\end{figure}

The fragment ions are calculated with the
\Rfunction{calculateFragments}, described in section
\ref{sec:calcfrag} on page \pageref{calcfrag}.

\subsection{Filtering identification data}

One can remove the features that have not been identified using
\Rfunction{removeNoId}. This function uses by default the
\Robject{pepseq} feature variable to search the presence of missing data
(\Robject{NA} values) and then filter these non-identified spectra.

<<msnexpIdentification4, echo=TRUE, cache=FALSE, tidy=FALSE>>=
fData(msexp)$pepseq
msexp <- removeNoId(msexp)
fData(msexp)$pepseq
idSummary(msexp)
@

Similarly, the \Rfunction{removeMultipleAssignment} method can be used
to filter out non-unique features, i.e. that have been assigned to
protein groups with more than one member. This function uses by
default the \Robject{nprot} feature variable.

\bigskip

Note that \Rfunction{removeNoId} and
\Rfunction{removeMultipleAssignment} methods can also be called on
\Rclass{MSnExp} instances.

\subsection{Calculate Fragments}\label{sec:calcfrag}

\Biocpkg{MSnbase} is able to calculate theoretical peptide fragments via
\Rfunction{calculateFragments}.

<<calculateFragments, echo=TRUE, cache=FALSE, tidy=FALSE>>=
calculateFragments("ACEK",
                   type = c("a", "b", "c", "x", "y", "z"))
@

It is also possible to match these fragments against an \Rclass{Spectrum2}
object.
<<msnexpcalculateFragments, echo=TRUE, cache=FALSE, tidy=FALSE>>=
pepseq <- fData(msexp)$pepseq[1]
calculateFragments(pepseq, msexp[[1]], type=c("b", "y"))
@

\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Quality control}\label{sec:qc}

The current section is not executed dynamically for package size and
processing time constrains.  The figures and tables have been
generated with the respective methods and included statically in the
vignette for illustration purposes.

\bigskip

\Biocpkg{MSnbase} allows easy and flexible access to the data, which
allows to visualise data features to assess it's quality. Some methods
are readily available, although many QC approaches will be experiment
specific and users are encourage to explore their data.

The \Rfunction{plot2d} method takes one \Rclass{MSnExp} instance as
first argument to produce retention time \textit{vs.} precursor MZ
scatter plots. Points represent individual MS2 spectra and can be
coloured based on precursor charge (with second argument
\Robject{z="charge"}), total ion count (\Robject{z="ionCount"}),
number of peaks in the MS2 spectra \Robject{z="peaks.count"}) or, when
multiple data files were loaded, file \Robject{z="file"}), as
illustrated on figure \ref{fig:plot2d}.  The lower right panel is
produced for only a subset of proteins.  See the method documentation
for more details.

\begin{figure}[htb]
    \includegraphics[width=\linewidth]{./plot2d-figure.png}
\caption{Illustration of the \Rfunction{plot2d} output. }
\label{fig:plot2d}
\end{figure}

The \Rfunction{plotDensity} method illustrates the distribution of
several parameters of interest (see figure \ref{fig:plotDensity}).
Similarly to \Rfunction{plot2d}, the first argument is an
\Rclass{MSnExp} instance.  The second is one of
\Robject{precursor.mz}, \Robject{peaks.count} or \Robject{ionCount},
whose density will be plotted.  An optional third argument specifies
whether the x axes should be logged.

\begin{figure}[htb]
    \includegraphics[width=\linewidth]{./plotDensity-figure.png}
\caption{ Illustration of the \Rfunction{plotDensity} output. }
\label{fig:plotDensity}
\end{figure}

%% xtable(preprocSelectionTable(velos))
%% % latex table generated in R 2.14.0 by xtable 1.5-6 package
%% % Thu May 19 10:22:29 2011
%% \begin{table}[ht]
%% \begin{center}
%% \begin{tabular}{rr}
%%   \hline
%%  & x \\
%%   \hline
%% 1 & 6836 \\
%%   2 &  59 \\
%%   3 &   5 \\
%%    \hline
%% \end{tabular}
%% \end{center}
%% \end{table}

%% matplot on iTRAQ5 quantitation (see allquant code chunk)

\bigskip

The \Rfunction{plotMzDelta} method\footnote{The code to generate the
  histograms has been contributed by Guangchuang Yu from Jinan
  University, China.} implements the $m/z$ delta plot from
\cite{Foster11} The $m/z$ delta plot illustrates the suitability of MS2
spectra for identification by plotting the $m/z$ differences of the most
intense peaks. The resulting histogram should optimally shown
outstanding bars at amino acid residu masses. More details and
parameters are described in the method documentation
(\Rfunction{?plotMzDelta}).  Figure \ref{fig:plotMzDelta-pride12011}
has been generated using the PRIDE experiment 12011, as in
\cite{Foster11}.

\begin{figure}[htb]
  \begin{center}
    \includegraphics[width=.7\linewidth]{./plotMzDelta-pride12011.pdf}
    \caption{Illustration of the \Rfunction{plotMzDelta} output for
      the PRIDE experiment 12011, as in figure 4A from
      \cite{Foster11}. }
\label{fig:plotMzDelta-pride12011}
  \end{center}
\end{figure}

\bigskip

In section \ref{sec:incompdissoc} on page \pageref{sec:incompdissoc},
we illustrate how to assess incomplete reporter ion dissociation.

\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Raw data processing}\label{sec:rawprocessing}

\subsection{Cleaning spectra}\label{sec:clean}

There are several methods implemented to perform basic raw data
processing and manipulation. Low intensity peaks can be set to 0 with
the \Rfunction{removePeaks} method from spectra or whole
experiments. The intensity threshold below which peaks are removed is
defined by the \Robject{t} parameter. \Robject{t} can be specified
directly as a numeric. The default value is the character
\texttt{"min"}, that will remove all peaks equal to the lowest non
null intensity in any spectrum.  We observe the effect of the
\Rfunction{removePeaks} method by comparing total ion count (i.e. the
total intensity in a spectrum) with the \Rfunction{ionCount} method
before (object \Robject{itraqdata}) and after (object
\Robject{experiment}) for spectrum \texttt{X55}. The respective
spectra are shown on figure \ref{fig:spectrum-clean-plot} (page
\pageref{fig:spectrum-clean-plot}).

<<removePeaks, echo=TRUE, cache=FALSE>>=
experiment <- removePeaks(itraqdata, t = 400, verbose = FALSE)
## total ion current
ionCount(itraqdata[["X55"]])
ionCount(experiment[["X55"]])
@

\begin{figure}[!ht]
<<spectrum-clean-plot, dev='pdf', fig.width=6, fig.height=3, echo=FALSE, fig.keep='high'>>=
p1 <- plot(itraqdata[["X55"]], full = TRUE, plot = FALSE) + theme_gray(5)
p2 <- plot(experiment[["X55"]], full = TRUE, plot = FALSE) + theme_gray(5)
grid.newpage()
pushViewport(viewport(layout = grid.layout(1, 2)))
vplayout <- function(x, y)
  viewport(layout.pos.row = x, layout.pos.col = y)
print(p1,vp=vplayout(1,1))
print(p2,vp=vplayout(1,2))
@
\caption{Same spectrum before (left) and after setting peaks <= 400 to 0.}
\label{fig:spectrum-clean-plot}
\end{figure}

Unlike the name might suggest, the \Rfunction{removePeaks} method does
not actually remove peaks from the spectrum; they are set to 0. This
can be checked using the \Rfunction{peaksCount} method, that returns
the number of peaks (including 0 intensity peaks) in a spectrum.  To
effectively remove 0 intensity peaks from spectra, and reduce the size
of the data set, one can use the \Rfunction{clean} method.  The effect
of the \Rfunction{removePeaks} and \Rfunction{clean} methods are
illustrated on figure \ref{fig:preproc} on page \pageref{fig:preproc}.

<<clean, echo=TRUE, cache=FALSE>>=
## number of peaks
peaksCount(itraqdata[["X55"]])
peaksCount(experiment[["X55"]])
experiment <- clean(experiment, verbose = FALSE)
peaksCount(experiment[["X55"]])
@

<<preprosp, cache=FALSE, echo=FALSE>>=
int <- c(0,1,1,3,1,1,0,0,0,1,3,7,3,1,0)
mz <- c(113.9,114.0,114.05,114.1,114.15,114.2,114.25,
        114.3,114.35,114.4,114.42,114.48,114.5,114.55,114.6)
ppsp <- new("Spectrum2",intensity=int,mz=mz,centroided=FALSE)
p1 <- plot(ppsp, full = TRUE, plot = FALSE) + theme_gray(5) +
  geom_point(size=3,alpha=I(1/3)) +
  geom_hline(yintercept=3,linetype=2) +
  ggtitle("Original spectrum")
p2 <- plot(removePeaks(ppsp,t=3), full=TRUE, plot = FALSE) +
  theme_gray(5) +
  geom_point(size=3,alpha=I(1/3)) +
  geom_hline(yintercept=3,linetype=2) +
  ggtitle("Peaks < 3 removed")
p3 <- plot(clean(removePeaks(ppsp,t=3)), full = TRUE, plot = FALSE) +
  theme_gray(5) +
  geom_point(size=3,alpha=I(1/3)) +
  geom_hline(yintercept=3,linetype=2) +
  ggtitle("Peaks < 3 removed and cleaned")
@

\begin{figure}[p]
\centering
<<preprocPlot, dev='pdf', fig.width=3, fig.height=6, echo=FALSE>>=
grid.newpage()
pushViewport(viewport(layout = grid.layout(3, 1)))
print(p1, vp=vplayout(1, 1))
print(p2, vp=vplayout(2, 1))
print(p3, vp=vplayout(3, 1))
@
\caption{This figure illustrated the effect or the \Rfunction{removePeaks}
  and \Rfunction{clean} methods. The left-most spectrum displays two peaks,
  of max height 3 and 7 respectively. The middle spectrum shows the result
  of calling \Rfunction{removePeaks} with argument \Robject{t=3}, which sets
  all data points of the first peak, whose maximum height is smaller or equal
  to \Robject{t} to 0. The second peak is unaffected. Calling \Rfunction{clean}
  after \Rfunction{removePeaks} effectively deletes successive 0 intensities from
  the spectrum, as shown on the right plot. }
\label{fig:preproc}
\end{figure}

\subsection{Focusing on specific MZ values}\label{sec:trim}

Another useful manipulation method is \Rfunction{trimMz}, that takes
as parameters and \Rclass{MSnExp} (or a \Rclass{Spectrum}) and a
numeric \Robject{mzlim}. MZ values smaller then \texttt{min(mzlim)} or
greater then \texttt{max(mzmax)} are discarded. This method is
particularly useful when one wants to concentrate on a specific MZ
range, as for reporter ions quantification, and generally results in
substantial reduction of data size. Compare the size of the full
trimmed experiment to the original \Sexpr{sz} Mb.
\label{trimMz-example}

<<trimMz, echo=TRUE, cache=FALSE>>=
range(mz(itraqdata[["X55"]]))
experiment <- filterMz(experiment, mzlim = c(112,120))
range(mz(experiment[["X55"]]))
experiment
@

As can be seen above, all processing performed on the experiment is
recorded and displayed as integral part of the experiment object.

\subsection{Spectrum processing}\label{sec:specproc}

\Rclass{MSnExp} and \Rclass{Spectrum2} instances also support standard
MS data processing such as smoothing and peak picking, as described in
the \Rfunction{smooth} and \Rfunction{pickPeak} manual pages. The
methods that either single spectra of experiments, process the
spectrum/spectra, and return a updated, processed, object. The
implementations originate from the \CRANpkg{MALDIquant} package
\cite{Gibb2012}.

\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{MS$^2$ isobaric tagging quantitation}\label{sec:isoquant}

\subsection{Reporter ions quantitation}\label{sec:quant}

Quantitation is performed on fixed peaks in the spectra, that are
specified with an \Rclass{ReporterIons} object. A specific peak is
defined by it's expected \Robject{mz} value and is searched for within
\Robject{mz} $\pm$ \Robject{width}.  If no data is found, \Robject{NA}
is returned.

<<reporters, echo=TRUE, cache=FALSE>>=
mz(iTRAQ4)
width(iTRAQ4)
@

The \Rfunction{quantify} method takes the following parameters: an
\Rclass{MSnExp} experiment, a character describing the quantification
\Robject{method}, the \Robject{reporters} to be quantified and a
\Robject{strict} logical defining whether data points ranging outside
of \Robject{mz} $\pm$ \Robject{width} should be considered for
quantitation.  Additionally, a progress bar can be displaying when
setting the \Robject{verbose} parameter to \Robject{TRUE}.  Three
quantification methods are implemented, as illustrated on figure
\ref{fig:quant-methods}: \Robject{trapezoidation} returns the area
under the peak of interest, \Robject{max} returns the apex of the peak
and \Robject{sum} returns the sum of all intensities of the peak.  See
\Rfunction{?quantify} for more details.

<<simplesp, cache=FALSE, echo=FALSE, fig.keep='none'>>=
int <- c(0, 1, 1, 3, 1, 1, 0)
mz <- c(113.9, 114.0, 114.05, 114.1, 114.15, 114.2, 114.25)
ssp <- new("Spectrum2", intensity = int, mz = mz, centroided = FALSE)
p <- plot(ssp, full = TRUE, plot = FALSE)
p <- p + theme_gray(5)
@

\begin{figure}[p]
\centering
<<quantitationPlot, dev='pdf',fig.width=5, fig.height=5, echo=FALSE>>=
grid.newpage()
pushViewport(viewport(layout = grid.layout(2, 2)))
vplayout <- function(x, y)
  viewport(layout.pos.row = x, layout.pos.col = y)
print(p + ggtitle("Quantitation using 'sum'") +
      geom_point(size = 3, alpha = I(1/3), colour = "red"),
      vp = vplayout(1, 1))
print(p + ggtitle("Quantitation using 'max'") +
      geom_point(aes(x = 114.1, y = 3),
                 alpha = I(1/18),
                 colour = "red", size = 3),
      vp = vplayout(1, 2))
print(p +
      ggtitle("Trapezoidation and strict=FALSE") +
      geom_polygon(alpha = I(1/5), fill = "red"),
      vp = vplayout(2, 1))
print(p +
      ggtitle("Trapezoidation and strict=TRUE") +
      geom_polygon(aes(x = c(NA, 114.05, 114.05, 114.1, 114.15, 114.15, NA),
                       y=c(NA, 0, 1, 3, 1, 0, NA)), fill = "red", alpha = I(1/5)),
      vp = vplayout(2,2))
@
\caption{The different quantitation methods are illustrated above.
  Quantitation using \Robject{sum} sums all the data points in the
  peaks to produce, for this example,
  \Sexpr{quantify(ssp,iTRAQ4[1],method="sum")[[1]]}, whereas method
  \Robject{max} only uses the peak's maximum intensity,
  \Sexpr{quantify(ssp,iTRAQ4[1],method="max")[[1]]}.
  \Robject{Trapezoidation} calculates the area under the peak taking
  the full with into account (using \Robject{strict=FALSE} gives
  \Sexpr{round(quantify(ssp,iTRAQ4[1],method="trap",strict=FALSE)[[1]],3)})
  or only the width as defined by the reporter (using
  \Robject{strict=TRUE} gives
  \Sexpr{round(quantify(ssp,iTRAQ4[1],method="trap",strict=TRUE)[[1]],3)}). }
\label{fig:quant-methods}
\end{figure}

\bigskip

The \Rfunction{quantify} method returns \Rclass{MSnSet} objects, that
extend the well-known \Rclass{eSet} class defined in the
\Biocpkg{Biobase} package. \Rclass{MSnSet} instances are very
similar to \Rclass{ExpressionSet} objects, except for the experiment
meta-data that captures MIAPE specific information.  The assay data is
a matrix of dimensions $n \times m$, where $m$ is the number of
features/spectra originally in the \Rclass{MSnExp} used as parameter
in \Rfunction{quantify} and $m$ is the number of reporter ions, that can
be accessed with the \Rfunction{exprs} method.  The meta data is
directly inherited from the \Rclass{MSnExp} instance.

<<quantify, echo=TRUE, cache=FALSE, tidy=FALSE>>=
qnt <- quantify(experiment,
                method = "trap",
                reporters = iTRAQ4,
                strict = FALSE,
                verbose = FALSE)
qnt
head(exprs(qnt))
@

Figure \ref{fig:tmt10} illustrated the quantitation of the TMT 10-plex
isobaric tags using the \Rfunction{quantify} method and the
\texttt{TMT10} reporter instance. The data on the $x$ axis has been
quantified using \texttt{method = "max"} and centroided data (as
generated using ProteoWizard's \texttt{msconvert} with vendor
libraries' peak picking); on the $y$ axis, the quantitation method was
\texttt{trapezoidation} and \texttt{strict = TRUE} (that's important
for TMT 10-plex) and the profile data. We observe a very good
correlation.

\begin{figure}[htb]
  \centering
  \includegraphics[width=.6\linewidth]{./tmt10comp.pdf}
\caption{TMT 10-plex quantitation. }
\label{fig:tmt10}
\end{figure}

If no peak is detected for a reporter ion peak, the respective
quantitation value is set to \texttt{NA}. In our case, there
is \Sexpr{sum(is.na(exprs(qnt)))} such case in row
\Sexpr{which(is.na(exprs(qnt))) %% nrow(qnt)}.
We will remove the offending line using the \Rfunction{filterNA} method.
The \Robject{pNA} argument defines the percentage of accepted missing
values per feature. As we do not expect any missing peaks, we set it
to be 0 (which is also the detault value).

<<filterNA, echo=TRUE>>=
table(is.na(qnt))
qnt <- filterNA(qnt, pNA = 0)
sum(is.na(qnt))
@

The filtering criteria for \Rfunction{filterNA} can also be defined as
a pattern of columns that can have missing values and columns that
must not exhibit any.  See \Rfunction{?filterNA} for details and
examples.

The infrastructure around the \Rclass{MSnSet} class allows flexible
filtering using the \Rfunction{[} sub-setting operator. Below, we
mimic the behaviour of \Rfunction{filterNA(, pNA = 0)} by calculating
the row indices that should be removed, i.e. those that have at least
one \Robject{NA} value and explicitly remove these rows.  This method
allows one to devise and easily apply any filtering strategy.

<<removeNa, echo=TRUE, eval=FALSE>>=
whichRow <- which(is.na((qnt))) %% nrow(qnt)
qnt <- qnt[-whichRow, ]
@

See also the \Rfunction{plotNA} method to obtain a graphical overview of
the completeness of a data set.

\subsection{Importing quantitation data}\label{sec:io2}

If quantitation data is already available as a spreadsheet, it can be
imported, along with additional optional feature and sample (pheno) meta data,
with the \Rfunction{readMSnSet} function. This function takes the
respective text-based spreadsheet (comma- or tab-separated) file names
as argument to create a valid \Rclass{MSnSet} instance.

Note that the quantitation data of \Rclass{MSnSet} objects can also
be exported to a text-based spreadsheet file using the
\Rfunction{write.exps} method.

\Biocpkg{MSnbase} also supports the \texttt{mzTab}
format\footnote{\url{https://github.com/HUPO-PSI/mzTab}}, a
light-weight, tab-delimited file format for proteomics
data. \texttt{mzTab} files can be read into \R with
\Rfunction{readMzTabData} to create and \Rclass{MSnSet} instance.

\bigskip

See the \texttt{MSnbase-io} vignette for a general overview of
\Biocpkg{MSnbase}'s input/ouput capabilites.

\subsection{Peak adjustments}\label{sec:purcor}

\paragraph{Single peak adjustment} In certain cases, peak intensities
need to be adjusted as a result of peak interferance. For example, the
$+1$ peak of the phenylalanine (F, Phe) immonium ion (with $m/z$ 120.03)
inteferes with the 121.1 TMT reporter ion.  Below, we calculate the
relative intensity of the +1 peaks compared to the main peak using the
\Biocpkg{Rdisop} package.

<<pheplus1, echo=TRUE, cache=FALSE>>=
library(Rdisop)
## Phenylalanine immonium ion
Fim <- getMolecule("C8H10N")
getMass(Fim)
isotopes <- getIsotope(Fim)
F1 <- isotopes[2, 2]
F1
@

If desired, one can thus specifically quantify the F immonium ion in
the MS2 spectrum, estimate the intensity of the +1 ion
(\Sexpr{round(F1,4)}\% of the F peak) and substract this calculated
value from the 121.1 TMT reporter intensity.

The above principle can also be generalised for a set of overlapping
peaks, as described below.

\paragraph{Reporter ions purity correction} Impurities in the reporter
reagents can also bias the results and can be corrected when
manufacturers provide correction coefficients.  These generally come
as percentages of each reporter ion that have masses differing by -2,
-1, +1 and +2 Da from the nominal reporter ion mass due to isotopic
variants.  The \Rfunction{purityCorrect} method applies such
correction to \Rclass{MSnSet} instances. It also requires a square
matrix as second argument, \Robject{impurities}, that defines the
relative percentage of reporter in the quantified each peak.  See
\Rfunction{?purityCorrect} for more details.

<<purityCorrect, echo=TRUE, cache=FALSE, tidy = FALSE>>=
impurities <- matrix(c(0.929, 0.059, 0.002, 0.000,
                       0.020, 0.923, 0.056, 0.001,
                       0.000, 0.030, 0.924, 0.045,
                       0.000, 0.001, 0.040, 0.923),
                     nrow = 4)
qnt.crct <- purityCorrect(qnt, impurities)
head(exprs(qnt))
head(exprs(qnt.crct))
@

The \Rfunction{makeImpuritiesMatrix} can be used to create impurity
matrices.  It opens a rudimentary spreadsheet that can be directly
edited.

\clearpage

\section{Processing quantitative data}\label{sec:qproc}

\subsection{Data imputation}\label{sec:imp}

A set of imputation methods are available in the \Rfunction{impute}
method: it takes an \Rclass{MSnSet} instance as input, the name of the
imputation method to be applied (one of \Sexpr{paste(imputeMethods(),
  collapse=", ")}), possible additional parameters and returns an
updated for \Rclass{MSnSet} without any missing values. Below, we
apply a deterministic minimum value imputation on the \Robject{naset}
example data:

<<impute>>=
## an example MSnSet containing missing values
data(naset)
table(is.na(naset))
## number of NAs per protein
table(fData(naset)$nNA) 
x <- impute(naset, "min")
processingData(x)
table(is.na(x))
@

As described in more details in \cite{Lazar:2016}, there are two types
of mechanisms resulting in missing values in LC/MSMS experiments.

\begin{itemize}
\item Missing values resulting from absence of detection of a feature,
  despite ions being present at detectable concentrations.  For
  example in the case of ion suppression or as a result from the
  stochastic, data-dependent nature of the MS acquisition
  method. These missing value are expected to be randomly distributed
  in the data and are defined as \emph{missing at random} (MAR) or
  \emph{missing completely at random} (MCAR).
\item Biologically relevant missing values, resulting from the
  \emph{absence} of the low abundance of ions (below the limit of
  detection of the instrument). These missing values are not expected
  to be randomly distributed in the data and are defined as
  \emph{missing not at random} (MNAR).
\end{itemize}

MAR and MCAR values can be reasonably well tackled by many imputation
methods. MNAR data, however, requires some knowledge about the
underlying mechanism that generates the missing data, to be able to
attempt data imputation. MNAR features should ideally be imputed with
a \emph{left-censor} (for example using a deterministic or
probabilistic minimum value) method. Conversely, it is recommended to
use \emph{hot deck} methods (for example nearest neighbour, maximum
likelihood, etc) when data are missing at random.

\begin{figure}[h]
  \centering
  <<miximpfig, echo=FALSE, dev='pdf', fig.width=8, fig.width=8>>=
  x <- impute(naset, "zero")
  exprs(x)[exprs(x) != 0] <- 1

  suppressPackageStartupMessages(library("gplots"))

  heatmap.2(exprs(x), col = c("lightgray", "black"),
              scale = "none", dendrogram = "none",
              trace = "none", keysize = 0.5, key = FALSE,
              RowSideColors = ifelse(fData(x)$randna, "orange", "brown"),
              ColSideColors = rep(c("steelblue", "darkolivegreen"), each = 8))
  @
\caption{Mixed imputation method. Black cells represent presence of
  quantitation values and light grey corresponds to missing data. The
  two groups of interest are depicted in green and blue along the
  heatmap columns. Two classes of proteins are annotated on the left:
  yellow are proteins with randomly occurring missing values (if any)
  while proteins in brown are candidates for non-random missing value
  imputation.}
\label{fig:miximp}
\end{figure}

It is anticipated that the identification of both classes of missing
values will depend on various factors, such as feature intensities and
experimental design. Below, we use perform mixed imputation, applying
nearest neighbour imputation on the \Sexpr{sum(fData(naset)$randna)}
features that are assumed to contain randomly distributed missing
values (if any) (yellow on figure \ref{fig:miximp}) and a
deterministic minimum value imputation on the
\Sexpr{sum(!fData(naset)$randna)} proteins that display a non-random
pattern of missing values (brown on figure \ref{fig:miximp}).

<<miximp>>=
x <- impute(naset, method = "mixed",
            randna = fData(naset)$randna,
            mar = "knn", mnar = "min")
x
@

Please read \Rfunction{?impute} for a description of the different
methods.

\clearpage

\subsection{Normalisation}\label{sec:norm}

A \Rclass{MSnSet} object is meant to be compatible with further
downstream packages for data normalisation and statistical
analysis. There is also a \Rfunction{normalise} (also available as
\Rfunction{normalize}) method for expression sets. The method takes
and instance of class \Rclass{MSnSet} as first argument, and a
character to describe the \Robject{method} to be used:

\begin{description}
\item{\Robject{quantiles}} Applies quantile normalisation
  \cite{Bolstad03} as implemented in the
  \Rfunction{normalize.quantiles} function of the
  \Biocpkg{preprocessCore} package.
\item{\Robject{quantiles.robust}} Applies robust quantile
  normalisation \cite{Bolstad03} as implemented in the
  \Rfunction{normalize.quantiles.robust} function of the
  \Biocpkg{preprocessCore} package.
\item{\Robject{vsn}} Applies variance stabilisation normalization
  \cite{Huber2002} as implemented in the \Rfunction{vsn2} function of
  the \Biocpkg{vsn} package.
\item{\Robject{max}} Each feature's reporter intensity is divided by
  the maximum of the reporter ions intensities.
\item{\Robject{sum}} Each feature's reporter intensity is divided by
  the sum of the reporter ions intensities.
\end{description}

See \Rfunction{?normalise} for more methods. A \Rfunction{scale}
method for \Rclass{MSnSet} instances, that relies on the
\Rfunction{base::scale} function.

<<normalise, echo=TRUE, cache=FALSE>>=
qnt.max <- normalise(qnt, "max")
qnt.sum <- normalise(qnt, "sum")
qnt.quant <- normalise(qnt, "quantiles")
qnt.qrob <- normalise(qnt, "quantiles.robust")
qnt.vsn <- normalise(qnt, "vsn")
@

The effect of these are illustrated on figure \ref{fig:norm-plot} and
figure \ref{fig:cv-plot} reproduces figure 3 of \cite{Karp2010} that
described the application of vsn on iTRAQ reporter data.

\begin{figure}[p]
\centering
<<normPlot, dev='pdf', fig.width=5, fig.height=7, echo=FALSE>>=
.plot <- function(x,ttl=NULL) {
  boxplot(exprs(x),
          main=ifelse(is.null(ttl),processingData(x)@processing[2],ttl),
          cex.main=.8,
          cex.lab=.5,
          cex.axis=.5,
          cex=.8)
  grid()
}
oldmar <- par()$mar
par(mfrow=c(3,2),mar=c(2.9,2.9,2.9,1))
.plot(qnt, ttl = "Non-normalised data")
.plot(qnt.max, ttl = "Maximum")
.plot(qnt.sum, ttl = "Sum")
.plot(qnt.quant, ttl = "Quantile")
.plot(qnt.qrob, ttl = "Robust quantile")
.plot(qnt.vsn, ttl = "vsn")
@
\caption{ Comparison of the normalisation \Rclass{MSnSet} methods.
  Note that vsn also glog-transforms the intensities. }
\label{fig:norm-plot}
\end{figure}

<<cvdata, echo=FALSE, cache=FALSE>>=
sd1 <- apply(log2(exprs(qnt))+10,1,sd)
mn1 <- apply(log2(exprs(qnt))+10,1,mean)
cv1 <- sd1/mn1
sd2 <- apply(exprs(qnt.vsn)+10,1,sd)
mn2 <- apply(exprs(qnt.vsn)+10,1,mean)
cv2 <- sd2/mn2
dfr <- rbind(data.frame(rank=order(mn1),cv=cv1,norm="raw"),
             data.frame(rank=order(mn2),cv=cv2,norm="vsn"))
library("zoo")
## rmed1 <- rollapply(cv1,7,function(x) median(x,na.rm=TRUE))
## rmed2 <- rollapply(cv2,7,function(x) median(x,na.rm=TRUE))
##
## Calling directly rollapply.zoo to make it zoo_1.6-4 compatible.
## The above requires zoo >= 1.7-0, which is as of 15 March 2011
## not yet available on CRAN (only on r-forge).
rmed1 <- zoo:::rollapply.zoo(cv1,7,function(x) median(x,na.rm=TRUE))
rmed2 <- zoo:::rollapply.zoo(cv2,7,function(x) median(x,na.rm=TRUE))
dfr2 <-
  rbind(data.frame(x=seq(1,70,by=(70/length(rmed1))),
                   y=rmed1,norm="raw"),
        data.frame(x=seq(1,70,by=(70/length(rmed1))),
                   y=rmed2,norm="vsn"))

p <- qplot(rank,cv,data=dfr,col=norm) +
  geom_line(data=dfr2,aes(x=x,y=y,colour=norm)) +
  theme_gray(7)
@

\begin{figure}[!ht]
\centering
<<cvPlot, dev='pdf', fig.width=5, fig.height=4, echo=FALSE>>=
print(p)
@
\caption{CV versus signal intensity comparison for log2 and vsn
  transformed data. Lines indicate running CV medians.}
\label{fig:cv-plot}
\end{figure}

Note that it is also possible to normalise individual spectra or whole
\Rclass{MSnExp} experiments with the \Rfunction{normalise} method
using the \Robject{max} method.  This will rescale all peaks between 0
and 1.  To visualise the relative reporter peaks, one should this
first trim the spectra using method \Rfunction{trimMz} as illustrated
in section \ref{sec:rawprocessing}, then normalise the \Rclass{MSnExp}
with \Rfunction{normalise} using \Robject{method="max"} as illustrated
above and plot the data using \Rfunction{plot} (figure
\ref{fig:msnexp-norm-plot}).

<<prepareMsnsetNormPlot, cache=FALSE, echo=FALSE, keep.fig='none'>>=
p <- plot(normalise(experiment[bsasel], "max"),
          reporters = iTRAQ4, full = FALSE, plot = FALSE)
p <- p + theme_gray(7)
@

\begin{figure}[p]
\centering
<<msnexpNormPlot, dev='pdf', fig.width=5, fig.height=6, echo=FALSE>>=
print(p)
@
\caption{Experiment-wide normalised MS2 spectra. The y-axes of the individual
  spectra is now rescaled between 0 and 1 (highest peak), as opposed to
  figure \ref{fig:msnexp-plot}. }
\label{fig:msnexp-norm-plot}
\end{figure}

\bigskip

Additional dedicated normalisation method are available for MS$^2$
label-free quantitation, as described in section \ref{sec:lf} and in
the \Rfunction{quantify} documentation.

\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Feature aggregation}\label{sec:feataggregation}

The above quantitation and normalisation has been performed on
quantitative data obtained from individual spectra. However, the
biological unit of interest is not the spectrum but the peptide or the
protein. As such, it is important to be able to summarise features
that belong to a same group, i.e. spectra from one peptide, peptides
that originate from one protein, or directly combine all spectra that
have been uniquely associated to one protein.

\Biocpkg{MSnbase} provides one function, \Rfunction{combineFeatures},
that allows to aggregate features stored in an \Rclass{MSnSet} using
build-in or user defined summary function and return a new
\Rclass{MSnSet} instance.  The three main arguments are described
below.  Additional details can be found in the method documentation.

<<makeGroups1,echo=FALSE,cache=FALSE>>=
gb <- fData(qnt)$ProteinAccession
@

\Rfunction{combineFeatures}'s first argument, \Robject{object}, is an
instance of class \Rclass{MSnSet}, as has been created in the section
\ref{sec:quant} for instance.  The second argument, \Robject{groupBy},
is a \Robject{factor} than has as many elements as there are features
in the \Rclass{MSnSet} \Robject{object} argument. The features
corresponding to the \Robject{groupBy} levels will be aggregated so
that the resulting \Rclass{MSnSet} output will have
\Rfunction{length(levels(groupBy))} features. Here, we will combine
individual MS2 spectra based on the protein they originate from.  As
shown below, this will result in \Sexpr{length(table(gb))} new
aggregated features.

<<makeGroups2,echo=TRUE,cache=FALSE>>=
gb <- fData(qnt)$ProteinAccession
table(gb)
length(unique(gb))
@

The third argument, \Robject{fun}, defined how to combine the
features. Predefined functions are readily available and can be
specified as strings (\Robject{fun="mean"}, \Robject{fun="median"},
\Robject{fun="sum"}, \Robject{fun="weighted.mean"} or
\Robject{fun="medianpolish"} to compute respectively the mean, media,
sum, weighted mean or median polish of the features to be aggregated).
Alternatively, is is possible to supply user defined functions with
\Robject{fun=function(x) \{ ... \}}. We will use the \Robject{median}
here.

<<combineFeatures, echo=TRUE, cache=FALSE>>=
qnt2 <- combineFeatures(qnt, groupBy = gb, fun = "median")
qnt2
@

%% Other intended ways to define aggregation groups is to import
%% identification data into R and collate it to the \Robject{MSnSet}
%% feature data, and subsequently use peptide sequences or protein assignments
%% to combine features.

\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Label-free MS$^2$ quantitation}\label{sec:lf}

\subsection{Peptide counting}

Note that if samples are not multiplexed, label-free MS$^2$
quantitation by spectral counting is possible using
\Biocpkg{MSnbase}. Once individual spectra have been assigned to
peptides and proteins (see section \ref{sec:id}), it becomes
straightforward to estimate protein quantities using the simple
peptide counting method, as illustrated in section
\ref{sec:feataggregation}.

<<count>>=
sc <- quantify(msexp, method = "count")
## lets modify out data for demonstration purposes
fData(sc)$accession[1] <- fData(sc)$accession[2]
fData(sc)$accession
sc <- combineFeatures(sc, groupBy = fData(sc)$accession,
                      fun = "sum")
exprs(sc)
@

Such count data could then be further analyses using dedicated count
methods (originally developed for high-throughput sequencing) and
directly available for \Rclass{MSnSet} instances in the
\Biocpkg{msmsTests} Bioconductor package.

\subsection{Spectral counting and intensity methods}

The spectral abundance factor (SAF) and the normalised form (NSAF)
\cite{Paoletti2006} as well as the spectral index (SI) and other
normalised variations (SI$_{GI}$ and SI$_N$) \cite{Griffin2010} are
also available. Below, we illustrate how to apply the normalised
SI$_N$ to the experiment containing identification data produced in
section \ref{sec:id}.

The spectra that did not match any peptide have already been remove
with the \Rfunction{removeNoId} method. As can be seen in the
following code chunk, the first spectrum could not be matched to any
single protein. Non-identified spectra and those matching multiple
proteins are removed automatically prior to any label-free
quantitation. Once can also remove peptide that do not match uniquely
to proteins (as defined by the \Robject{nprot} feature variable column)
with the \Rfunction{removeMultipleAssignment} method.

<<labelfree>>=
fData(msexp)[, c("accession", "nprot")]
@

Note that the label-free methods implicitely apply feature aggregation
(section \ref{sec:feataggregation}) and normalise (section
\ref{sec:norm}) the quantitation values based on the total sample
intensity and or the protein lengths (see \cite{Paoletti2006} and
\cite{Griffin2010} for details).

Let's now proceed with the quantitation using the
\Rfunction{quantify}, as in section \ref{sec:quant}, this time however
specifying the method of interest, \Robject{SIn} (the \Robject{reporters}
argument can of course be ignored here). The required peptide-protein
mapping and protein lengths are extracted automatically from the
feature meta-data using the default \Robject{accession} and \Robject{length}
feature variables.

<<SIn>>=
siquant <- quantify(msexp, method = "SIn")
processingData(siquant)
exprs(siquant)
@

Other label-free methods can be applied by specifiying the appropriate
\Rfunction{method} argument. See \Rfunction{?quantify} for more details.

\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Spectra comparison}

\subsection{Plotting two spectra}

\Biocpkg{MSnbase} provides functionality to compare spectra against
each other. The first notable function is \Rfunction{plot}. If two
\Rclass{Spectrum2} objects are provided \Rfunction{plot} will draw
two plots: the upper and lower panel contain respectively the first
and second spectrum. Common peaks are drawn in a slightly darker
colour.

\begin{figure}[!ht]
\centering
<<plotSpectrumSpectrum, dev='pdf', fig.width=6.5, fig.height=6.5>>=
centroided <- pickPeaks(itraqdata, verbose = FALSE)
(k <- which(fData(centroided)[, "PeptideSequence"] == "TAGIQIVADDLTVTNPK"))
mzk <- precursorMz(centroided)[k]
zk <- precursorCharge(centroided)[k]
mzk * zk
plot(centroided[[k[1]]], centroided[[k[2]]])
@
\caption{Comparing two MS$^2$ spectra.}
\label{fig:compms2plot}
\end{figure}

\subsection{Comparison metrics}

Currently \Biocpkg{MSnbase} supports three different metrics to
compare spectra against each other: \Robject{common} to calculate the
number of common peaks, \Robject{cor} to calculate the Pearson
correlation and \Robject{dotproduct} to calculate the dot product. See
\Rfunction{?compareSpectra} to apply other arbitrary metrics.

<<compareSpectra, tidy=FALSE>>=
compareSpectra(centroided[[2]], centroided[[3]],
               fun = "common")
compareSpectra(centroided[[2]], centroided[[3]],
               fun = "cor")
compareSpectra(centroided[[2]], centroided[[3]],
               fun = "dotproduct")
@

\Rfunction{compareSpectra} supports \Rclass{MSnExp} objects as well.

<<msnexpcompareSpectra>>=
compmat <- compareSpectra(centroided, fun="cor")
compmat[1:10, 1:5]
@

Below, we illustrate how to compare a set of spectra using a
hierarchical clustering.

\begin{figure}[!ht]
\centering
<<hclustcompareSpectra, dev='pdf', fig.width=5, fig.height=6>>=
plot(hclust(as.dist(compmat)))
@
\label{fig:dendo}
\end{figure}

\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Quantitative assessment of incomplete dissociation}
\label{sec:incompdissoc}

Quantitation using isobaric reporter tags assumes complete
dissociation between the reporter group (red on figure
\ref{fig:itraqchem}), balance group (blue) and peptide (the peptide
reactive group is drawn in green).  However, incomplete dissociation
does occur and results in an isobaric tag (i.e reporter and balance
groups) specific peaks.

\begin{figure}[h]
  \begin{center}
    \includegraphics[width=4cm]{./itraqchem.pdf}
    \caption{ iTRAQ 4-plex isobaric tags reagent consist of three
      parts: (1) a charged reporter group (MZ of 114, 115, 116 and
      117) that is unique to each of the four reagents (red), (2) an
      uncharged mass balance group (28-31 Da) (blue)and (3) a peptide
      reactive group (NHS ester) that binds to the peptide.  In case
      of incomplete dissociation, the reporter and balance groups
      produce a specific peaks at MZ 145.  }
\label{fig:itraqchem}
  \end{center}
\end{figure}

\Biocpkg{MSnbase} provides, among others, a \Rclass{ReporterIons}
object for iTRAQ 4-plex that includes the 145 peaks, called
\Rclass{iTRAQ5}. This can then be used to quantify the experiment as
show in section \ref{sec:quant} to estimate incomplete dissociation
for each spectrum.

<<incompdiss, echo=TRUE, cache=FALSE, tidy=FALSE>>=
iTRAQ5
incompdiss <- quantify(itraqdata,
                       method = "trap",
                       reporters = iTRAQ5,
                       strict = FALSE,
                       verbose = FALSE)
head(exprs(incompdiss))
@

Figure \ref{fig:incompdiss} compares these intensities for the whole experiment.

\begin{figure}[!hbt]
\centering
<<incompdissPlot, dev='pdf', fig.width=5, fig.height=2.5, echo=FALSE, warning=FALSE>>=
dfr <- melt(exprs(incompdiss))
colnames(dfr) <- c("feature", "reporters", "intensity")
dfr$reporters <- sub("iTRAQ5.", "", dfr$reporters)
repsum <- apply(exprs(incompdiss), 1, function(x) sum(x[1:4]))
dfr2 <- data.frame(iTRAQ1to4 = repsum, iTRAQ5 = exprs(incompdiss)[,5])
p1 <- ggplot(data = dfr, aes(x = reporters,y = log10(intensity))) +
  geom_boxplot() +
  theme_gray(6)
p2 <- ggplot(data = dfr2, aes(x = log10(iTRAQ1to4), y = log10(iTRAQ5))) +
  geom_point(alpha = I(1/2)) +
  geom_abline(intercept = 0, slope = 1, linetype = "dotted") +
  stat_smooth(method = lm, se = FALSE) +
  xlab(label = expression(log[10]~sum~114~to~117)) +
  ylab(label = expression(log[10]~145)) +
  theme_gray(6)
grid.newpage()
pushViewport(viewport(layout = grid.layout(1, 2)))
print(p1, vp = vplayout(1, 1))
print(p2, vp = vplayout(1, 2))
@
\caption{Boxplot and scatterplot comparing intensities of the 4
  reporter ions (or their sum, on the right) and the incomplete
  dissociation specific peak. }
\label{fig:incompdiss}
\end{figure}

\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Combining \Rclass{MSnSet} instances}\label{sec:combine}

Combining mass spectrometry runs can be done in two different ways
depending on the nature of these runs. If the runs represent repeated
measures of identical samples, for instance multiple fractions, the
data has to be combine along the row of the quantitation matrix: all
the features (along the rows) represent measurements of the same set
of samples (along the columns). In this situation, described in
section \ref{sec:comb1}, two experiments of dimensions $n_1$ (rows) by
$m$ (columns and $n_2$ by $m$ will produce a new experiment of
dimensions $n_1 + n_2$ by $m$.

When however, different sets of samples have been analysed in
different mass spectrometry runs, the data has to be combined along
the columns of the quantitation matrix: some features will be shared
across experiments and should thus be aligned on a same row in the new
data set, whereas unique features to one experiment should be set as
missing in the other one.  In this situation, described in section
\ref{sec:comb2}, two experiments of dimensions $n_1$ by $m_1$ and
$n_2$ by $m_2$ will produce a new experiment of dimensions
$unique_{n_1} + unique_{n_2} + shared_{n_1, n_2}$ by $m_1 + m_2$.  The
two first terms of the first dimension will be complemented by
\Robject{NA} values.

Default \Rclass{MSnSet} feature names (\Robject{X1}, \Robject{X2},
\ldots) and sample names (\Robject{iTRAQ4.114}, \Robject{iTRAQ4.115},
\Robject{iTRAQ4.116}, \ldots) are not informative.  The features and
samples of these anonymous quantitative data-sets should be updated
before being combined, to guide how to meaningfully merge them.

\subsection{Combining identical samples}\label{sec:comb1}

To simulate this situation, let us use quantiation data from the
\Robject{itraqdata} object that is provided with the package as
experiment 1 and the data from the \Robject{rawdata} \Rclass{MSnExp}
instance created at the very beginning of this document. Both
experiments share the \emph{same} default iTRAQ 4-plex reporter names
as default sample names, and will thus automatically be combined along
rows.

<<makeexp12, echo=TRUE, cache=FALSE, tidy = FALSE>>=
exp1 <- quantify(itraqdata, reporters = iTRAQ4,
                 verbose = FALSE)
sampleNames(exp1)
centroided(rawdata) <- FALSE
exp2 <- quantify(rawdata, reporters = iTRAQ4,
                 verbose = FALSE)
sampleNames(exp2)
@

It important to note that the features of these independent
experiments share the same default anonymous names: X1, X2, X3,
\ldots, that however represent quantitation of distinct physical
analytes.  If the experiments were to be combined as is, it would
result in an error because data points for the same \emph{feature}
name (say \Robject{X1}) and the same \emph{sample name} (say
\Robject{iTRAQ4.114}) have different values.  We thus first update the
feature names to explicitate that they originate from different
experiment and represent quantitation from different spectra using the
convenience function \Rfunction{updateFeatureNames}.  Note that
updating the names of one experiment would suffice here.

<<updateFnames, echo=TRUE, cache=FALSE>>=
head(featureNames(exp1))
exp1 <- updateFeatureNames(exp1)
head(featureNames(exp1))
head(featureNames(exp2))
exp2 <- updateFeatureNames(exp2)
head(featureNames(exp2))
@

The two experiments now share the same sample names and have different
feature names and will be combined along the row. Note that all
meta-data is correctly combined along the quantitation values.

<<comb1, echo=TRUE, cache=FALSE>>=
exp12 <- combine(exp1, exp2)
dim(exp1)
dim(exp2)
dim(exp12)
@

\subsection{Combine different samples}\label{sec:comb2}

Lets now create two \Rclass{MSnSet}s from the same raw data to
simulate two different independent experiments that share some
features.  As done previously (see section \ref{sec:feataggregation}),
we combine the spectra based on the proteins they have been identified
to belong to.  Features can thus naturally be named using protein
accession numbers.  Alternatively, if peptide sequences would have
been used as grouping factor in \Rfunction{combineFeatures}, then
these would be good feature name candidates.

<<make2exps2, echo=TRUE, cache=FALSE, tidy=FALSE>>=
set.seed(1)
i <- sample(length(itraqdata), 35)
j <- sample(length(itraqdata), 35)
exp1 <- quantify(itraqdata[i], reporters = iTRAQ4,
                 verbose = FALSE)
exp2 <- quantify(itraqdata[j], reporters = iTRAQ4,
                 verbose = FALSE)
exp1 <- droplevels(exp1)
exp2 <- droplevels(exp2)
table(featureNames(exp1) %in% featureNames(exp2))
exp1 <- combineFeatures(exp1,
                        groupBy = fData(exp1)$ProteinAccession)
exp2 <- combineFeatures(exp2,
                        groupBy = fData(exp2)$ProteinAccession)
head(featureNames(exp1))
head(featureNames(exp2))
@

The \Rfunction{droplevels} drops the unused \Robject{featureData}
levels.  This is required to avoid passing absent levels as
\Robject{groupBy} in \Rfunction{combineFeatures}.  Alternatively, one
could also use \Robject{factor(fData(exp1)\$ProteinAccession)} as
\Robject{groupBy} argument.

The feature names are updated automatically by
\Rfunction{combineFeatures}, using the \Robject{groupBy} argument.
Proper feature names, reflecting the nature of the features (spectra,
peptides or proteins) is critical when multiple experiments are to be
combined, as this is done using common features as defined by their
names (see below).

\bigskip

Sample names should also be updated to replace anonymous reporter
names with relevant identifiers; the individual reporter data is
stored in the \Robject{phenoData} and is not lost.  A convenience
function \Rfunction{updateSampleNames} is provided to append the
\Rclass{MSnSet}'s variable name to the already defined names,
although in general, biologically relevant identifiers are preferred.

<<renameSamples, echo=TRUE, cache=FALSE>>=
sampleNames(exp1)
exp1 <- updateSampleNames(exp1)
sampleNames(exp1)
sampleNames(exp1) <- c("Ctrl1", "Cond1", "Ctrl2", "Cond2")
sampleNames(exp2) <- c("Ctrl3", "Cond3", "Ctrl4", "Cond4")
@

At this stage, it is not yet possible to combine the two experiments,
because their feature data is not compatible yet; they share the same
feature variable labels, i.e. the feature data column names
(\Sexpr{paste(head(fvarLabels(exp1), n = 3), collapse=", ")}, \ldots),
but the part of the content is different because the original data was
(in particular all the spectrum centric data: identical peptides in
different runs will have different retention times, precursor
intensities, \ldots).  Feature data with identical labels (columns in
the data frame) and names (row in the data frame) are expected to have
the same data and produce an error if not conform.

<<fdatanames, echo=TRUE, cache=FALSE>>=
stopifnot(all(fvarLabels(exp1) == fvarLabels(exp2)))
fData(exp1)["BSA", 1:4]
fData(exp2)["BSA", 1:4]
@

Instead of removing these identical feature data columns, one can use
a second convenience function, \Rfunction{updateFvarLabels}, to update
feature labels based on the experiements variable name and maintain
all the metadata.

<<renameFvars, echo=TRUE, cache=FALSE>>=
exp1 <- updateFvarLabels(exp1)
exp2 <- updateFvarLabels(exp2)
head(fvarLabels(exp1))
head(fvarLabels(exp2))
@

It is now possible to combine \Robject{exp1} and \Robject{exp2},
including all the meta-data, with the \Rfunction{combine} method.  The
new experiment will contain the union of the feature names of the
individual experiments with missing values inserted appropriately.

<<combine, echo=TRUE, cache=FALSE>>=
exp12 <- combine(exp1, exp2)
dim(exp12)
pData(exp12)
exprs(exp12)[25:28, ]
exp12
@

\bigskip

In summary, when experiments with different samples need to be
combined (along the columns), one needs to (1) clarify the sample
names using \Rfunction{updateSampleNames} or better manually, for
biological relevance and (2) update the feature data variable labels
with \Rfunction{updateFvarLabels}.  The individual experiments (there
can be more than 2) can then easily be combined with the
\Rfunction{combine} method while retaining the meta-data.

If runs for the same sample (different fractions for example) need to
be combines, one needs to (1) differentiate the feature provenance
with \Rfunction{updateFeatureNames} prior to use \Rfunction{combine}.

\subsection{Splitting and unsplitting \Rclass{MSnSet} instances}\label{sec:split}

A single \Rclass{MSnSet} can also be split along the features/rows or
samples/columns using the \Rfunction{split} method and a factor
defining the splitting groups, resulting in an instance of class
\Rclass{MSnSetList}:

<<split>>=
data(dunkley2006)
head(pData(dunkley2006))
split(dunkley2006, dunkley2006$replicate)
## or, defining the appropriate annotation variable name
dun <- split(dunkley2006, "replicate")
@

Above, we split along the columns/samples, but the function would
equally work with a factor of length equal to the number of rows of
the \Rclass{MSnSet} (or a feature variable name) to split along the
rows/features.

Finally, the effect of \Rfunction{split} can be reverted by
\Rfunction{unsplit}.

<<unsplit>>=
dun2 <- unsplit(dun, pData(dunkley2006)$replicate)
compareMSnSets(dunkley2006, dun2)
@

See \Rfunction{?MSnSetList} for more details about the class,
\Rfunction{split} and \Rfunction{unsplit} and comments about storing
multiple assays pertaining the same experiment.

\subsection{Averaging \Rclass{MSnSet} instances}\label{sec:avg}

It is sometimes useful to average a set of replicated experiments to
facilitate their visualisation. This can be easily achieved with the
\Rfunction{averageMSnSet} function, which takes a list of valid
\Rclass{MSnSet} instances as input and creates a new object whose
expression values are an average of the original values. A value of
dispersion (\Robject{disp}) and a count of missing values (\Robject{nNA}) is
recorded in the feature metadata slot. The average and dispersion are
computed by default as the median and (non-parametric) coefficient of
variation (see \Rfunction{?npcv} for details), although this can easily be
parametrised, as described in \Rfunction{?averageMSnSet}.

The next code chunk illustrates the averaging function using three
replicated experiments from \cite{Tan2009} available in the
\Biocpkg{pRolocdata} package.

<<avg>>=
library("pRolocdata")
data(tan2009r1)
data(tan2009r2)
data(tan2009r3)
msnl <- MSnSetList(list(tan2009r1, tan2009r2, tan2009r3))
avgtan <- averageMSnSet(msnl)
head(exprs(avgtan))
head(fData(avgtan)$disp)
head(fData(avgtan)$nNA)
@

We are going to visualise the average data on a principle component
(PCA) plot using the \Rfunction{plot2D} function from the
\Biocpkg{pRoloc} package \cite{Gatto2014}. In addition, we are going
to use the measure of dispersion to highlight averages with high
variability by taking, for each protein, the maximum observed
dispersion in the 4 samples. Note that in the default implementation,
dispersions estimated from a single measurement (i.e. that had 2
missing values in our example) are set to 0; we will set these to the
overal maximum observed dispersion.

\begin{figure}[!ht]
\centering
<<avg2, dev='pdf', fig.height=5, fig.width=5>>=
disp <- rowMax(fData(avgtan)$disp)
disp[disp == 0] <- max(disp)
range(disp)
library("pRoloc")
plot2D(avgtan, cex = 3 * disp)
@
\caption{PCA plot of the averaged \Rclass{MSnSet}. The point sizes
  are proportional to the dispersion of the protein quantitation
  across the averaged data.}
\label{fig:plot2Davg}
\end{figure}

\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{MS$^E$ data processing}\label{sec:mse}

\Biocpkg{MSnbase} can also be used for MS$^E$ data independent
acquisition from Waters instrument.  The MS$^E$ pipeline depends on
the Bioconductor \Biocpkg{synapter} package \cite{Bond2013} that
produces \Rclass{MSnSet} instances for indvidual acquisitions. The
\Biocpkg{MSnbase} infrastructure can subsequently be used to further
combine experiments, as shown in section \ref{sec:comb2} and apply
\textit{top3} quantitation using the \Rfunction{topN} method.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% \section{Conclusion}\ref{sec:ccl}

\section{Session information}\label{sec:sessionInfo}
<<sessioninfo, results='asis', echo=FALSE, cache=FALSE>>=
toLatex(sessionInfo())
@

\bibliography{MSnbase}

\end{document}

### 2016-12-18 docs/articles/benchmarking.Rmd
udpate docs

---
title: "MSnbase2 benchmarking"
output:
  BiocStyle::html_document2:
     toc: false
vignette: >
  % \VignetteIndexEntry{MSnbase benchmarking}
  % \VignetteEngine{knitr::rmarkdown}
  % \VignetteKeyword{proteomics, mass spectrometry}
  % \VignettePackage{MSnbase}
---

```{r style, echo = FALSE, results = 'asis', message=FALSE}
BiocStyle::markdown()
```

```{r env, echo=FALSE, message=FALSE}
suppressPackageStartupMessages(library("MSnbase"))
suppressPackageStartupMessages(library("BiocParallel"))
```

**Package**: `r Biocpkg("MSnbase")`<br />
**Authors**: Laurent Gatto and Johannes Rainer<br />
**Modified:** `r file.info("benchmarkding.Rmd")$mtime`<br />
**Compiled**: `r date()`

# Introduction

In this vignette, we will document various timings and benchmarkings
of the recent `r Biocpkg("MSnbase")` development (aka `MSnbase2`),
that focuses on *on-disk* data access (as opposed to
*in-memory*). More details about the new implementation will be
documented elsewhere.

As a benchmarking dataset, we are going to use a subset of an TMT
6-plex experiment acquired on an LTQ Orbitrap Velos, that is
distributed with the `r Biocexptpkg("msdata")` package

```{r msdata}
library("msdata")
f <- msdata::proteomics(full.names = TRUE, pattern = "TMT_Erwinia")
basename(f)
```

We need to load the `r Biocpkg("MSnbase")` package and set the
session-wide verbosity flag to `FALSE`.

```{r verb}
library("MSnbase")
setMSnbaseVerbose(FALSE)
```

# Reading data

We first read the data using the original `readMSData` function that
generates an in-memory representation of the MS2-level raw data and
measure the time needed for this operation.

```{r read1}
system.time(inmem <- readMSData(f, msLevel = 2,
                                centroided = TRUE))
```

Next, we use the `readMSData2` function to generate an on-disk
representation of the same data.

```{r read2}
system.time(ondisk <- readMSData2(f, msLevel = 2,
                                  centroided = TRUE))
```

Creating the on-disk experiment is considerable faster and scales to
much bigger, multi-file data, both in terms of object creation time,
but also in terms of object size (see next section). We must of course
make sure that these two datasets are equivalent:

```{r equal12}
all.equal(inmem, ondisk)
```

# Data size

To compare the size occupied in memory of these two objects, we are
going to use the `object_size` function from the `r CRANpkg("pryr")`
package, which accounts for the data (the spectra) in the `assayData`
environment (as opposed to the `object.size` function from the `utils`
package).

```{r}
library("pryr")
object_size(inmem)
object_size(ondisk)
```

The difference is explained by the fact that for `ondisk`, the spectra
are not created and stored in memory; they are access on disk when
needed, such as for example for plotting:

```{r plot0, eval=FALSE}
plot(inmem[[200]], full = TRUE)
plot(ondisk[[200]], full = TRUE)
```

```{r plot1, echo=FALSE, fig.wide=TRUE, fig.cap = "Plotting in-memory and on-disk spectra"}
suppressMessages(requireNamespace("gridExtra"))
gridExtra::grid.arrange(plot(inmem[[200]],  full = TRUE),
                        plot(ondisk[[200]], full = TRUE),
                        ncol = 2)

```

# Accessing spectra

The drawback of the on-disk representation is when the spectrum data
has to actually be accessed. To compare access time, we are going to
use the `r CRANpkg("microbenchmark")` and repeat access 10 times to
compare access to all `r length(inmem)` and a single spectrum
in-memory (i.e. pre-loaded and constructed) and on-disk
(i.e. on-the-fly access).

```{r mb, cache=TRUE}
library("microbenchmark")
mb <- microbenchmark(spectra(inmem),
                     inmem[[200]],
                     spectra(ondisk),
                     ondisk[[200]],
                     times = 10)
mb
```

While it takes order or magnitudes more time to access the data on-the-fly
rather than a pre-generated spectrum, accessing all spectra is only marginally
slower than accessing all spectra, as most of the time is spent preparing the
file for access, which is done only once.

On-disk access performance will depend on the read throughput of the
disk. A comparison of the data import of the above file from an
internal solid state drive and from an USB3 connected hard disk showed
only small differences for the `readMSData2` call (1.07 *vs* 1.36
seconds), while no difference were observed for accessing individual
or all spectra. Thus, for this particular setup, performance was about
the same for SSD and HDD. This might however not apply to setting in
which data import is performed in parallel from multiple files.

<!-- ```{r ssdVsSata, eval = FALSE, echo = FALSE} -->
<!-- ## Reading the file from the internal SSD harddisk. -->
<!-- file_name <- "TMT_Erwinia_1uLSike_Top10HCD_isol2_45stepped_60min_01.mzML.gz" -->
<!-- f_1 <- paste0("/Users/jo/tmp/", file_name) -->
<!-- system.time(ondisk_1 <- readMSData2(f_1, -->
<!--                                     centroided = TRUE)) -->
<!-- ## Reading the same file from an external USB3 connected HD -->
<!-- f_2 <- paste0("/Volumes/jodata/tmp/", file_name) -->
<!-- system.time(ondisk_2 <- readMSData2(f_2, -->
<!--                                     centroided = TRUE)) -->

<!-- mb <- microbenchmark(spectra(ondisk_1), -->
<!--                      ondisk_1[[200]], -->
<!--                      spectra(ondisk_2), -->
<!--                      ondisk_2[[200]], -->
<!--                      times = 10) -->
<!-- ``` -->

Data access does not prohibit interactive usage, such as
plotting, for example, as it is about 1/2 seconds, which is an
operation that is relatively rare, compared to subsetting and
filtering, which are faster for on-disk data:

```{r subset}
i <- sample(length(inmem), 100)
system.time(inmem[i])
system.time(ondisk[i])
```

Operations on the spectra data, such as peak picking, smoothing,
cleaning, ... are cleverly cached and only applied when the data is
accessed, to minimise file access overhead. Finally, specific
operations such as for example quantitation (see next section) are
optimised for speed.

# MS2 quantitation

Below, we perform TMT 6-plex reporter ions quantitation on the first
100 spectra and verify that the results are identical (ignoring
feature names).

```{r qnt, cache=TRUE}
system.time(eim <- quantify(inmem[1:100], reporters = TMT6,
                            method = "max"))
system.time(eod <- quantify(ondisk[1:100], reporters = TMT6,
                            method = "max"))
all.equal(eim, eod, check.attributes = FALSE)
```

# Conclusions

This document focuses on speed and size improvements of the new
on-disk `MSnExp` representation. The extend of these improvements will
substantially increase for larger data.

For general functionality about the on-disk `MSnExp` data class and
`r Biocpkg("MSnbase")` in general, see other vignettes available with

```{r vigs, eval=FALSE}
vignette(package = "MSnbase")
```

### 2016-12-23 README.Rmd
Merge branch 'master' into devel
* master: (27 commits)
  Fix MS level test in quantify,OnDiskMSnExp to support MS2+ level quantitation
  add onDisk vs inMem addId data unit test
  add covr r dep
  remove codecov bash script as build fails/timeouts
  run codecov bash script
  trigger check and codecov
  add codecov shield
  udpate readme with build status
  add link to coc
  remove some latex pkgs not present in repo
  add code of conduct
  all latex package deps
  add BiocStyle dep to hopefully address all latex deps
  add parnotes dep
  add tex dependency
  but before_install first
  ass tlm

@@
+[![Build Status](https://travis-ci.org/lgatto/MSnbase.svg?branch=master)](https://travis-ci.org/lgatto/MSnbase) [![codecov.io](https://codecov.io/github/lgatto/MSnbase/coverage.svg?branch=master)](https://codecov.io/github/lgatto/MSnbase?branch=master)
+
 ```{r env, echo=FALSE}
+suppressPackageStartupMessages(library("BiocStyle"))
 knitr::opts_chunk$set(comment=NA)
@@
 
-`MSnbase` is an R/Bioconductor package that provides infrastructure
-for plotting, data manipulation and processing of MS-based proteomics
-data. The project was started by
+`r Biocpkg("MSnbase")` is an R/Bioconductor package that provides
+infrastructure for plotting, data manipulation and processing of
+MS-based proteomics data. The project was started by
 [Laurent Gatto](http://lgatto.github.io/) in October 2010 (Mon Oct 4
 23:35:23 2010, according to the git log) and has, since then,
-benefited from various contributions.
-
-```{r, echo=FALSE}
-auth <- packageDescription("MSnbase")$Author
-auth <- sub("<lg390@cam.ac.uk> ", "", auth)
-auth <- gsub("\n\t", "\n", auth)
-auth <- gsub("\t", " ", auth)
-cat(auth, "\n")
-```
+benefited from
+[various contributions](http://lgatto.github.io/msnbase-contribs/).
 
@@
 
-To install `MSnbase`
+To install `r Biocpkg("MSnbase")`
 
@@
 
-If you need the github version
+If you need the github version (not recommended unless you know what
+you are doing)
 
@@
 
-## Build status
-
-[![Build Status](https://travis-ci.org/lgatto/MSnbase.svg?branch=master)](https://travis-ci.org/lgatto/MSnbase) [![codecov.io](https://codecov.io/github/lgatto/MSnbase/coverage.svg?branch=master)](https://codecov.io/github/lgatto/MSnbase?branch=master)
-
 ## Questions
@@
 [Bioconductor support forum](https://support.bioconductor.org/), using
-`MSnbase` to tag the question.
+`MSnbase` to tag the question. Feel also free to open a GitHub
+[issue](https://github.com/lgatto/MSnbase/issues).
 
@@
 
-To cite the MSnbase package in publications, please use:
+To cite the `MSnbase` package in publications, please use:
 
@@
 any suggestions. The
-[`MSnbase` development vignette](http://bioconductor.org/packages/devel/bioc/vignettes/MSnbase/inst/doc/MSnbase-development.pdf)
+[`MSnbase` development vignette](http://lgatto.github.io/MSnbase/articles/MSnbase-development.html)
 gives some background on the class infrastructure.

### 2016-12-23 README.Rmd
update docs

@@
 
-`r Biocpkg("MSnbase")` is an R/Bioconductor package that provides
-infrastructure for plotting, data manipulation and processing of
-MS-based proteomics data. The project was started by
-[Laurent Gatto](http://lgatto.github.io/) in October 2010 (Mon Oct 4
-23:35:23 2010, according to the git log) and has, since then,
-benefited from
+[**MSbase**](http://lgatto.github.io/MSnbase/) is an R/Bioconductor
+package that provides infrastructure for plotting, data manipulation
+and processing of mass spectrometry and proteomics data. The project
+was started by [Laurent Gatto](http://lgatto.github.io/) in October
+2010 (Mon Oct 4 23:35:23 2010, according to the git log) and has,
+since then, benefited from
 [various contributions](http://lgatto.github.io/msnbase-contribs/).

### 2016-12-23 README.Rmd
fix typos from d342c70 in README.Rmd

--- 2016-12-23 README.Rmd
@@
 
-`r Biocpkg("MSnbase")` is an R/Bioconductor package that provides
-infrastructure for plotting, data manipulation and processing of
-MS-based proteomics data. The project was started by
-[Laurent Gatto](http://lgatto.github.io/) in October 2010 (Mon Oct 4
-23:35:23 2010, according to the git log) and has, since then,
-benefited from
+[**MSnbase**](http://lgatto.github.io/MSnbase/) is an R/Bioconductor
+package that provides infrastructure for plotting, data manipulation
+and processing of mass spectrometry and proteomics data. The project
+was started by [Laurent Gatto](http://lgatto.github.io/) in October
+2010 (Mon Oct 4 23:35:23 2010, according to the git log) and has,
+since then, benefited from
 [various contributions](http://lgatto.github.io/msnbase-contribs/).

### 2016-12-23 docs/articles/Figures/MSnbase-io-out.tex
Merge branch 'master' into devel
* master: (24 commits)
  fix typos from d342c70 in README.Rmd
  update installation ignore
  fix typo [ci skip]
  udpate docs with verbose is FALSE
  set default MSnbase verbose to FALSE
  update docs
  update docs
  update news with new news format
  Move NEWS to NEWS.md
  Migrate io and dev vignettes to BiocStyle's html_document2 style
  removes notes
  mv devel vignette to BiocStyle html2
  use BiocStyle's html_docuement2 for benchmarking vignette
  Amend show method for MSnExp
  Small change to show method for MSnExp
  udpate docs
  update docs with Seb's s

--- 2013-01-15 inst/doc/MSnbase-io/MSnbase-io.Rnw
@@
-%\VignetteIndexEntry{MSnbase IO capabilities}
-%\VignetteKeywords{Mass Spectrometry, Proteomics, Infrastructure }
-%\VignettePackage{MSnbase-io}
-\documentclass{article}
+\documentclass{minimal}
+
 \usepackage{tikz}
-\usetikzlibrary{shapes,arrows,shadows,fit}
-\usepackage{hyperref}
-\usepackage[authoryear,round]{natbib}
+\usetikzlibrary{shapes,arrows,shadows,fit,automata}
 
-\newcommand{\R}{\texttt{R} }
-\newcommand{\Rfunction}[1]{{\texttt{#1}}}
-\newcommand{\Robject}[1]{{\texttt{#1}}}
-\newcommand{\Rpackage}[1]{{\mbox{\normalfont\textsf{#1}}}}
-\newcommand{\email}[1]{\href{mailto:#1}{\normalfont\texttt{#1}}}
-%% colors
-\definecolor{Red}{rgb}{0.7,0,0}
-\definecolor{Blue}{rgb}{0,0,0.8}
+%%% <
+\usepackage{verbatim}
+\usepackage[active,tightpage]{preview}
+\PreviewEnvironment{tikzpicture}
+\setlength\PreviewBorder{5pt}%
+%%% >
 
-\usepackage{geometry}
-\geometry{verbose,
-  tmargin = 2.5cm,
-  bmargin = 2.5cm,
-  lmargin = 3.0cm,
-  rmargin = 3.0cm}
-
-\usepackage{hyperref}
-\usepackage{breakurl}
-\hypersetup{%
-  pdfusetitle,
-  bookmarks = {true},
-  bookmarksnumbered = {true},
-  bookmarksopen = {true},
-  bookmarksopenlevel = 2,
-  unicode = {true},
-  breaklinks = {false},
-  hyperindex = {true},
-  colorlinks = {true},
-  linktocpage = {true},
-  plainpages = {false},
-  linkcolor = {Blue},
-  citecolor = {Blue},
-  urlcolor = {Red},
-  pdfstartview = {Fit},
-  pdfpagemode = {UseOutlines},
-  pdfview = {XYZ null null null}
-}
+\begin{comment}
+  :Title:  MSnbase Output
+  :Author: Laurent Gatto
+\end{comment}
 
@@
 
-\input{Author.tex}
-
 \begin{document}
 
-\title{\Rpackage{MSnbase} input/output capabilities}
-
-\maketitle
-
-%% Abstract and keywords %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\vskip 0.3in minus 0.1in
-\hrule
-\begin{abstract}
-  This vignette describes \Rpackage{MSnbase}'s input and output 
-  capabilites.
-\end{abstract}
-\textit{Keywords}: Mass Spectrometry (MS), proteomics, infrastructure, IO.
-\vskip 0.1in minus 0.05in
-\hrule
-\vskip 0.2in minus 0.1in
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%% \tableofcontents
-
-\vspace{10mm}
-
-<<knitr, include=FALSE, cache=FALSE>>=
-library(knitr)
-opts_chunk$set(fig.align = 'center', 
-               fig.show = 'hold', 
-               par = TRUE,
-               prompt = TRUE,
-               comment = NA)
-options(replace.assign = TRUE, 
-        width = 70)
-@
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Section
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%\input{Foreword.tex}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% section
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Overview}
-
-\Rpackage{MSnbase}'s aims are to facilitate the reproducible analysis of 
-mass spectrometry data within the \R environment, from raw data import and processing, 
-feature quantification, quantification and statistical analysis of the results.
-Data import functions for several formats are provided and intermediate or final 
-results can also be saved or exported. 
-These capabilites are presented below.
-
-\section{Data input}
-
-\paragraph{Raw data} Data stored in one of the published \texttt{XML}-based 
-formats. i.e. \texttt{mzXML} \citep{Pedrioli2004}, \texttt{mzData} \citep{Orchard2007} 
-or \texttt{mzML} \citep{Martens2010}, 
-can be imported with the \Rfunction{readMSData} method, which makes use of 
-the \Rpackage{mzR} package to create \Robject{MSnExp} objects. 
-The files can be in profile or centroided mode. 
-See \Rfunction{?readMSData} for details.
-
-\paragraph{Peak lists} Peak lists in the \texttt{mgf} 
-format\footnote{\url{http://www.matrixscience.com/help/data\_file\_help.html\#GEN}} 
-can be imported using the \Rfunction{readMgfData}. 
-In this case, the peak data has generally been pre-processed by other software.
-See \Rfunction{?readMgfData} for details.
-
-\paragraph{Quantitation data} Third party software can be used to generate 
-quantitative data and exported as a spreadsheet (generally comma or tab separated format). 
-This data as well as any additional metadata can be imported with the 
-\Rfunction{readMSnSet} function. See \Rfunction{?readMSnSet} for details.
-
-\Rpackage{MSnbase} also supports the \texttt{mzTab} 
-format\footnote{\url{http://code.google.com/p/mztab/}}, a light-weight, tab-delimited 
-file format for proteomics data developed within the Proteomics Standards Initiative (PSI).
-\texttt{mzTab} files can be read into \R with \Rfunction{readMzTabData} to create and \Robject{MSnSet} instance.
-
-\begin{figure}[!htb] %% input
-  \begin{center}
-    \begin{tikzpicture}[node distance = 2cm, auto]
-      % Place nodes
-      \node [input] (raw) {Raw data in an open \texttt{XML} format};
-      \node [fun, left of=raw, node distance=4cm] (readMSData) {\Rfunction{readMSData}};
-      \node [input, below of=raw] (mgf) {Peak list in \texttt{mgf} format};
-      \node [fun, left of=mgf, node distance=4cm] (readMgfData) {\Rfunction{readMgfData}};
-      \node [input, below of=mgf] (spreadsheet) {Quantitation data as a spreadsheet};
-      \node [input, below of=spreadsheet] (mztab) {\texttt{mzTab} format};
-      \node [fun, left of=spreadsheet, node distance=4cm] (readMSnSet) {\Rfunction{readMSnSet}};
-      \node [fun, left of=mztab, node distance=4cm] (readMzTabData) {\Rfunction{readMzTabData}};
-      \node [obj, left of=readMSData, yshift=-2.5mm, node distance=4cm] (MSnExp) {\Robject{MSnExp}};
-      \node [obj, left of=readMSnSet, node distance=4cm] (MSnSet) {\Robject{MSnSet}};
-      \node [fun, below of=MSnExp] (quantify) {\Rfunction{quantify}};
-      % Background
-      \begin{pgfonlayer}{background}
-        \node [fill=yellow!20,rounded corners, draw=black!50, dashed, fit=(MSnExp) (quantify) (MSnSet)] {};
-      \end{pgfonlayer}
-      % Draw edges
-      \draw  (raw) -- (readMSData);
-      \draw  (mgf) -- (readMgfData);
-      \draw  (spreadsheet) -- (readMSnSet);
-      \draw [->] (readMSData) -- (MSnExp);
-      \draw [->] (readMgfData) -- (MSnExp);
-      \draw [->] (readMSnSet) -- (MSnSet);
-      \draw  (MSnExp) -- (quantify);
-      \draw [->] (quantify) -- (MSnSet);
-      \draw (mztab) -- (readMzTabData);
-      \draw [->] (readMzTabData) -- (MSnSet);
-    \end{tikzpicture}
-    \caption{Illustration of \texttt{MSnbase} input capabilities. 
-      The white and red boxes represent \R functions/methods and objects respectively. 
-      The blue boxes represent different disk storage formats.}
-    \label{fig:input}
-  \end{center}
-\end{figure}
-
-\section{Data output}
-
-\paragraph{RData files} \R objects can most easily be stored on disk with the \Rfunction{save} function.
-It creates compressed binary images of the data representation that can later be read back from the file 
-with the \Rfunction{load} function.
-
-\paragraph{Peak lists} \Robject{MSnExp} instances as well as individual spectra can be written 
-as \texttt{mgf} files with the \Rfunction{writeMgfData} method. Note that the metadata in the original 
-\R object can not be included in the file. See \Rfunction{?writeMgfData} for details.
-
-\paragraph{Quantitation data} Quantitation data can be exported to spreadsheet files with 
-the \Rfunction{write.exprs} method. Feature metadata can be appended to the feature 
-intenstiy values. See \Rfunction{?writeMgfData} for details.
-
-\Robject{MSnSet} instances can also be exported to \texttt{mzTab} files using the 
-\Rfunction{writeMzTabData} function.
-
-\begin{figure}[!htb] %% out
-  \begin{center}
 \begin{tikzpicture}[node distance = 2cm, auto]
@@
   \node [input, below of=raw] (mgf) {Peak list in \texttt{mgf} format};
-      \node [fun, right of=mgf, node distance=4cm, yshift=1cm] (writeMgfData) {\Rfunction{writeMgfData}};
+  \node [fun, right of=mgf, node distance=4cm, yshift=1cm] (writeMgfData) {\texttt{writeMgfData}};
   \node [input, below of=mgf] (spreadsheet) {Quantitation data as a spreadsheet};
-      \node [input, below of=spreadsheet] (mztab) {\texttt{mzTab} format};
-      \node [fun, right of=spreadsheet, node distance=4cm] (writeexprs) {\Rfunction{write.exprs}};
-      \node [fun, right of=mztab, node distance=4cm] (writeMzTabData) {\Rfunction{writeMzTabData}};
-      \node [obj, right of=writeMgfData, yshift=1.25cm, node distance=4cm] (MSnExp) {\Robject{MSnExp}};
-      \node [obj, right of=writeexprs, node distance=4cm] (MSnSet) {\Robject{MSnSet}};
-      \node [fun, below of=MSnExp] (quantify) {\Rfunction{quantify}};
+  % \node [input, below of=spreadsheet] (mztab) {\texttt{mzTab} format};
+  \node [fun, right of=spreadsheet, node distance=4cm] (writeexprs) {\texttt{write.exprs}};
+  % \node [fun, right of=mztab, node distance=4cm] (writeMzTabData) {\texttt{writeMzTabData}};
+  \node [obj, right of=writeMgfData, yshift=1.25cm, node distance=4cm] (MSnExp) {\texttt{MSnExp}};
+  \node [obj, right of=writeexprs, node distance=4cm] (MSnSet) {\texttt{MSnSet}};
+  \node [fun, below of=MSnExp] (quantify) {\texttt{quantify}};
   % Background
@@
   \draw [->] (quantify) -- (MSnSet);
-      \draw (MSnSet) -- (writeMzTabData);
-      \draw [->] (writeMzTabData) -- (mztab);
+  % \draw (MSnSet) -- (writeMzTabData);
+  % \draw [->] (writeMzTabData) -- (mztab);
 \end{tikzpicture}
-    \caption{Illustration of \texttt{MSnbase} output capabilities.
-      The white and red boxes represent \R functions/methods and objects respectively. 
-      The blue boxes represent different disk storage formats.}
-    \label{fig:output}
-  \end{center}
-\end{figure}
-
-\clearpage
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% section
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \section{Session information}
-%% \label{sec:sessionInfo}
-
-%% <<label=sessioninfo,results='asis',echo=FALSE,cache=FALSE>>=
-%% toLatex(sessionInfo())
-%% @
-
-\bibliographystyle{plainnat}
-\bibliography{MSnbase}
 

### 2016-12-23 docs/articles/MSnbase-development.Rnw
Merge branch 'master' into devel
* master: (24 commits)
  fix typos from d342c70 in README.Rmd
  update installation ignore
  fix typo [ci skip]
  udpate docs with verbose is FALSE
  set default MSnbase verbose to FALSE
  update docs
  update docs
  update news with new news format
  Move NEWS to NEWS.md
  Migrate io and dev vignettes to BiocStyle's html_document2 style
  removes notes
  mv devel vignette to BiocStyle html2
  use BiocStyle's html_docuement2 for benchmarking vignette
  Amend show method for MSnExp
  Small change to show method for MSnExp
  udpate docs
  update docs with Seb's s

%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{MSnbase development}
%\VignetteKeywords{Mass Spectrometry, Proteomics, Infrastructure }
%\VignettePackage{MSnbase-development}

\documentclass[12pt, oneside]{article}

<<style, eval=TRUE, echo=FALSE, results="asis">>=
BiocStyle::latex2()
@

\author{
  Laurent Gatto and Johannes Rainer
}

\bioctitle[\Biocpkg{MSnbase} devel]{\Biocpkg{MSnbase} development }

\begin{document}

\maketitle

\begin{abstract}
  This vignette describes the classes implemented in \Biocpkg{MSnbase}
  package.  It is intended as a starting point for developers or users
  who would like to learn more or further develop/extend mass
  spectrometry and proteomics data structures.
\end{abstract}

<<'setup', include = FALSE, cache = FALSE>>=
library("knitr")
opts_chunk$set(fig.align = 'center',
               fig.show = 'hold',
               par = TRUE,
               prompt = FALSE,
               comment = NA)
options(replace.assign = TRUE,
        width = 65)
@

<<environment, cache=FALSE, echo=FALSE>>=
suppressPackageStartupMessages(library(MSnbase))
@

\newpage

\tableofcontents

\newpage

\newpage

\input{Foreword.tex}

\input{Bugs.tex}

\newpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

This document is not a replacement for the individual manual pages, that
document the slots of the \Biocpkg{MSnbase} classes. It is a centralised high-level
description of the package design.

\Biocpkg{MSnbase} aims at being compatible with the \Biocpkg{Biobase}
infrastructure \cite{Gentleman2004}.  Many meta data structures that
are used in \Rclass{eSet} and associated classes are also used
here. As such, knowledge of the \textit{Biobase development and the
  new eSet} vignette would be beneficial; the vignette can directly be
accessed with
\texttt{vignette("BiobaseDevelopment", package="Biobase")}.

The initial goal is to use the \Biocpkg{MSnbase} infrastructure for
MS$^2$ labelled (iTRAQ \cite{Ross2004} and TMT \cite{Thompson2003})
and label-free (spectral counting, index and abundance ) quantitation
- see the documentation for the \Rfunction{quantify} function for
details.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{\Biocpkg{MSnbase} classes}

All classes have a \Rclass{.\_\_classVersion\_\_} slot, of class
\Rclass{Versioned} from the \Biocpkg{Biobase} package. This slot
documents the class version for any instance to be used for debugging
and object update purposes. Any change in a class implementation
should trigger a version change.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{\Rclass{pSet}: a virtual class for raw mass spectrometry data and meta data}
\label{sec:pset}

This virtual class is the main container for mass spectrometry data,
i.e spectra, and meta data. It is based on the \Rclass{eSet}
implementation for genomic data. The main difference with
\Rclass{eSet} is that the \Robject{assayData} slot is an environment
containing any number of \Rclass{Spectrum} instances (see section
\ref{sec:spectrum}).

One new slot is introduced, namely \Robject{processingData}, that
contains one \Rclass{MSnProcess} instance (see section
\ref{sec:msnprocess}).  and the \Robject{experimentData} slot is now
expected to contain \Rclass{MIAPE} data (see section \ref{sec:miape}).
The \Robject{annotation} slot has not been implemented, as no prior
feature annotation is known in shotgun proteomics.

<<pSet>>=
getClass("pSet")
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{\Rclass{MSnExp}: a class for MS experiments}
\label{sec:msnexp}

\Rclass{MSnExp} extends \Rclass{pSet} to store MS experiments.
It does not add any new slots
to \Rclass{pSet}. Accessors and setters are all inherited from
\Rclass{pSet} and new ones should be implemented for \Rclass{pSet}.
Methods that manipulate actual data in experiments are implemented
for \Rclass{MSnExp} objects.

<<MSnExp>>=
getClass("MSnExp")
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{\Rclass{OnDiskMSnExp}: a on-disk implementation of the \Rclass{MSnExp} class}
\label{sec:ondiskmsnexp}

The \Rclass{OnDiskMSnExp} class extends \Rclass{MSnExp} and inherits
all of its functionality but is aimed to use as little memory as
possible based on a balance between memory demand and
performance. Most of the spectrum-specific data, like retention time,
polarity, total ion current are stored within the object's
\Robject{featureData} slot. The actual M/Z and intensity values from
the individual spectra are, in contrast to \Rclass{MSnExp} objects,
not kept in memory (in the \Robject{assayData} slot), but are fetched
from the original files on-demand. Because mzML files are indexed,
using the \Biocpkg{mzR} package to read the relevant spectrum data is
fast and only moderately slower than for in-memory
\Rclass{MSnExp}\footnote{The \textit{benchmarking} vignette compares
  data size and operation speed of the two implementations.}.

To keep track of data manipulation steps that are applied to spectrum
data (such as performed by methods \Rfunction{removePeaks} or
\Rfunction{clean}) a \textit{lazy execution} framework was
implemented. Methods that manipulate or subset a spectrum's M/Z or
intensity values can not be applied directly to a
\Rclass{OnDiskMSnExp} object, since the relevant data is not kept in
memory. Thus, any call to a processing method that changes or subset
M/Z or intensity values are added as \Rclass{ProcessingStep} items to
the object's \Robject{spectraProcessingQueue}. When the spectrum data
is then queried from an \Rclass{OnDiskMSnExp}, the spectra are read in
from the file and all these processing steps are applied on-the-fly to
the spectrum data before being returned to the user.

% This has obviously a negative impact on the performance (e.g. for
% repeated function calls), would however also allow to remove certain
% processing steps at a later stage.

The operations involving extracting or manipulating spectrum data are
applied on a per-file basis, which enables parallel processing. Thus,
all corresponding method implementations for \Rclass{OnDiskMSnExp}
objects have an argument \Robject{BPPARAM} and users can set a
\Rcode{PARALLEL\_THRESH} option flag \footnote{see
  \Rfunction{?MSnbaseOptions} for details.} that enables to define how
and when parallel processing should be performed (using the
\Biocpkg{BiocParallel} package).

Note that all data manipulations that are not applied to M/Z or
intensity values of a spectrum (e.g. sub-setting by retention time
etc) are very fast as they operate directly to the object's
\Rclass{featureData} slot.

<<OnDiskMSnExp>>=
getClass("OnDiskMSnExp")
@

The distinction between \Rclass{MSnExp} and \Rclass{OnDiskMSnExp} is
often not explicitly stated as it should not matter, from a user's
perspective, which data structure they are working with, as both
behave in equivalent ways. Often, they are referred to as
\textit{in-memory} and \textit{on-disk} \Rclass{MSnExp}
implementations.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{\Rclass{MSnSet}: a class for quantitative proteomics data}
\label{sec:msnset}

This class stores quantitation data and meta data after running
\Rfunction{quantify} on an \Rclass{MSnExp} object or by creating an
\Rclass{MSnSet} instance from an external file, as described in the
\texttt{MSnbase-io} vignette and in \Rfunction{?readMSnSet},
\Rfunction{readMzTabData}, etc. The quantitative data is in form of a
$m \times n$ matrix, where $m$ is the number of features/spectra
originally in the \Rclass{MSnExp} used as parameter in
\Rfunction{quantify} and $m$ is the number of reporter ions (see
section \ref{sec:reporterions}). If read from an external file, $n$
corresponds to the number of features (protein groups, proteins,
peptides, spectra) in the file and $m$ is the number of columns with
quantitative data (samples) in the file.

This prompted to keep a similar implementation as the
\Rclass{ExpressionSet} class, while adding the proteomics-specific
annotation slot introduced in the \Rclass{pSet} class, namely
\Robject{processingData} for objects of class \Rclass{MSnProcess} (see
section \ref{sec:msnprocess}).

<<MSnSet>>=
getClass("MSnSet")
@

The \Rclass{MSnSet} class extends the virtual \Rclass{eSet} class to
provide compatibility for \Rclass{ExpressionSet}-like behaviour.  The
experiment meta-data in \Robject{experimentData} is also of class
\Rclass{MIAPE} (see section \ref{sec:miape}).  The
\Robject{annotation} slot, inherited from \Rclass{eSet} is not
used. As a result, it is easy to convert \Rclass{ExpressionSet} data
from/to \Rclass{MSnSet} objects with the coersion method
\Rfunction{as}.

<<as>>=
data(msnset)
class(msnset)
class(as(msnset, "ExpressionSet"))

data(sample.ExpressionSet)
class(sample.ExpressionSet)
class(as(sample.ExpressionSet, "MSnSet"))
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{\Rclass{MSnProcess}: a class for logging processing meta data}
\label{sec:msnprocess}

This class aims at recording specific manipulations applied to
\Rclass{MSnExp} or \Rclass{MSnSet} instances. The \Robject{processing}
slot is a \Robject{character} vector that describes major
processing. Most other slots are of class \Robject{logical} that
indicate whether the data has been centroided, smoothed, \ldots
although many of the functionality is not implemented yet.  Any new
processing that is implemented should be documented and logged here.

It also documents the raw data file from which the data originates
(\Robject{files} slot) and the \Biocpkg{MSnbase} version that was in
use when the \Rclass{MSnProcess} instance, and hence the
\Rclass{MSnExp}/\Rclass{MSnSet} objects, were originally created.

<<MSnProcess>>=
getClass("MSnProcess")
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{\Rclass{MIAPE}: Minimum Information About a Proteomics Experiment}
\label{sec:miape}

The Minimum Information About a Proteomics
Experiment \cite{Taylor2007, Taylor2008} \Rclass{MIAPE} class
describes the experiment, including contact details,
information about the mass spectrometer and control
and analysis software.

<<MIAPE>>=
getClass("MIAPE")
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{\Rclass{Spectrum} \textit{et al.}: classes for MS spectra}
\label{sec:spectrum}

\Rclass{Spectrum} is a virtual class that defines
common attributes to all types of spectra. MS1 and MS2 specific attributes
are defined in the \Rclass{Spectrum1} and \Rclass{Spectrum2} classes,
that directly extend \Rclass{Spectrum}.

%% The choices of attributes has been dictated by the \Rfunction{xcms:::rampRawData}
%% and \Rfunction{xcms:::rampRawDataMSn} functions and what data from the
%% \texttt{mzXML} file they gave access to.
%% It is expected that some hopefully minor changes might
%% come up here when migrating to other data import packages, that allow random access
%% to \texttt{mzXML} data and support \texttt{mzML} \cite{Martens2010}.

<<Spectrum>>=
getClass("Spectrum")
@

<<Spectrum1>>=
getClass("Spectrum1")
@

<<Spectrum2>>=
getClass("Spectrum2")
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{\Rclass{ReporterIons}: a class for isobaric tags}
\label{sec:reporterions}

The iTRAQ and TMT (or any other peak of interest)
are implemented \Rclass{ReporterIons} instances,
that essentially defines an expected MZ position for the peak and a width around
this value as well a names for the reporters.

<<ReporterIons>>=
getClass("ReporterIons")
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{\Rclass{NAnnotatedDataFrame}: multiplexed \Rclass{AnnotatedDataFrame}s}
\label{sec:nannotateddataframe}

The simple expansion of the \Rclass{AnnotatedDataFrame}
classes adds the \Robject{multiplex} and \Robject{multiLabel} slots to
document the number and names of multiplexed samples.

<<NAnnotatedDF>>=
getClass("NAnnotatedDataFrame")
@

\subsection{Other classes}\label{sec:other}

\subsubsection*{Lists of \Rclass{MSnSet} instances}\label{sec:msnl}

When several \Rclass{MSnSet} instances are related to each other and
should be stored together as different objects, they can be grouped as
a list into and \Rclass{MSnSetList} object. In addition to the actual
\Rclass{list} slot, this class also has basic logging functionality
and enables iteration over the \Rclass{MSnSet} instances using a
dedicated \Rfunction{lapply} methods.

<<msl>>=
getClass("MSnSetList")
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Miscellaneous}

\paragraph{Unit tests} \Biocpkg{MSnbase} implements unit tests with
the \CRANpkg{testthat} package.

\paragraph{Processing methods} Methods that process raw data, i.e. spectra
should be implemented for \Rclass{Spectrum} objects first and then
\texttt{eapply}'ed (or similar) to the \Robject{assayData} slot of an
\Rclass{MSnExp} instance in the specific method.

%% \paragraph{Why this MIAME slot?} \label{misc:whymiame} While it
%% would have been possible to transfer all data stored in
%% \Biocpkg{Biobase}'s \Rclass{MIAME} to a new \Rclass{MIAPE} class
%% and use the latter for \Robject{experimentData} slots in the
%% \Rclass{pSet} class, it would not have been possible to directly
%% transfer this to \Rclass{MSnSet} instances, as \Rclass{MSnSet}
%% classes directly inherit from the \Rclass{ExpressionSet}, whose
%% \Robject{experimentData} slot must be of class \Rclass{MIAME}.

%% \input{NoteAboutSpeedAndMemory.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Session information}
\label{sec:sessionInfo}

<<sessioninfo, results='asis', echo=FALSE>>=
toLatex(sessionInfo())
@

\bibliography{MSnbase}

\end{document}

### 2016-12-23 docs/articles/NoteAboutSpeedAndMemory.tex
Merge branch 'master' into devel
* master: (24 commits)
  fix typos from d342c70 in README.Rmd
  update installation ignore
  fix typo [ci skip]
  udpate docs with verbose is FALSE
  set default MSnbase verbose to FALSE
  update docs
  update docs
  update news with new news format
  Move NEWS to NEWS.md
  Migrate io and dev vignettes to BiocStyle's html_document2 style
  removes notes
  mv devel vignette to BiocStyle html2
  use BiocStyle's html_docuement2 for benchmarking vignette
  Amend show method for MSnExp
  Small change to show method for MSnExp
  udpate docs
  update docs with Seb's s

--- 2011-03-24 inst/doc/NoteAboutSpeedAndMemory.tex
@@
 data containers can easily grow very large and thus require large amounts of RAM. 
-This requirement may be tackled in the future by avoiding to load the raw data into memory
-and using random access to the content of \texttt{mzXML}/\texttt{mzML} data files on demand. 
+This requirement is being tackled by avoiding to load the raw data into memory
+and using on-disk random access to the content of \texttt{mzXML}/\texttt{mzML} data files on demand. 
 When focusing on reporter ion quantitation, a direct solution for this is to trim the 
 spectra using the \Robject{trimMz} method to select the area of interest and thus 
-substantially reduce the size of the \Robject{Spectrum} objects. This is illustrated on page 
-\pageref{trimMz-example} of the \texttt{MSnbase-demo} vignette.
+substantially reduce the size of the \Robject{Spectrum} objects. This is illustrated in 
+section \ref{sec:trim} on page \pageref{trimMz-example} of the \texttt{MSnbase-demo} vignette.
 
 The independent handling of spectra is ideally suited for parallel processing. 
-This will be added soon.
+The \Rfunction{quantify} method now performs reporter peaks quantitation in parallel. 
+More functions are being updated.

### 2016-12-23 R/zzz.R
Merge branch 'master' into devel
* master: (24 commits)
  fix typos from d342c70 in README.Rmd
  update installation ignore
  fix typo [ci skip]
  udpate docs with verbose is FALSE
  set default MSnbase verbose to FALSE
  update docs
  update docs
  update news with new news format
  Move NEWS to NEWS.md
  Migrate io and dev vignettes to BiocStyle's html_document2 style
  removes notes
  mv devel vignette to BiocStyle html2
  use BiocStyle's html_docuement2 for benchmarking vignette
  Amend show method for MSnExp
  Small change to show method for MSnExp
  udpate docs
  update docs with Seb's s

@@ -13,6 +13,6 @@
 .onLoad <- function(libname, pkgname) {
     ## Add MSnbase options.
     msOps <- list(PARALLEL_THRESH = 1000,
-                  verbose = TRUE)
+                  verbose = FALSE)
     options(MSnbase = msOps)
 }

### 2016-12-24 README.Rmd

--- 2016-12-23 README.Rmd
@@
 
-`r Biocpkg("MSnbase")` is an R/Bioconductor package that provides
-infrastructure for plotting, data manipulation and processing of
-MS-based proteomics data. The project was started by
-[Laurent Gatto](http://lgatto.github.io/) in October 2010 (Mon Oct 4
-23:35:23 2010, according to the git log) and has, since then,
-benefited from
-[various contributions](http://lgatto.github.io/msnbase-contribs/).
+[**MSnbase**](http://lgatto.github.io/MSnbase/) is an R/Bioconductor
+package that provides infrastructure for plotting, data manipulation
+and processing mass spectrometry and proteomics data. The project was
+started by [Laurent Gatto](http://lgatto.github.io/) in October 2010
+(Mon Oct 4 23:35:23 2010, according to the git log) and has, since
+then, benefited from [various
+contributions](http://lgatto.github.io/msnbase-contribs/).
 

### 2016-12-31 README.Rmd
remove link to doi that doesn't render properly
upate readme shields [skip ci]

@@
-[![Build Status](https://travis-ci.org/lgatto/MSnbase.svg?branch=master)](https://travis-ci.org/lgatto/MSnbase) [![codecov.io](https://codecov.io/github/lgatto/MSnbase/coverage.svg?branch=master)](https://codecov.io/github/lgatto/MSnbase?branch=master)
+Travis-CI:
+[![Build Status](https://travis-ci.org/lgatto/MSnbase.svg?branch=master)](https://travis-ci.org/lgatto/MSnbase) 
+[![codecov.io](https://codecov.io/github/lgatto/MSnbase/coverage.svg?branch=master)](https://codecov.io/github/lgatto/MSnbase?branch=master)
+Bioconductor:
+[![Bioconductor devel build Status](http://bioconductor.org/shields/build/devel/bioc/MSnbase.svg)](http://bioconductor.org/packages/devel/bioc/html/MSnbase.html) 
+[![Test coverage](https://bioconductor.org/shields/coverage/release/MSnbase.svg)](https://bioconductor.org/shields/coverage/devel/MSnbase.svg)
 
@@
 > quantitation*. Bioinformatics. 2012 Jan
-> 15;28(2):288-9. doi:[10.1093/bioinformatics/btr645](http://bioinformatics.oxfordjournals.org/content/28/2/288.long). Epub
-> 2011 Nov 22.  PubMed
+> 15;28(2):288-9. doi:10.1093/bioinformatics/btr645. PubMed
 > [PMID:22113085](http://www.ncbi.nlm.nih.gov/pubmed/22113085).

### 2016-12-31 README.Rmd
upate readme shields [skip ci]

--- 2016-12-24 README.Rmd
@@
-[![Build Status](https://travis-ci.org/lgatto/MSnbase.svg?branch=master)](https://travis-ci.org/lgatto/MSnbase) [![codecov.io](https://codecov.io/github/lgatto/MSnbase/coverage.svg?branch=master)](https://codecov.io/github/lgatto/MSnbase?branch=master)
+[![Build Status](https://travis-ci.org/lgatto/MSnbase.svg?branch=master)](https://travis-ci.org/lgatto/MSnbase) 
+[![codecov.io](https://codecov.io/github/lgatto/MSnbase/coverage.svg?branch=master)](https://codecov.io/github/lgatto/MSnbase?branch=master)
+[![Bioconductor devel build Status](http://bioconductor.org/shields/build/devel/bioc/MSnbase.svg)](http://bioconductor.org/packages/devel/bioc/html/MSnbase.html) 
+[![Test coverage](https://bioconductor.org/shields/coverage/release/MSnbase.svg)](https://bioconductor.org/shields/coverage/devel/MSnbase.svg)
 

### 2017-01-04 README.Rmd
merged with master
From: Laurent <lg390@cam.ac.uk>

git-svn-id: https://hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/MSnbase@125626 bc3139a8-67e5-0310-9ffc-ced21a209358

--- 2016-12-24 README.Rmd
@@
-[![Build Status](https://travis-ci.org/lgatto/MSnbase.svg?branch=master)](https://travis-ci.org/lgatto/MSnbase) [![codecov.io](https://codecov.io/github/lgatto/MSnbase/coverage.svg?branch=master)](https://codecov.io/github/lgatto/MSnbase?branch=master)
+[![Build Status](https://travis-ci.org/lgatto/MSnbase.svg?branch=master)](https://travis-ci.org/lgatto/MSnbase) 
+[![codecov.io](https://codecov.io/github/lgatto/MSnbase/coverage.svg?branch=master)](https://codecov.io/github/lgatto/MSnbase?branch=master)
+[![Bioconductor devel build Status](http://bioconductor.org/shields/build/devel/bioc/MSnbase.svg)](http://bioconductor.org/packages/devel/bioc/html/MSnbase.html) 
+[![Test coverage](https://bioconductor.org/shields/coverage/devel/MSnbase.svg)](https://codecov.io/github/Bioconductor-mirror/MSnbase/branch/master)
 

### 2017-01-20 docs/articles/MSnbase-demo.Rnw

@@
 
-\input{NoteAboutSpeedAndMemory.tex}
+\subsection{Speed and memory requirements}
+
+Raw mass spectrometry file are generally several hundreds of MB large
+and most of this is used for binary raw spectrum data. As such, data
+containers can easily grow very large and thus require large amounts
+of RAM.  This requirement is being tackled by avoiding to load the raw
+data into memory and using on-disk random access to the content of
+\texttt{mzXML}/\texttt{mzML} data files on demand.  When focusing on
+reporter ion quantitation, a direct solution for this is to trim the
+spectra using the \Robject{trimMz} method to select the area of
+interest and thus substantially reduce the size of the
+\Robject{Spectrum} objects. This is illustrated in section
+\ref{sec:trim} on page \pageref{trimMz-example}.
+
+The independent handling of spectra is ideally suited for parallel
+processing.  The \Rfunction{quantify} method now performs reporter
+peaks quantitation in parallel.  More functions are being updated.
+
+Finally, recent developmenets in version 2 of the package have solved
+the memory issue by implementing and \textit{on-disk} version the of
+data class storing raw data (\Rclass{MSnExp}, see
+section~\ref{sec:msnexp}), where the spectra a accessed on-disk only
+when required. The \textit{benchmarking} vignette compares the on-disk
+and in-memory implemenatations.
 
@@
 \begin{figure}[htb]
-    \includegraphics[width=\linewidth]{./plot2d-figure.png}
+    \includegraphics[width=\linewidth]{./Figures/plot2d-figure.png}
 \caption{Illustration of the \Rfunction{plot2d} output. }
@@
 \begin{figure}[htb]
-    \includegraphics[width=\linewidth]{./plotDensity-figure.png}
+    \includegraphics[width=\linewidth]{./Figures/plotDensity-figure.png}
 \caption{ Illustration of the \Rfunction{plotDensity} output. }
@@
   \begin{center}
-    \includegraphics[width=.7\linewidth]{./plotMzDelta-pride12011.pdf}
+    \includegraphics[width=.7\linewidth]{./Figures/plotMzDelta-pride12011.pdf}
     \caption{Illustration of the \Rfunction{plotMzDelta} output for
@@
   \centering
-  \includegraphics[width=.6\linewidth]{./tmt10comp.pdf}
+  \includegraphics[width=.6\linewidth]{./Figures/tmt10comp.pdf}
 \caption{TMT 10-plex quantitation. }
@@
   \begin{center}
-    \includegraphics[width=4cm]{./itraqchem.pdf}
+    \includegraphics[width=4cm]{./Figures/itraqchem.pdf}
     \caption{ iTRAQ 4-plex isobaric tags reagent consist of three

### 2017-04-18 docs/articles/Figures/MSnbase-io-in.tex
add docs dir

--- 2013-01-15 inst/doc/MSnbase-io/MSnbase-io.Rnw
@@
-%\VignetteIndexEntry{MSnbase IO capabilities}
-%\VignetteKeywords{Mass Spectrometry, Proteomics, Infrastructure }
-%\VignettePackage{MSnbase-io}
-\documentclass{article}
+\documentclass{minimal}
+
 \usepackage{tikz}
-\usetikzlibrary{shapes,arrows,shadows,fit}
-\usepackage{hyperref}
-\usepackage[authoryear,round]{natbib}
+\usetikzlibrary{shapes,arrows,shadows,fit,automata}
 
-\newcommand{\R}{\texttt{R} }
-\newcommand{\Rfunction}[1]{{\texttt{#1}}}
-\newcommand{\Robject}[1]{{\texttt{#1}}}
-\newcommand{\Rpackage}[1]{{\mbox{\normalfont\textsf{#1}}}}
-\newcommand{\email}[1]{\href{mailto:#1}{\normalfont\texttt{#1}}}
-%% colors
-\definecolor{Red}{rgb}{0.7,0,0}
-\definecolor{Blue}{rgb}{0,0,0.8}
+%%% <
+\usepackage{verbatim}
+\usepackage[active,tightpage]{preview}
+\PreviewEnvironment{tikzpicture}
+\setlength\PreviewBorder{5pt}%
+%%% >
 
-\usepackage{geometry}
-\geometry{verbose,
-  tmargin = 2.5cm,
-  bmargin = 2.5cm,
-  lmargin = 3.0cm,
-  rmargin = 3.0cm}
-
-\usepackage{hyperref}
-\usepackage{breakurl}
-\hypersetup{%
-  pdfusetitle,
-  bookmarks = {true},
-  bookmarksnumbered = {true},
-  bookmarksopen = {true},
-  bookmarksopenlevel = 2,
-  unicode = {true},
-  breaklinks = {false},
-  hyperindex = {true},
-  colorlinks = {true},
-  linktocpage = {true},
-  plainpages = {false},
-  linkcolor = {Blue},
-  citecolor = {Blue},
-  urlcolor = {Red},
-  pdfstartview = {Fit},
-  pdfpagemode = {UseOutlines},
-  pdfview = {XYZ null null null}
-}
+\begin{comment}
+  :Title:  MSnbase Input
+  :Author: Laurent Gatto
+\end{comment}
 
@@
 
-\input{Author.tex}
-
 \begin{document}
 
-\title{\Rpackage{MSnbase} input/output capabilities}
-
-\maketitle
-
-%% Abstract and keywords %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\vskip 0.3in minus 0.1in
-\hrule
-\begin{abstract}
-  This vignette describes \Rpackage{MSnbase}'s input and output 
-  capabilites.
-\end{abstract}
-\textit{Keywords}: Mass Spectrometry (MS), proteomics, infrastructure, IO.
-\vskip 0.1in minus 0.05in
-\hrule
-\vskip 0.2in minus 0.1in
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%% \tableofcontents
-
-\vspace{10mm}
-
-<<knitr, include=FALSE, cache=FALSE>>=
-library(knitr)
-opts_chunk$set(fig.align = 'center', 
-               fig.show = 'hold', 
-               par = TRUE,
-               prompt = TRUE,
-               comment = NA)
-options(replace.assign = TRUE, 
-        width = 70)
-@
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Section
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%\input{Foreword.tex}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% section
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Overview}
-
-\Rpackage{MSnbase}'s aims are to facilitate the reproducible analysis of 
-mass spectrometry data within the \R environment, from raw data import and processing, 
-feature quantification, quantification and statistical analysis of the results.
-Data import functions for several formats are provided and intermediate or final 
-results can also be saved or exported. 
-These capabilites are presented below.
-
-\section{Data input}
-
-\paragraph{Raw data} Data stored in one of the published \texttt{XML}-based 
-formats. i.e. \texttt{mzXML} \citep{Pedrioli2004}, \texttt{mzData} \citep{Orchard2007} 
-or \texttt{mzML} \citep{Martens2010}, 
-can be imported with the \Rfunction{readMSData} method, which makes use of 
-the \Rpackage{mzR} package to create \Robject{MSnExp} objects. 
-The files can be in profile or centroided mode. 
-See \Rfunction{?readMSData} for details.
-
-\paragraph{Peak lists} Peak lists in the \texttt{mgf} 
-format\footnote{\url{http://www.matrixscience.com/help/data\_file\_help.html\#GEN}} 
-can be imported using the \Rfunction{readMgfData}. 
-In this case, the peak data has generally been pre-processed by other software.
-See \Rfunction{?readMgfData} for details.
-
-\paragraph{Quantitation data} Third party software can be used to generate 
-quantitative data and exported as a spreadsheet (generally comma or tab separated format). 
-This data as well as any additional metadata can be imported with the 
-\Rfunction{readMSnSet} function. See \Rfunction{?readMSnSet} for details.
-
-\Rpackage{MSnbase} also supports the \texttt{mzTab} 
-format\footnote{\url{http://code.google.com/p/mztab/}}, a light-weight, tab-delimited 
-file format for proteomics data developed within the Proteomics Standards Initiative (PSI).
-\texttt{mzTab} files can be read into \R with \Rfunction{readMzTabData} to create and \Robject{MSnSet} instance.
-
-\begin{figure}[!htb] %% input
-  \begin{center}
 \begin{tikzpicture}[node distance = 2cm, auto]
@@
   \node [input] (raw) {Raw data in an open \texttt{XML} format};
-      \node [fun, left of=raw, node distance=4cm] (readMSData) {\Rfunction{readMSData}};
+  \node [fun, left of=raw, node distance=4cm] (readMSData) {\texttt{readMSData}};
   \node [input, below of=raw] (mgf) {Peak list in \texttt{mgf} format};
-      \node [fun, left of=mgf, node distance=4cm] (readMgfData) {\Rfunction{readMgfData}};
+  \node [fun, left of=mgf, node distance=4cm] (readMgfData) {\texttt{readMgfData}};
   \node [input, below of=mgf] (spreadsheet) {Quantitation data as a spreadsheet};
   \node [input, below of=spreadsheet] (mztab) {\texttt{mzTab} format};
-      \node [fun, left of=spreadsheet, node distance=4cm] (readMSnSet) {\Rfunction{readMSnSet}};
-      \node [fun, left of=mztab, node distance=4cm] (readMzTabData) {\Rfunction{readMzTabData}};
-      \node [obj, left of=readMSData, yshift=-2.5mm, node distance=4cm] (MSnExp) {\Robject{MSnExp}};
-      \node [obj, left of=readMSnSet, node distance=4cm] (MSnSet) {\Robject{MSnSet}};
-      \node [fun, below of=MSnExp] (quantify) {\Rfunction{quantify}};
+  \node [fun, left of=spreadsheet, node distance=4cm] (readMSnSet) {\texttt{readMSnSet}};
+  \node [fun, left of=mztab, node distance=4cm] (readMzTabData) {\texttt{readMzTabData}};
+  \node [obj, left of=readMSData, yshift=-2.5mm, node distance=4cm] (MSnExp) {\texttt{MSnExp}};
+  \node [obj, left of=readMSnSet, node distance=4cm] (MSnSet) {\texttt{MSnSet}};
+  \node [fun, below of=MSnExp] (quantify) {\texttt{quantify}};
   % Background
@@
 \end{tikzpicture}
-    \caption{Illustration of \texttt{MSnbase} input capabilities. 
-      The white and red boxes represent \R functions/methods and objects respectively. 
-      The blue boxes represent different disk storage formats.}
-    \label{fig:input}
-  \end{center}
-\end{figure}
-
-\section{Data output}
-
-\paragraph{RData files} \R objects can most easily be stored on disk with the \Rfunction{save} function.
-It creates compressed binary images of the data representation that can later be read back from the file 
-with the \Rfunction{load} function.
-
-\paragraph{Peak lists} \Robject{MSnExp} instances as well as individual spectra can be written 
-as \texttt{mgf} files with the \Rfunction{writeMgfData} method. Note that the metadata in the original 
-\R object can not be included in the file. See \Rfunction{?writeMgfData} for details.
-
-\paragraph{Quantitation data} Quantitation data can be exported to spreadsheet files with 
-the \Rfunction{write.exprs} method. Feature metadata can be appended to the feature 
-intenstiy values. See \Rfunction{?writeMgfData} for details.
-
-\Robject{MSnSet} instances can also be exported to \texttt{mzTab} files using the 
-\Rfunction{writeMzTabData} function.
-
-\begin{figure}[!htb] %% out
-  \begin{center}
-    \begin{tikzpicture}[node distance = 2cm, auto]
-      % Place nodes
-      \node [input] (raw) {Raw data in an open \texttt{XML} format};
-      \node [input, below of=raw] (mgf) {Peak list in \texttt{mgf} format};
-      \node [fun, right of=mgf, node distance=4cm, yshift=1cm] (writeMgfData) {\Rfunction{writeMgfData}};
-      \node [input, below of=mgf] (spreadsheet) {Quantitation data as a spreadsheet};
-      \node [input, below of=spreadsheet] (mztab) {\texttt{mzTab} format};
-      \node [fun, right of=spreadsheet, node distance=4cm] (writeexprs) {\Rfunction{write.exprs}};
-      \node [fun, right of=mztab, node distance=4cm] (writeMzTabData) {\Rfunction{writeMzTabData}};
-      \node [obj, right of=writeMgfData, yshift=1.25cm, node distance=4cm] (MSnExp) {\Robject{MSnExp}};
-      \node [obj, right of=writeexprs, node distance=4cm] (MSnSet) {\Robject{MSnSet}};
-      \node [fun, below of=MSnExp] (quantify) {\Rfunction{quantify}};
-      % Background
-      \begin{pgfonlayer}{background}
-        \node [fill=yellow!20,rounded corners, draw=black!50, dashed, fit=(MSnExp) (quantify) (MSnSet)] {};
-      \end{pgfonlayer}
-      % Draw edges
-      \draw  (MSnExp) -- (writeMgfData);
-      \draw [->] (writeMgfData) -- (mgf);
-      \draw (MSnSet) -- (writeexprs);
-      \draw [->] (writeexprs) -- (spreadsheet);
-      \draw  (MSnExp) -- (quantify);
-      \draw [->] (quantify) -- (MSnSet);
-      \draw (MSnSet) -- (writeMzTabData);
-      \draw [->] (writeMzTabData) -- (mztab);
-    \end{tikzpicture}
-    \caption{Illustration of \texttt{MSnbase} output capabilities.
-      The white and red boxes represent \R functions/methods and objects respectively. 
-      The blue boxes represent different disk storage formats.}
-    \label{fig:output}
-  \end{center}
-\end{figure}
-
-\clearpage
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% section
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \section{Session information}
-%% \label{sec:sessionInfo}
-
-%% <<label=sessioninfo,results='asis',echo=FALSE,cache=FALSE>>=
-%% toLatex(sessionInfo())
-%% @
-
-\bibliographystyle{plainnat}
-\bibliography{MSnbase}
 

### 2017-06-08 R/quantitation-MS2-isobaric.R
merged master
From: Laurent <lg390@cam.ac.uk>

git-svn-id: file:///home/git/hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/MSnbase@130177 bc3139a8-67e5-0310-9ffc-ced21a209358

@@ -84,7 +84,7 @@ quantify_MSnExp <- function(object, method,
 ## spectra spi in file f using max of peak (irrespective if spectrum
 ## is centroided or profile mode)
 fastquant_max <- function(f, pk, spi, wd = 0.5) {
-    ramp <- openMSfile(f)
+    ramp <- .openMSfile(f)
     on.exit(close(ramp))
     pks <- peaks(ramp, spi)
     if (length(spi) == 1)

### 2017-08-03 R/methods-MSnExp.R
new  readMzIdData and filterIdenticicationDataFrame functions

@@ -270,11 +270,13 @@ setMethod("addIdentificationData", c("MSnExp", "character"),
              desc = "DatabaseDescription",
              pepseq = "sequence",
              key = "spectrumID",
+             rank = "rank",
+             decoy = "isDecoy",
              verbose = isMSnbaseVerbose(),
              ...)
-        .addCharacterIdentificationData(object, id, fcol, icol,
-                                        acc, desc, pepseq, key,
-                                        verbose, ...))
+        .addCharacterIdentificationData(object, id, fcol, icol, acc,
+                                        desc, pepseq, key, rank,
+                                        decoy, verbose, ...))
 
 setMethod("addIdentificationData", c("MSnExp", "mzRident"),
         function(object, id,
@@ -284,11 +286,13 @@ setMethod("addIdentificationData", c("MSnExp", "mzRident"),
                  desc = "DatabaseDescription",
                  pepseq = "sequence",
                  key = "spectrumID",
+                 rank = "rank",
+                 decoy = "isDecoy",                 
                  verbose = isMSnbaseVerbose(),
                  ...)
             .addMzRidentIdentificationData(object, id, fcol, icol,
                                            acc, desc, pepseq, key,
-                                           verbose, ...))
+                                           rank, decoy, verbose, ...))
 
 setMethod("addIdentificationData", c("MSnExp", "mzIDClasses"),
         function(object, id,
@@ -298,21 +302,23 @@ setMethod("addIdentificationData", c("MSnExp", "mzIDClasses"),
                  desc = "description",
                  pepseq = "pepseq",
                  key = "spectrumid",
+                 rank = "rank",
+                 decoy = "isdecoy",
                  verbose = isMSnbaseVerbose(),
                  ...)
             .addMzIDIdentificationData(object, id, fcol, icol, acc,
-                                       desc, pepseq, key, verbose,
-                                       ...))
+                                       desc, pepseq, key, rank, decoy,
+                                       verbose,...))
 
 setMethod("addIdentificationData", c("MSnExp", "data.frame"),
           function(object, id,
                    fcol = c("spectrum.file", "acquisition.number"),
-                   icol, acc, desc, pepseq, key,
-                   verbose = isMSnbaseVerbose(),
-                   ...)
-            .addDataFrameIdentificationData(object, id, fcol, icol, acc,
-                                            desc, pepseq, key, verbose,
-                                            ...))
+                   icol, acc, desc, pepseq, key, rank, decoy,
[…]

### 2017-08-03 R/methods-MSnSet.R

@@ -664,11 +664,13 @@ setMethod("addIdentificationData", c("MSnSet", "character"),
              desc = "DatabaseDescription",
              pepseq = "sequence",
              key = "spectrumID",
+             rank = "rank",
+             decoy = "isDecoy", 
              verbose = isMSnbaseVerbose(),
              ...)
-        .addCharacterIdentificationData(object, id, fcol, icol,
-                                        acc, desc, pepseq, key,
-                                        verbose, ...))
+        .addCharacterIdentificationData(object, id, fcol, icol, acc,
+                                        desc, pepseq, key, rank,
+                                        decoy, verbose, ...))
 
 setMethod("addIdentificationData", c("MSnSet", "mzRident"),
         function(object, id,
@@ -678,11 +680,13 @@ setMethod("addIdentificationData", c("MSnSet", "mzRident"),
                  desc = "DatabaseDescription",
                  pepseq = "sequence",
                  key = "spectrumID",
+                 rank = "rank",
+                 decoy = "isDecoy",
                  verbose = isMSnbaseVerbose(),
                  ...)
             .addMzRidentIdentificationData(object, id, fcol, icol,
                                            acc, desc, pepseq, key,
-                                           verbose, ...))
+                                           rank, decoy, verbose, ...))
 
 setMethod("addIdentificationData", c("MSnSet", "mzIDClasses"),
         function(object, id,
@@ -692,21 +696,23 @@ setMethod("addIdentificationData", c("MSnSet", "mzIDClasses"),
                  desc = "description",
                  pepseq = "pepseq",
                  key = "spectrumid",
+                 rank = "rank",
+                 decoy = "isdecoy", 
                  verbose = isMSnbaseVerbose(),
                  ...)
             .addMzIDIdentificationData(object, id, fcol, icol, acc,
-                                       desc, pepseq, key, verbose,
-                                       ...))
+                                       desc, pepseq, key, rank, decoy,
+                                       verbose, ...))
 
 setMethod("addIdentificationData", c("MSnSet", "data.frame"),
           function(object, id,
                    fcol = c("spectrum.file", "acquisition.number"),
-                   icol, acc, desc, pepseq, key,
-                   verbose = isMSnbaseVerbose(),
-                   ...)
-            .addDataFrameIdentificationData(object, id, fcol, icol, acc,
-                                            desc, pepseq, key, verbose,
-                                            ...))
+                   icol, acc, desc, pepseq, key, rank, decoy,
[…]

### 2017-08-10 README.Rmd
fix link to sticker
link to raw sticker url
add sticker
From: Laurent <lg390@cam.ac.uk>

git-svn-id: file:///home/git/hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/MSnbase@127880 bc3139a8-67e5-0310-9ffc-ced21a209358
add sticker
From: Laurent <lg390@cam.ac.uk>

git-svn-id: file:///home/git/hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/MSnbase@127881 bc3139a8-67e5-0310-9ffc-ced21a209358
200 pixels
From: Laurent <lg390@cam.ac.uk>

git-svn-id: file:///home/git/hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/MSnbase@127883 bc3139a8-67e5-0310-9ffc-ced21a209358
only show github (not B

--- 2016-12-31 README.Rmd
@@
 [![codecov.io](https://codecov.io/github/lgatto/MSnbase/coverage.svg?branch=master)](https://codecov.io/github/lgatto/MSnbase?branch=master)
-[![Bioconductor devel build Status](http://bioconductor.org/shields/build/devel/bioc/MSnbase.svg)](http://bioconductor.org/packages/devel/bioc/html/MSnbase.html) 
-[![Test coverage](https://bioconductor.org/shields/coverage/release/MSnbase.svg)](https://bioconductor.org/shields/coverage/devel/MSnbase.svg)
 
@@
 ```
+
+# The `MSnbase` package
+
+<img align = "right" src="https://raw.githubusercontent.com/Bioconductor/BiocStickers/master/MSnbase/MSnbase.png" height="200">
 

### 2017-08-10 tests/testthat.R
readMSData gets a mode - see issue #233

@@ -12,23 +12,23 @@ f <- msdata::proteomics(full.names = TRUE,
                         pattern = "TMT_Erwinia_1uLSike_Top10HCD_isol2_45stepped_60min_01.mzML.gz")
 tmt_erwinia_in_mem_ms1 <- readMSData(f, msLevel = 1)
 tmt_erwinia_in_mem_ms2 <- readMSData(f, msLevel = 2)
-tmt_erwinia_on_disk <- readMSData2(f)
-tmt_erwinia_on_disk_ms1 <- readMSData2(f, msLevel = 1)
-tmt_erwinia_on_disk_ms2 <- readMSData2(f, msLevel = 2)
+tmt_erwinia_on_disk <- readMSData(f, mode = "onDisk")
+tmt_erwinia_on_disk_ms1 <- readMSData(f, msLevel = 1, mode = "onDisk")
+tmt_erwinia_on_disk_ms2 <- readMSData(f, msLevel = 2, mode = "onDisk")
 
 ## microtofq
 f <- c(system.file("microtofq/MM14.mzML", package = "msdata"),
        system.file("microtofq/MM8.mzML", package = "msdata"))
 microtofq_in_mem_ms1 <- readMSData(f, msLevel = 1)
-microtofq_on_disk_ms1 <- readMSData2(f, msLevel = 1)
-microtofq_on_disk <- readMSData2(f)
+microtofq_on_disk_ms1 <- readMSData(f, msLevel = 1, mode = "onDisk")
+microtofq_on_disk <- readMSData(f, mode = "onDisk")
 
 ## extdata mzML
 f <- dir(system.file(package = "MSnbase", dir = "extdata"),
          full.name = TRUE, pattern = "mzXML$")
 extdata_mzXML_in_mem_ms2 <- readMSData(f, verbose = FALSE, centroided. = FALSE)
-extdata_mzXML_on_disk <- readMSData2(f, centroided. = FALSE)
-extdata_mzXML_on_disk_ms2 <- readMSData2(f, msLevel = 2, centroided. = FALSE)
+extdata_mzXML_on_disk <- readMSData(f, centroided. = FALSE, mode = "onDisk")
+extdata_mzXML_on_disk_ms2 <- readMSData(f, msLevel = 2, centroided. = FALSE, mode = "onDisk")
 
 
 test_check("MSnbase")

### 2017-08-10 tests/testthat/test_OnDiskMSnExp.R

@@ -308,7 +308,7 @@ test_that("chromatogram,OnDiskMSnExp works", {
     file.copy(mzf[2], paste0(tmpd, "b.mzML"))
     mzf <- c(mzf, paste0(tmpd, c("a.mzML", "b.mzML")))
     
-    onDisk <- readMSData2(files = mzf, msLevel. = 1, centroided. = TRUE)
+    onDisk <- readMSData(files = mzf, msLevel. = 1, centroided. = TRUE, mode = "onDisk")
 
     ## Full rt range.
     mzr <- matrix(c(100, 120), nrow = 1)

### 2017-08-11 README.Rmd
Merge branch 'master' into devel
* master: (100 commits)
  use precise with vignette and tests
  use precise
  no tests nor vignettes to try to build on travis
  update pkgdown
  readMSData2 is now deprecated
  fix typo in example
  update pkddown docs
  fix introduced bug
  remove trailing spaces
  Fix signature for pData<-,Chromatogram (issue #244)
  update scripts to set mode when reading data
  add alias from pData<-,MSnSet,data.frame
  readMSData gets a mode - see issue #233
  check MSnSet validity when replacing pData - closes issue #242
  update README and home page
  only show github (

@@
 [**MSnbase**](http://lgatto.github.io/MSnbase/) is an R/Bioconductor
-package that provides infrastructure for plotting, data manipulation
-and processing mass spectrometry and proteomics data. The project was
+package that provides infrastructure for plotting, manipulation and
+processing mass spectrometry and proteomics data. The project was
 started by [Laurent Gatto](http://lgatto.github.io/) in October 2010
 (Mon Oct 4 23:35:23 2010, according to the git log) and has, since
-then, benefited from [various
-contributions](http://lgatto.github.io/msnbase-contribs/).
+then, benefited
+from
+[various contributions](http://lgatto.github.io/msnbase-contribs/), in
+particular [Sebastian Gibb](http://sebastiangibb.de/)
+and [Johannes Rainer](https://github.com/jotsetung/).
 
 The official page is the Bioconductor landing page
-([release](http://www.bioconductor.org/packages/release/bioc/html/MSnbase.html)
-or
-[devel](http://www.bioconductor.org/packages/devel/bioc/html/MSnbase.html)
-versions). The [github page](https://github.com/lgatto/MSnbase) page
-is for sharing, testing, issue tracking and forking/pulling
-purposes. The
-[Bioconductor svn server](https://hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/),
-is the official repository for the working source code.
+([release](http://www.bioconductor.org/packages/release/bioc/html/MSnbase.html) or
+[devel](http://www.bioconductor.org/packages/devel/bioc/html/MSnbase.html) versions). The
+[github page](https://github.com/lgatto/MSnbase) page is for active
+development, issue tracking and forking/pulling
+purposes.
 
@@
 contribute to this package, you should follow the same conventions as
-the rest of the functions whenever it makes sense to do so. Feel free
-to get in touch (preferable opening a
-[github issue](https://github.com/lgatto/MSnbase/issues/)) to discuss
-any suggestions. The
-[`MSnbase` development vignette](http://lgatto.github.io/MSnbase/articles/MSnbase-development.html)
-gives some background on the class infrastructure.
+the rest of the functions. Please do get in touch (preferable opening
+a [github issue](https://github.com/lgatto/MSnbase/issues/)) to
+discuss any
+suggestions. The
+[`MSnbase` development vignette](http://lgatto.github.io/MSnbase/articles/MSnbase-development.html) gives
+some background on the class infrastructure.
 

### 2017-08-11 R/functions-Chromatogram.R

@@ -132,3 +132,27 @@ Chromatogram <- function(rtime = numeric(), intensity = numeric(),
     plot(x = rtime(x), y = intensity(x), main = main, col = col, lty = lty,
          type = type, xlab = xlab, ylab = ylab, ...)
 }
+
+#' @aliases aggregationFun
+#'
+#' @description \code{aggregationFun,aggregationFun<-} get or set the
+#'     aggregation function.
+#' 
+#' @rdname Chromatogram-class
+aggregationFun <- function(object) {
+    if (!is(object, "Chromatogram"))
+        stop("'object' is supposed to be a 'Chromatogram' class")
+    object@aggregationFun
+}
+
+#' @aliases productMz
+#' 
+#' @description \code{productMz} get the mz of the product chromatogram/ion. The
+#'     function returns a \code{numeric(2)} with the lower and upper mz value.
+#' 
+#' @rdname Chromatogram-class
+productMz <- function(object) {
+    if (!is(object, "Chromatogram"))
+        stop("'object' is supposed to be a 'Chromatogram' class")
+    object@productMz
+}

### 2017-08-11 R/functions-mzR.R

@@ -110,3 +110,67 @@ plotMzDelta_list <- function(object,            ## peakLists
         print(p)
     invisible(p)
 }
+
+##' A function to convert the identification data contained in an
+##' \code{mzRident} object to a \code{data.frame}. Each row represents
+##' a scan, which can however be repeated several times if the PSM
+##' matches multiple proteins and/or contains two or more
+##' modifications. To reduce the \code{data.frame} so that rows/scans
+##' are unique and use semicolon-separated values to combine
+##' information pertaining a scan, use
+##' \code{\link[=reduce,data.frame-method]{reduce}}.
+##'
+##' See also the \emph{Tandem MS identification data} section in the
+##' \emph{MSnbase-demo} vignette.
+##' 
+##' @title Coerce identification data to a \code{data.frame}
+##' @param from An object of class \code{mzRident} defined in the
+##'     \code{mzR} package.
+##' @return A \code{data.frame}
+##' @author Laurent Gatto
+##' @name as
+##' @rdname mzRident2dfr
+##' @aliases as.data.frame.mzRident
+##' @examples
+##' ## find path to a mzIdentML file
+##' identFile <- dir(system.file(package = "MSnbase", dir = "extdata"),
+##'                  full.name = TRUE, pattern = "dummyiTRAQ.mzid")
+##' library("mzR")
+##' x <- openIDfile(identFile)
+##' x
+##' as(x, "data.frame")
+setAs("mzRident", "data.frame",
+      function(from) {
+          ## peptide spectrum matching
+          iddf <- factorsAsStrings(psms(from))
+          ## add file raw and mzid provenances      
+          iddf$spectrumFile <- basename(sourceInfo(from))
+          iddf$idFile <- basename(fileName(from))
+          ## add scores
+          scores <- factorsAsStrings(score(from))
+          stopifnot(identical(iddf[, 1], scores[, 1]))
+          iddf <- cbind(iddf, scores[, -1])
+          ## add modification
+          mods <- factorsAsStrings(modifications(from))
+          names(mods)[-1] <- makeCamelCase(names(mods), prefix = "mod")[-1]
+          iddf <- merge(iddf, mods,
+                        by.x = c("spectrumID", "sequence"),
+                        by.y = c("spectrumID",  "modSequence"),
+                        suffixes = c("", ".y"),
+                        all = TRUE, sort = FALSE)
+          iddf[, "spectrumID.y"] <- NULL
[…]

### 2017-08-11 R/methods-Chromatograms.R

@@ -1,6 +1,6 @@
 #' @rdname Chromatograms-class
 #'
-#' @param object For \code{show}: a \code{Chromatograms} object.
+#' @param object a \code{Chromatograms} object.
 setMethod("show", "Chromatograms", function(object) {
     nr <- nrow(object)
     nc <- ncol(object)
@@ -31,6 +31,7 @@ setMethod("show", "Chromatograms", function(object) {
         rownames(out) <- c("", rn)        
         print(out, quote = FALSE, right = TRUE)
     }
+    cat("phenoData with", length(varLabels(object@phenoData)), "variables\n")
 })
 
 #' @rdname Chromatograms-class
@@ -40,7 +41,12 @@ setMethod("show", "Chromatograms", function(object) {
 #'     can be replaced using e.g. \code{x[1, 1] <- value} where \code{value}
 #'     has to be a \code{Chromatogram} object or a \code{list} of such objects.
 #'
-#' @param x For \code{[}: the \code{Chromatograms} object to subset.
+#' @note Subsetting with \code{[} will always return a \code{Chromatograms}
+#'     object (with the exception of extracting a single element)
+#'     unless \code{drop = TRUE} is specified. This is different from the
+#'     default subsetting behaviour of \code{matrix}-like objects.
+#' 
+#' @param x For all methods: a \code{Chromatograms} object.
 #' 
 #' @param i For \code{[}: \code{numeric}, \code{logical} or \code{character}
 #'     defining which row(s) to extract.
@@ -49,16 +55,28 @@ setMethod("show", "Chromatograms", function(object) {
 #'     defining which columns(s) to extract.
 #'
 #' @param drop For \code{[}: \code{logical(1)} whether to drop the
-#'     dimensionality of the returned object (if possible).
+#'     dimensionality of the returned object (if possible). The default is
+#'     \code{drop = FALSE}, i.e. each subsetting returns a \code{Chromatograms}
+#'     object (or a \code{Chromatogram} object if a single element is
+#'     extracted).
 #'
 #' @return For \code{[}: the subset of the \code{Chromatograms} object. If a
 #'     single element is extracted (e.g. if \code{i} and \code{j} are of length
-#'     1) a \code{\link{Chromatogram}} object is returned. Otherwise a
-#'     \code{list} of \code{\link{Chromatogram}} objects is returned, except
-#'     if \code{drop = FALSE} in which case \emph{always} a \code{Chromatograms}
[…]

### 2017-08-11 R/nav.R

@@ -19,7 +19,7 @@
 ##' @author Laurent Gatto
 ##' @examples
 ##' f <- msdata::proteomics(full.names = TRUE, pattern = "MS3")
-##' x <- readMSData2(f, centroided = c(FALSE, TRUE, FALSE))
+##' x <- readMSData(f, centroided. = c(FALSE, TRUE, FALSE), mode = "onDisk")
 ##' (sp <- which(msLevel(x) == 3)[2]) ## 2nd MS3 spectrum
 ##' x[[sp]] ## curent MS3
 ##' MSnbase:::nextMS(sp, x) ## next MS3

### 2017-08-11 R/quantitation-MS2-isobaric.R

@@ -41,8 +41,8 @@ quantify_MSnExp <- function(object, method,
     ## Creating new phenoData slot or creating one
     .phenoData <- new("AnnotatedDataFrame",
                       data = data.frame(mz = reporters@mz,
-                          reporters = reporters@name,
-                          row.names = reporters@reporterNames))
+                                        reporters = reporters@name,
+                                        row.names = reporters@reporterNames))
     if (nrow(pData(object)) > 0) {
         if (nrow(pData(object)) == length(reporters)) {
             .phenoData <- combine(phenoData(object), .phenoData)
@@ -182,6 +182,7 @@ quantify_OnDiskMSnExp_max <- function(object, reporters,
                featureData = featureData(object),
                phenoData = .phenoData)
     fData(ans)$reporterMzs <- mzs
+    ans@processingData <- object@processingData
     ans <- logging(ans, paste0("Fast ", names(reporters),
                                " quantitation by max"))
     if (validObject(ans)) ans

### 2017-08-11 inst/scripts/torture.R

@@ -12,13 +12,353 @@ for (i in 1:10000) {
     sp <- spectra(res)
 }
 
-## readMSData2
+## readMSData onDisk
 for (i in 1:7500) {
     if (i %% 200 == 0)
         cat(i, "\n")
-    res <- readMSData2(f[1], verbose = FALSE)
+    res <- readMSData(f[4], verbose = FALSE, mode = "onDisk")
     sp <- res[[6]]
-    res <- readMSData2(f[2], verbose = FALSE)
+    res <- readMSData(f[4], verbose = FALSE, mode = "onDisk")
     sp <- spectra(res)
 }
+## OK without gc()
 
+## spectrapply2
+res <- readMSData(f[4], verbose = FALSE, mode = "onDisk")
+for (i in 1:10000) {
+    if (i %% 200 == 0)
+        cat(i, "\n")
+    ## sp <- res[[6]]
+    sp <- spectrapply(res)
+    sp <- spectrapply(res, FUN = mz)
+    tmp <- filterRt(res, rt = c(1108, 1109))
+    sp <- spectrapply(tmp)
+}
+## OK without gc()
+
+## spectrapply2
+res <- readMSData(f[4], verbose = FALSE, mode = "onDisk")
+for (i in 1:10000) {
+    if (i %% 200 == 0)
+        cat(i, "\n")
+    ## sp <- res[[6]]
+    res <- readMSData(f[4], verbose = FALSE, mode = "onDisk")
+    sp <- MSnbase:::spectrapply2(res)
+    sp <- MSnbase:::spectrapply2(res, FUN = mz)
+    tmp <- filterRt(res, rt = c(1108, 1109))
+    sp <- MSnbase:::spectrapply2(tmp)
+}
+## OK without gc()
+
+## Issue #170, i.e. random "[MSData::Spectrum::getMZIntensityPairs()] Sizes do
+## not match." with the pwiz backend (segfault with ramp). The solution was to
+## read call mzR::header() before the mzR::peaks, but this significantly slows
+## things down.
+
+library(MSnbase)
+
+torturing <- function(x) {
+    tmp <- readMSData(x, msLevel. = 1, mode = "onDisk")
+    register(SerialParam())
+    for (i in 1:10) {
+        cat("--- ", i, " ---", "\n")
+        cat("first spectrapply\n")
+        sp <- MSnbase::spectrapply(tmp, FUN = function(z) {max(mz(z))})
+        ## sp has a size of 21.9GB!!!
+        rm(sp)
[…]

### 2017-08-11 tests/testthat/test_methods-Chromatograms.R

@@ -9,40 +9,91 @@ test_that("[,Chromatograms works", {
     ints <- abs(rnorm(40, sd = 34))
     ch3 <- Chromatogram(rtime = 1:length(ints), ints)
     chrs <- Chromatograms(list(ch, ch1, ch2, ch3), nrow = 2)
-    colnames(chrs) <- c("a", "b")
     
-    ## Subset using indices
-    ## extract single element
+    ## o Subset using indices
     expect_true(is(chrs[1, 1], "Chromatogram"))
-    expect_true(is(chrs[1, 1, drop = FALSE], "Chromatograms"))
     expect_equal(chrs[1, 2], ch2)
-    ## extract a row
-    expect_equal(chrs[1, ], list(a = ch, b = ch2))
-    expect_equal(unname(chrs[1, , drop = FALSE]),
-                 Chromatograms(list(ch, ch2), nrow = 1))    
-    ## extract a column
-    expect_equal(chrs[, 2], list(ch2, ch3))
-    expect_equal(unname(chrs[, 2, drop = FALSE]),
-                 Chromatograms(list(ch2, ch3), ncol = 1))
-
-    ## Subset using logical
+    ##   extract a row
+    expect_equal(chrs[1, , drop = TRUE], list(`1` = ch, `2` = ch2))
+    expect_equal(chrs[1, , drop = FALSE], Chromatograms(list(ch, ch2), nrow = 1))
+    ##   Test the default
+    expect_equal(chrs[1, ], Chromatograms(list(ch, ch2), nrow = 1))    
+    ##   extract a column
+    expect_equal(chrs[, 2, drop = TRUE], list(ch2, ch3))
+    res <- chrs[, 2, drop = FALSE]
+    res_exp <- Chromatograms(list(ch2, ch3), ncol = 1,
+                             dimnames = list(NULL, "2"))
+    ## Have to re-place the rownames of pheno data othewise we compare numeric
+    ## against character
+    rownames(pData(res)) <- rownames(pData(res))
+    expect_equal(res, res_exp)
+    ##   Repeat with colnames:
+    colnames(chrs) <- c("a", "b")
+    expect_true(is(chrs[1, 1], "Chromatogram"))
+    expect_equal(chrs[1, 2], ch2)
[…]

### 2017-08-11 tests/testthat/test_nav.R

@@ -5,7 +5,7 @@ f <- proteomics(full.names = TRUE, pattern = "MS3TMT10")
 
 
 test_that("nextMS and prevMS", {
-    x <- readMSData2(f)
+    x <- readMSData(f, mode = "onDisk")
     sp <- "F1.S009"
     expect_identical(acquisitionNum(x[[sp]]), 32926L)
     expect_identical(acquisitionNum(MSnbase:::nextMS(sp, x, 3L)), 32929L)

### 2017-08-11 tests/testthat/test_readMSData2.R

@@ -1,18 +1,18 @@
-context("readMSData2")
+context("readMSData onDisk mode")
 
 test_that("msLevel set correctly", {
     f <- msdata::proteomics(full.names = TRUE,
                             pattern = "TMT_Erwinia_1uLSike_Top10HCD_isol2_45stepped_60min_01.mzML.gz")
-    x <- readMSData2(f, verbose = FALSE)
+    x <- readMSData(f, verbose = FALSE, mode = "onDisk")
     expect_equivalent(centroided(x), rep(NA, length(x)))
-    x <- readMSData2(f, centroided = TRUE, verbose = FALSE)
+    x <- readMSData(f, centroided = TRUE, verbose = FALSE, mode = "onDisk")
     expect_equivalent(centroided(x), rep(TRUE, length(x)))
-    x <- readMSData2(f, centroided = FALSE, verbose = FALSE)
+    x <- readMSData(f, centroided = FALSE, verbose = FALSE, mode = "onDisk")
     expect_equivalent(centroided(x), rep(FALSE, length(x)))
-    x <- readMSData2(f, centroided = c(FALSE, TRUE), verbose = FALSE)
+    x <- readMSData(f, centroided = c(FALSE, TRUE), verbose = FALSE, mode = "onDisk")
     expect_true(all(centroided(filterMsLevel(x, msLevel = 2))))
-    expect_true(all(!centroided(filterMsLevel(x, msLevel = 1))))
-    x2 <- readMSData2(f, centroided = c(FALSE, TRUE, NA, NA), verbose = FALSE)
+    expect_true(all(!centroided(filterMsLevel(x, msLevel = 1))))    
+    x2 <- readMSData(f, centroided = c(FALSE, TRUE, NA, NA), verbose = FALSE, mode = "onDisk")
     expect_identical(centroided(x), centroided(x2))
 })
 
@@ -36,8 +36,8 @@ test_that("msLevel set correctly", {
 
 mzf <- .getMzMLFiles(TRUE)[1:2]
 
-## Load the data with readMSData2
-odmse <- readMSData2(files = mzf, centroided = TRUE)
+## Load the data with readMSData mode = onDisk
+odmse <- readMSData(files = mzf, centroided = TRUE, mode = "onDisk")
 
 test_that("Constructor performance and test for MS1 only", {
     featDat <- fData(odmse)
@@ -55,7 +55,7 @@ test_that("Constructor performance and test for MSn", {
     ## Get the test data file.
[…]

### 2017-08-21 docs/articles/MSnbase-development.Rmd
update docs

---
title: "A short introduction to *MSnbase* development"
author: 
- name: Laurent Gatto
  affiliation: Computational Proteomics Unit, Cambridge, UK.
- name: Johannes Rainer
  affiliation: Center for Biomedicine, EURAC, Bolzano, Italy.
- name: Sebastian Gibb
  affiliation: Department of Anesthesiology and Intensive Care, University Medicine Greifswald, Germany.
package: MSnbase
abstract: >
  This vignette describes the classes implemented in \Biocpkg{MSnbase}
  package.  It is intended as a starting point for developers or users
  who would like to learn more or further develop/extend mass
  spectrometry and proteomics data structures.
bibliography: MSnbase.bib
output:
  BiocStyle::html_document:
    toc_float: true
vignette: >
  %\VignetteIndexEntry{A short introduction to `MSnbase` development}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteKeywords{Mass Spectrometry, Proteomics, Infrastructure }
  %\VignetteEncoding{UTF-8}
---

```{r environment, echo=FALSE}
suppressPackageStartupMessages(library("MSnbase"))
suppressPackageStartupMessages(library("BiocStyle"))
```

```{r include_forword, echo=FALSE, results="asis"}
cat(readLines("./Foreword.md"), sep = "\n")
```

```{r include_bugs, echo=FALSE, results="asis"}
cat(readLines("./Bugs.md"), sep = "\n")
```

# Introduction

This document is not a replacement for the individual manual pages,
that document the slots of the `r Biocpkg("MSnbase")` classes. It is a
centralised high-level description of the package design.

`r Biocpkg("MSnbase")` aims at being compatible with the 
`r Biocpkg("Biobase")` infrastructure [@Gentleman2004].  Many meta data
structures that are used in `eSet` and associated classes are also
used here. As such, knowledge of the *Biobase development and the new
eSet* vignette would be beneficial; the vignette can directly be
accessed with `vignette("BiobaseDevelopment", package="Biobase")`.

The initial goal is to use the `r Biocpkg("MSnbase")` infrastructure
for MS2 labelled (iTRAQ [@Ross2004] and TMT [@Thompson2003]) and
label-free (spectral counting, index and abundance) quantitation
- see the documentation for the `quantify` function for details. The
infrastructure is currently extended to support a wider range of
technologies, including metabolomics.

# `r Biocpkg("MSnbase")` classes

All classes have a `.__classVersion__` slot, of class `Versioned` from
the `r Biocpkg("Biobase")` package. This slot documents the class
version for any instance to be used for debugging and object update
purposes. Any change in a class implementation should trigger a
version change.

## `pSet`: a virtual class for raw mass spectrometry data and meta data

This virtual class is the main container for mass spectrometry data,
i.e spectra, and meta data. It is based on the `eSet` implementation
for genomic data. The main difference with `eSet` is that the
`assayData` slot is an environment containing any number of
`Spectrum` instances (see the [`Spectrum` section](#Spectrum)).

One new slot is introduced, namely `processingData`, that contains one
`MSnProcess` instance (see the [`MSnProcess` section](#MSnProcess)).
and the `experimentData` slot is now expected to contain `MIAPE` data.
The `annotation` slot has not been implemented, as no prior feature
annotation is known in shotgun proteomics.

```{r pSet}
getClass("pSet")
```

## `MSnExp`: a class for MS experiments

`MSnExp` extends `pSet` to store MS experiments.  It does not add any
new slots to `pSet`. Accessors and setters are all inherited from
`pSet` and new ones should be implemented for `pSet`.  Methods that
manipulate actual data in experiments are implemented for
`MSnExp` objects.

```{r MSnExp}
getClass("MSnExp")
```

## `OnDiskMSnExp`: a on-disk implementation of the `MSnExp` class

The `OnDiskMSnExp` class extends `MSnExp` and inherits all of its
functionality but is aimed to use as little memory as possible based
on a balance between memory demand and performance. Most of the
spectrum-specific data, like retention time, polarity, total ion
current are stored within the object's `featureData` slot. The actual
M/Z and intensity values from the individual spectra are, in contrast
to `MSnExp` objects, not kept in memory (in the `assayData` slot), but
are fetched from the original files on-demand. Because mzML files are
indexed, using the `r Biocpkg("mzR")` package to read the relevant
spectrum data is fast and only moderately slower than for in-memory
`MSnExp`^[The *benchmarking* vignette compares data size and operation speed of the two implementations.].

To keep track of data manipulation steps that are applied to spectrum
data (such as performed by methods `removePeaks` or `clean`) a *lazy
execution* framework was implemented. Methods that manipulate or
subset a spectrum's M/Z or intensity values can not be applied
directly to a `OnDiskMSnExp` object, since the relevant data is not
kept in memory. Thus, any call to a processing method that changes or
subset M/Z or intensity values are added as `ProcessingStep` items to
the object's `spectraProcessingQueue`. When the spectrum data is then
queried from an `OnDiskMSnExp`, the spectra are read in from the file
and all these processing steps are applied on-the-fly to the spectrum
data before being returned to the user.

The operations involving extracting or manipulating spectrum data are
applied on a per-file basis, which enables parallel processing. Thus,
all corresponding method implementations for `OnDiskMSnExp` objects
have an argument `BPPARAM` and users can set a `PARALLEL_THRESH`
option flag^[see `?MSnbaseOptions` for details.] that enables to
define how and when parallel processing should be performed (using the
`r Biocpkg("BiocParallel")` package).

Note that all data manipulations that are not applied to M/Z or
intensity values of a spectrum (e.g. sub-setting by retention time
etc) are very fast as they operate directly to the object's
`featureData` slot.

```{r OnDiskMSnExp}
getClass("OnDiskMSnExp")
```

The distinction between `MSnExp` and `OnDiskMSnExp` is often not
explicitly stated as it should not matter, from a user's perspective,
which data structure they are working with, as both behave in
equivalent ways. Often, they are referred to as *in-memory* and
*on-disk* `MSnExp` implementations.

## `MSnSet`: a class for quantitative proteomics data

This class stores quantitation data and meta data after running
`quantify` on an `MSnExp` object or by creating an `MSnSet` instance
from an external file, as described in the *MSnbase-io* vignette and
in `?readMSnSet`, `readMzTabData`, etc. The quantitative data is in
form of a *n* by *p* matrix, where *n* is the number of
features/spectra originally in the `MSnExp` used as parameter in
`quantify` and *p* is the number of reporter ions. If read from an
external file, *n* corresponds to the number of features (protein
groups, proteins, peptides, spectra) in the file and $p$ is the number
of columns with quantitative data (samples) in the file.

This prompted to keep a similar implementation as the `ExpressionSet`
class, while adding the proteomics-specific annotation slot introduced
in the `pSet` class, namely `processingData` for objects of class
`MSnProcess`.

```{r MSnSet}
getClass("MSnSet")
```

The `MSnSet` class extends the virtual `eSet` class to provide
compatibility for `ExpressionSet`-like behaviour.  The experiment
meta-data in `experimentData` is also of class `MIAPE` .  The
`annotation` slot, inherited from `eSet` is not used. As a result, it
is easy to convert `ExpressionSet` data from/to `MSnSet` objects with
the coersion method `as`.

```{r as}
data(msnset)
class(msnset)
class(as(msnset, "ExpressionSet"))

data(sample.ExpressionSet)
class(sample.ExpressionSet)
class(as(sample.ExpressionSet, "MSnSet"))
```

## `MSnProcess`: a class for logging processing meta data {#MSnProcess}

This class aims at recording specific manipulations applied to
`MSnExp` or `MSnSet` instances. The `processing`
slot is a `character` vector that describes major
processing. Most other slots are of class `logical` that
indicate whether the data has been centroided, smoothed, \ldots
although many of the functionality is not implemented yet.  Any new
processing that is implemented should be documented and logged here.

It also documents the raw data file from which the data originates
(`files` slot) and the `r Biocpkg("MSnbase")` version that was in
use when the `MSnProcess` instance, and hence the
`MSnExp`/`MSnSet` objects, were originally created.

```{r MSnProcess}
getClass("MSnProcess")
```

## `MIAPE`: Minimum Information About a Proteomics Experiment

The Minimum Information About a Proteomics Experiment
[@Taylor2007; @Taylor2008] `MIAPE` class describes the experiment,
including contact details, information about the mass spectrometer and
control and analysis software.

```{r MIAPE}
getClass("MIAPE")
```

## `Spectrum`  *et al.*: classes for MS spectra {#Spectum}

`Spectrum` is a virtual class that defines common attributes to all
types of spectra. MS1 and MS2 specific attributes are defined in the
`Spectrum1` and `Spectrum2` classes, that directly extend `Spectrum`.

```{r Spectrum}
getClass("Spectrum")
```

```{r Spectrum1}
getClass("Spectrum1")
```

```{r Spectrum2}
getClass("Spectrum2")
```

## `ReporterIons`: a class for isobaric tags

The iTRAQ and TMT (or any other peak of interest) are implemented
`ReporterIons` instances, that essentially defines an expected MZ
position for the peak and a width around this value as well a names
for the reporters.

```{r ReporterIons}
getClass("ReporterIons")
```

## `NAnnotatedDataFrame`: multiplexed `AnnotatedDataFrame`s

The simple expansion of the `AnnotatedDataFrame` classes adds the
`multiplex` and `multiLabel` slots to document the number and names of
multiplexed samples.

```{r NAnnotatedDF}
getClass("NAnnotatedDataFrame")
```

## `Chromatogram` and `Chromatograms`: classes to handle chromatographic data

The `Chromatogram` class represents chromatographic MS data, i.e. retention time
and intensity duplets for one file/sample. The `Chromatograms` class allows to
arrange multiple `Chromatogram` instances in a two-dimensional grid, with
columns supposed to represent different samples and rows two-dimensional areas
in the plane spanned by the m/z and retention time dimensions from which the
intensities are extracted (e.g. an extracted ion chromatogram for a specific
ion). The `Chromatograms` class extends the base `matrix` class. `Chromatograms`
objects can be extracted from an `MSnExp` or `OnDiskMSnExp` object using the
`chromatogram` method.

```{r Chromatogram}
getClass("Chromatogram")
```

```{r Chromatograms}
getClass("Chromatograms")
```

## Other classes

### Lists of `MSnSet` instances {-}

When several `MSnSet` instances are related to each other and should
be stored together as different objects, they can be grouped as a list
into and `MSnSetList` object. In addition to the actual `list` slot,
this class also has basic logging functionality and enables iteration
over the `MSnSet` instances using a dedicated `lapply` methods.

```{r msl}
getClass("MSnSetList")
```

# Miscellaneous

#### Unit tests {-}

`r Biocpkg("MSnbase")` implements unit tests with the
`r CRANpkg("testthat")` package.

#### Processing methods {-}

Methods that process raw data, i.e. spectra should be implemented for
`Spectrum` objects first and then `eapply`ed (or similar) to the
`assayData` slot of an `MSnExp` instance in the specific method.

# Session information

```{r si}
sessionInfo()
```

# References

### 2017-08-21 README.Rmd
update readme

@@
 
+To get an overview of the package, see
+the
+[*MSnbase-demo*](https://lgatto.github.io/MSnbase/articles/MSnbase-demo.html) vignette. More
+vignettes are available in the *Articles* tab.
+
 ## Installation
@@
 
-General questions should be asked on the
-[Bioconductor support forum](https://support.bioconductor.org/), using
-`MSnbase` to tag the question. Feel also free to open a GitHub
-[issue](https://github.com/lgatto/MSnbase/issues).
+General questions should be asked on
+the [Bioconductor support forum](https://support.bioconductor.org/),
+using `MSnbase` to tag the question. Feel also free to open a
+GitHub [issue](https://github.com/lgatto/MSnbase/issues), in
+particular for bug reports.
 

### 2017-09-07 docs/articles/MSnbase-io.Rmd

---
title: "MSnbase IO capabilities"
author: 
- name: Laurent Gatto
  affiliation: Computational Proteomics Unit, Cambridge, UK.
package: MSnbase
abstract: >
  This vignette describes *MSnbase*'s input and output capabilities.
bibliography: MSnbase.bib
output:
  BiocStyle::html_document:
    toc_float: true
vignette: >
  %\VignetteIndexEntry{MSnbase IO capabilities}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteKeywords{Mass Spectrometry, Proteomics, Infrastructure }
  %\VignetteEncoding{UTF-8}
---

```{r env, echo=FALSE}
suppressPackageStartupMessages(library("BiocStyle"))
suppressPackageStartupMessages(library("MSnbase"))
suppressPackageStartupMessages(library("pRolocdata"))
```

```{r include_forword, echo=FALSE, results="asis"}
cat(readLines("./Foreword.md"), sep = "\n")
```

```{r include_bugs, echo=FALSE, results="asis"}
cat(readLines("./Bugs.md"), sep = "\n")
```

# Overview

`r Biocpkg("MSnbase")`'s aims are to facilitate the reproducible
analysis of mass spectrometry data within the R environment, from raw
data import and processing, feature quantification, quantification and
statistical analysis of the results [@Gatto2012].  Data import
functions for several formats are provided and intermediate or final
results can also be saved or exported.  These capabilities are
presented below.

# Data input

#### Raw data {-}

Data stored in one of the published `XML`-based formats. i.e. `mzXML`
[@Pedrioli2004], `mzData` [@Orchard2007] or `mzML` [@Martens2010], can
be imported with the `readMSData` method, which makes use of the 
`r Biocpkg("mzR")` package to create `MSnExp` objects.  The files can be
in profile or centroided mode.  See `?readMSData` for details.

#### Peak lists {-}

Peak lists in the `mgf`
format^[http://www.matrixscience.com/help/data_file_help.html]
can be imported using the `readMgfData`.  In this case, the peak data
has generally been pre-processed by other software.  See
`?readMgfData` for details.

#### Quantitation data {-}

Third party software can be used to generate quantitative data and
exported as a spreadsheet (generally comma or tab separated format).
This data as well as any additional meta-data can be imported with the
`readMSnSet` function. See `?readMSnSet` for details.

`r Biocpkg("MSnbase")` also supports the `mzTab`
format^[https://github.com/HUPO-PSI/mzTab], a light-weight,
tab-delimited file format for proteomics data developed within the
Proteomics Standards Initiative (PSI).  `mzTab` files can be read into
R with `readMzTabData` to create and `MSnSet` instance.

![*MSnbase* input capabilities.  The white and red boxes represent R functions/methods and objects respectively.  The blue boxes represent different disk storage formats.](./Figures/MSnbase-io-in.png)

# Data output

#### RData files {-}

R objects can most easily be stored on disk with the `save` function.
It creates compressed binary images of the data representation that
can later be read back from the file with the `load` function.

#### Peak lists {-}

`MSnExp` instances as well as individual spectra can be written as
`mgf` files with the `writeMgfData` method. Note that the meta-data in
the original R object can not be included in the file. See
`?writeMgfData` for details.

#### Quantitation data {-}

Quantitation data can be exported to spreadsheet files with the
`write.exprs` method. Feature meta-data can be appended to the feature
intensity values. See `?writeMgfData` for details.

**Deprecated** `MSnSet` instances can also be exported to `mzTab`
files using the `writeMzTabData` function.

![*MSnbase* output capabilities. The white and red boxes represent R functions/methods and objects respectively. The blue boxes represent different disk storage formats.](./Figures/MSnbase-io-out.png)

# Creating `MSnSet` from text spread sheets

This section describes the generation of `MSnSet` objects using data
available in a text-based spreadsheet. This entry point into R and
`r Biocpkg("MSnbase")` allows to import data processed by any of the
third party mass-spectrometry processing software available and
proceed with data exploration, normalisation and statistical analysis
using functions available in \R and the numerous Bioconductor
packages.

## A complete work flow

The following section describes a work flow that uses three input
files to create the `MSnSet`. These files respectively describe the
quantitative expression data, the sample meta-data and the feature
meta-data.  It is taken from the `r Biocpkg("pRoloc")` tutorial and
uses example files from the `r Biocpkg("pRolocdat")` package.

We start by describing the `csv` to be used as input using the
`read.csv` function.

```{r readCsvData0}
## The original data for replicate 1, available
## from the pRolocdata package
f0 <- dir(system.file("extdata", package = "pRolocdata"), 
          full.names = TRUE, 
          pattern = "pr800866n_si_004-rep1.csv")
csv <- read.csv(f0)
```

The three first lines of the original spreadsheet, containing the data
for replicate one, are illustrated below (using the function
`head`). It contains `r nrow(csv)` rows (proteins) and `r ncol(csv)`
columns, including protein identifiers, database accession numbers,
gene symbols, reporter ion quantitation values, information related to
protein identification, ...

```{r showOrgCsv}
head(csv, n=3)
```

Below read in turn the spread sheets that contain the quantitation
data (`exprsFile.csv`), feature meta-data (`fdataFile.csv`) and sample
meta-data (`pdataFile.csv`).

```{r readCsvData1}
## The quantitation data, from the original data
f1 <- dir(system.file("extdata", package = "pRolocdata"), 
          full.names = TRUE, pattern = "exprsFile.csv")
exprsCsv <- read.csv(f1)
## Feature meta-data, from the original data
f2 <- dir(system.file("extdata", package = "pRolocdata"), 
          full.names = TRUE, pattern = "fdataFile.csv")
fdataCsv <- read.csv(f2)
## Sample meta-data, a new file
f3 <- dir(system.file("extdata", package = "pRolocdata"), 
          full.names = TRUE, pattern = "pdataFile.csv")
pdataCsv <- read.csv(f3)
```

`exprsFile.csv` contains the quantitation (expression) data for the
`r nrow(exprsCsv)` proteins and 4 reporter tags.

```{r showExprsFile}
head(exprsCsv, n = 3)
```

`fdataFile.csv` contains meta-data for the `r nrow(fdataCsv)`
features (here proteins).

```{r showFdFile}
head(fdataCsv, n = 3)
```

`pdataFile.csv` contains samples (here fractions) meta-data. This
simple file has been created manually.

```{r showPdFile}
pdataCsv
```

The self-contained `MSnSet` can now easily be generated using the
`readMSnSet` constructor, providing the respective `csv` file names
shown above and specifying that the data is comma-separated (with `sep
= ","`). Below, we call that object `res` and display its content.

```{r makeMSnSet}
library("MSnbase")
res <- readMSnSet(exprsFile = f1,
                  featureDataFile = f2,
                  phenoDataFile = f3,
                  sep = ",")
res
```

### The `MSnSet` class

Although there are additional specific sub-containers for additional
meta-data (for instance to make the object MIAPE compliant), the
feature (the sub-container, or slot `featureData`) and sample (the
`phenoData` slot) are the most important ones. They need to meet the
following validity requirements (see figure below):

- the number of row in the expression/quantitation data and feature
  data must be equal and the row names must match exactly, and

- the number of columns in the expression/quantitation data and number
  of row in the sample meta-data must be equal and the column/row
  names must match exactly.

A detailed description of the `MSnSet` class is available by typing
`?MSnSet` in the R console.

![Dimension requirements for the respective expression, feature and sample meta-data slots.](./Figures/msnset.png)

The individual parts of this data object can be accessed with their respective accessor methods:

- the quantitation data can be retrieved with `exprs(res)`,
- the feature meta-data with `fData(res)` and 
- the sample meta-data with `pData(res)`.

## A shorter work flow

The `readMSnSet2` function provides a simplified import workforce.  It
takes a single spreadsheet as input (default is `csv`) and extract the
columns identified by `ecol` to create the expression data, while the
others are used as feature meta-data. `ecol` can be a `character` with
the respective column labels or a numeric with their indices. In the
former case, it is important to make sure that the names match
exactly. Special characters like `'-'` or `'('` will be transformed by
R into `'.'` when the `csv` file is read in.  Optionally, one can also
specify a column to be used as feature names.  Note that these must be
unique to guarantee the final object validity.

```{r readMSnSet2}
ecol <- paste("area", 114:117, sep = ".")
fname <- "Protein.ID"
eset <- readMSnSet2(f0, ecol, fname)
eset
```

The `ecol` columns can also be queried interactively from R using the
`getEcols` and `grepEcols` function. The former return a character
with all column names, given a splitting character, i.e. the
separation value of the spreadsheet (typically `","` for `csv`, `"\t"`
for `tsv`, ...). The latter can be used to grep a pattern of interest
to obtain the relevant column indices.

```{r ecols}
getEcols(f0, ",")
grepEcols(f0, "area", ",")
e <- grepEcols(f0, "area", ",")
readMSnSet2(f0, e)
```

The `phenoData` slot can now be updated accordingly using the
replacement functions `phenoData<-` or `pData<-` (see `?MSnSet` for
details).

# Session information

```{r}
sessionInfo()
```

# References {-}

### 2018-02-08 R/combineFeatures.R
Quantification using QuantificationParam

@@ -22,6 +22,15 @@ combineFeatures <- function(object,
         stopifnot(fcol %in% fvarLabels(object))
         groupBy <- fData(object)[, fcol]
     }
+
+    ## Handle missing values in groupBy
+    if (anyNA(groupBy)) {
+        msg <- paste0(sum(is.na(groupBy)), " out of ", length(groupBy),
+                      " values are missing in the 'groupBy' argument. ",
+                      "Either remove these features or assign a dummy category.")
+        stop(paste0(strwrap(msg), sep = "\n  "))
+    }
+    
     if (is.list(groupBy)) {
         if (length(groupBy) != nrow(object))
             stop("'length(groupBy)' must be equal to 'nrow(object)': ",
@@ -110,12 +119,10 @@ combineFeaturesV <- function(object,   ## MSnSet
                                                   verbose = verbose,
                                                   ...))
     }
-    ## takes the first occurences and omit NAs
-    gbsel <- !duplicated(groupBy) & !is.na(groupBy)
-    fdata <- fData(object)[gbsel, , drop = FALSE]
+    ## ordering fdata according to groupBy factor
+    fdata <- fData(object)[!duplicated(groupBy), , drop = FALSE]
     ## ordering fdata according to groupBy factor    
-    fdata <- fdata[order(unique(groupBy[gbsel])), ,
-                   drop = FALSE] 
+    fdata <- fdata[order(unique(groupBy)), , drop = FALSE] 
     rownames(matRes) <- rownames(fdata)
     colnames(matRes) <- sampleNames(object)
     if (cv)

### 2018-02-08 R/methods-MSnExp.R

@@ -152,6 +152,12 @@ setMethod("trimMz",
               return(filterMz(object, mz = mzlim, msLevel.))
           })
 
+setMethod("quantify", c("MSnExp", "QuantitationParam"),
+          function(object, method,
+                   BPPARAM,
+                   verbose = isMSnbaseVerbose(),
+                   ...) quantify2(object, method, BPPARAM, verbose, ...))
+
 setMethod("quantify",
           signature = c("MSnExp", "character"),
           function(object,

### 2018-02-08 R/methods-QuantitationParam.R

@@ -33,6 +33,7 @@ IsobaricTagging <- function(reporters,
 ##' @param pepseq See class slot.
 ##' @param dbaccess See class slot.
 ##' @param plength See class slot.
+##' @param nprot See class slot.
 SpectralCounting <- function(method =
                                  c("count", "tic",
                                    "SI", "SIgi",
@@ -40,15 +41,16 @@ SpectralCounting <- function(method =
                                    "NSAF"),
                              pepseq = "sequence", 
                              dbaccess = "DatabaseAccess",
-                             plength = "DBseqLength") {
+                             plength = "DBseqLength",
+                             nprot = "nprot") {
     method <- match.arg(method)
     .QuantitationParam(msLevel = 2L,
                        method = method,
                        name = "SpectralCounting",
                        pepseq = pepseq,     ## needed for count
                        dbaccess = dbaccess, ## needed for SI* and *SAF
-                       plength = plength)   ## needed for SI* and *SAF
-
+                       plength = plength,   ## needed for SI* and *SAF
+                       nprot = nprot)       ## needed for SI* and *SAF
 }
 
 ##' @rdname QuantitationParam-class

### 2018-07-18 README.Rmd
synonymous changes

@@
 (Mon Oct 4 23:35:23 2010, according to the git log) and has, since
-then, benefited
-from
+then, benefited from
 [various contributions](http://lgatto.github.io/msnbase-contribs/), in
@@
 [github page](https://github.com/lgatto/MSnbase) page is for active
-development, issue tracking and forking/pulling
-purposes.
+development, issue tracking and forking/pulling purposes.
 
-To get an overview of the package, see
-the
+To get an overview of the package, see the
 [*MSnbase-demo*](https://lgatto.github.io/MSnbase/articles/MSnbase-demo.html) vignette. More
@@
 ```{r install, eval = FALSE}
-library("BiocInstaller")
-biocLite("MSnbase")
+install.packages("BiocManager")
+
 ```
@@
 If you need the github version (not recommended unless you know what
-you are doing)
+you are doing), use
 
 ```{r installgh, eval = FALSE}
-biocLite("lgatto/MSnbase")
+BiocManager::install("lgatto/MSnbase")
 ```
@@
 a [github issue](https://github.com/lgatto/MSnbase/issues/)) to
-discuss any
-suggestions. The
+discuss any suggestions. The
 [`MSnbase` development vignette](http://lgatto.github.io/MSnbase/articles/MSnbase-development.html) gives

### 2019-04-08 README.Rmd

@@
 install.packages("BiocManager")
-
+BiocManager::install("MSnbase")
 ```
@@
 discuss any suggestions. The
-[`MSnbase` development vignette](http://lgatto.github.io/MSnbase/articles/MSnbase-development.html) gives
+[`MSnbase` development vignette](http://lgatto.github.io/MSnbase/articles/v05-MSnbase-development.html) gives
 some background on the class infrastructure.

### 2019-04-10 R/functions-Chromatograms.R
remove NAnnotatedDataFrame

@@ -34,7 +34,7 @@
 #' @param data A \code{list} of \code{\link{Chromatogram}} objects.
 #'
 #' @param phenoData either a \code{data.frame}, \code{AnnotatedDataFrame} or
-#'     \code{NAnnotatedDataFrame} describing the phenotypical information of the
+#'     \code{AnnotatedDataFrame} describing the phenotypical information of the
 #'     samples.
 #'
 #' @param featureData either a \code{data.frame} or \code{AnnotatedDataFrame}
@@ -60,8 +60,6 @@ Chromatograms <- function(data, phenoData, featureData, ...) {
     ## Convert phenoData...
     if (is(phenoData, "data.frame"))
         phenoData <- AnnotatedDataFrame(phenoData)
-    if (is(phenoData, "AnnotatedDataFrame"))
-        phenoData <- as(phenoData, "NAnnotatedDataFrame")
     if (missing(featureData))
         featureData <- annotatedDataFrameFrom(datmat, byrow = TRUE)
     if (nrow(datmat) != nrow(featureData))

### 2019-04-10 R/methods-Spectra.R

@@ -557,7 +557,7 @@ setAs("Spectra", "MSnExp", function(from) {
     assaydata <- list2env(as(from, "list"))
     lockEnvironment(assaydata, bindings = TRUE)
     new("MSnExp", assayData = assaydata,
-        phenoData = new("NAnnotatedDataFrame", pd), featureData = fd,
+        phenoData = new("AnnotatedDataFrame", pd), featureData = fd,
         processingData = process)
 })
 

### 2019-04-10 tests/testthat/test_NAnnotatedDataFrame.R

@@ -1,22 +1,0 @@
-context("NAnnotatedDataFrame")
-
-test_that("NAnnotatedDataFrame validity", {
-    na <- new("NAnnotatedDataFrame")
-    expect_true(validObject(na))
-    dnms <- c(rowNames = 0, columnNames = 0, multiNames = 1)
-    expect_true(all.equal(dim(na), dnms))
-    pd <- data.frame(pvarA = "A", pvarB = "B")
-    na <- new("NAnnotatedDataFrame", data = pd) 
-    expect_true(validObject(na))
-    dnms <- c(rowNames = 1, columnNames = 2, multiNames = 1)
-    expect_true(all.equal(dim(na), dnms))
-    na <- new("NAnnotatedDataFrame",
-              multiplex = 4, multiLabels = reporterNames(iTRAQ4),
-              data = pd)
-    expect_true(validObject(na))
-    dnms <- c(rowNames = 1, columnNames = 2, multiNames = 4)
-    expect_true(all.equal(dim(na), dnms))
-
-    expect_true(multiplex(na) == 4)
-    expect_true(all.equal(multiLabels(na), reporterNames(iTRAQ4)))
-})

### 2019-04-10 tests/testthat/test_functions-Chromatograms.R

@@ -28,11 +28,11 @@ test_that("Chromatograms works", {
     ## Chromatograms with a phenoData.
     pheno <- AnnotatedDataFrame(data.frame(idx = 1:4, name = letters[1:4]))
     chs <- Chromatograms(ch_list, nrow = 2, phenoData = pheno)
-    expect_equal(chs@phenoData, as(pheno, "NAnnotatedDataFrame"))
+    expect_equal(chs@phenoData, as(pheno, "AnnotatedDataFrame"))
     expect_equal(colnames(chs), as.character(1:ncol(chs)))
     rownames(pheno) <- letters[1:4]
     chs <- Chromatograms(ch_list, nrow = 2, phenoData = pheno)
-    expect_equal(chs@phenoData, as(pheno, "NAnnotatedDataFrame"))
+    expect_equal(chs@phenoData, as(pheno, "AnnotatedDataFrame"))
     expect_equal(colnames(chs), letters[1:4])
     ## Chromatograms with a featureData.
     fd <- data.frame(name = c("a", "b"), mzmin = 1:2)
@@ -58,7 +58,7 @@ test_that(".validChromatograms works", {
     ch_list <- list(ch, ch, ch, ch, ch, ch, ch, ch)
     chs <- Chromatograms(ch_list, nrow = 2)
     expect_true(MSnbase:::.validChromatograms(chs))
-    chs@phenoData <- as(AnnotatedDataFrame(), "NAnnotatedDataFrame")
+    chs@phenoData <- as(AnnotatedDataFrame(), "AnnotatedDataFrame")
     expect_error(validObject(chs))
     chs <- Chromatograms(ch_list, nrow = 2)
     chs@featureData <- AnnotatedDataFrame()

### 2020-05-15 README.Rmd
fix: link to demo vignette
doc: add citation and refs to 2020 MSnbase2 preprint

@@
 To get an overview of the package, see the
-[*MSnbase-demo*](https://lgatto.github.io/MSnbase/articles/MSnbase-demo.html) vignette. More
-vignettes are available in the *Articles* tab.
+[*MSnbase-demo*](http://lgatto.github.io/MSnbase/articles/v01-MSnbase-demo.html)
+vignette. More vignettes are available in the *Articles* tab.
 
@@
 
+> *`MSnbase`, efficient and elegant R-based processing and
+> visualisation of raw mass spectrometry data*. Laurent Gatto,
+> Sebastian Gibb, Johannes Rainer. bioRxiv 2020.04.29.067868; doi:
+> https://doi.org/10.1101/2020.04.29.067868
+
 ## Contributing
